### 2015-10-15 reasoning/huiswerk/inl1019.tex
tried beginning with one but having troubles

\documentclass{article}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

\begin{document}
\author{Jappie Klooster}
\title{On proving things}
\maketitle

\section{Proof GKC rules}
\subsection{$L\to$}
To proof that if one of the premises of $L\to$ of $GKC$ is not true
(under all valuations) than the conclusion not true either.

The definition of $L\to$ in $GKC$:

\begin{scprooftree}{1}
	\AxiomC{$\Gamma \Rightarrow A, \Delta$}
	\AxiomC{$\Gamma, B \Rightarrow \Delta$}
	\BinaryInfC{$\Gamma, A \to B \Rightarrow \Delta$}
\end{scprooftree}

Let $I(x)$ be the interpretation function that determines if sequent x
is true.

Assume one of the premises is false while the conclusion remains true.
So $I(\Gamma \Rightarrow A, \Delta) = 0$
$I(\Gamma, A \to B \Rightarrow \Delta) = 1$

To get $I((\Gamma \Rightarrow A, \Delta) = 0$ if $\Gamma$ should have 
the value of $\neg A$
\subsection{$R\to$}
\end{document}

### 2015-10-15 reasoning/huiswerk/inl1019.tex
finished 1.1

@@
 
-Let $I(x)$ be the interpretation function that determines if sequent x
+Let $V(x)$ be the valuation function that determines if sequent x
 is true.
 
-Assume one of the premises is false while the conclusion remains true.
-So $I(\Gamma \Rightarrow A, \Delta) = 0$
-$I(\Gamma, A \to B \Rightarrow \Delta) = 1$
+Let $\Gamma \Rightarrow A, \Delta$ be $P_1$,
+let $\Gamma, B \Rightarrow \Delta$ be $P_2$ and
+let $\Gamma, A \to B \Rightarrow \Delta$ be $C$
 
-To get $I((\Gamma \Rightarrow A, \Delta) = 0$ if $\Gamma$ should have 
-the value of $\neg A$
+To proof $V(P_1) = 0$ or
+$V(P_2) = 0$ then
+$V(C) = 0$. We will try to proof this by searching for a case
+where one of the $P$'s is false and $C$ is true.
+
+Both $V(P_1)$ and $V(P_2)$ can only be 0 if the result of
+$V(\Gamma) \not= V(\Delta)$
+so we assume $V(\Gamma) = 1$ and $V(\Delta) = 0$.
+
+If $V(A) = 1$ in this case $V(P_1)$ becomes true so $V(P_2)$ should be false.
+The only way to make $P_2$ false is by having $V(B) = 1$.
+But once you do this $V(A \to B) = 1$ which means $V(C) = 0$
+
+The other option is to make $V(A) = 0$ Now $V(P_1)$ is false, so $V(P_2)$ can
+be true or false. The value of $B$ doesn't matter because in either case
+$V(A \to B) = 1$ which means $V(C)$ will be false.
+
+The same technique can be applied for so we assume $V(\Gamma) = 0$ and $V(\Delta) = 1$.
+
 \subsection{$R\to$}

### 2015-10-15 reasoning/huiswerk/inl1019.tex
started with 2a

@@
 	\AxiomC{$\Gamma, B \Rightarrow \Delta$}
+	\RightLabel{$L\to$}
 	\BinaryInfC{$\Gamma, A \to B \Rightarrow \Delta$}
@@
 is true.
-
 Let $\Gamma \Rightarrow A, \Delta$ be $P_1$,
@@
 
-The same technique can be applied for so we assume $V(\Gamma) = 0$ and $V(\Delta) = 1$.
+The same technique can be applied for $V(\Gamma) = 0$ and $V(\Delta) = 1$.
+
+There are no other valuations left, so if one of the premises of L $\to$ is
+false then so is the conclusion
 
 \subsection{$R\to$}
+To proof that if one of the premises of $R\to$ of $GKC$ is not true
+(under all valuations) than the conclusion not true either.
+
+The definition of $R\to$ in $GKC$:
+
+\begin{scprooftree}{1}
+	\AxiomC{$\Gamma, A \Rightarrow B, \Delta$}
+	\RightLabel{$R\to$}
+	\UnaryInfC{$\Gamma \Rightarrow A \to B, \Delta$}
+\end{scprooftree}
+Let $V(x)$ be the valuation function that determines if sequent x
+is true.
+Let $\Gamma, A \Rightarrow B, \Delta$ be $P$ and
+let $\Gamma \Rightarrow A \to B, \Delta$ be $C$
+
+$V(P)$ can only be 0 if the result of
+$V(\Gamma) \not= V(\Delta)$
+
+If $V(\Gamma) = 1$, $V(A)$ also has to be 1 otherwise the premise would be true.
+$V(B)$ has to be 0 because otherwise the premise would be true.
+
+With these valuation $V(C)$ will be false.
+
+The same result for $V(\Gamma) = 0$
+
+There are no other valuations left, so if one of the premises of R $\to$ is
+false then so is the conclusion
+
+\section{Positivity and negativity}
+\subsection{Prove C keeps occuring in premise and conclusion}
+To prove that if formula $C$ occurs positively in one of the premises of a rule,
+then the same formula occurs positively in the conclusion fro all rules of GKC
+
+This will be done on a rule by rule basis.
+\subsubsection{$R\wedge$}
+Suppose C occurs positively in one of the premisses
+($\Gamma \Rightarrow A, \Delta$) or ($\Gamma \Rightarrow B,\Delta$),
+say in the first one. Thus it occurs positively in a formula in $\Delta$ or
+$A$ or negatively in a formula in $\Gamma$. In the first and the last case,
+the same holds for the conclusion ($\Gamma \Rightarrow A \wedge B, \Delta)$.
+In case C occurs positively in $A$ then by definition it occurs positively
+in $A \wedge B$ (even in the case that $C = A$). Hence it occurs positively
+in ($\Gamma \Rightarrow A \wedge B, \Delta$). This is what had to be shown.
+
 \end{document}

### 2015-10-16 reasoning/huiswerk/inl1019.tex
finsihed 2a, yes all the texts

@@
 To prove that if formula $C$ occurs positively in one of the premises of a rule,
-then the same formula occurs positively in the conclusion fro all rules of GKC
+then the same formula occurs positively in the conclusion for all rules of GKC
 
 This will be done on a rule by rule basis.
+
 \subsubsection{$R\wedge$}
@@
 $A$ or negatively in a formula in $\Gamma$. In the first and the last case,
-the same holds for the conclusion ($\Gamma \Rightarrow A \wedge B, \Delta)$.
+the same holds for the conclusion ($\Gamma \Rightarrow A \wedge B, \Delta$).
 In case C occurs positively in $A$ then by definition it occurs positively
@@
 
+\subsubsection{$L\vee$}
+Suppose C occurs positively in one of the premisses
+($\Gamma, A \Rightarrow \Delta$) or ($\Gamma, B \Rightarrow \Delta$),
+say in the first one. Thus it occurs positively in a formula in $\Delta$ or 
+negatively in a formula in $\Gamma$ or $A$. In the first and the last case,
+the same holds for the conclusion ($\Gamma, A \vee B \Rightarrow \Delta$).
+In case C occurs negatively in $A$ then by definition it occurs negatively
+in $A \vee B$ (even in the case that $C = A$). Hence it occurs negatively
+in ($\Gamma, A \vee B \Rightarrow \Delta$). This is what had to be shown.
+
+\subsubsection{$L\to$}
+Suppose C occurs positively in one of the premisses
+($\Gamma \Rightarrow A, \Delta$) or ($\Gamma, B \Rightarrow \Delta$),
+say in the first one. Thus it occurs positively in a formula in $\Delta$ or
+$A$ or negatively in a formula in $\Gamma$. In the first and the last case,
+the same holds for the conclusion ($\Gamma A \to B \Rightarrow \Delta$).
+In case C occurs positively in $A$ then by definition it occurs positively
+in $A \to B$ (even in the case that $C = A$). Hence it occurs positively
+in ($\Gamma A \to B \Rightarrow \Delta$). This is what had to be shown.
+
+\subsubsection{$L\wedge$}
+Suppose C occurs positively the premiss
+($\Gamma, A, B \Rightarrow \Delta$). Thus it occurs positively in a formula
+in $\Delta$ or negatively in a formula in $\Gamma$, $A$ or $B$.`
+In the first and the last case,
+the same holds for the conclusion ($\Gamma, A \wedge B \Rightarrow \Delta$).
+In case C occurs negatively in $A$ then by definition it occurs negatively
+in $A \wedge B$ (even in the case that $C = A$). Hence it occurs negatively
+in ($\Gamma, A \wedge B \Rightarrow \Delta$). This is what had to be shown.
+
+\subsubsection{$R\vee$}
+Suppose C occurs positively the premiss
+($\Gamma \Rightarrow A, B, \Delta$). Thus it occurs positively in a formula
+in $\Delta$, $A$ or $B$ or negatively in a formula in $\Gamma$.
+In the first and the last case,
+the same holds for the conclusion ($\Gamma \Rightarrow A \vee B, \Delta$).
+In case C occurs positively in $A$ then by definition it occurs positively
+in $A \vee B$ (even in the case that $C = A$). Hence it occurs positively
+in ($\Gamma, A \wedge B \Rightarrow \Delta$). This is what had to be shown.
+
+\subsubsection{$R\to$}
+Suppose C occurs positively the premiss
+($\Gamma, A \Rightarrow B, \Delta$). Thus it occurs positively in a formula
+in $\Delta$ or $B$ or negatively in a formula in $\Gamma$ or $A$.
+In the first and the last case,
+the same holds for the conclusion ($\Gamma \Rightarrow A \to B, \Delta$).
+In case C occurs negatively in $A$ then by definition it occurs negatively
+in $A \to B$ (even in the case that $C = A$). Hence it occurs negatively
+in ($\Gamma, A \to B \Rightarrow \Delta$). This is what had to be shown.
+
+\subsubsection{$L\neg$}
+Suppose C occurs positively the premiss
+($\Gamma \Rightarrow A, \Delta$). Thus it occurs positively in a formula
+in $\Delta$ or $A$ or negatively in a formula in $\Gamma$.
+In the first and the last case,
+the same holds for the conclusion ($\Gamma, \neg A \Rightarrow \Delta$).
+In case C occurs positively in $A$ then by definition it occurs positively
+in $\neg A$ (even in the case that $C = A$). Hence it occurs positively
+in ($\Gamma, \neg A \Rightarrow \Delta$). This is what had to be shown.
+
+\subsubsection{$R\neg$}
+Suppose C occurs positively the premiss
+($\Gamma, A \Rightarrow \Delta$). Thus it occurs positively in a formula
+in $\Delta$ or negatively in a formula in $\Gamma$ or $A$.
+In the first and the last case,
+the same holds for the conclusion ($\Gamma \Rightarrow \neg A, \Delta$).
+In case C occurs negatively in $A$ then by definition it occurs negatively
+in $\neg A$ (even in the case that $C = A$). Hence it occurs negatively
+in ($\Gamma \Rightarrow \neg A, \Delta$). This is what had to be shown.
+\subsection{Proof introduction of $A\wedge B$ remains positive}
 \end{document}

### 2015-10-16 reasoning/huiswerk/inl1019.tex
finshed all of 2 I think....

@@
 in ($\Gamma \Rightarrow \neg A, \Delta$). This is what had to be shown.
-\subsection{Proof introduction of $A\wedge B$ remains positive}
+
+\subsection{Proof introduction remains positive}
+
+\subsubsection{$A\wedge B$}
+To prove that in any proof in GKC of a sequent $S$, if the formula
+$A \wedge B$ is introduced in an application of the $R\wedge$ rule,
+then that $A \wedge B$ occurs positively in the endsequent $S$of the
+proof.
+
+Let $A \wedge B$ be C, as shown in 2.1 if formula C occurs positively in
+one of the premises of a rule, then the same formula occurs positively in
+the conclusion for all rules of GKC, now we have proven that C will
+occur in the endsequent.
+
+\subsubsection{$A\vee B$}
+To prove that in any proof in GKC of a sequent $S$, if the formula
+$A \vee B$ is introduced in an application of the $R\vee$ rule,
+then that $A \vee B$ occurs positively in the endsequent $S$of the
+proof.
+
+Let $A \vee B$ be C, as shown in 2.1 if formula C occurs positively in
+one of the premises of a rule, then the same formula occurs positively in
+the conclusion for all rules of GKC, now we have proven that C will
+occur in the endsequent.
+
+\subsection{Single branch}
+To prove that every provable sequent that does not contains negations and
+implications and in which every subformula of the form $A \wedge B$ occurs
+negatively and every subformula of the form $A \vee B$ occurs positively,
+has a proof in GKC that consists of a single branch.
+
+Because the sequent does not contain negations and implications the rules
+$R\to$, $L\to$, $R\neg$ and $L\neg$ can't be applied.
+$R\wedge$ can't be applied because $A \wedge B$ occurs negatively,
+which by definition means $A \wedge B$ occurs on the $L$ side of the
+$\wedge$ rules.
+$L\vee$ can't be
+applied because $A \vee B$ occurs positively, which by definition means
+that the $A \vee B$ formula is on the $R$ side of the $\vee$ rules.
+
+The only applicable rules left in GKC are the $L\wedge$ and $R\vee$ rules.
+These are single branch by definition. Thus this entire proof has a single
+branch, this is what had to be shown.
+
+\subsection{$S$ consists of \ldots}
+Supose that $S$ is a provable sequent that does not contain negations and
+implications and which every subformula of the form $A\wedge B$ occurs
+negatively and every subformula of the form $A \vee B$ occurs positively.
+Show that if the sum of all the occurrences of connectives in $S$ is $n$.
+then $S$ has a proof in $GKC$ that consists of at most $n + 1$ sequents.
+
+Because the sequent $S$ does not contain negations and implications the rules
+$R\to$, $L\to$, $R\neg$ and $L\neg$ can't be applied.
+$R\wedge$ can't be applied because $A \wedge B$ occurs negatively,
+which by definition means $A \wedge B$ occurs on the $L$ side of the
+$\wedge$ rules.
+$L\vee$ can't be
+applied because $A \vee B$ occurs positively, which by definition means
+that the $A \vee B$ formula is on the $R$ side of the $\vee$ rules.
+
+The only applicable rules left in GKC are the $L\wedge$ and $R\vee$ rules.
+Both the $L\wedge$ and $R\vee$ rules add only 1 connective to the conclusion
+once applied to the premise, creating a new sequent in the process. The proof
+of $S$ also has just a single branch, because $L\wedge$ and $R\vee$ only allow
+a single premise.
+So the
+sequent count in the proof of $S$ is at least $n$. It is $n$ + 1 because $S$ is also
+a sequent by definition. This is what had to be shown.
 \end{document}

### 2015-10-17 reasoning/huiswerk/inl1019.tex
finished 3??

@@
 \usepackage{pifont}
+\usepackage{qtree}
 \newenvironment{scprooftree}[1]%
@@
 This will be done on a rule by rule basis.
+
+Everywhere where postive is used negative can also be used as long
+that everywhere where negative is used is replaced by positive.
 
@@
 a sequent by definition. This is what had to be shown.
+\section{Proof of $GK_{\{\neg, \wedge\}}$ is complete}
+Let $GK_{\{\neg, \wedge\}}$ be the calcules consisting of the axiom
+$(\Gamma, A \Rightarrow A, \Delta)$ where $A$ is an atom, and the rules:
+
+\begin{prooftree}
+	\AxiomC{$\Gamma \Rightarrow A, \Delta$}
+	\RightLabel{$L\neg$}
+	\UnaryInfC{$\Gamma, \neg A \Rightarrow \Delta$}
+\end{prooftree}
+\begin{prooftree}
+	\AxiomC{$\Gamma, A \Rightarrow \Delta$}
+	\RightLabel{$R\neg$}
+	\UnaryInfC{$\Gamma \Rightarrow \neg A, \Delta$}
+\end{prooftree}
+
+\begin{prooftree}
+	\AxiomC{$\Gamma, A, B \Rightarrow \Delta$}
+	\RightLabel{$L\wedge$}
+	\UnaryInfC{$\Gamma, A \wedge B \Rightarrow \Delta$}
+\end{prooftree}
+
+\begin{prooftree}
+	\AxiomC{$\Gamma \Rightarrow A, \Delta$}
+	\AxiomC{$\Gamma, \Rightarrow B, \Delta$}
+	\RightLabel{$R\wedge$}
+	\BinaryInfC{$\Gamma \Rightarrow A \wedge B, \Delta$}
+\end{prooftree}
+To prove that $GK_{\{\neg, \wedge\}}$ is complete with respect to
+$CPC_{\{\neg, \wedge\}}$.
+
+The completness of $GK_{\{\neg, \wedge\}}$ can be proven by showing
+that If $\nvdash S$, then $v(I(S)) = 0$
+
+$\Gamma \cup \Delta$  consists of atoms and $\Gamma \cap \Delta \not= \o$
+
+Valuation $v(p) = 1$  if $p \in \Gamma$ and $v(p) = 0$ if $p \in \Delta$
+
+$S$ should not be derivable, so there is a leaf that is not an axiom.
+This means $V(I(l)) = 0$ where l is the leaf.
+
+If the valuation of the premise is 0, then so is the conclusion for any of
+the rules in $GK_{\{\neg, \wedge\}}$. This is shown in 2.1.
+
+So $V(I(S))$ is also 0 because one of the leafs is 0, which eventually
+will find its way to the final conclusion.
+
 \end{document}

### 2015-10-24 reasoning/huiswerk/inl1026.tex
started with the homework of 26th

\documentclass{article}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\usepackage{qtree}
\usepackage{tikz}

\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

\begin{document}
\author{Jappie Klooster}
\title{On proving things}
\maketitle

\section{Prove $R_k$}
Prove that the rule $R_k$ of the calculus $GKC_k$ is sound with respect to
Kripke models: If $I(\Gamma \Rightarrow A)$ holds in all Kripke models, then
so does $I(\Pi, \Box \Gamma \Rightarrow \Box A, \Delta)$

\begin{prooftree}
	\AxiomC{$\Gamma \Rightarrow A$}
	\RightLabel{$R_k$}
	\UnaryInfC{$\Pi, \Box \Gamma \Rightarrow \Box A, \Delta$}
\end{prooftree}

Take an arbitrary kripke model $C$. If $C$ $I(\Gamma \Rightarrow A)=0$ then
so does $I(\Pi, \Box \Gamma \Rightarrow \Box A, \Delta) = 0$ For example:

\begin{tikzpicture}
[scale=.8,auto=left,every node/.style={circle,fill=blue!20}]
\draw (3,3) circle (3.5);
\node (n1) at (1,3) {$A \Rightarrow A$};
\node (n2) at (3,5) {$B \Rightarrow A$};
\node (n3) at (4,2)  {$C \wedge B, A \Rightarrow A$};
\foreach \from/\to in {n1/n2,n2/n3}
\draw (\from) -- (\to);

\end{tikzpicture}

, if
 If $C$ $I(\Gamma \Rightarrow A)=1$ then
so does $I(\Pi, \Box \Gamma \Rightarrow \Box A, \Delta) = 1$ for example:
\begin{tikzpicture}
[scale=.8,auto=left,every node/.style={circle,fill=blue!20}]
\draw (3,3) circle (3.5);
\node (n1) at (1,3) {$A \Rightarrow A$};
\node (n2) at (3,5) {$B, A \Rightarrow A$};
\node (n3) at (4,2)  {$C \wedge B, A \Rightarrow A$};
\foreach \from/\to in {n1/n2,n2/n3}
\draw (\from) -- (\to);

\end{tikzpicture}

Therefore $R_k$ is sound.

% I just don't know how to explain this crap.

\section{}
\subsection{Prove $\Box p \Rightarrow \Box \Box p$}
Prove that the sequent $\Box p \Rightarrow \Box \Box p$, where $p$ is a
propositional variable, is not derivable in $GKC_k$.

A propositional variable is either true or false (boolean). To prove 
$\Box p \Rightarrow \Box \Box p$ is not derivable in $GKC_k$ we assume
it is derivable and try to construct a prove towards it from the axioms.

The Ax ($\Gamma, A \Rightarrow A \Delta$) Axiom cannot be used because it
doesn not contain propositional vraibles. So we're left with the 
$L\bot(\Gamma \bot \Rightarrow \Delta)$ and $R\top(\Gamma \top
\Rightarrow \Delta)$ axioms
\subsubsection{$L\bot$}
\begin{prooftree}
\end{prooftree}

\subsection{Prove $\Box A \to \Box \Box A$ }
Prove that $\Box A \to \Box \Box A$ holds in all transitive Kripke models.

\subsection{Explain why 2.1 and 2.2 imply incompletness}
Explain why 2a and 2b imply that K is not complete with respect to
transitive Kripke models.
\section{Prove, for any propositional}
Let G denote the calculus GKC extended by the rule
\begin{prooftree}
	\AxiomC{$\Box \Gamma, \Gamma \Rightarrow A$}
	\RightLabel{$R$}
	\UnaryInfC{$\Box \Gamma \Rightarrow \Box A$}
\end{prooftree}
Prove, for any variable $p$, that $\Box p \Rightarrow \Box \Box p$ is
not provable in G. Prove that not every sequent provable in $GKC_{K4}$
is provable in G. 
\subsection{Does the converse hold?}
%Yes or no, maybe a because
\section{Prove, using soundness}
Prove using sondness, that $GKC_K$ does not derive $(\Box A \Rightarrow A)$.

\section{Prove, that if $\vdash_{GKC_K}$}
Prove that if $\vdash_{GKC_K}(\Rightarrow \Box A)$, then 
$\vdash_{GKC_K}(\Rightarrow A)$

\end{document}

### 2015-10-24 reasoning/huiswerk/inl1026.tex
added more texts...

@@
 doesn not contain propositional vraibles. So we're left with the
-$L\bot(\Gamma \bot \Rightarrow \Delta)$ and $R\top(\Gamma \top
+$L\bot(\Gamma, \bot \Rightarrow \Delta)$ and $R\top(\Gamma \top
 \Rightarrow \Delta)$ axioms
 \subsubsection{$L\bot$}
+$(\Gamma, \bot \Rightarrow \bot, \Delta)$ will be used as starting point
+because it is a valid axiom, and the p needs to be on both sides.
 \begin{prooftree}
+	\AxiomC{$\Gamma, \bot \Rightarrow \bot, \Delta$}
+	\RightLabel{$R_k$}
+	\UnaryInfC{$\Gamma, \Box \bot \Rightarrow \Box \bot, \Delta$}
 \end{prooftree}
+There are after this point however no rules available to add a single box
+on 1 side, so the $L\bot$ axiom can't be used to derive
+\subsubsection{$L\top$}
+$(\Gamma, \top \Rightarrow \top, \Delta)$ will be used as starting point
+because it is a valid axiom, and the p needs to be on toph sides.
+\begin{prooftree}
+	\AxiomC{$\Gamma, \top \Rightarrow \top, \Delta$}
+	\RightLabel{$R_k$}
+	\UnaryInfC{$\Gamma, \Box \top \Rightarrow \Box \top, \Delta$}
+\end{prooftree}
+There are after this point however no rules available to add a single box
+on 1 side, so the $L\top$ axiom can't be used to derive
+$\Box p \Rightarrow \Box \Box p$
+
+\subsubsection{Conclusion}
+Because there are no axioms available for the sequent
+($\Box p \Rightarrow \Box \Box p$) to work towards, it has been
+proven that the sequant is not derivable.
 
@@
 
+Transitivity means $uRw$ and $vRw$ then $uRw$. So if $\Box A$, A is in every
+state. Thus you know that every state is accessible to another state because
+its transitive. So you can say $\Box \Box A$
+
 \subsection{Explain why 2.1 and 2.2 imply incompletness}
@@
 transitive Kripke models.
+
+Becasue if you create a sequent of 2.2 like this:
+$\Rightarrow \Box A \to \Box \Box A$ and apply the $L\to$ opperation,
+the formula is essentlially the same as 2.1. But in 2.1 we've shown
+that the formula is *not* derivable. So we have some semantic which is
+true but a system that says its not true, which indicates incompleteness.
+
 \section{Prove, for any propositional}
@@
 \end{prooftree}
-Prove, for any variable $p$, that $\Box p \Rightarrow \Box \Box p$ is
-not provable in G. Prove that not every sequent provable in $GKC_{K4}$
+\subsection{Prove p}
+Prove, for any propositional variable $p$,
+that $\Box p \Rightarrow \Box \Box p$ is not provable in G.
+
+The Ax ($\Gamma, A \Rightarrow A \Delta$) Axiom cannot be used because it
+doesn not contain propositional vraibles. So we're left with the
+$L\bot(\Gamma, \bot \Rightarrow \Delta)$ and $R\top(\Gamma \top
+\Rightarrow \Delta)$ axioms
+\subsubsection{$L\top$}
+$(\Gamma, \top \Rightarrow \top, \Delta)$ will be used as starting point
+because it is a valid axiom, and the p needs to be on toph sides.
+\begin{prooftree}
+	\AxiomC{}
+	\RightLabel{$L\top$}
+	\UnaryInfC{$\Box \Box \top, \Box \top, \top \Rightarrow \top$}
+	\RightLabel{$R$}
+	\UnaryInfC{$\Box \Box \top, \Box \top\Rightarrow \Box \top$}
+	\RightLabel{$R$}
+	\UnaryInfC{$\Box \Box \top\Rightarrow \Box \Box \top$}
+\end{prooftree}
+There are after this point however no rules available to remove a single box
+on the left side, so the $L\top$ axiom can't be used to derive
+$\Box p \Rightarrow \Box \Box p$.
+
+\subsubsection{$L\bot$}
+$(\Gamma, \bot \Rightarrow \bot, \Delta)$ will be used as starting point
+because it is a valid axiom, and the p needs to be on both sides.
+\begin{prooftree}
+	\AxiomC{}
+	\RightLabel{$L\bot$}
+	\UnaryInfC{$\Box \Box \bot, \Box \bot, \bot \Rightarrow \bot$}
+	\RightLabel{$R$}
+	\UnaryInfC{$\Box \Box \bot, \Box \bot\Rightarrow \Box \bot$}
+	\RightLabel{$R$}
+	\UnaryInfC{$\Box \Box \bot\Rightarrow \Box \Box \bot$}
+\end{prooftree}
+There are after this point however no rules available to remove a single box
+on the left side, so the $L\bot$ axiom can't be used to derive
+$\Box p \Rightarrow \Box \Box p$.
+
+\subsubsection{Conclusion}
+Because there are no axioms available for the sequent
+($\Box p \Rightarrow \Box \Box p$) to work towards, it has been
+proven that the sequant is not derivable.
+
+\subsection{Prove s from $GKC_{K4}$ in $G$}
+Prove that not every sequent provable in $GKC_{K4}$
 is provable in G.
+
+To proof this one sequent will be shown to be provable in $GKC_{K4}$
+and then will be shown this sequant is not provable in $G$
+
+G = GKC and:
+\begin{prooftree}
+	\AxiomC{$\Box \Gamma, \Gamma \Rightarrow A$}
+	\RightLabel{$R$}
+	\UnaryInfC{$\Box \Gamma \Rightarrow \Box A$}
+\end{prooftree}
+
+$GKC_{K4}$ = GKC and:
+\begin{prooftree}
+	\AxiomC{$\Box \Gamma, \Gamma \Rightarrow A$}
+	\RightLabel{$R$}
+	\UnaryInfC{$\Pi, \Box \Gamma \Rightarrow \Box A, \Delta$}
+\end{prooftree}
+
+As an intial observation the rules of $GKC_{K4}$ and $G$ are functionally
+the same. They're both intendid to remove the box on the right side and to
+hide it away in the left side in one of the `garbage' spots $(\Pi$ and 
+$\Delta)$ of the axioms.
+
+However the $GKC_{K4}$ has also garbage spots in the conclusion. So to show
+something is proofable in $GKC_{K4}$ and not in $G$ all you need to do is
+fill in one of those garbage spots. Consider the following sequant:
+
+\[A, \Box B \Rightarrow \Box B, A\]
+
+This can be solved in $GKC_{K4}$ in the following way:
+
+\begin{prooftree}
+	\AxiomC{}
+	\RightLabel{Ax}
+	\UnaryInfC{$\Box B, B \Rightarrow B$}
+	\RightLabel{$R$}
+	\UnaryInfC{$A, \Box B \Rightarrow \Box B, A$}
+\end{prooftree}
+
+However this sequant is not solvable in G, the rule for dealing with the 
+boxes cannot match the `garbage'
+
 \subsection{Does the converse hold?}
+yes because the `garbage' spots are allowed to be empty.
 %Yes or no, maybe a because
 \section{Prove, using soundness}
-Prove using sondness, that $GKC_K$ does not derive $(\Box A \Rightarrow A)$.
+Prove using soundness, that $GKC_K$ does not derive $(\Box A \Rightarrow A)$.
+
+$(\Box A \Rightarrow A)$ can't be derived in $GKC_K$ because
+$(\Box A \Rightarrow A)$ isn't valid in kripke. $\Box A$ inidicates that the
+state should be able to get to A. But it does not mean that the current
+state has an $A$ as this sequance indicates. So $(\Box A \Rightarrow A)$
+shouldn't be derivable in $GKC_K$ otherwise $GKC_K$ wouldn't be sound.
+This is what thad to be shown.
 
 \section{Prove, that if $\vdash_{GKC_K}$}
+
 Prove that if $\vdash_{GKC_K}(\Rightarrow \Box A)$, then

### 2015-10-27 reasoning/summery.tex
started to write a big big summery

\documentclass{article}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\usepackage{qtree}
\usepackage{tikz}
\usepackage{listings}
\usepackage{hyperref}
\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

\begin{document}
\lstset{language=Java}
\author{Jappie Klooster}
\title{A Summery about reasoning with logic}
\maketitle

\section{Introduction}
This summery is written for people that try to understand logicians on a master
level *but* they're comming from `hbo informatica'

\section{On proofing things}
If you were lucky like me you'll be asked to write a proof in the first week
which will be graded. I got a terible grade, but learning I did.

So there are 2 kinds of proving. The first one is writing a proof in the system
this should be rather easy, although you have to be carefull to follow the
rules to the letter. I'm going to ignore this one since its just using the
rules, which you've done tons of times as a programmer, so it shouldn't be
that strange.

The second one is the more troublsome, its proving the system itself. This
is also split up into 3 different aspects. They all have in commen that
it should be done in natural langauge. But, there is a structure to it
I'm going to try and give a complete overview of this structure.

First you have soundness, then there is completness, and finally you can be
asked to proof certain properties of a system.

\subsection{functions}
Proving things is sometimes done with a function,
now you should be familiar with functions.
But they are defined in natural langauge, which may seem a little starnge
so as an example I'll define the valuation as first time I came accross it:

\[v:\mathcal{F}_{prp} \to \{0,1\}\]

the prototype in Java is like this (I assume you never can be sure of course,
see \autoref{notations}):

\begin{lstlisting}[frame=single] 
	public boolean V(Formula f);
\end{lstlisting}

Now in logic you describe the datatypes in natural langauge, so first they will
say `let', `define' or ` v is a function' \ldots

The other function I came across was the I function. Which stands for 
interpation, what it does is just replacing the commas for the propper
connectives.

\section{proving soundness}
This basically means your system doesn't modify the value you put in.
If you put in truth, apply a bunch of rules from your system, then you should
get out truth.

\subsection{Structure of your proof}
First write down your overal strategy of the proof \autoref{strategies}. For
soundness this is often `proof by induction' \autoref{induction}.

Then write down the lemmas you're gonna use. A lemma is a little proof, often
derived from the properties of the system you're proving. You could see it
as a basic argument. When proving soundness you want to connect this to the
semantics, for classical systems the semantics are truth tables, for kripke
models it are kripke frames. The semantics are the thing your talking about
in your logical system.
%I can see how I dind't understand all this crap the first time I saw it,
% its freaking packed with definitions, and even if you understand all the
% definiontions, its still hard
Make sure that you've proven the soundness of the axioms *and* of the rules
with help of the semantics.

Finally ty up the lemmas together in the conclusion, this is done with help
of the proofing strategy.

Finish with a sentence like: `This had to be shown'.

\section{proving strategies}
\label{strategies}
There are many different ways of writing proofs, in here I'll sumerize the ones
I came accross during the course.

\subsection{Induction}
\label{induction}

\section{notations}
\label{notations}
There is no standart in notating things. Every logician seems to use a
slightly different notations. Keep this in mind. But I think the connectives
are mostly the same. You probably need to memorize these.

\section{Systems}
GKC, Gentzen sequent calculy.

\end{document}

### 2015-10-27 reasoning/summery.tex
added the connectives

@@
 as a basic argument. When proving soundness you want to connect this to the
-semantics, for classical systems the semantics are truth tables, for kripke
+semantics, for propositional systems the semantics are truth tables, for kripke
 models it are kripke frames. The semantics are the thing your talking about
@@
 
+\noindent
+\begin{tabular}{@{}lll@{}}
+Symbols&
+In natural language &
+Technical name \\ \toprule
+$\neg$ 			& not 				& negation \\
+$\wedge$,\& 	& and 				& conjunction \\
+$\vee$ 			& or 				& disjunction\\
+$\to$	& if \ldots then 	& implication\\
+$\leftrightarrow$	& if and only if 	& equivalent\\ \bottomrule
+\end{tabular}
+
+Then there are some other symbols used:
+
+\noindent
+\begin{tabular}{@{}ll@{}}
+Symbols&
+In natural language \\ \toprule
+$\vdash$ 					& Part of system \\
+$\vdash, \Rightarrow$ 		& Seperator between antecedent and succedent\\ \bottomrule
+\end{tabular}
+
 \section{Systems}
-GKC, Gentzen sequent calculy.
+\begin{itemize}
+	\item $H$  Hilbert, has a huge set of axioms, and only 1 rule. (also called fredge)
+	\item $CPC$ Clasicial Propositional logic, formulas in here are always true
+(tautologies)
+	\item $GKC$ Gentzen calculus, also called sequent calcules
+	\item $GKC_K$  Getnzen + R 
+\end{itemize}
 

### 2015-10-27 reasoning/summery.tex
added more texts, made a crazy text on how to proof

@@
 rules, which you've done tons of times as a programmer, so it shouldn't be
-that strange.
+that strange. Although I'm making an exception for Hilbert and Grammar,
+Hilbert isn't that obvious and Grammar isn't obvious either.
 
@@
 
-First you have soundness, then there is completness, and finally you can be
-asked to proof certain properties of a system.
+First you have soundness, then there is completness, these are both just
+properties of a system, so there are many more things to proof, but these
+are some of the most important because they decide if a system is usable.
 
@@
 
+The technical defintion is: The argument should be valid and all
+its premisses are true. Premisses in a system are the axioms.
+So they should be valid (valution function returns 1). The arguments
+in a system are the rules, so they should at least preserve truth.
+
 \subsection{Structure of your proof}
@@
 
+\section{proving completeness}
+Every validity should be proofable in the system. So there is usually no 
+obvious way of proving this, but for the system treated in the course
+there are some
+
 \section{proving strategies}
@@
 \label{induction}
+In here you try to prove something by starting with a single case n.
+
+For all n (*), n=1, supose (*) for n, then show (*) for n+1.
+
+So you start with an assumption, usually based on the question.
+For example \emph{show that if $\vdash_x A$ then A is a bird.}
+
+So you start with the base case, ie: \emph{We assume that $\vdash_x A$}
+You can do this because of propositional logic, whenever you see
+if x then y or x implies y you can just assume x. Continuing:
+\emph{Which means there is a proof $A_1,\ldots,A_n$ of A}.
+Here you're just saying that because A is part of X, there consists
+a bunch of steps that lead to A. \emph{We prove$A_i$ is a
+`birdlike property' with induction to $i$} Here you're saying
+that the steps that lead to A have all `birdlike properties' (or whatever
+property you're using) and you're gonna use induction to show that.
+\emph{As $A_n = A$ this implies A is a `birdlike poperty'}
+
+Now you're using the properties of the system and you call them lemmas.
+These are the arguments and you connect them to the semantic.
+\emph{Lemma 1, we have to proof that all axioms are `birdlike properties'
+	This is done by showing with help of the semantics of x, in case of
+system x this is a sleeve.
+Bellow you can see the sleeve of axiom D which clearly indicates that
+axiom D is birdlike
+}
+
+\[sleevefigure\]
+
+\emph{For axiom Z you can also see its birdlike:}
+\[(slightly altered) sleevefigure\]
+
+You do this for all the axioms.
+
+The next lemma is about the rules, these also have to be shown to keep
+track of birdlikeness in system x:
+
+\emph{Lemma 2, all the rules of x preserve birdlikeness:
+	For the rule argument ad hominum, supose that the precedents
+	are bird like
+}
+Precedents are the stuff you put into the rule.
+You can supose this because you've already shown in lemma 1 all the axioms are
+birdlike, so for the induction step it can't be anything else.
+\emph{
+	With the following sleeve it is shown that not under all valuations 
+	birdlikeness is preserved, but only for the ones were birdlikeness was
+	already there
+}
+\[(uninteresting) sleevefigure\]
+
+\emph{this finises the proof for lemma 2}
+
+Now the final step of induction is doing the induction step with help of the
+just proven lemmas.
+
+\subsection{tabloo or truth tree}
+In this strategy you create a tree, starting from the axioms which is either
+true or false, you go down step by step and say what happens. In most systems
+if one of the axioms is false it means the conclusion is false to.
 
@@
 In natural language \\ \toprule
-$\vdash$ 					& Part of system \\
+$\vdash$ 					& Part of system, ie, is derivable \\
 $\vdash, \Rightarrow$ 		& Seperator between antecedent and succedent\\ \bottomrule
@@
 	\item $GKC_K$  Getnzen + R 
+	\item $GKP$ Genzen + structural rules.
 \end{itemize}
 
+\section{Proving in Hilbert}
+Hilbert is the most simple system. However because of its simplicity proving
+things in it isn't very obvious. There are multiple ways of writing proovs
+
+\section{Proving in Grammar Logic}
+Grammar logic is a (failed) attempt to make sense of natural langauge. It
+tries to organize words into types, and groups of those words into compositional
+types.
+Connection words such as ` that' will have complex types.
+
 \end{document}

### 2015-10-28 agents/summery.tex
another sumarry for another part

\documentclass{article}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\usepackage{qtree}
\usepackage{tikz}
\usepackage{listings}
\usepackage{hyperref}
\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

\begin{document}
\lstset{language=Java}
\author{Jappie Klooster}
\title{Agents and you, a brief summary}
\maketitle

\section{Introduction}
This will be an introduction to agents coming from someone who followed
` hbo informatica' as pre study.

In here I will write down the things I think are important.

\section{Modal logic}

\section{APL}
Medhi dastani thought about the agent programming langauge (APL)

\subsection{Agent data structure}
This represents the mental state of a agent. \\
\begin{tabular}{ll}
	Name & Does what \\ \toprule
	Beliefs & Information available to the agent \\
	Goals & Objectives that the agent wants to achieve \\
	Events & Obsevations of (environmental) changes \\
	Capabilities & Actions that the agent can perform \\
	Plans & Procedures to achieve objectives \\
	Reasoning rules & Reason about goal and plans \\
\end{tabular} \\

\noindent
The rules can be:

\begin{itemize}
	\item Planning rules (goal $\to$ plan)
	\item Procedural rules (events $\to$ plan)
	\item Plan repair rules (plan $\to$ plan)
\end{itemize}

\subsection{Processing mental states}

\begin{itemize}
	\item Generate plans for received events.
	\item Generate plans for goals.
	\item Process exceptions and handle failures.
	\item Repair plans.
	\item Select plans for execution.
	\item Execute plans.
\end{itemize}

\section{Markov}

\end{document}

### 2015-10-28 agents/summery.tex
added configuration and operational semantics

@@
 
+\subsection{Operational semantics}
+Operational semantics defines the computation steps a program
+configuration may take. Operational semantics allows studying
+of programming constructs in a rigorous manner. It facilitates proving
+general properties about a language. It lies close to the implementation
+of an interpreter. It facilitates model checking.
+
+\subsubsection{Basic syntax}
+The configuration $C$ evolves into configuration $C'$:
+
+\[C \to C'\]
+
+If premise $P$ holds transition $C \to C'$ can be derived:
+
+\begin{prooftree}
+	\AxiomC{$P$}
+	\UnaryInfC{$C \to C'$}
+\end{prooftree}
+
+\subsection{Configuration}
+configurations only mention things that change.
+Multi-agent configuration where $A$ is an agent and $\chi$ is the
+environment:
+
+\[\langle \{A_1, \cdots, A_n\}, \chi\rangle \]
+
+Individual agent configuration where $i$ is the agent number, $\sigma$ the
+believes, $\gamma$ the goals, $\Pi$ the plans $\theta$ the assignments and
+$\xi$ the events:
+
+\[A_i = \langle i, \sigma_i, \gamma_i, \Pi_i, \theta_i, \xi_i \rangle \]
+
 \section{Markov}

### 2015-10-28 agents/summery.tex
added the update transition and comments.

@@
 
+\subsection{Transitions}
+Execution of a belief update action $\alpha$ modifies the belief and goal
+bases. $Update$ is an update function that is specified by the believe
+update actions and $\gamma'=\gamma - \{\psi \in \gamma | \sigma' \vDash \psi\}$.
+Which basically says that $\gamma'$ is equal to $\gamma$ minus the update.
+
+\begin{prooftree}
+	\AxiomC{$Update(\sigma, \alpha\theta)=\sigma'$}
+	\UnaryInfC{$\langle i, \sigma, \gamma, \{(\alpha,id)\}, \theta, 
+		\xi\rangle \to 
+	\langle \langle i, \sigma', \gamma', \{\}, \theta, \xi\rangle$}
+
+\end{prooftree}
 \section{Markov}

### 2015-10-29 agents/summery.tex
finished 2apl notes I think

@@
 \subsection{Transitions}
+The confiugrations described \emph{what} can change where transition
+will describe \emph{how} they can change.
+
+$\vDash$ Describes a first order entailment relation. As I understand
+this means $A \vDash B$ there exsits a $B$ in $A$ which has an entailment
+to $A$. So $A$ is a set of things and a $B$ like thing exists in $A$.
+\subsubsection{Belief update}
 Execution of a belief update action $\alpha$ modifies the belief and goal
@@
 \end{prooftree}
+
+So on the left side of the arrow, the plan is to do a believeUpdate $\alpha$
+on the right side this is done so its removed from the plans and the believes
+\emph{and} goals are modified. The goals are modified to because if a believe
+with the same name as a goal becomes true than the goal is removed. (this is
+a design decision btw, its just how 2apl works).
+
+\subsubsection{Adopta goal}
+Execution of a goal adopt action adopta(g) adds the goal to the beginning of
+the goal base. If the goal $g$ trough assignment $\theta$ does not already
+exists in goalbase.
+
+\begin{prooftree}
+	\AxiomC{$\sigma \nvDash g\theta$}
+	\UnaryInfC{$\langle 
+		i, \sigma, [\gamma_1, \cdots, y_n], \{(adopta(g),id)\}, \theta, \xi, 
+	\rangle \to \langle 
+		i, \sigma, [g\theta,\gamma_1, \cdots, y_n], \{\}, \theta, \xi, 
+	\rangle$}
+\end{prooftree}
+
+So the agent comes accross a adopta method in the plan base $\Pi$, if
+the associated believe of the goal is not true then you can start adding
+it.
+
+Note that in the slides the environment was also part of this action. But I
+think this was a mistake, because in the paper of goaldynamic actions there
+isn't a mention of an environment. Besides it doesn't make sense to include
+the environment in an agent configuration.
+
+\subsubsection{Applying planning goal rules}
+The structure of a PG-rule in 2apl is as follows:
+
+\[\kappa <- \beta | \pi \]
+
+Where $\kappa$ is the name of the plan.
+$\beta$ is the condition that is applied before
+executing the plan (basically a big if around the plan). $\pi$ is
+the body of the plan, a set of instructions.
+
+\begin{prooftree}
+	\AxiomC{$\gamma \vDash \kappa \tau_1 \quad \& \quad
+	\sigma \vDash \beta \tau_1 \tau_2$}
+	\UnaryInfC{$\langle
+		i, \sigma, \gamma, \Pi, \theta, \xi
+	\rangle \to \langle
+	i, \sigma, \gamma, \Pi \cup \{ \pi \tau_1 \tau_2 \}, \theta, \xi
+	\rangle$}
+\end{prooftree}
+
+The transition can be executed if $\kappa$ is an active goal (this valuation
+is stored in $\tau_1$ \emph{and} $\beta$ is true according to the believe 
+base (this eveluation is stored in $\tau_2$). If all these things are true
+then $\pi$ is added to the plan base $\Pi$ \emph{but} it will only be executed
+if $\tau_1$ and $\tau_2$ return true.
+
 \section{Markov}

### 2015-10-29 agents/summery.tex
decided to add the notes on the possible exam questions.
grammar

@@
 \section{APL}
-Medhi dastani thought about the agent programming langauge (APL)
-
+Medhi dastani taught about the agent programming langauge (APL).
 \subsection{Agent data structure}
@@
 if $\tau_1$ and $\tau_2$ return true.
+\subsection{Exam}
+some notes I gathered for possible exam questions:
+
+What happens if $\gamma = pos(2,3)$ and $\sigma = pos(4,5)$, resolve with
+state transitions.
 
@@
 
+\subsection{Exam}
+
 \end{document}

### 2015-10-29 agents/summery.tex
added some markov things
added the probability craP??

@@
 \section{Markov}
+Computerised decision makers must deal with an uncertain world.
+
+\subsection{Expectation}
+Excpectation is the average value you excpect to happen. For example if you
+throw a dice the average value you should get is 3.5. Because:
+
+\[1 * 1/6 + 2 * 1/6 + 3 * 1/6 + 4 * 1/6 + 5 * 1/6 + 6 * 1/6 = 3.5\]
+
+You multiple all possible values time the chance of them happening and then
+calculate the sum of them. Formally it is written like this:
+
+\[ E[X]=Def \sum_{all x} xP\{X=x\} \]
+
+So you can fill it in like:
+\[E[X]= 1 * P\{X=1\} + \cdots + 6 *P\{X=6\}\]
+\[E[X]= 1 * 1/6 + \cdots + 6 *1/6 = 3.5\]
+
+\subsubsection{Properties of expectation}
+
+\begin{itemize}
+	\item $E[c] = c$
+	\item $E[X+Y] = E[X] + E[Y]$
+	\item $E[cX] = cE[X]$
+	\item $E[XY] = E[X]E[Y]$ if and only if $X$ and $Y$ are independent.
+	\item $E[X] \le E[Y]$, if $X \le Y$ almost surely.
+\end{itemize}
+
+\paragraph{Law of the onconcious statistician}
+You have to prove this, its \emph{not} a property:
+
+\[E[g(X)]=\sum_{all x} g(x)P{X=x}\]
+
+where $g$ is any real-valued function. (It returns a value that is part
+of the real set of numbers).
+
+\subsection{Conditional expectation}
+The expectation of $X | Y = y$ is called the \emph{conditional expectation}
+of $X$ given $Y$, and is denoted by:
+
+\[E[X|Y=y]\]
+
+$E[X|Y=y]$ Is a number that depends on $y \in Y$. So you only get an
+expectation $X$ if you have an y that satisfies the condition.
+This yields a random variable $E[X|Y]$. Because of this it has all the
+properties of a random variable.
+
+\subsubsection{Probability mass functions}
+The standart notation for probability mass functions is:
+
+\[p_x(x): D(X) \to [0,1]: x \mapsto P\{X=x\} \]
+\[p_y(y): D(Y) \to [0,1]: y \mapsto P\{Y=y\} \]
+\[p_{xy}(x,y): D(X) \times D(Y) \to [0,1]: (x,y) \mapsto P\{X=x,Y=y\} \]
+\[p_{x|y}(x|y): D(X) \times D(Y) \to [0,1]: (x,y) \mapsto P\{X=x|Y=y\} \]
+
+For example:
+
+\[ P\{X=x|Y=y\}=\frac{P\{X=x,Y=y\}}{P\{Y=y\}}\]
+
+Is often written as:
+
+\[P_{x|y}(x|y)=\frac{P_{xy}(x,y)}{P_y(y)}\]
 
 \subsection{Exam}
+ For a given decision problem with a certain
+discount factor.A pure policy sufices.
+Do policy iteration, (a single iteration), do value iteration
+(single iteration).
 

### 2015-10-31 reasoning/summery.tex
added a section about jargon and deleted the proof ad absurdom started explaining induction with an almost trivial example.

--- 2015-10-27 reasoning/summery.tex
@@
 
+\newcommand{\brcell}[2][l]{%
+	\begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
+
 \begin{document}
@@
 
-So there are 2 kinds of proving. The first one is writing a proof in the system
-this should be rather easy, although you have to be carefull to follow the
-rules to the letter. I'm going to ignore this one since its just using the
-rules, which you've done tons of times as a programmer, so it shouldn't be
-that strange.
+Proving is in meta space. Its a little like programming in that you have to
+be specific. There are lots of ways to prove things. In this report I will
+discuss a select few to give an idea about what is going on. But first
+I'll start covering with some basics.
 
-The second one is the more troublsome, its proving the system itself. This
-is also split up into 3 different aspects. They all have in commen that
-it should be done in natural langauge. But, there is a structure to it
-I'm going to try and give a complete overview of this structure.
+\subsection{Jargon}
+In here I will discuss some jargon logicians use. It may seem silly but knowing
+the precise defintion of these words makes understanding proofs a lot easier.
 
-First you have soundness, then there is completness, and finally you can be
-asked to proof certain properties of a system.
+\begin{tabular}{ll}
+	Word & definition \\ \toprule
+	Formula & The letters, but also letters combined with conectives (its
+	recursive) \\
+	Axiom & Something which is undoubtibly valid. proofs work from a
+	complex formula towards axioms. \\
+	Theorom & the thing you're trying to proof. \\
+	Lemma & \brcell{a hypothosis which you're trying to prove, you usually
+		define \\
+	a lemma to reference to later. Its like a little theorom} \\
+	Sequent & A thing with a sequent connective (in my case either
+	$\vdash$ or $\Rightarrow$) \\
+	Derivable & if X is derivable in system A, then you can proof x with
+	system A
+\end{tabular}
 
@@
 
-\section{proving soundness}
-This basically means your system doesn't modify the value you put in.
-If you put in truth, apply a bunch of rules from your system, then you should
-get out truth.
+\section{Structure of proofs}
+\label{structure of proofs}
+Now of course different poeple have different kinds of notations. The logicians
+don't really seem to adhere to standarts. But that's fine, you just have to
+get used to conflicting symbols.
 
-\subsection{Structure of your proof}
-First write down your overal strategy of the proof \autoref{strategies}. For
-soundness this is often `proof by induction' \autoref{induction}.
+\subsection{Find references}
+Before you do anything check if you can't refer to any existing proofs.
+This saves a lot of time. You can just ignore writing a proof if you can
+write down a reference. (I bet this is allowed on the exam as long you
+can refer to the stuff you wrote down yourself).
 
-Then write down the lemmas you're gonna use. A lemma is a little proof, often
-derived from the properties of the system you're proving. You could see it
-as a basic argument. When proving soundness you want to connect this to the
-semantics, for propositional systems the semantics are truth tables, for kripke
-models it are kripke frames. The semantics are the thing your talking about
-in your logical system.
-%I can see how I dind't understand all this crap the first time I saw it,
-% its freaking packed with definitions, and even if you understand all the
-% definiontions, its still hard
-Make sure that you've proven the soundness of the axioms *and* of the rules
-with help of the semantics.
+\subsection{Head of a proof}
+In the head of the proof you don't do any work, you just state what we're
+talking about and what you're going to do. So you start with the definitions.
+Then you say what you're going to proof (by using the defintions), so here
+you just repeat the question. And finally you mention your strategy.
 
-Finally ty up the lemmas together in the conclusion, this is done with help
-of the proofing strategy.
+\subsection{Body of the proof}
+In here you work out the proof with help of the strategy.
+See the \autoref{strategies} section.
 
-Finish with a sentence like: `This had to be shown'.
+\subsection{End of the proof}
+Just write down ` QED' or ` this is what had to be shown' even if you're not
+certain what you're doing.
 
@@
 \label{induction}
+\subsubsection{When to use it?}
+You can use induction if you have to proof something for countibly infinite
+things, but there exists a base method that can validate if its true.
+So basically all thigns that can and make sense be resolved with a
+recursive function.
 
-\section{notations}
+\subsubsection{How to use it?}
+In here you try to prove something by starting with a single case n.
+For all n (*), n=1, supose (*) for n, then show (*) for n+1.
+
+Fat chance you'll understand it based on that. Its way to abstract.
+But I'll show how to use it with help of several examples. \emph{cursive
+text will be comments}
+
+\paragraph{Proof that formulas always have an even number of brackets.}
+\emph{So space cowboy, we have to write a proof, before panicking remember
+	the boiler plate in \autoref{structure of proofs}, we may get some
+points for doing this, its at least polite}
+
+\subparagraph{Definitions}
+A formula can be a propositional variable, in which case it has zero brackets.
+A formula can also be multpile proporsitional variables, connected with connectives,
+in this case it has the connective count * 2 brackets, because for every connective
+2 brackets are added.
+\subparagraph{Strategy} Every formula has a even number of brackets, this will
+be proven with help of iniduction.
+\subparagraph{Proof}
+
+\subsection{Tabloo or truth tree}
+\subsubsection{When to use it?}
+The tabloo method comes to mind if you have to proof things about systems
+that use trees, and where the values of the leaves of the tree trincle down
+to the root of the tree.
+
+\subsubsection{How to use it?}
+In this strategy you create a tree, starting from the axioms which is either
+true or false, you go down step by step and say what happens. In most systems
+if one of the axioms is false it means the conclusion is false to.
+
+\section{Notations}
 \label{notations}

### 2015-10-31 reasoning/summery.tex
finished with the even number of brackets.

@@
 \begin{tabular}{ll}
-	Word & definition \\ \toprule
+	Jargon & Meaning\\ \toprule
 	Formula & The letters, but also letters combined with conectives (its
 	recursive) \\
-	Axiom & Something which is undoubtibly valid. proofs work from a
-	complex formula towards axioms. \\
+	Axiom & \brcell{Something which is undoubtibly valid. Proofs work from a \\
+	complex formula towards axioms.} \\
 	Theorom & the thing you're trying to proof. \\
@@
 
-\paragraph{Proof that formulas always have an even number of brackets.}
+\paragraph{Proof that propositional formulas always have an even number of brackets.}
 \emph{So space cowboy, we have to write a proof, before panicking remember
@@
 \subparagraph{Definitions}
-A formula can be a propositional variable, in which case it has zero brackets.
-A formula can also be multpile proporsitional variables, connected with connectives,
-in this case it has the connective count * 2 brackets, because for every connective
-2 brackets are added.
-\subparagraph{Strategy} Every formula has a even number of brackets, this will
-be proven with help of iniduction.
+A propositional formula can be a propositional variable,
+in which case it has zero brackets.
+A propositional formula can also be multpile proporsitional variables,
+connected with connectives, in this case it has the connective count * 2
+brackets, because for every connective 2 brackets are added.
+
+\subparagraph{Strategy}
+For all propositional formula the number of brackets is even.
+This will be proven with help of iniduction.
+
 \subparagraph{Proof}
+\emph{So this is where the induction starts, first do the base case}
+The base case where n = 0,
+a propositional formula is a single propositional variable. This means
+it has 0 brackets according to definition 1, and 0 is an even number, so
+the amount of brackets in the base case is even.
+
+\emph{Induction step}
+For the case n+1, asume $A$ and $B$ have a even number of brackets,
+then $A$ connective $B$ has an even number of brackets to because you have
+to add an even number of brackets according to the definition of adding
+connectives.
+
+\emph{Because this can be used for the case of propositional variables,
+and all complexer formulas that can be constructed with connectives, you're
+done. All posibilities have been treated now.}
+
+QED
+
+\paragraph{Proof that all natural numbers are either odd or even}
+\emph{Another quite obvious prove, but this one has a more complex induction
+step, so I'll treat it here}
+
+\subparagraph{Definitions}
+Natural numbers are all integers bigger than 0.
+\[Even(n) = \left\{ \begin{matrix}
+	n=0 &\\ 
+	Even(n-2)& 
+	\end{matrix}
+\right \]
 

### 2015-10-31 reasoning/summery.tex
added another induction example.

@@
 Natural numbers are all integers bigger than 0.
-\[Even(n) = \left\{ \begin{matrix}
-	n=0 &\\ 
-	Even(n-2)& 
-	\end{matrix}
-\right \]
+
+\[
+Even(n) := \left\{
+	\begin{array}{l}
+		n= 0 \\
+		Even(n-2)
+	\end{array}
+\right.
+\]
+\[
+Odd(n) := \left\{
+	\begin{array}{l}
+		n= 1 \\
+		Odd(n-2)
+	\end{array}
+\right.
+\]
+
+\subparagraph{Strategy}
+All Natural numbers are either odd or even, this will be proven by induction.
+
+\subparagraph{Proof} \emph{In here we look at the definitions and try to see
+the base case(s)}. In case n=0, n is even according to the definition of
+even, so n is either postivie or negative. \emph{reflecting back on what
+we're trying to prove, saying its the case}
+
+In case n=1, n is odd according to the definition of odd. So n is either
+odd or even.
+
+\emph{After the base case, first assume, then push with help of the definition}
+Assume n is odd or even, then n+2 is also odd or even according to the
+definition of odd or even.
+
+QED
+
+\paragraph{Proof that $\vdash_{gkc} (\Gamma \Rightarrow \Delta)$ is
+closed under right weakening}
+\emph{I'll treat this one because its a rather simple case, but still
+touches the concepts of logic a bit deeper}
+
+\subparagraph{Definitions}
+$GKC$ is gentzen calulus. $\vdash_x$ means something is derivable in system x.
+We assume $\Gamma \Rightarrow \Delta$ is derivable in $GKC$. This means
+there exists an axiom in $\Gamma \Rightarrow \Delta$, for example
+$\Gamma', B \Rightarrow B, \Delta'$.
+Closed under right weakening means that it doesn't matter if
+$\Gamma', B \Rightarrow B, \Delta'$ or $\Gamma', B \Rightarrow B, \Delta', A$
+is used since they're both axioms.
+
+\subparagraph{Strategy} To proof for all sequents in $GKC$ right weakening
+is closed. To proof with iniduction.
+\subparagraph{Proof}
+Take an arbitrary proof $P$ that is derivable in $GKC$,
+this proof has several sequants $S$ to the depth $n$, such that $S_n = P$.
+
+on depth $n=0$ the sequent is an axiom, because $P$ is derivable in $GKC$.
+
+Assume $P_n$ is closed under right weakening. Then $P_n-1$ is closed under
+right weakening to
 

### 2015-10-31 reasoning/summery.tex
removed the right weakening proof because I don't understand it and can't explain it. It doesn't confirm well to the structure I impose on induction.

--- 2015-10-31 reasoning/summery.tex
@@
 Natural numbers are all integers bigger than 0.
-\[Even(n) = \left\{ \begin{matrix}
-	n=0 &\\ 
-	Even(n-2)& 
-	\end{matrix}
-\right \]
+
+\[
+Even(n) := \left\{
+	\begin{array}{l}
+		n= 0 \\
+		Even(n-2)
+	\end{array}
+\right.
+\]
+\[
+Odd(n) := \left\{
+	\begin{array}{l}
+		n= 1 \\
+		Odd(n-2)
+	\end{array}
+\right.
+\]
+
+\subparagraph{Strategy}
+All Natural numbers are either odd or even, this will be proven by induction.
+
+\subparagraph{Proof} \emph{In here we look at the definitions and try to see
+the base case(s)}. In case n=0, n is even according to the definition of
+even, so n is either postivie or negative. \emph{reflecting back on what
+we're trying to prove, saying its the case}
+
+In case n=1, n is odd according to the definition of odd. So n is either
+odd or even.
+
+\emph{After the base case, first assume, then push with help of the definition}
+Assume n is odd or even, then n+2 is also odd or even according to the
+definition of odd or even.
+
+QED
 

### 2015-10-31 agents/summery.tex
finished the basic probablity stuff, on to the real markof stuffs.
I was right, slides were modified, so I deleted the note to avoid confusion.

@@
 it.
-
-Note that in the slides the environment was also part of this action. But I
-think this was a mistake, because in the paper of goaldynamic actions there
-isn't a mention of an environment. Besides it doesn't make sense to include
-the environment in an agent configuration.
-
 \subsubsection{Applying planning goal rules}
@@
 
+\subsection{Total expectation}
+The tower property, which is awesome:
+
+Let $X$ and $Y$ be random variables then:
+
+\[ E[E[X|Y]] = E[X] \]
+
+The proof is in the slides.
 \subsection{Exam}

### 2015-10-31 agents/summery.tex
added a markov example and how to put it in a matrix

@@
 \documentclass{article}
+\usepackage{amsmath}
 \usepackage{txfonts}
@@
 \usepackage{tikz}
+\usetikzlibrary{automata,arrows,positioning,calc}
 \usepackage{listings}
@@
 The proof is in the slides.
+
+\subsection{Markov to matrix}
+A finite Markov process consisting of $n$ states can be represented by an 
+$n\times n$ \emph{probablity transition matrix}.
+
+Consider the following markov process:
+\begin{center}
+\begin{tikzpicture}[->, >=stealth', auto, semithick, node distance=3cm]
+\tikzstyle{every state}=[fill=white,draw=black,thick,text=black,scale=1]
+\node[state]    (A)                     {A};
+\node[state]    (B)[above right of=A]   {B};
+\node[state]    (C)[below right of=A]   {C};
+\path
+(A) edge[loop left]     node{0.5}         (A)
+    edge[bend left]     node{0.3}     (B)
+    edge[bend right, left]     node{0.2}      (C)
+(B) edge[bend left,right]    node{0.9}      (A)
+    edge[bend left,right]    node{0.1}      (C)
+(C) edge[bend right, right]    node{1}      (A);
+\end{tikzpicture}
+
+The probability matrix belonging to this process looks like the following:
+
+\[
+\begin{matrix}
+		& A		& B		& C \\
+	A	& 0.5	& 0.3	& 0.2 \\
+	B	& 0.9	& 0		& 0.1 \\
+	C	& 1		& 0		& 0 \\
+\end{matrix}
+\]
+
+\end{center}
+
+Adding the $A,B,C$ is optional, but I find it gives more overview. The first
+cell reads, the probobality of going from a to a.
+
+\subsubsection{Stochastic}
+A matrix where every row sums to1 is called a \emph{stochastic matrix}. Every
+transition matrix is a \emph{stochastic matrix}.
+
+\subsection{Sample path}
+A \emph{sample path} of a markov chain is a specific specific realistion of
+that chain.
+
+\[ X_0,X_1,X_2,\dots\]
+
+The $X_i$ are random variables.
+
+\subsubsection{Markov process}
+A \emph{Markov process} is a stochastic process $X_0, X_1, X_2, \dots$ such 
+that:
+
+\[ P\{ X_{n+1} X_n, X_{n-1}, \dots\}=P\{X_{n+1}|X_n\}\]
+for $n \ge 0$
 \subsection{Exam}

### 2015-10-31 agents/summery.tex
Finished another slide.

@@
 
-\subsubsection{Markov process}
+\subsection{Markov process}
 A \emph{Markov process} is a stochastic process $X_0, X_1, X_2, \dots$ such 
@@
 for $n \ge 0$
+
+\subsubsection{Remarks}
+If states are numbers, then often state space is denoted by $I$, with 
+typical elements $i,j,k$.
+
+A markov process is sometimes called a markov chain. These concepts are
+equal.
+
+If $S$ is countably infinite, then the probability transition matrix
+can be represented as an ``open ended'' stochastic matrix.
+\[
+	P =
+\begin{pmatrix}
+	p_{1,1}	& p_{2,1}	& p_{3,1} & \cdots\\
+	p_{1,2}	& p_{2,2}	& p_{3,2}	& \cdots\\
+	p_{1,3}	& p_{2,3}	& p_{3,3}	& \cdots\\
+	\vdots	& \vdots	& \vdots		& \ddots \\
+\end{pmatrix}
+\]
+With markov proceci it is extremly important to beclear about the
+\emph{cardinality of the state space} (finite, discrete, continues) and
+the \emph{nature of time} (discrete, continuous)
+
+A \emph{dead state} or a \emph{dead end} is a state that does not have
+any outgoing arcs. An \emph{absorbing  state} is a state that has precisely
+one outgoing arc with probability 1 to itself.
+
+\subsection{Chapman-Kolmogorov equations}
+
+\[P^{(n)}_{ij} = \sum_{all\mbox{\ }k} P^{(n_1)}_{ik} P^{(n_2)}_{kj} \]
+
+It is convenient to list the $n$-step transition probabilities in a matrix.
+\[
+	P^{(n)} =
+\begin{pmatrix}
+	p^{(n)}_{1,1}	& \cdots	& p^{(n)}_{n,3} 	& \\
+	\vdots	& \ddots	& \vdots	& \\
+	p^{(n)}_{1,n}	& \cdots	& p^{(n)}_{n,n}	& \\
+\end{pmatrix}
+\]
+
+In particiular $P^{(1)} = P$.
+
+\subsubsection{Lemma}
+Suppose $k+l=n$ Where $k,l\ge1$. Then $P^{(k)} \cdot P^{(l)} = P^{(n)}$
+
+(I think this was proven)
+
+$P^{(n)} = P^n$, this was definitivley proven. Not sure why you would
+prove removing brackets, but whatever.
+
+\subsection{Classes}
+No this aren't default private structures (its a c++ reference).
+
+Classes make dealing with large markov chains easier by grouping nodes
+together. \emph{Open classes} allow the process to escape the class once
+it enters a \emph{closed class} it will remain there forever.
+A class is \emph{Transient} if a process can never reenter it, a
+state is recurrent if it can. A process is multichain if there are
+multiple recurrent classes, if there is just one reccurent class
+its uni-chain.
+
+A critical transition is a transition between classes.
+
+A closed class doesn't have to be recurrent if there are infinite cases.
+Otherwise a closed class is always recurrent.
+An openclass is always transient.
+
 \subsection{Exam}

### 2015-11-01 cogniveProcessing/summery.tex
added a cognitive start...

\documentclass{article}
\usepackage{amsmath}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\usepackage{qtree}
\usepackage{tikz}
\usetikzlibrary{automata,arrows,positioning,calc}
\usepackage{listings}
\usepackage{hyperref}
\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

\begin{document}
\lstset{language=Java}
\author{Jappie Klooster}
\title{Thoughts about cognition}
\maketitle

\section{Introduction}
This will be an introduction to cognitive proccessing
coming from someone who followed
` hbo informatica' as pre study.

In here I will write down the things I think are important.

\section{Research and design}
This chapter will talk about how to set up experiments and how to draw
conclusions from them. It will also discuss what tools are available.
Matlab was also discussed, but considering my background I will not spend
time on this subject.

\subsection{Brains}
The human and monkey brain are rather similiar (in shape mostly).

\subsubsection{Why do we have brains?}
Reproduction is an important factor, there exists animals that consume their
own brain ones they start making their offspring.

\subsection{Common denominator}
The soma or ``cell body'' is the bulbous end of a
neuron, containing the cell nucleus.

Dendrites are the branched projections of a neuron that act to
propagate the electrochemical stimulation received from other neural
cells to the cell body, or soma, of the neuron from which the
dendrites project

\end{document}

### 2015-11-01 agents/summery.tex
added some stuffs about belman the operator.

@@
 \usepackage{tikz}
+\usepackage{amssymb}
 \usetikzlibrary{automata,arrows,positioning,calc}
@@
 
+\subsection{Markov reward}
+A \emph{Markov reward process} is a markov process with rewards on
+transitions. these rewards are real valued.
+
+Markov rewards processes are also known as \emph{markov reward chain} or
+\emph{valued Markov chains}.
+
+example
+
+\[
+	P =
+\begin{pmatrix}
+	0.2	& 0.8 & 0 	& \\
+	0.3	& 0	& 0.7	& \\
+	0	& 0 & 1	& \\
+\end{pmatrix}
+\]
+With reward matrix:
+\[
+	R =
+\begin{pmatrix}
+	1	& 2 & 0 	& \\
+	0	& 0	& 3	& \\
+	0	& 0 & 0	& \\
+\end{pmatrix}
+\]
+
+\subsubsection{Total reward}
+
+Let $\omega$ be a realisation of a Markov process:
+
+\[\omega = s_0, s_1, \dots\]
+
+where $s_0$ is a start state.
+The rewards that accumulate are $r_1, r_2, \dots$:
+
+\[ s_0 \xrightarrow[\mbox{reward } r_1]{} s_1 \xrightarrow[\mbox{reward } r_2]{} s_2\]
+
+\paragraph{Definition}
+
+$r_k | (s_0=1) : $ reward at kth step, given $s_0 = i$.
+$R|(s_0=i),$ the \emph{total reward} : $(r_1 + r_2 \dots)|s_0=i,$
+$V(i),$ the \emph{value} of state i : $E[R|s_0=i]$.
+
+\subsection{The bellman equation}
+
+\[V(i)=\sum_{all\mbox{\ }j} p_{ij}[r_{ij} + V(j)]\]
+
+This system of linear equations enables us to compute state values in two
+ways, by using algebra or by using iteration. Iteration you just wait until
+convergence within a specified range. For this course iteration is enough.
+
+\subsection{Utility pump}
+A utility pump is a MRP in which at least one recurrent state possesses a
+non-zero reward.
+\subsubsection{Discount factor}
+A \emph{dicsount factor} is a real number $0 \le \gamma < 1$. it represents
+the probability to survive from one state to the next.
+
+\paragraph{remarks} The discount factor is constant. The discount factor
+does not impose a fixed cuttoff bound. It is a probablistic concept.
+
+\subsubsection{Discounted reward}
+Suppose a discount factor $0 \le \gamma < 1$. Then
+
+\[V(i)=\sum_{all\mbox{\ }j} p_{ij}[r_{ij} + \gamma V(j)]\]
+
+for every state $i$.
+
+\paragraph{definition} A \emph{valuation} is a vector
+
+\[v=Def
+\begin{pmatrix}
+	v(1) \\
+	\vdots \\
+	v(n)
+\end{pmatrix}
+\in \mathbb{R}^n
+\]
+
+\paragraph{definition} The \emph{expected immediate reward} at state $i$ is
+the by $P_{ij}$ - probabilities weighted averaga of the $r_{ij}$ rewards:
+
+\[C(i)=Def\sum_{all \mbox{\ }j} p_{ij} r_{ij} \]
+
+The (vertical) vector of expected immediate rewards is denoted by $C$.
+
+\paragraph{definition} The \emph{Bellman operator}:
+
+\[B_p : \mathbb{R}^n \to \mathbb{R}^n : v \mapsto C + \gamma Pv \]
 \subsection{Exam}

### 2015-11-01 agents/summery.tex
added soem info abut fixed points...

@@
 
-\subsection{Utility pump}
+\subsubsection{Utility pump}
 A utility pump is a MRP in which at least one recurrent state possesses a
@@
 \[B_p : \mathbb{R}^n \to \mathbb{R}^n : v \mapsto C + \gamma Pv \]
+
+\subsection{Fixed point}
+The value vector, $v^*$, of a discounted reward Markov process is a fixed
+point of the $B_p$ operator:
+
+\[ B_p(v)=v\]
+
+\[
+	v_0 \to B_p(v_0) \to B_p B_p (v_0) \to \dots \to B^n_P v_0
+	\to \dots \to v^*
+\]
+
 \subsection{Exam}

### 2015-11-01 cogniveProcessing/summery.tex
added a lot of information about BRAINZ

@@
 \subsection{Common denominator}
-The soma or ``cell body'' is the bulbous end of a
-neuron, containing the cell nucleus.
+All the brains have neurons.
 
-Dendrites are the branched projections of a neuron that act to
-propagate the electrochemical stimulation received from other neural
-cells to the cell body, or soma, of the neuron from which the
-dendrites project
+Incoming singals arrive at denrites, and are integrated to trigger an outgoing.
+signal (or not) at the axon hillock.
+Cell firing action potential is sent down the axon, which is often insulated
+with meylin to increase conduction peed, jumping between the nodes of ranvier.
+
+When the action potential arrives at the terminal bouton, neurotransmitters
+are realeased to activate the next cell.
+
+\subsubsection{Action potentials}
+At the baseline, neuron potential difference between the outside and inside
+of the cell is about -70mv. Incoming singals(arrows) cause changes that
+make the inside of teh cell more positive. When a treshold is reached a
+cascade of events is triggeredthat make the inside of the cell very positive,
+very quickly. This discharge is the action potential and starts at the hillock.
+It travels trough the axon to the next cell.
+
+So the charge level slowly built until a treshold, and then releases. This
+causes chemicals to jump at the terminals. Which is treated as an incoming
+singal to the next neuron.
+
+\subsubsection{Neuro transmision}
+Neuro transmission is done with neurotranmitters, there are two main
+transmitters, Glu, which makes the next cell more likely to fire, its
+exictatory. GABA makes the next cell less likely to fire, its inhibitory.
+
+Neurotransmitters (NT) are stored in vesicles. Electrical signal causes
+vesicles to bind with the membrane of the terminal and thus release NT
+out of the cell into synaptic cleft. NT spread and ome bind to next
+cell at particluar locations: receptors. Binding to receptor causes
+changes in the cell that may result in it signaling/firing itself.
+
+The synaptic cleft is the gap between the axon terminal and
+the dendritic spine.
+
+\paragraph{Receptors} Usually a protein channel. NT can bind to it
+and change its shape. When open: Ions can flow trough it and
+chagne the electrical potential of the cell (try to excite or
+inhibit it).
+
+Many receptors come in the form of `channels'. This is a chanal,
+aka potential opening, between the in and outside of the cell.
+Depending on how it's folded it may allow ions to pass or not.
+Its shape also makes it selecive: Only some ions can pass.
+
+There are many different kinds (Glu, Gaba, dopamine etc).
+Sensory receptors repsond to the outside world instead of signals
+from other neurons: Light, pressure, taste or smell molucles vibration etc.
+
+Dopamine and serotine are very common. Not main transmitter but important
+modulators of activity. Not all recptors are channals. We also have sensory
+recptors. They're special and can be very different. Complex molecules bind
+to chemical recptors in the nose and on the tongue. Hair cels in the cockle
+(ear) respond to vibration of a particular frequency.
+
+\subsection{Sensory systems}
+Converting energy from outside the body to neural signals trough: stimulus.
+Modification of signal, transduction or transfer.
+
+Flavour is a multi sensoy phonomon. Olfaction and ustation strong connections
+to limbic system. Chemo signalling affects behoviour, perception and emmotions.
+
+Chemical senses appear to have a strong link to social behaviour.
+
+\subsubsection{The pupil}
+Allows more or less light to enter the eye. More light usually not a good idea.
+Dilation: arousal, Contrsiction: Relaxation.
+
+Sensory aquisition \& rejection: emotrion affects pupil respones.
+
+\subsubsection{Smell}
+
+MHC: Major histocompatibility complex. Genes with product that is part of
+immume systems Having more different. Smelling mice and vice versa.
+
+Humans can small differences beteween mouse families only differing in MHC.
+mice can recognise human MH-types from urine.
+
+\paragraph{Do we use it?} Contraceptives seem to interfere with natural mate
+selection: A big deal because similiar MHC tends to reduce succesfull
+pregnacies.
+
+\subsubsection{Receptive fields}
+Description depends on modality. Somatosensory, Auditory, Visual, Gustation.
+
+Pain for example spreads from 1 point to the entire surrounding area, even
+if its not affected.
+
+\section{Langauge}
+Trying to figure out what langauge is.
+
+\section{Langauge and Context}
+In here we will focus on trying to reduce the influence of context or
+at least try to explain it.
 

### 2015-11-02 agents/summery.tex
added a bunch of information, halfway the last slide of markov

@@
 {\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }
+
+\newcommand{\brcell}[2][l]{%
+	\begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
 
@@
 
+\subsection{Example Markov reward process}
+Consider the following markov reard process. Suppose a discount factor
+$\gamma = 0.5$.
+
+\[ \left\{ \begin{matrix}
+		s: & s(r=1, p=0.4) & ; & a(r=4, p=0.6), \\
+		a: & s(r=3, p=0.3) & ; & a(r=8, p=0.7), \\
+		b: & s(r=2, p=0.5) & ; & b(r=5, p=0.5). \\
+	\end{matrix}
+\right. \]
+
+\subsubsection{Transition diagram}
+\begin{tikzpicture}[->, >=stealth', auto, semithick, node distance=3cm]
+\tikzstyle{every state}=[fill=white,draw=black,thick,text=black,scale=1]
+\node[state]    (S)                     {s};
+\node[state]    (A)[right of=S]   {a};
+\node[state]    (B)[left of=S]   {b};
+\path
+(S) edge[loop above]     	node{p=0.4, r=1}(S)
+    edge[above]     		node{p=0.6, r=4}(A)
+(A) edge[bend left,below]   node{p=0.3, r=3}(S)
+	edge[loop above]     	node{p=0.7, r=8}(A)
+(B) edge[above]    			node{p=0.5, r=2}(S)
+	edge[loop above]     	node{p=0.5, r=5}(B);
+\end{tikzpicture}
+\subsubsection{Matrici}
+\paragraph{Probability transition matrix}
+\[
+	P =
+\begin{pmatrix}
+	0.4	& 0.6 & 0 	& \\
+	0.3	& 0.7 & 0	& \\
+	0.5	& 0 & 0.5	& \\
+\end{pmatrix}
+\]
+\paragraph{Immediate reward matrix}
+\[
+	R =
+\begin{pmatrix}
+	1	& 4 & 0 	& \\
+	3	& 8	& 3	& \\
+	2	& 0 & 5	& \\
+\end{pmatrix}
+\]
+\subsubsection{Value evalutaion}
+apply:
+\[V(i)=\sum_{all\mbox{\ }j} p_{ij}[r_{ij} + V(j)]\]
+
+\[ \left\{ \begin{matrix}
+		V(s) = 0.4(1+\gamma V(s)) + 0.6(4+\gamma V(a)), \\
+		V(a) = 0.3(3+\gamma V(s)) + 0.7(8+\gamma V(a)), \\
+		V(b) = 0.5(2+\gamma V(s)) + 0.5(5+\gamma V(b)).
+	\end{matrix}
+\right. \]
+
+With abusing notation:
+
+\[ \left\{ \begin{matrix}
+		s = 0.4(1+\gamma s) + 0.6(4+\gamma a), \\
+		a = 0.3(3+\gamma s) + 0.7(8+\gamma a), \\
+		b = 0.5(2+\gamma s) + 0.5(5+\gamma b).
+	\end{matrix}
+\right. \]
+
+Just fill in the values recursevily to perform value iteration.
+
+\subsection{Decision processes}
+I'm gonna cover this subject poorly because its dead simple coming from an
+IT background.
+
+\begin{tabular}{ll}
+	word & definition \\ \toprule
+	decision tree & a simple directed tree stucture \\
+	states & represented by verticies.\\
+	Immediate rewards & represented by numbers along the edges. \\
+	Decision alternatives & the outgoing edges, also called possible actions. \\
+	Decision point & A state where a decision can be made. \\
+	Finite decision tree & \brcell{A tree with a finite number of states, \\
+		{\footnotesize Makes you wonder if there could be infinite number 
+		of decisions in such a tree.}} \\
+\end{tabular}
+\\
+The set of possible actions at state $i \in I$ is denoted by $A_i$.
+The immediate reward for action $j$ at $i$ is denoted by $r_{ij}$.
+
+\subsubsection{Policy}
+A \emph{policy} specifies for every decision point $s$ a probability
+distribution over $A_s$. A policy is typically denoted by $\pi, \tau\dots$.
+If at $s$ the probability distribution puts all weight on one choice, the
+policy is said to be \emph{pure} at $s$. Otherwise the policy is said to
+be \emph{mixed} at $s$.
+
+A \emph{run} of a policy is the \emph{actual execution} of that policy.
+An \emph{realisation}, or \emph{sample path} of a policy is the trace of
+that run.
+
+The \emph{payoff} or \emph{total reward} of a run is the total sum of rewards
+acuumulated in its realisation.
+
+\subsubsection{Value function}
+A \emph{value function} is any function $V: S \to \mathbb{R}$. The
+\emph{expected payoff of a policy $\pi$}, given a certain starting state $s$:
+	\[E[\pi|s_0=s]\]
+	is called the value of $s$ under $\pi$, and is denoted by $V^\pi(s)$.
+Value functions are compared in the following way:
+
+\[ V_1 \le V_2 \Leftrightarrow_{Def} V_1(s) \le V_2(s), \mbox{for all states } s.\]
+
+Policy $\pi$ is \emph{as good as} policy $\tau$ if $V^\tau \le V^\pi$. A policy
+is \emph{optimal} if it is as good as any other policy. A value function is
+\emph{optimal} if it is induced by a policy, and is as large as any other
+such value function.
+
+\subsubsection{Policy = strategy}
+% this is the first what came to mind when I saw the policies.
+In game theory, a policy is called a \emph{strategy} and a pure strategy is
+sometimes called \emph{stationary} or \emph{deterministic}. A mixed strategy
+is sometimes called \emph{probabilistic, non-stationary,} or
+\emph{non-deterministic}.
+% Imaginative with their non-*
+
+An example of a pure policy is
+\[ \tau = \{(r,b),(a,d),(b,e),(d,h),(f,i)\} \]
+
+All the decisions are made, even if you already cut them off.
+So you say (Start, End) for each node with choice.
+
+Mixed policy:
+\[ \tau = \{(r,\{(a, 0.2), (b,0.8)\}),(a,d),(b,\{(e, 0.7), (f,0.3)\}),
+(d,h),(f,i)\} \]
+% yes some really poor langauge design over here.
+
+\subsubsection{Expected payoff}
+$\pi$'s expected payoff, $E[\pi|s_0=r]$, can be computed recursivly with the
+law of total expectation.
+
+\subsubsection{Backward induction}
+A decision between alternatives that leads to certain outcomes, is a certainty
+itself.
+
+Backward induction is applicable in \emph{directed acyclic graphs}(DAGs).
+
+\subsubsection{Decision problem}
+A \emph{decision problem} amounts to determining the set $\Pi^*$, where:
+
+\[ \Pi = max_{\pi \in \Pi}(payoff(\pi))\]
+
+Where $\Pi$ is the set of all policies.
+
+For finite decision trees: Supppose at least one optimal policy exists.
+If this policy is mixed, it may be replaced by an optimal pure policy,
+such that, at every state, the payoff of the pure policy equals the expected
+payoff of the mixed policy.
+
+\subsubsection{Problems}
+\paragraph{infinite width} no best policy, because you're valuating forever.
+\paragraph{infinite depth} How to compare infinite total rewards (can't).
+\paragraph{discount factor} The right decision is very dependent on the
+discount factor. Counter intiuitive choices might be better for example.
+
+\subsection{Utility pumps}
+If there is no discount factor policy is irrelivant (the reward will be
+infinite no matter what you do). A tool is needed to kill infinity.
+For example a finiti horizon, discount factor, or long-run average
+payoffs.
+
+In this course we will just use discount factors.
+
 \subsection{Exam}

### 2015-11-03 agents/summery.tex
added all the belman and markov things, going on to modal logics.

@@
 
-\[ \Pi = max_{\pi \in \Pi}(payoff(\pi))\]
+\[ \Pi = argmax_{\pi \in \Pi}(payoff(\pi))\]
 
@@
 
-\subsection{Utility pumps}
+\paragraph{Utility pumps}
 If there is no discount factor policy is irrelivant (the reward will be
@@
 
+\paragraph{Investment} 
+At cycle $a$, it is possible to `pump' 1 utitlity per iteration for as
+long as a run lasts. At cycle $b$ it is possbile to `pump' $2$ utilities
+per iteration.
+In the long term it is better to be at $b$, but starting from $a$ this
+requires an investment of $K$ rewardless transitions to get there.
+
+\begin{center}
+\begin{tikzpicture}[->, >=stealth', auto, semithick, node distance=3cm]
+\tikzstyle{every state}=[fill=white,draw=black,thick,text=black,scale=1]
+\node[state]    (S)                     {s};
+\node[state]    (A)[right of=S]   {$b_1$};
+\node[state]    (B)[right of=A]   {$b_{k-1}$};
+\node[state]    (C)[right of=B]   {$b_k$};
+\path
+(S) edge[loop above]     	node{1}(S)
+    edge[above]     		node{0}(A)
+(A)    edge[above]    node{0}(B)
+(B)    edge[above]     		node{0}(C)
+(C)    edge[loop above]		node{1}(C);
+\end{tikzpicture}
+\end{center}
+
+If $K \ne 0$ it is best to stay left if $\gamma < 2^{-1/k}$. (for the
+proof see the slides, its to much math to copy in here).
+
+\subsection{Policy improvement}
+
+Recall the bellman equations for a Markov reward process:
+
+\[ v = C+\gamma P v\]
+
+To emphasise that $v$ depends on a probability transition matrix $P$:
+
+\[ v^P = C^P + \gamma P v^P\]
+
+In a \emph{decision problem}, $P$ is not chosen by nature, but by a 
+\emph{decision maker}. In that case, $P$ is called a \emph{ixed Markov
+policy}. and is denoted by $\pi$:
+
+\[ v^\pi = C^\pi + \gamma P^\pi v^\pi\]
+
+Each policy $\pi$ gives a value function $v^\pi$. THe objective is to find a
+policy $\pi$ that maximises the value function $v^\pi$.
+
+\subsubsection{First sketch}
+A first sketch of a policy iteration algorithm:
+\begin{enumerate}
+	\item Start with an arbitrary policy $\pi$
+	\item Compute $V_\pi$. (know as \emph{policy evaluation}).
+	\item Improve $\pi$ locally to $\tau$.
+	\item If $\tau \ne \pi$, then $\pi := \tau$, and go to 2.
+	\item $\pi$ is optimal.
+\end{enumerate}
+
+\paragraph{Problem} Therer are to many ways to improve a policy.
+For example if a state $s$ enables two actions, then any
+mixed policy $(p, 1-p)$ may be tried locally. This gives
+uncountably many (i.e., $|\mathbb{R}|$) alternatives!
+
+Fortunally for discounted decision problems, locl search can be
+restricted to pure policies only.
+
+Let $\pi$ be apossibly mixed policy. For each state choose a decision
+that is locally maximal:
+
+% I'm not sure where the J is supposed to go.
+\[M(i) \in argmax\{r_{ij} + \gamma V_\pi(j)\} \]
+
+Define the pure policy:
+
+\[ \tau_{ij} =_{Def} \delta_{M(i),j} \]
+
+Then $\tau$ is a global improvement of $\pi$.
+
+\subsubsection{Partial correctness}
+
+Correctness  partial correctness + termination.
+
+Partial correctness means: correct if the algrotihm terminates.
+\emph{Policy iteration is partially correct:} any strict
+local improvement induces a global improvment.
+\emph{Policy iteration terminates:} each time a policy is improved,
+the new policy is strictly better than the previous.
+
+\subsection{Bellman optimallity}
+
+If $\pi$ is optimal, it follows from the policy improvement thearom that
+
+\[ V^\pi(i) = \underset{all\mbox{\ }j}{max}[r_{ij} + \gamma V^\pi(j)] \]
+
+Removing explicit reference to the optimal policy automatically (more elegant).
+
+\subsubsection{operator}
+%seriously what is meant with operator?
+
+% I'm assuming he means with exact the algebraic method and
+% the aproximation is the itteration method.
+For both methods (the exact as well as the approximation method),
+a compacter notation is needed.
+
+\[B: \mathbb{R}^n \to \mathbb{R}^n : v \mapsto
+	\underset{all\mbox{\ }j}{max}[C_j + \gamma V(j)] \]
+
+Where $C_j$ is the vector of rewards that emanate from state $i$.
+Antoher way to write $B:$
+\[\underset{all\mbox{\ }j}{max}[r_{ij} + \gamma V(j)] \]
+
+$B$ does not depend on a policy or transition matrix. ($B_p$ does.)
+
+\subsubsection{Fixed point}
+The value-vector, $v^*$, of a decision problem is a fixed point of the
+$B$-operator:
+\[ B(v) = v\]
+
 \subsection{Exam}
@@
 
+% That's all she wrote.
+
 \end{document}

### 2015-11-03 agents/summery.tex
added some very important comments.
added some text about kripke!

@@
 \documentclass{article}
+% I just added all the things, if you have problems build, try deleting the
+% naggin packages, fat chance I don't even use it.
 \usepackage{amsmath}
@@
 \section{Modal logic}
+% Before starting I just want to lay out the idea here, everytime I hear
+% kripke, I imagine someone waving his arm and shouting KRIPKE in anger.
+% Not that anyone will ever read these comments but its a funny idea.
+
+This course was thought by John-Jules Meyer. The `theoretic' part focused
+on the ideas behind the logics, while the `practical' focussed on writing
+propper proofs. I guess I'll treat both in here.
+
+There are 2 main sources of information I'll use, the slides by John Jules
+and the modal logic paper for artificial intelligence by rosja mastop to help
+me with the practical part. 
+% In truth I didn't understand much of the practicals, just giving a bunch
+% of examples is not a good way of teaching how to write proofs.
+% This paper seems to explain it in a more structural way.
+
+\subsection{Basic modal logic}
+Express intensional (context/situation-sensitve) notations such as: 
+Knowledge, Belief, Obligation, Action and Time.
+
+\subsubsection{Modal langauge}
+
+Imports all the things from propositional logic and is extended
+with the modal operators ($\Box\phi,\Diamond\phi$) which can be read as:
+
+\begin{tabular}{ll}
+	Symbol & Definition \\ \toprule
+	$\Box\phi$ & It is \emph{necesarry} that $\phi$ \\
+	$\Diamond\phi$ & it is \emph{possible} that $\phi$
+
+\end{tabular}
+
+\noindent
+Also $\neg\Box\neg\phi = \Diamond\phi$ and
+$\neg\Diamond\neg\phi = \Box\phi$. So you could theoratically
+drop one to create the other.
+% where $\phi$ is an object I guess?
+
+\paragraph{Semantics} Kripke %KRIPKE, *waves arm*.
+model/frame $M=\langle S,\pi,R \rangle$
+
+\begin{tabular}{ll}
+	Symbol & Definition \\ \toprule
+	% worlds implies mutability, states, well a state is by definition 
+	% imutable. I'm pretty sure we want immutability in these kind 
+	% off logics
+	$S$ & Is a set of \emph{states (or worlds)} \\
+	$\pi$ & Is a truth assignment function $\pi : S \times AT \to {tt, ff}$ \\
+	$R$	& $R \subseteq S \times S$ is an \emph{accesibility relation}
+\end{tabular}
 

### 2015-11-03 agents/summery.tex
added most of the basics of KRIPKE!!!

@@
 	$\pi$ & Is a truth assignment function $\pi : S \times AT \to {tt, ff}$ \\
-	$R$	& $R \subseteq S \times S$ is an \emph{accesibility relation}
+	$R$	& $R \subseteq S \times S$ is an \emph{accesibility relation}\\
 \end{tabular}
+
+\subsubsection{Interpertation}
+
+\begin{center}
+\begin{tikzpicture}[auto, >=stealth', auto, semithick, node distance=3cm]
+\tikzstyle{every state}=[fill=white,draw=black,thick,text=black,scale=1]
+\node[state]    (S)            {$\neg p, q$};
+\node[state]    (A)[below of=S]{$p, \neg q$};
+\node[state]    (B)[right of=S]{$\neg p, \neg q$};
+\node[state]    (C)[right of=B]{$p, q$};
+\node[state]    (E)[below of=B]{$\neg p, q$};
+\path
+(S)   edge[above]     		(A)
+(S)   edge[above]     		(E)
+(S)   edge[above]     		(B)
+(B)    edge[above]     		(E)
+(E)    edge[above]     		(C);
+\end{tikzpicture}
+\end{center}
+
+Like classical propositional logic, but now realtive to a model and a state:
+
+\[ M,s \vDash \phi \]
+
+thus:
+
+\begin{tabular}{lll}
+	Modal & $\Leftrightarrow$ & Interpertation \\ \toprule
+$M,s \vDash p$ & $\Leftrightarrow$ & $\pi(s,p) = tt$ \\
+$M,s \vDash \phi \wedge \varphi $ & $\Leftrightarrow$ & $ M,s \vDash \phi
+\mbox{ and }M,s\vDash \varphi$ \\
+$M,s \vDash \Box \phi $ & $\Leftrightarrow$ & $ M,t \vDash \phi $ for 
+\emph{every} $t$ such that $R(s,t)$ \\
+$M,s \vDash \Diamond \phi $ & $\Leftrightarrow$ & $ M,t \vDash \phi $ for 
+\emph{some} $t$ such that $R(s,t)$ \\
+\end{tabular}
+
+\subsubsection{Validity}
+\begin{tabular}{llll}
+	Definition & Denotation & $\Leftrightarrow$ & Interpertation \\ \toprule
+
+$\phi$ is \emph{valid} in a model $M= \langle S, \pi, R \rangle$ &
+$M\vDash\phi$ & $\Leftrightarrow$ & $M,s \vDash \phi$ for all $s \in S$ \\
+
+$\phi$ is \emph{valid}  &
+$\vDash\phi$ & $\Leftrightarrow$ & $M\vDash\phi$ for all Kripke models $M$ \\
+
+$\phi$ is \emph{valid} wrt class $\zeta$ &
+$\zeta\vDash\phi$ & $\Leftrightarrow$ & $M\vDash\phi$ for all Kripke models
+$M \in \zeta$ \\
+\end{tabular}
+
+The last one doesn't occur often.
+
+\subsection{System K}
+We try to  axiomatize validities.
+
+\paragraph{Axioms} We import all (or enough) propositional tautologies.
+The K axiom ($(\Box\phi \wedge \Box (\phi \to \varphi))\to\Box\varphi$) is added.
+
+\paragraph{Rules} Modes ponens
+\begin{prooftree}
+	\AxiomC{$\phi$}
+	\AxiomC{$\phi \to \varphi$}
+	\BinaryInfC{$\varphi$}
+\end{prooftree}
+
+Necisitation rule:
+\begin{prooftree}
+	\AxiomC{$\phi$}
+	\UnaryInfC{$\Box\phi$}
+\end{prooftree}
 

### 2015-11-05 cogniveProcessing/summery.tex
added some more, givig up

@@
 pregnacies.
+\subsubsection{Transduction}
+converts physical (in real world, such as sound or pressure) signals to neural
+It can use chemical mechanical both or neither.
 
-\subsubsection{Receptive fields}
+\subsubsection{Transfer}
+From the transduction, to the brain, or CPU.
+%I guess a brain counts as a
+%processing unit and its pretty central so, yeah CPU fits fine.
+
+\subsection{Receptive fields}
 Description depends on modality. Somatosensory, Auditory, Visual, Gustation.
@@
 if its not affected.
+
+\subsubsection{space}
+The relation to space is well maintained in early sensory processing.
+
+\subsubsection{response profiles}
+What is the relation between a stimulus and a response?  Its an S curve.
+How is the strenth of an external stimulus related to its proccessing?
+
+\subsection{Adoptation}
+Sensory responses are not stable over time.
+
+The frequency of output of a snesory cell goes down over time. After
+excitiation
+the cell becomes more negative, becuase it pushed out all these postive ions
+that entered during the spike. In the period just after the spike no positive
+ions are allowed in. This makes it harder to reach the threshold for an
+action potential.
+
+Neurons adapted to their prefered stimulus are hyperpolarized. Repatitive
+presynaptic stimulation depletes a readily releasable vescile pool.
+It takes time to refill the readily releasable pool.
+
+\subsection{Simple \& complex signals}
+Neurons are used for information processing. They take information from
+one or multiple sources and trn it into something else. Usually a more
+complex form of information.
+
+\paragraph{But} perception does not coincide with reality. Think of all the
+illusions you may have seen on the web.
+
+\subsubsection{Light reaches the retina}
+Beofre light reaches the lens it already gets transformed. The cornea
+(outside harder part) filters UV-light. Another interseting feature is the
+layout. The axons of ganglion cells, which sends onfio to the brain,
+are actually in front of the recoptors that decode light.
+This already has a profound effect on vision the blindspot.
+
+\subsubsection{Responding to light}
+1 type of photo recoptors: rods and cones. They differ in the type of light
+they respond to. They prefer different wave lengths. The cones will later
+be used to figure out color. The rods are more sensitive to light, and
+are used mostly in dark envioronments. They respond to light rather simliarly.
+Both are filled with pgiments and this pigment is what gets effected by light.
+Photons kind of destroy it.
 

### 2015-11-05 cogniveProcessing/summery.tex
added more texts :*(

@@
 
+\subsection{Processing vision}
+\subsubsection{Horizontal cells}
+It processes it with cirlces and there is some basic logic
+going on when the signals leave the eyes (nand gates and such).
+
+\subsubsection{V1}
+Processes light and dark paterns, the cells aren't to pickey (always give
+of some signal) but some have preference for certain orientations.
+
+\subsubsection{V2}
+% A heuristic layer?
+Responds to known rules of the visual world. Patterns we are used to seeing.
+
+Receptive fields go from simple and small to progresivley complex and large.
+
+\subsubsection{After V2}
+Things become very complex very fast.
+
+\subsection{Scanning techniques}
+
+\subsubsection{MRI}
+Magnatic resonanse imaging, fMRI is the same but uses
+the diff of 2 images. MRI always uses blood oxygion
+levels to get an image.
+
+\paragraph{DPRI}
+Same as MRI but measures water instead of oxygian.
+This can potantioally image communcation between parts
+of the brain. But its a work in progress.
+
+\subsection{EEG}
+Measure the activity in the brain, great temporal res,
+but bad spatial res.
+
+\paragraph{ECoG}
+EEG with open skull. Great temporal and good spatial res.
+But its invasive, so there is little opertunity to o this.
+
+\subsubsection{MEG}
+Best thing, has high temporal and spatial resolution.
+But is very expensive.
+
+\subsubsection{TMS}
+The zapping. It does not give an image but it interferes
+with brain activity instead by sending and EMP.
+% I'm going to stay away from this one.
+
+\paragraph{rTMS} Long term TMS by using repeated zaps.
+It may work as an antidrepresant, but drug companys try to fight these 
+results.
+
+\subsubsection{Drugs}
+
+You can also use drugs to mess with the brain.
+
 \section{Langauge}

### 2015-11-10 evolutionaryComputation/notes.tex
decided to throw all the homework just in one git repo,
Saves a lot of repo creation on github, and thus a lot of time

# seminars

talk about it half an hour
discuss 15 minutes after it.

Slides should be good enough.

First come first serve, pick top 3
In groups of ~3

# Practical assignments
Artifical benchmarks.
... next thursday, groups 1 or 2 students
Not very complicated, can use any programing langauge.

prefered to do it in groups.

Link, can download C code.

# What is it?
No domain knowlegde of your problem.
Black box algorithms,
	good thing, it works on a lot of problems
	bad thing its slow.

There exists hybred form which combines evolutionary
with specialised local algrotims.
	These are good at jumping out of local optimums

# Key concepts of darwanian system
Vaiation should exists in the copies, not to much not to little
Can add a little change or combine a solutions.

Inheritance, if the parents are good, then the childrens are to probably.

You have to destroy things. otherwise you never end up with something new.

Compitition, there should be competition otherwise there is no evolution.

# Evolutionary algoritm
1. Generate purly randomly or get a good start solution.
2. if you start with a 'good start' then you isolate your searchspace. if the
population stay similiar.

hybred no mutation, only selection compitition and inheritance.

while Not terminated do:
	1. select mates(T)
	2. x= generate offspring (T)
	3. evaluate population (x)
	4. T = select fittest.
return T

Don't select to *hard*, then after x generation your population will converge.

These algoritms can be very creative.

#Selection
##Tournamnet selection:
	S1 f1
	s2 f2 <- third
	s3 f3 <- best one
	s4 f4 <- second

	$p(t) -> p^s(t)$

	s3 f3
	s4 f3
	s3 f3
	s2 f2

The best will have 2, the wors 0, and the medium 1.
selection based on a grid, only compete against those next to you.

##Truncation selection:
Sort on fitness, take the top 1/x, and give them x copies for the new
population.

##Proportional select

$P^{select}(s_1) = frac{F_i}{\sum_{j=1}F_j}$

The next generation will be proportionally filled with how good the previous
was. This has the problem that the population will converge very fast.

Its not really for pragmatics, but better for theorticall analises.
So you should always use tournament or truncation.

You don't need to have all darwenin properties,
Inheritance, representation, compitition.

IRL function evaluation is usually expensive.

#Varation
The closer you can stay to your representation, then the crosover operators
preservers it. So binary representation is bad.

## MUtation
Just pick something and change it.

## crosover
The child inherints what the parents agree on,
where not the child can use random.
### uniform
flip a coin for every position

### 2point
pick 2 random points and cross them over.
positionally biased.

### 2015-11-10 inteligentagents/notes.tex
decided to throw all the homework just in one git repo,
Saves a lot of repo creation on github, and thus a lot of time

#Exam

You can use all material except that about eleborated excersizes.
You should be able to work with it. You can mark stuff.

#known
Stuff in mehthods of AI should be known
Modal logic is very important.

meyer & van der hoek, epistamic logic for AI and computer science, cambridge
univ press.
Very usefull for self study. Because it has excersizes with eleborations in
the back.

#Inteligent agents

Has to do with behaviour, try to mimic bahavior from the natural world in
software.

Agents, has many definitions but in this course its something with a
coginitive state.
Believe desires intentions - in a methorphorical stance, so not really

## Complete autonomy
Is kindoff useless, we need limited autonomy.
We need some kind of initiative or proactive or goal-directed.

##Mostly described in having mental states.
BDI 'strong notion of agency'

Show informational and motivational attitudes. an agent has a
motivation system of its own. It can be programmed, but then
the agent is on its own.

#Properties of agents.

##Situated
It can sense and reason about the environment.

They don't just give advice like exper systems

##reactive

## proactive

persuin goals, make plans.

## social
compete with eachorther.

#Objects vs agents

objects have some autonomy, but agents its more involved.

Agents are more 'subject' than 'object' in linguistic

objects can just be used, can be wrong but ok. An agent however
can only ask, so it can be refused. This isn't failing
its just the way it works.

... agents are objects with an attitude... wil of its own.
This is just an idea and there is an entire course about how we can do this.
It should be logical.

#Multi agent systems
After christmess.

There are quite difficult logics to describe their behavior.
You want multiple agents to wrok on a complex task, so you want to devide
the labour. if one fails others can take over.
Gracefull degradation. <- if somethign is wrong not everything is over

## why MAS

also paralism
roles of which a task can be done

## application

Softbots

### Robots
The rottordam harbour, the cranes are agent oriented. work in progress though.
Self driving cars, that could comminucate....

Replace the controller by an agent. Cognotive robotics, robots that think.
A robot steared by an agent.

Itel?? agents/robots exchanging goals and believes.

### Industrial
Airtraffic, one of the first applications of agents.
Manufactiring more agile, which can change easily

### Comericial
Automatic auctions. specifiy limits and the agent will negotiate.

### Entertainment
Virtual characters, more intelligent. Common sense.
Cinema <-- titanic was the first one that used agents.

# courses
##Philosophy
sloman: <- emotional robots

###Artitsotle
Practical sylogism, reasoning about actions.

### Denet
Strong AI, the computer can implement thinking like a human, as complex the
human brain.

###BRatman intetnions
Something to do with desires. But he thought about how to implement it with
software.

##Logical foundations
It needs to be combined. Quite complicated.

##Reasoning about action and frame
Frame problem, implementation of actions does not change much, but the rest does.

Situation calculus, <- to controll robots

##Agent architectures

### 2015-11-10 methods/agents/summery.tex
decided to throw all the homework just in one git repo,
Saves a lot of repo creation on github, and thus a lot of time

--- 2015-11-03 agents/summery.tex
@@
 $M,s \vDash \Diamond \phi $ & $\Leftrightarrow$ & $ M,t \vDash \phi $ for 
-\emph{some} $t$ such that $R(s,t)$ \\
+\emph{some} $t$ such that $R(s,t)$ \\\bottomrule
 \end{tabular}
@@
 $\zeta\vDash\phi$ & $\Leftrightarrow$ & $M\vDash\phi$ for all Kripke models
-$M \in \zeta$ \\
+$M \in \zeta$ \\\bottomrule
 \end{tabular}
@@
 
+\subparagraph{Caution} the necisitation is different from the invalid
+assertion: $\phi \to \Box \phi$.
+
+\subsubsection{Derivability in K}
+A \emph{derivation} of a formula $\phi$ is a finite sequence of formulas
+$\phi_1, \dots \phi_n = \phi$. where each $\phi_i$ for $1 \ge i \ge n$
+is either an instance of the acioms (or rather axiom schemes), or the
+conclusion of one of the rules of which the premises have
+been derived already, i.e{.} appear as $\phi_j$ in the sequance with
+$j < i$.
+
+% sad thing is I understand this crap now.
+When we can derive an epistemic formula $\phi$ by using the axioms
+and rules of $K_{(m)}$ we write $K_{(m)} \vdash \phi$
+
+System $K_{(m)}$ is \emph{sound and complete}. This means
+that exactly ll valid modal assertions can be obtained by derivations in
+system $K_{(m)}$.
+
+\[\vDash \phi \Leftrightarrow K_{(m)} \vdash \phi\]
+
+% I leave out the (non) theoroms, it seems overkill.
+
+\subsection{Dynamic logic}
+An example of an (indexed) version of system $K$ is dynamic logic,
+wehre the $\Box$ modality is associated with the execution results of
+a program / action.
+
+$\Box_\alpha$, normally written $[\alpha]$ 
+with reading $[\alpha]\phi$: after exuction of $\alpha$ it holds that $\phi$
+
+\[<\alpha>\phi = \neg [\alpha] \neg \phi \]
+
+\subsubsection{Semantics}
+Accessibility relation $R_\alpha$ for every action $\alpha$
+
+\begin{tabular}{llll}
+	% I added the natural collumn, because otherwise I won't remember.
+	% Translating something in a different notation is stupid.
+	notation & = & definition & natural \\ \toprule
+	$R_{\alpha;\beta}$ & = & $R_\alpha \circ R_\beta$ &
+	execute from $R_\alpha$ until including $R_\beta$\\
+	$R_{\alpha+\beta}$ & = & $R_\alpha \cup R_\beta$ &
+	execute $R_\alpha$ or $R_\beta$\\
+	$R_{\alpha*}$ & = & $R_\alpha*$ &
+	Execute an arbitrary times \\ \bottomrule
+\end{tabular}
+
+\subsubsection{Interpertation}
+\begin{tabular}{lll}
+	Modal & $\Leftrightarrow$ & Interpertation \\ \toprule
+	$M,s \vDash [\alpha]\phi$ & $\Leftrightarrow$ &
+	for all $s'$ with $R_\alpha(s,s'): M,s' \vDash \phi$ \\
+	$M,s \vDash <\alpha>\phi$ & $\Leftrightarrow$ &
+	for some $s'$ with $R_\alpha(s,s'): M,s' \vDash \phi$ \\\bottomrule
+\end{tabular}
+
+\subsection{Special properties of accessibility relations}
+\begin{tabular}{lll}
+	R is \dots& if\dots & natural language \\ \toprule
+	reflexive & $\forall s \in S(s,s) \in R$ & all s are linked with 
+	themselves\\
+
+	transitive & $\forall s,t,u \in S:(s,t) \in R \& (t,u) \in R \Rightarrow
+	(s,u) \in R$ & \brcell{if s is related to t, and t is related to u,  \\
+		then s has a relation with u, \\ in other words, you can make shortcuts}\\
+
+	symmetrical &$\forall s,t \in S:(s,t) \in R \Rightarrow (t,s) \in R$ &
+	undirected, can go back and forth\\
+	euclidiean & $\forall s,t,u \in S:(s,t) \in R \& (s,u) \in R \Rightarrow
+	(t,u) \in R$ & \brcell{if s has a relation with t and u, then u and t \\
+	have a relation to, you can make triangles.} \\
+	serial & $\forall s \in S \exists t \in S (s,t) \in R$&
+	\brcell{All states in S are connected to \\ at least one other state in S}\\
+	an equivalance relation& R is \emph{reflexive, transitive and 
+	symetrical}& \brcell{the states R links to have something in \\ 
+	common (not equal?)} \\ \bottomrule
+\end{tabular}
 \section{APL}

### 2015-11-10 methods/cogniveProcessing/freeasociations/report.tex
decided to throw all the homework just in one git repo,
Saves a lot of repo creation on github, and thus a lot of time

\documentclass{article}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\usepackage{qtree}
\usepackage{tikz}

\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

\begin{document}
\author{Jappie Klooster}
\title{Free asociations}
\maketitle

\section{Thoughts on results}
The first friend I asked didn't took the
test serious (thats why the associations don't make much sense), I'm not
sure if I should remove this data or just keep it.

The personal category was added for words that really don't have any
relation to the other word, this is justified because some poeple might
have a strong memory with a particular word. For example if your best
friend just died and sombody tries to do this test and says friend,
death might be the first response.

Coolocation and coordination are by far the most common responses.

It should be noted that results 3 till 10 are done on my classmates,
so they are biased. Although I don't think it really matters for this
kind of test, because you should say the first thing that comes to mind.
If people want to give a different anwser they're gonna give a different
anwser. Regarldess of training.

\section{Reactions on colecting}
The first 2 participants I asked where some friends of my. I did this trough
Whatsapp. They were both quite curious about why I did it. They were also
interested in the `results'.

I think people often think you are testing them personally rather than
just collecting some data. At least thats how I felt when I got tested in class.
Because of this I think people will try to do extra `their' best.
Or in case of my first friend, just troll it.

The other 8 participants were classmates. So they required a lot less explaining.

\section{Comments on the assignment}
It is rather difficult to find subjects. But once you managed to find the
subject the assignment is quite easy. I'm not sure I did it right but I
think I did.

I think that the assignment small for group work,
giving it a week before the exams isn't great either.
\end{document}

### 2015-11-10 methods/cogniveProcessing/summery.tex
decided to throw all the homework just in one git repo,
Saves a lot of repo creation on github, and thus a lot of time

--- 2015-11-05 cogniveProcessing/summery.tex
@@
 
-\section{Langauge}
+\section{Langauge structure}
 Trying to figure out what langauge is.
@@
 
+\subsection{Felicitous}
+Helpfull to the conversation. A sentence can be this or not. This
+works better than a simple truth value, because it can also handle
+state changing sentences, such as: ``I declare war''.
+
+\subsection{Cooperative principle}
+A conversation has a goal, that the speaker wants to contribute to, assuming
+the hearer does the same.
+
+\subsubsection{Quality}
+Speak the truth, and don't say things you don't have evidence for.
+
+\paragraph{Example} B has stolen A's chocoletes
+A: ``have you stolen my chocolates?''\\
+B1: {\color{green} ``Yes.''}\\
+B2: {\color{red} ``No.''}
+
+\subsubsection{Quantity}
+Don't give to much or too little information
+
+\paragraph{Example} B knows A's chocolates are on the table.
+A: ``Where are my chocolates?''\\
+B1: {\color{green} ``On the table.''}\\
+B2: {\color{red} ``Somewhere.''}\\
+B2: {\color{red} ``On the table. Which is an oaken item of furniture
+located on the other end of this living room''}
+
+\subsubsection{Relevance}
+Make relevant contributions.
+
+\paragraph{Example} B has not stolen A's chocolates.
+A: ``Someone has stolen my chocolates?''\\
+B1: {\color{green} ``That's terrible.''}\\
+B2: {\color{red} ``The weather's really nice today, don't you agree?.''}\\
+
+\subsubsection{Manner}
+Be clear and bried, don't do obfuscation.
+
+\paragraph{Example} 
+A: ``Tell me about the performance''\\
+B1: {\color{green} ``Mrs Smith sang \emph{Bridge over Troubled Water.}''}\\
+B2: {\color{red} ``Mrs Smith produced a series of soundwaves that resembled
+the melody of \emph{Bridge over Troubled Water}.''}\\
+
+\subsection{Impliciture}
+A: ``Where cna I find a service station?''\\
+$B_1$: ``Around the corner. \emph{My mother wokrs there''}.\\
+$B_2$: ``Around the corner. \emph{But I'm not sure wether it's open''}.
+
+Both $B_1$ and $B_2$ violate quantity, but unlike $B_1$, $B_2$ is 
+felicitous. Why?
+
+If B had known the service station would be closed and useless
+to A his remark would have violated relavance.
+
+So A concludes: ``B thinks that the service station is open''.
+
+B, knowing this preemts this conclusion by adding ``but I'm not sure it's 
+open'', so violating a maxim cna be communicatively meaningfull (to preemt
+wrong conclusions).
+
+So B thinks about how his message will be received by A, and decides that
+it will be interpeted wrongly so he adds information.
+
+An inference like $A's$ (``$B$ thinks the service station is open'')
+is called an \emph{implicature}
+
+\subsubsection{Two flavours}
+There are two flavours of implicuture: \emph{conevnetional} and 
+conversational.
+
+\paragraph{Conversational} Implicatures: the information that conversational
+partners infer from each other's utterances, based on the gricean maxims.
+
+\subparagraph{Not entailment!} Conversational implicatures can be canceled.
+
+\paragraph{Conventional} implicatures: non-cancellable implicutres
+that are part of the conceptual meaning of a word.
+
 \end{document}

### 2015-11-10 methods/reasoning/huiswerk/inl1026.tex
decided to throw all the homework just in one git repo,
Saves a lot of repo creation on github, and thus a lot of time

--- 2015-10-24 reasoning/huiswerk/inl1026.tex
@@
 \draw (3,3) circle (3.5);
-\node (n1) at (1,3) {$A \Rightarrow A$};
-\node (n2) at (3,5) {$B \Rightarrow A$};
-\node (n3) at (4,2)  {$C \wedge B, A \Rightarrow A$};
+\node (n1) at (1,3) {$A$};
+\node (n2) at (3,5) {$B$};
+\node (n3) at (4,2)  {$C \wedge B$};
 \foreach \from/\to in {n1/n2,n2/n3}
 \draw (\from) -- (\to);
-
 \end{tikzpicture}
 
-, if
+In this model you can see that the preceding sequant does not automaticly
+imply an A. So the premise ($\Gamma \Rightarrow A$) is false, 
+which means the conclusion is also false,
+there is no path available form every state to an A.
+
  If $C$ $I(\Gamma \Rightarrow A)=1$ then
 so does $I(\Pi, \Box \Gamma \Rightarrow \Box A, \Delta) = 1$ for example:
+\\
 \begin{tikzpicture}
@@
 \draw (3,3) circle (3.5);
-\node (n1) at (1,3) {$A \Rightarrow A$};
-\node (n2) at (3,5) {$B, A \Rightarrow A$};
-\node (n3) at (4,2)  {$C \wedge B, A \Rightarrow A$};
+\node (n1) at (1,3) {$A$};
+\node (n2) at (3,5) {$A \wedge B$};
+\node (n3) at (4,2)  {$A \wedge (C \vee B)$};
 \foreach \from/\to in {n1/n2,n2/n3}
 \draw (\from) -- (\to);
-
 \end{tikzpicture}
+\\
+In this model every state A holds at least.
+So there is also always a path available to a state were A holds.
+This means the premise ($\Gamma \Rightarrow A$) and conclusion
+($\Pi, \Box \Gamma \Rightarrow \Box A, \Delta$) are true.
 
@@
 the same. They're both intendid to remove the box on the right side and to
-hide it away in the left side in one of the `garbage' spots $(\Pi$ and 
-$\Delta)$ of the axioms.
+hide it away in the left side in one of the `garbage' spot ($\Gamma$) of
+the axioms.
 
-However the $GKC_{K4}$ has also garbage spots in the conclusion. So to show
+However the $GKC_{K4}$ has also garbage spots $(\Pi$ and
+$\Delta)$ in the conclusion. So to show
 something is proofable in $GKC_{K4}$ and not in $G$ all you need to do is
@@
 
-\[A, \Box B \Rightarrow \Box B, A\]
+\[A, \Box B \Rightarrow \Box B\]
 
@@
 	\RightLabel{$R$}
-	\UnaryInfC{$A, \Box B \Rightarrow \Box B, A$}
+	\UnaryInfC{$A, \Box B \Rightarrow \Box B$}
 \end{prooftree}
@@
 
+$(\Rightarrow \Box A)$ is derivable in $GKC_K$ this was given.
+To proof $(\Rightarrow A)$ is derivable in $GKC_K$, this can be done by
+tranforming $(\Rightarrow \Box A)$ with the rules of $GKC_K$ into 
+$(\Rightarrow A)$.
+
+\begin{prooftree}
+	\AxiomC{$\Rightarrow A$}
+	\RightLabel{$R_K$}
+	\UnaryInfC{$\Rightarrow \Box A$}
+\end{prooftree}
+
+This is what had to be shown.
+
 \end{document}

### 2015-11-10 methods/reasoning/notes.tex
decided to throw all the homework just in one git repo,
Saves a lot of repo creation on github, and thus a lot of time

\documentclass{report}
\usepackage{booktabs}
\usepackage{graphicx}
\graphicspath{{images/}}

\usepackage{amsmath}

\begin{document}
\author{J. T. Klooster}
\title{Reasoning track notes}
\maketitle
\tableofcontents

\chapter{Examples}

\begin{itemize}
	\item Software for medical diagonostics
			Figure out how and what diseases a patiant has.
	\item Software for games
			This is different from diagnostics.
\end{itemize}

there is often a common core. After this course it should become easier to
work with logic because you've developed your own as a refernce pont

\chapter{Logic}
stuff you need to know

\begin{itemize}
		\item Classical logic, important
		\item Epistemic logics
		\item modal logics, not really that important
		\item logics for linguistics
		\item fuzzy logics, seems important?
\end{itemize}

Logic should be decidable, can you solve the problem, for example:
if you have $L$, does $A$ hold in $L$?
does $L$ hold in $A$ \\

It might be difficult to combine this with expressive power.

\section{Predicate logic}

Predicate logic is undeidable. Propositional logic is but its weaker. \\

Description logic, is between predicate and propositional. But it isn't
used in this course. Description logic has been devloped to be used in practice,
meaning it wasn't really developed my mathematicans. But its very succesfull but.

\subsection{proofs}

If it is asked that something is proofed take effort to do it. No sloppy work!

\subsection{propoistional logic}

\begin{tabular}{@{}lll@{}}
$\Rightarrow$ & implication & if A is true, then B is true \\
$\iff$ & bi-implication

\end{tabular}

\subsubsection{CPC}
Clasical propositional logic $\wedge$ clasical propositional computation

\section{decidable}

Logic is decidable if there is a turing machine when presented with a formula
$\phi$ as input, outpus 1 if $\phi$ holds in the logic and 0 otherwise. \\

Valdity in classical propoistional logic CPC is decidable. \\
Validity in claical logic CPQ is undicidable \\

Something with P vs NP

\section{sound and complete}
A proof system is sound and complete for a sementics if and only if formula
is true acording to the semantics exeactly when it is dervable in the proof
system.

\section{Hilbert}

hilbert is difficult to work with. Despair in 5. 4. 3. \dots

There are a bunch of axioms which just define the operators I think.
And there is one rule the Modus Ponens.

\subsection{axioms}

all the actioms have a name.

W says weakining and E says Exchange.

\subsection{nasty part}

finding the right axiom

\subsection{premises}
also called assumption.

\subsection{Lemma}

\section{Proof}
A proof of $A$ is $A_1 \dots A_n$ such that:
\begin{itemize}
	\item $A_n=A$
	\item For all $1<= i <= n$ $A$ is an axiam in system \\
		there are $j,h<i$ such that $A_j A_h / A_i$
\end{itemize}

\end{document}

### 2015-11-10 methods/reasoning/summery.tex
decided to throw all the homework just in one git repo,
Saves a lot of repo creation on github, and thus a lot of time

--- 2015-10-31 reasoning/summery.tex
@@
 talking about and what you're going to do. So you start with the definitions.
+Remeber the definitions should be stupid, only add obvious things, don't do
+deductions until the body of the proof.
 Then you say what you're going to proof (by using the defintions), so here
@@
 
+So for the base case the tabloo is method is used to show that the valuation
+trinceling down happens for the rules, then induction is used for the entire
+proof tree to show that it works for the entire system.
+
 \subsubsection{How to use it?}
@@
 if one of the axioms is false it means the conclusion is false to.
+
+\paragraph{Proof that for $L\to$ of GKC that if one of the premises is not
+true (under all valuations) than so is the conclusion}
+
+\subparagraph{Definitions} GKC is gentzen calculy, in which $L\to$ is defined
+as follows:
+
+\begin{prooftree}
+	\AxiomC{$\Gamma \Rightarrow  A, \Delta$}
+	\AxiomC{$\Gamma, B \Rightarrow \Delta$}
+	\RightLabel{$ L\to$ }
+	\BinaryInfC{$\Gamma, A \to B \Rightarrow \Delta$}
+\end{prooftree}
+
+Where $\Gamma$ is a collection of formulas connected by $\wedge$ and $\Delta$
+is a collection of formulas connected by $\vee$.
+A and B are singular formulas.
+I(x) is a function that changes the $,$ into proper connectives. So in case
+of the antecedent it will become a $\wedge$ and in the succedent a $\vee$.
+
+\subparagraph{Strategy} Under all valuations, if one of the premises is not
+true of $L\to$ then so is the conclusion. This will be proven with help
+of the tabloo method.
+\subparagraph{Proof}
+There are two cases to consider first $V(I(\Gamma \Rightarrow  A, \Delta))=0$
+and secondly $V(I(\Gamma, B \Rightarrow  \Delta))=0$
+
+$V(I(\Gamma \Rightarrow  A, \Delta))=0$ implies $V(A)=0$ and for all formulas
+where $C\in \Delta$ and $D \in \Gamma$, $V(C) = 0$ and $V(D)=1$. In this case
+$V(A \to B) = 1$ and thus $V(I(\Gamma, A \to B \Rightarrow \Delta))=0$.
+
+\emph{So here I say if the $A$ premise is false then the conclusion is false
+	because propositional logic, you need the $I$ function to deal with
+	commas. And you need the V function to execute a true false valuation
+(ergo execute propositional logic)}
+
+$V(I(\Gamma, B \Rightarrow \Delta))=0$ implies $V(B)=1$ and for all formulas
+where $C\in \Delta$ and $D \in \Gamma$, $V(C) = 0$ and $V(D)=1$. In this case
+$V(A \to B) = 1$ and thus $V(I(\Gamma, A \to B \Rightarrow \Delta))=0$.
+
+\emph{So step by step, say what the valuation for the premis implies for the
+	individual elements (if its a set of elements, just do it on a member of
+the set, then show it for the conclusion with help of propositional logic}
 

### 2015-11-10 inteligentagents/kripke.tex
KRIPKE!!!

--- 2015-11-10 methods/agents/summery.tex
@@
 \author{Jappie Klooster}
-\title{Agents and you, a brief summary}
+\title{Logic, Kripke and Chaos}
 \maketitle
@@
 \section{Introduction}
-This will be an introduction to agents coming from someone who followed
+This will be an introduction to modal logic coming from someone who followed
 ` hbo informatica' as pre study.
 
-In here I will write down the things I think are important.
+In here I will write down the things I think are important for understanding
+modal logic and how to proof things in that system.
 
-\section{Modal logic}
 % Before starting I just want to lay out the idea here, everytime I hear
@@
 
-\subsection{Basic modal logic}
+\section{Basic modal logic}
 Express intensional (context/situation-sensitve) notations such as: 
@@
 
-\subsubsection{Modal langauge}
+\subsection{Modal langauge}
 
@@
 
-\subsubsection{Interpertation}
+\subsection{Interpertation}
 
@@
 
-\subsubsection{Validity}
+\subsection{Validity}
 \begin{tabular}{llll}
@@
 
-\subsection{System K}
+\section{System K}
 We try to  axiomatize validities.
@@
 
-\subsubsection{Derivability in K}
+\subsection{Derivability in K}
 A \emph{derivation} of a formula $\phi$ is a finite sequence of formulas
@@
 
-\subsection{Dynamic logic}
+\section{Dynamic logic}
 An example of an (indexed) version of system $K$ is dynamic logic,
@@
 
-\subsubsection{Semantics}
+\subsection{Semantics}
 Accessibility relation $R_\alpha$ for every action $\alpha$
@@
 
-\subsubsection{Interpertation}
+\subsection{Interpertation}
 \begin{tabular}{lll}
@@
 
-\subsection{Special properties of accessibility relations}
+\section{Special properties of accessibility relations}
 \begin{tabular}{lll}
@@
 \end{tabular}
-\section{APL}
-Medhi dastani taught about the agent programming langauge (APL).
-\subsection{Agent data structure}
-This represents the mental state of a agent. \\
-\begin{tabular}{ll}
-	Name & Does what \\ \toprule
-	Beliefs & Information available to the agent \\
-	Goals & Objectives that the agent wants to achieve \\
-	Events & Obsevations of (environmental) changes \\
-	Capabilities & Actions that the agent can perform \\
-	Plans & Procedures to achieve objectives \\
-	Reasoning rules & Reason about goal and plans \\
-\end{tabular} \\
-
-\noindent
-The rules can be:
-
-\begin{itemize}
-	\item Planning rules (goal $\to$ plan)
-	\item Procedural rules (events $\to$ plan)
-	\item Plan repair rules (plan $\to$ plan)
-\end{itemize}
-
-\subsection{Processing mental states}
-
-\begin{itemize}
-	\item Generate plans for received events.
-	\item Generate plans for goals.
-	\item Process exceptions and handle failures.
-	\item Repair plans.
-	\item Select plans for execution.
-	\item Execute plans.
-\end{itemize}
-
-\subsection{Operational semantics}
-Operational semantics defines the computation steps a program
-configuration may take. Operational semantics allows studying
-of programming constructs in a rigorous manner. It facilitates proving
-general properties about a language. It lies close to the implementation
-of an interpreter. It facilitates model checking.
-
-\subsubsection{Basic syntax}
-The configuration $C$ evolves into configuration $C'$:
-
-\[C \to C'\]
-
-If premise $P$ holds transition $C \to C'$ can be derived:
-
-\begin{prooftree}
-	\AxiomC{$P$}
-	\UnaryInfC{$C \to C'$}
-\end{prooftree}
-
-\subsection{Configuration}
-configurations only mention things that change.
-Multi-agent configuration where $A$ is an agent and $\chi$ is the
-environment:
-
-\[\langle \{A_1, \cdots, A_n\}, \chi\rangle \]
-
-Individual agent configuration where $i$ is the agent number, $\sigma$ the
-believes, $\gamma$ the goals, $\Pi$ the plans $\theta$ the assignments and
-$\xi$ the events:
-
-\[A_i = \langle i, \sigma_i, \gamma_i, \Pi_i, \theta_i, \xi_i \rangle \]
-
-\subsection{Transitions}
-The confiugrations described \emph{what} can change where transition
-will describe \emph{how} they can change.
-
-$\vDash$ Describes a first order entailment relation. As I understand
-this means $A \vDash B$ there exsits a $B$ in $A$ which has an entailment
-to $A$. So $A$ is a set of things and a $B$ like thing exists in $A$.
-\subsubsection{Belief update}
-Execution of a belief update action $\alpha$ modifies the belief and goal
-bases. $Update$ is an update function that is specified by the believe
-update actions and $\gamma'=\gamma - \{\psi \in \gamma | \sigma' \vDash \psi\}$.
-Which basically says that $\gamma'$ is equal to $\gamma$ minus the update.
-
-\begin{prooftree}
-	\AxiomC{$Update(\sigma, \alpha\theta)=\sigma'$}
-	\UnaryInfC{$\langle i, \sigma, \gamma, \{(\alpha,id)\}, \theta, 
-		\xi\rangle \to 
-	\langle \langle i, \sigma', \gamma', \{\}, \theta, \xi\rangle$}
-
-\end{prooftree}
-
-So on the left side of the arrow, the plan is to do a believeUpdate $\alpha$
-on the right side this is done so its removed from the plans and the believes
-\emph{and} goals are modified. The goals are modified to because if a believe
-with the same name as a goal becomes true than the goal is removed. (this is
-a design decision btw, its just how 2apl works).
-
-\subsubsection{Adopta goal}
-Execution of a goal adopt action adopta(g) adds the goal to the beginning of
-the goal base. If the goal $g$ trough assignment $\theta$ does not already
-exists in goalbase.
-
-\begin{prooftree}
-	\AxiomC{$\sigma \nvDash g\theta$}
-	\UnaryInfC{$\langle 
-		i, \sigma, [\gamma_1, \cdots, y_n], \{(adopta(g),id)\}, \theta, \xi, 
-	\rangle \to \langle 
-		i, \sigma, [g\theta,\gamma_1, \cdots, y_n], \{\}, \theta, \xi, 
-	\rangle$}
-\end{prooftree}
-
-So the agent comes accross a adopta method in the plan base $\Pi$, if
-the associated believe of the goal is not true then you can start adding
-it.
-\subsubsection{Applying planning goal rules}
-The structure of a PG-rule in 2apl is as follows:
-
-\[\kappa <- \beta | \pi \]
-
-Where $\kappa$ is the name of the plan.
-$\beta$ is the condition that is applied before
-executing the plan (basically a big if around the plan). $\pi$ is
-the body of the plan, a set of instructions.
-
-\begin{prooftree}
-	\AxiomC{$\gamma \vDash \kappa \tau_1 \quad \& \quad
-	\sigma \vDash \beta \tau_1 \tau_2$}
-	\UnaryInfC{$\langle
-		i, \sigma, \gamma, \Pi, \theta, \xi
-	\rangle \to \langle
-	i, \sigma, \gamma, \Pi \cup \{ \pi \tau_1 \tau_2 \}, \theta, \xi
-	\rangle$}
-\end{prooftree}
-
-The transition can be executed if $\kappa$ is an active goal (this valuation
-is stored in $\tau_1$ \emph{and} $\beta$ is true according to the believe 
-base (this eveluation is stored in $\tau_2$). If all these things are true
-then $\pi$ is added to the plan base $\Pi$ \emph{but} it will only be executed
-if $\tau_1$ and $\tau_2$ return true.
-\subsection{Exam}
-some notes I gathered for possible exam questions:
-
-What happens if $\gamma = pos(2,3)$ and $\sigma = pos(4,5)$, resolve with
-state transitions.
-
-\section{Markov}
-Computerised decision makers must deal with an uncertain world.
-
-\subsection{Expectation}
-Excpectation is the average value you excpect to happen. For example if you
-throw a dice the average value you should get is 3.5. Because:
-
-\[1 * 1/6 + 2 * 1/6 + 3 * 1/6 + 4 * 1/6 + 5 * 1/6 + 6 * 1/6 = 3.5\]
-
-You multiple all possible values time the chance of them happening and then
-calculate the sum of them. Formally it is written like this:
-
-\[ E[X]=Def \sum_{all x} xP\{X=x\} \]
-
-So you can fill it in like:
-\[E[X]= 1 * P\{X=1\} + \cdots + 6 *P\{X=6\}\]
-\[E[X]= 1 * 1/6 + \cdots + 6 *1/6 = 3.5\]
-
-\subsubsection{Properties of expectation}
-
-\begin{itemize}
-	\item $E[c] = c$
-	\item $E[X+Y] = E[X] + E[Y]$
-	\item $E[cX] = cE[X]$
-	\item $E[XY] = E[X]E[Y]$ if and only if $X$ and $Y$ are independent.
-	\item $E[X] \le E[Y]$, if $X \le Y$ almost surely.
-\end{itemize}
-
-\paragraph{Law of the onconcious statistician}
-You have to prove this, its \emph{not} a property:
-
-\[E[g(X)]=\sum_{all x} g(x)P{X=x}\]
-
-where $g$ is any real-valued function. (It returns a value that is part
-of the real set of numbers).
-
-\subsection{Conditional expectation}
-The expectation of $X | Y = y$ is called the \emph{conditional expectation}
-of $X$ given $Y$, and is denoted by:
-
-\[E[X|Y=y]\]
-
-$E[X|Y=y]$ Is a number that depends on $y \in Y$. So you only get an
-expectation $X$ if you have an y that satisfies the condition.
-This yields a random variable $E[X|Y]$. Because of this it has all the
-properties of a random variable.
-
-\subsubsection{Probability mass functions}
-The standart notation for probability mass functions is:
-
-\[p_x(x): D(X) \to [0,1]: x \mapsto P\{X=x\} \]
-\[p_y(y): D(Y) \to [0,1]: y \mapsto P\{Y=y\} \]
-\[p_{xy}(x,y): D(X) \times D(Y) \to [0,1]: (x,y) \mapsto P\{X=x,Y=y\} \]
-\[p_{x|y}(x|y): D(X) \times D(Y) \to [0,1]: (x,y) \mapsto P\{X=x|Y=y\} \]
-
-For example:
-
-\[ P\{X=x|Y=y\}=\frac{P\{X=x,Y=y\}}{P\{Y=y\}}\]
-
-Is often written as:
-
-\[P_{x|y}(x|y)=\frac{P_{xy}(x,y)}{P_y(y)}\]
-
-\subsection{Total expectation}
-The tower property, which is awesome:
-
-Let $X$ and $Y$ be random variables then:
-
-\[ E[E[X|Y]] = E[X] \]
-
-The proof is in the slides.
-
-\subsection{Markov to matrix}
-A finite Markov process consisting of $n$ states can be represented by an 
-$n\times n$ \emph{probablity transition matrix}.
-
-Consider the following markov process:
-\begin{center}
-\begin{tikzpicture}[->, >=stealth', auto, semithick, node distance=3cm]
-\tikzstyle{every state}=[fill=white,draw=black,thick,text=black,scale=1]
-\node[state]    (A)                     {A};
-\node[state]    (B)[above right of=A]   {B};
-\node[state]    (C)[below right of=A]   {C};
-\path
-(A) edge[loop left]     node{0.5}         (A)
-    edge[bend left]     node{0.3}     (B)
-    edge[bend right, left]     node{0.2}      (C)
-(B) edge[bend left,right]    node{0.9}      (A)
-    edge[bend left,right]    node{0.1}      (C)
-(C) edge[bend right, right]    node{1}      (A);
-\end{tikzpicture}
-
-The probability matrix belonging to this process looks like the following:
-
-\[
-\begin{matrix}
-		& A		& B		& C \\
-	A	& 0.5	& 0.3	& 0.2 \\
-	B	& 0.9	& 0		& 0.1 \\
-	C	& 1		& 0		& 0 \\
-\end{matrix}
-\]
-
-\end{center}
-
-Adding the $A,B,C$ is optional, but I find it gives more overview. The first
-cell reads, the probobality of going from a to a.
-
-\subsubsection{Stochastic}
-A matrix where every row sums to1 is called a \emph{stochastic matrix}. Every
-transition matrix is a \emph{stochastic matrix}.
-
-\subsection{Sample path}
-A \emph{sample path} of a markov chain is a specific specific realistion of
-that chain.
-
-\[ X_0,X_1,X_2,\dots\]
-
-The $X_i$ are random variables.
-
-\subsection{Markov process}
-A \emph{Markov process} is a stochastic process $X_0, X_1, X_2, \dots$ such 
-that:
-
-\[ P\{ X_{n+1} X_n, X_{n-1}, \dots\}=P\{X_{n+1}|X_n\}\]
-for $n \ge 0$
-
-\subsubsection{Remarks}
-If states are numbers, then often state space is denoted by $I$, with 
-typical elements $i,j,k$.
-
-A markov process is sometimes called a markov chain. These concepts are
-equal.
-
-If $S$ is countably infinite, then the probability transition matrix
-can be represented as an ``open ended'' stochastic matrix.
-\[
-	P =
-\begin{pmatrix}
-	p_{1,1}	& p_{2,1}	& p_{3,1} & \cdots\\
-	p_{1,2}	& p_{2,2}	& p_{3,2}	& \cdots\\
-	p_{1,3}	& p_{2,3}	& p_{3,3}	& \cdots\\
-	\vdots	& \vdots	& \vdots		& \ddots \\
-\end{pmatrix}
-\]
-With markov proceci it is extremly important to beclear about the
-\emph{cardinality of the state space} (finite, discrete, continues) and
-the \emph{nature of time} (discrete, continuous)
-
-A \emph{dead state} or a \emph{dead end} is a state that does not have
-any outgoing arcs. An \emph{absorbing  state} is a state that has precisely
-one outgoing arc with probability 1 to itself.
-
-\subsection{Chapman-Kolmogorov equations}
-
-\[P^{(n)}_{ij} = \sum_{all\mbox{\ }k} P^{(n_1)}_{ik} P^{(n_2)}_{kj} \]
-
-It is convenient to list the $n$-step transition probabilities in a matrix.
-\[
-	P^{(n)} =
-\begin{pmatrix}
-	p^{(n)}_{1,1}	& \cdots	& p^{(n)}_{n,3} 	& \\
-	\vdots	& \ddots	& \vdots	& \\
-	p^{(n)}_{1,n}	& \cdots	& p^{(n)}_{n,n}	& \\
-\end{pmatrix}
-\]
-
-In particiular $P^{(1)} = P$.
-
-\subsubsection{Lemma}
-Suppose $k+l=n$ Where $k,l\ge1$. Then $P^{(k)} \cdot P^{(l)} = P^{(n)}$
-
-(I think this was proven)
-
-$P^{(n)} = P^n$, this was definitivley proven. Not sure why you would
-prove removing brackets, but whatever.
-
-\subsection{Classes}
-No this aren't default private structures (its a c++ reference).
-
-Classes make dealing with large markov chains easier by grouping nodes
-together. \emph{Open classes} allow the process to escape the class once
-it enters a \emph{closed class} it will remain there forever.
-A class is \emph{Transient} if a process can never reenter it, a
-state is recurrent if it can. A process is multichain if there are
-multiple recurrent classes, if there is just one reccurent class
-its uni-chain.
-
-A critical transition is a transition between classes.
-
-A closed class doesn't have to be recurrent if there are infinite cases.
-Otherwise a closed class is always recurrent.
-An openclass is always transient.
-
-\subsection{Markov reward}
-A \emph{Markov reward process} is a markov process with rewards on
-transitions. these rewards are real valued.
-
-Markov rewards processes are also known as \emph{markov reward chain} or
-\emph{valued Markov chains}.
-
-example
-
-\[
-	P =
-\begin{pmatrix}
-	0.2	& 0.8 & 0 	& \\
-	0.3	& 0	& 0.7	& \\
-	0	& 0 & 1	& \\
-\end{pmatrix}
-\]
-With reward matrix:
-\[
-	R =
-\begin{pmatrix}
-	1	& 2 & 0 	& \\
-	0	& 0	& 3	& \\
-	0	& 0 & 0	& \\
-\end{pmatrix}
-\]
-
-\subsubsection{Total reward}
-
-Let $\omega$ be a realisation of a Markov process:
-
-\[\omega = s_0, s_1, \dots\]
-
-where $s_0$ is a start state.
-The rewards that accumulate are $r_1, r_2, \dots$:
-
-\[ s_0 \xrightarrow[\mbox{reward } r_1]{} s_1 \xrightarrow[\mbox{reward } r_2]{} s_2\]
-
-\paragraph{Definition}
-
-$r_k | (s_0=1) : $ reward at kth step, given $s_0 = i$.
-$R|(s_0=i),$ the \emph{total reward} : $(r_1 + r_2 \dots)|s_0=i,$
-$V(i),$ the \emph{value} of state i : $E[R|s_0=i]$.
-
-\subsection{The bellman equation}
-
-\[V(i)=\sum_{all\mbox{\ }j} p_{ij}[r_{ij} + V(j)]\]
-
-This system of linear equations enables us to compute state values in two
-ways, by using algebra or by using iteration. Iteration you just wait until
-convergence within a specified range. For this course iteration is enough.
-
-\subsubsection{Utility pump}
-A utility pump is a MRP in which at least one recurrent state possesses a
-non-zero reward.
-\subsubsection{Discount factor}
-A \emph{dicsount factor} is a real number $0 \le \gamma < 1$. it represents
-the probability to survive from one state to the next.
-
-\paragraph{remarks} The discount factor is constant. The discount factor
-does not impose a fixed cuttoff bound. It is a probablistic concept.
-
-\subsubsection{Discounted reward}
-Suppose a discount factor $0 \le \gamma < 1$. Then
-
-\[V(i)=\sum_{all\mbox{\ }j} p_{ij}[r_{ij} + \gamma V(j)]\]
-
-for every state $i$.
-
-\paragraph{definition} A \emph{valuation} is a vector
-
-\[v=Def
-\begin{pmatrix}
-	v(1) \\
-	\vdots \\
-	v(n)
-\end{pmatrix}
-\in \mathbb{R}^n
-\]
-
-\paragraph{definition} The \emph{expected immediate reward} at state $i$ is
-the by $P_{ij}$ - probabilities weighted averaga of the $r_{ij}$ rewards:
-
-\[C(i)=Def\sum_{all \mbox{\ }j} p_{ij} r_{ij} \]
-
-The (vertical) vector of expected immediate rewards is denoted by $C$.
-
-\paragraph{definition} The \emph{Bellman operator}:
-
-\[B_p : \mathbb{R}^n \to \mathbb{R}^n : v \mapsto C + \gamma Pv \]
-
-\subsection{Fixed point}
-The value vector, $v^*$, of a discounted reward Markov process is a fixed
-point of the $B_p$ operator:
-
-\[ B_p(v)=v\]
-
-\[
-	v_0 \to B_p(v_0) \to B_p B_p (v_0) \to \dots \to B^n_P v_0
-	\to \dots \to v^*
-\]
-
-\subsection{Example Markov reward process}
-Consider the following markov reard process. Suppose a discount factor
-$\gamma = 0.5$.
-
-\[ \left\{ \begin{matrix}
-		s: & s(r=1, p=0.4) & ; & a(r=4, p=0.6), \\
-		a: & s(r=3, p=0.3) & ; & a(r=8, p=0.7), \\
-		b: & s(r=2, p=0.5) & ; & b(r=5, p=0.5). \\
-	\end{matrix}
-\right. \]
-
-\subsubsection{Transition diagram}
-\begin{tikzpicture}[->, >=stealth', auto, semithick, node distance=3cm]
-\tikzstyle{every state}=[fill=white,draw=black,thick,text=black,scale=1]
-\node[state]    (S)                     {s};
-\node[state]    (A)[right of=S]   {a};
-\node[state]    (B)[left of=S]   {b};
-\path
-(S) edge[loop above]     	node{p=0.4, r=1}(S)
-    edge[above]     		node{p=0.6, r=4}(A)
-(A) edge[bend left,below]   node{p=0.3, r=3}(S)
-	edge[loop above]     	node{p=0.7, r=8}(A)
-(B) edge[above]    			node{p=0.5, r=2}(S)
-	edge[loop above]     	node{p=0.5, r=5}(B);
-\end{tikzpicture}
-\subsubsection{Matrici}
-\paragraph{Probability transition matrix}
-\[
-	P =
-\begin{pmatrix}
-	0.4	& 0.6 & 0 	& \\
-	0.3	& 0.7 & 0	& \\
-	0.5	& 0 & 0.5	& \\
-\end{pmatrix}
-\]
-\paragraph{Immediate reward matrix}
-\[
-	R =
-\begin{pmatrix}
-	1	& 4 & 0 	& \\
-	3	& 8	& 3	& \\
-	2	& 0 & 5	& \\
-\end{pmatrix}
-\]
-\subsubsection{Value evalutaion}
-apply:
-\[V(i)=\sum_{all\mbox{\ }j} p_{ij}[r_{ij} + V(j)]\]
-
-\[ \left\{ \begin{matrix}
-		V(s) = 0.4(1+\gamma V(s)) + 0.6(4+\gamma V(a)), \\
-		V(a) = 0.3(3+\gamma V(s)) + 0.7(8+\gamma V(a)), \\
-		V(b) = 0.5(2+\gamma V(s)) + 0.5(5+\gamma V(b)).
-	\end{matrix}
-\right. \]
-
-With abusing notation:
-
-\[ \left\{ \begin{matrix}
-		s = 0.4(1+\gamma s) + 0.6(4+\gamma a), \\
-		a = 0.3(3+\gamma s) + 0.7(8+\gamma a), \\
-		b = 0.5(2+\gamma s) + 0.5(5+\gamma b).
-	\end{matrix}
-\right. \]
-
-Just fill in the values recursevily to perform value iteration.
-
-\subsection{Decision processes}
-I'm gonna cover this subject poorly because its dead simple coming from an
-IT background.
-
-\begin{tabular}{ll}
-	word & definition \\ \toprule
-	decision tree & a simple directed tree stucture \\
-	states & represented by verticies.\\
-	Immediate rewards & represented by numbers along the edges. \\
-	Decision alternatives & the outgoing edges, also called possible actions. \\
-	Decision point & A state where a decision can be made. \\
-	Finite decision tree & \brcell{A tree with a finite number of states, \\
-		{\footnotesize Makes you wonder if there could be infinite number 
-		of decisions in such a tree.}} \\
-\end{tabular}
-\\
-The set of possible actions at state $i \in I$ is denoted by $A_i$.
-The immediate reward for action $j$ at $i$ is denoted by $r_{ij}$.
-
-\subsubsection{Policy}
-A \emph{policy} specifies for every decision point $s$ a probability
-distribution over $A_s$. A policy is typically denoted by $\pi, \tau\dots$.
-If at $s$ the probability distribution puts all weight on one choice, the
-policy is said to be \emph{pure} at $s$. Otherwise the policy is said to
-be \emph{mixed} at $s$.
-
-A \emph{run} of a policy is the \emph{actual execution} of that policy.
-An \emph{realisation}, or \emph{sample path} of a policy is the trace of
-that run.
-
-The \emph{payoff} or \emph{total reward} of a run is the total sum of rewards
-acuumulated in its realisation.
-
-\subsubsection{Value function}
-A \emph{value function} is any function $V: S \to \mathbb{R}$. The
-\emph{expected payoff of a policy $\pi$}, given a certain starting state $s$:
-	\[E[\pi|s_0=s]\]
-	is called the value of $s$ under $\pi$, and is denoted by $V^\pi(s)$.
-Value functions are compared in the following way:
-
-\[ V_1 \le V_2 \Leftrightarrow_{Def} V_1(s) \le V_2(s), \mbox{for all states } s.\]
-
-Policy $\pi$ is \emph{as good as} policy $\tau$ if $V^\tau \le V^\pi$. A policy
-is \emph{optimal} if it is as good as any other policy. A value function is
-\emph{optimal} if it is induced by a policy, and is as large as any other
-such value function.
-
-\subsubsection{Policy = strategy}
-% this is the first what came to mind when I saw the policies.
-In game theory, a policy is called a \emph{strategy} and a pure strategy is
-sometimes called \emph{stationary} or \emph{deterministic}. A mixed strategy
-is sometimes called \emph{probabilistic, non-stationary,} or
-\emph{non-deterministic}.
-% Imaginative with their non-*
-
-An example of a pure policy is
-\[ \tau = \{(r,b),(a,d),(b,e),(d,h),(f,i)\} \]
-
-All the decisions are made, even if you already cut them off.
-So you say (Start, End) for each node with choice.
-
-Mixed policy:
-\[ \tau = \{(r,\{(a, 0.2), (b,0.8)\}),(a,d),(b,\{(e, 0.7), (f,0.3)\}),
-(d,h),(f,i)\} \]
-% yes some really poor langauge design over here.
-
-\subsubsection{Expected payoff}
-$\pi$'s expected payoff, $E[\pi|s_0=r]$, can be computed recursivly with the
-law of total expectation.
-
-\subsubsection{Backward induction}
-A decision between alternatives that leads to certain outcomes, is a certainty
-itself.
-
-Backward induction is applicable in \emph{directed acyclic graphs}(DAGs).
-
-\subsubsection{Decision problem}
-A \emph{decision problem} amounts to determining the set $\Pi^*$, where:
-
-\[ \Pi = argmax_{\pi \in \Pi}(payoff(\pi))\]
-
-Where $\Pi$ is the set of all policies.
-
-For finite decision trees: Supppose at least one optimal policy exists.
-If this policy is mixed, it may be replaced by an optimal pure policy,
-such that, at every state, the payoff of the pure policy equals the expected
-payoff of the mixed policy.
-
-\subsubsection{Problems}
-\paragraph{infinite width} no best policy, because you're valuating forever.
-\paragraph{infinite depth} How to compare infinite total rewards (can't).
-\paragraph{discount factor} The right decision is very dependent on the
-discount factor. Counter intiuitive choices might be better for example.
-
-\paragraph{Utility pumps}
-If there is no discount factor policy is irrelivant (the reward will be
-infinite no matter what you do). A tool is needed to kill infinity.
-For example a finiti horizon, discount factor, or long-run average
-payoffs.
-
-In this course we will just use discount factors.
-
-\paragraph{Investment} 
-At cycle $a$, it is possible to `pump' 1 utitlity per iteration for as
-long as a run lasts. At cycle $b$ it is possbile to `pump' $2$ utilities
-per iteration.
-In the long term it is better to be at $b$, but starting from $a$ this
-requires an investment of $K$ rewardless transitions to get there.
-
-\begin{center}
-\begin{tikzpicture}[->, >=stealth', auto, semithick, node distance=3cm]
-\tikzstyle{every state}=[fill=white,draw=black,thick,text=black,scale=1]
-\node[state]    (S)                     {s};
-\node[state]    (A)[right of=S]   {$b_1$};
-\node[state]    (B)[right of=A]   {$b_{k-1}$};
-\node[state]    (C)[right of=B]   {$b_k$};
-\path
-(S) edge[loop above]     	node{1}(S)
-    edge[above]     		node{0}(A)
-(A)    edge[above]    node{0}(B)
-(B)    edge[above]     		node{0}(C)
-(C)    edge[loop above]		node{1}(C);
-\end{tikzpicture}
-\end{center}
-
-If $K \ne 0$ it is best to stay left if $\gamma < 2^{-1/k}$. (for the
-proof see the slides, its to much math to copy in here).
-
-\subsection{Policy improvement}
-
-Recall the bellman equations for a Markov reward process:
-
-\[ v = C+\gamma P v\]
-
-To emphasise that $v$ depends on a probability transition matrix $P$:
-
-\[ v^P = C^P + \gamma P v^P\]
-
-In a \emph{decision problem}, $P$ is not chosen by nature, but by a 
-\emph{decision maker}. In that case, $P$ is called a \emph{ixed Markov
-policy}. and is denoted by $\pi$:
-
-\[ v^\pi = C^\pi + \gamma P^\pi v^\pi\]
-
-Each policy $\pi$ gives a value function $v^\pi$. THe objective is to find a
-policy $\pi$ that maximises the value function $v^\pi$.
-
-\subsubsection{First sketch}
-A first sketch of a policy iteration algorithm:
-\begin{enumerate}
-	\item Start with an arbitrary policy $\pi$
-	\item Compute $V_\pi$. (know as \emph{policy evaluation}).
-	\item Improve $\pi$ locally to $\tau$.
-	\item If $\tau \ne \pi$, then $\pi := \tau$, and go to 2.
-	\item $\pi$ is optimal.
-\end{enumerate}
-
-\paragraph{Problem} Therer are to many ways to improve a policy.
-For example if a state $s$ enables two actions, then any
-mixed policy $(p, 1-p)$ may be tried locally. This gives
-uncountably many (i.e., $|\mathbb{R}|$) alternatives!
-
-Fortunally for discounted decision problems, locl search can be
-restricted to pure policies only.
-
-Let $\pi$ be apossibly mixed policy. For each state choose a decision
-that is locally maximal:
-
-% I'm not sure where the J is supposed to go.
-\[M(i) \in argmax\{r_{ij} + \gamma V_\pi(j)\} \]
-
-Define the pure policy:
-
-\[ \tau_{ij} =_{Def} \delta_{M(i),j} \]
-
-Then $\tau$ is a global improvement of $\pi$.
-
-\subsubsection{Partial correctness}
-
-Correctness  partial correctness + termination.
-
-Partial correctness means: correct if the algrotihm terminates.
-\emph{Policy iteration is partially correct:} any strict
-local improvement induces a global improvment.
-\emph{Policy iteration terminates:} each time a policy is improved,
-the new policy is strictly better than the previous.
-
-\subsection{Bellman optimallity}
-
-If $\pi$ is optimal, it follows from the policy improvement thearom that
-
-\[ V^\pi(i) = \underset{all\mbox{\ }j}{max}[r_{ij} + \gamma V^\pi(j)] \]
-
-Removing explicit reference to the optimal policy automatically (more elegant).
-
-\subsubsection{operator}
-%seriously what is meant with operator?
-
-% I'm assuming he means with exact the algebraic method and
-% the aproximation is the itteration method.
-For both methods (the exact as well as the approximation method),
-a compacter notation is needed.
-
-\[B: \mathbb{R}^n \to \mathbb{R}^n : v \mapsto
-	\underset{all\mbox{\ }j}{max}[C_j + \gamma V(j)] \]
-
-Where $C_j$ is the vector of rewards that emanate from state $i$.
-Antoher way to write $B:$
-\[\underset{all\mbox{\ }j}{max}[r_{ij} + \gamma V(j)] \]
-
-$B$ does not depend on a policy or transition matrix. ($B_p$ does.)
-
-\subsubsection{Fixed point}
-The value-vector, $v^*$, of a decision problem is a fixed point of the
-$B$-operator:
-\[ B(v) = v\]
-
-\subsection{Exam}
- For a given decision problem with a certain
-discount factor.A pure policy sufices.
-Do policy iteration, (a single iteration), do value iteration
-(single iteration).
-
-% That's all she wrote.
-
 \end{document}

### 2015-11-11 methods/agents/summery.tex
added a bunch of extra info about different modal logics

@@
 \paragraph{Semantics} Kripke %KRIPKE, *waves arm*.
-model/frame $M=\langle S,\pi,R \rangle$
+model $M=\langle S,\pi,R \rangle$, kripke frame $M=\langle S,\pi,R \rangle$
+(truth assignment function is abstracted away)
 
@@
 	$R_{\alpha+\beta}$ & = & $R_\alpha \cup R_\beta$ &
-	execute $R_\alpha$ or $R_\beta$\\
+	execute $R_\alpha$ or $R_\beta$ or both\\
 	$R_{\alpha*}$ & = & $R_\alpha*$ &
@@
 \end{tabular}
+\subsection{Classes of models}
+\begin{tabular}{lll}
+	Class & properties & axioms \\ \toprule
+	$K$	& No constraints & All propositional +
+	$\{(\Box\phi \wedge \Box(\phi \to \psi)) \to \Box \phi\}$ \\
+	$T$	& reflexive	& K + $\{\Box \phi \to \phi\}$ \\
+	$S4$	& reflexive transative	& T + $\{\Box \Box \phi \to \Box \phi\}$ \\
+	$S5$	& equivalence & S4 + $\{\neg \Box \phi \to \Box\neg\Box \phi\}$ \\
+	$KD$	& serial & K + $\{\neg \Box \bot\}$ \\
+	$KD45$	& serial, transitive and euclidean & S5 + $\{\neg \Box \bot\}$ \\
+\end{tabular}
+Alternative formulation of the 5-axiom:
+
+\begin{prooftree}
+	\AxiomC{$\neg \Box \phi \to \Box\neg\Box \phi$}
+	\RightLabel{$\Box \phi = \neg\Diamond\neg\phi$}
+	\UnaryInfC{$\Diamond \neg \phi \to \Box\Diamond \neg \phi$}
+	\RightLabel{Substitute $\neg \phi$ for $\psi$}
+	\UnaryInfC{$\Diamond \psi \to \Box\Diamond \psi$}
+	\RightLabel{Substitute $\psi$ for $\phi$}
+	\UnaryInfC{$\Diamond \phi \to \Box\Diamond \phi$}
+\end{prooftree}
+
+\subsection{Deontic logic}
+The system $KD$ is also known as $SDL$ (standart deontic logic)
+% sdl = standart library as everyone knows.
+Deontic logic is the logic of \emph{obligation, prohibition and permission}
+or rather: the logic of \emph{ideal vs actual sistuations}
+
+\subsubsection{Rules}
+\begin{itemize}
+	\item Propositional calculus (including modus ponens)
+	\item $(O\phi \wedge O(\phi \to \psi)) \to O \phi$ (K axiom)
+	\item $\frac{\phi}{O\phi}$ (necisitation rule)
+	\item $\neg O \top $ (D-axiom: obligation is consistent)
+	\item $F\phi \leftrightarrow O\neg \phi$ (forbidden is obliged to not)
+	\item $P\phi \leftrightarrow \neg F \phi$ (permitted is not forbidden)
+\end{itemize}
+
+\subsection{Temporal Logic}
+Basic linear time logic (LTL), Time as accisibility relation, its reflexive
+and transitive.
+
+Viewed in this way: LTL= S4, $\Box$ stands for `always in the future'.
+By convention the present is included in the future.
+
+\subsubsection{Rules}
+\begin{itemize}
+	\item Propositional calculus (including modus ponens)
+	\item $(\Box\phi \wedge \Box(\phi \to \psi)) \to \Box \phi$ (K axiom)
+	\item $\frac{\phi}{\Box\phi}$ (necisitation rule)
+	\item $\Box \phi \to \phi$ (always implies now)
+	\item $\Box \phi \to \Box\Box\phi$ (always implies always)
+\end{itemize}
+
+\subsection{Epistemic \& doxastic logic}
+For \emph{knowledge} we take the relation $R_i$: reflexive, (transative) and
+eculidean, i.e. An equivalance relation.
+
+For \emph{believe} we take the relation $R_i$: serial, transitive and 
+euclidean.
+
+\subsubsection{Epistemic rules}
+\begin{itemize}
+	\item Propositional calculus (including modus ponens)
+	\item $(K\phi \wedge K(\phi \to \psi)) \to K \phi$ (K axiom)
+	\item $\frac{\phi}{K\phi}$ (necisitation rule)
+	\item $K \phi \to \phi$ (knowledge is true)
+	\item $K \phi \to KK\phi$ (positive introspection)
+	\item $\neg K \phi \to K\neg K\phi$ (negative introspection)
+\end{itemize}
+
+\subsubsection{Doxastic rules}
+\begin{itemize}
+	\item Propositional calculus (including modus ponens)
+	\item $(B\phi \wedge B(\phi \to \psi)) \to B \phi$ (B axiom)
+	\item $\frac{\phi}{B\phi}$ (necisitation rule)
+	\item $\neg B \bot$ (belief is consistent)
+	\item $B \phi \to BB\phi$ (positive introspection)
+	\item $\neg B \phi \to B\neg B\phi$ (negative introspection)
+\end{itemize}
 \section{APL}

### 2015-11-11 inteligentagents/notes2.tex
added notes about the lecture

#Bratman

Commitment, if you have a real intention, you want to go for it.
intentions should be consistent, otherwise you do something which could be
impossible
But this is debatable...

## Intentions
Screen of admissibility, filter maybe later intentions that could be
conflicitng

# Cohen& levesque
tried to formalise Bratman.

## Intention = choice + commitment
try to stretch it to formal logic.

##
agent beleives P is possible : a sort of self conficience of the agent

## side effects
agnets need not intend all the expected side effects of their intentions.
Going to the dentist implies pain.
then going to the dentist should be that you intend to have pain.

$\vDash \phi \to \psi \to \vDash \Box \phi \to \Box \psi$

#Rao & Georgeff BDI theory
Not only proposed a logic, but also the first implementation of an agent and
also the first application.

#Rodney Brooks anti-cogn.robotics
subsumption architecture: no cognition, just a reactive system.
Should be emergent behaviour rather than modeling.

# Aaron Sloman: How to build emotional agents
Looks at properties of humans, such as emotions, and tries to put it into an
AI.

20 years ago, there are patiens that don't have emotions, expect: in problem
solving such as puzzels. They go depth first, so they go on, and on and fail
misserably.

Archicterus of him are not the most enlighted, "messy".

# Towards social agents/robots
also research goes to sociology.

Oppertunisitic behavior, sellar buying a cup, but cup is broken,
the seller wants to sell it, but the buyer wants to use it. How
to moddel this in Agents

## Normative systems
the way of societies. you shouldn't violate the norm, but sometimes it can
be usefull to.

### Asimov
Laws of him. 4th one is stupid

# Theory of mind
The psycology one.
The false believe test.

# Logial foundations of actions

The terminology is not stabels.
desires and goals may be differetn or mean the same.

Agenda -> latin things that should happen.

# Modal logic
Logic of $\Box$ and $\Diamond$

## K axiom
true for any modal logic

## R
means its accesible. No restriction on R in arbitrary Kripke

### Constraints
you can add constraints on R, but with restrictions things certain
things become true.

serial you can go on.

## Frames
Models where the true assignment functions are abstracted from.

#naive BDI logic
seperate blocks, it misses connections.
No dymancis, changes of BDI in time.

# Cohen & levesque : langauge
$\alpha$ are complex actions and a's are primitive actions
We can't say multiple agents do the same action at the same time.

a is interpeted as a sequance of primitive events even though a is
atomic.

atomic action is a special class, which depends on the domain.

; = sequence
? = test if its true (groovy truth)
+ = is a non detrmenistic choice, collect them in an autcome set.
* = is done a finite but arbitrary number of times, including 0

if $\phi$ then $\alpha_1$ else
$\alpha_2$ = ($\phi? ; \alpha_1) + (\neg \phi?;\alpha)$

while $\phi$ do $\alpha$ = $(\phi?;\alpha)^*;\neg\phi?$

Pz = z has property P, a predicate attom

# Semantics
\omega = variables assginemnts
P = is the set of omega \in \omega
E = is a set of primitive events
Agt = Agt \in [E \to p] specifies the agent of an event [E \to P] is a
function that takes an E that gives a P.
T \subset [Z \to E]; a set of possible worlds, give a time number and get
an event

### 2015-11-12 evolutionaryComputation/notes2.tex
notes from the class

#dominio convergence...

linear function = no interaction between the bits.

functions with nonlinear interaction are problematic, because
you can't do bitlfips to see if the situation improves.

If a GA wants to be better than a bit flipper it has to be able to
interact with higher order blocks.

# Schemata
is a  Similirarity subset

Look at schemata to see how a GA works

$m(h, t)$, m counts the number of instances that start with h, t is the
generations.

## schema growth by selection
### Porpotionselect
DOn't use proportion select!

### tournement select
don't have nice forumula

### Schema distruption by mutation
Mutation is destructive on schemata
h=1xx0xx1xxx
with probability P I'm gonna flip one of h's bits

the survival probaility of a schemata is the probaility of not flipping
those bits

A schema with a higher order can be distrupted more easily

#### 1 point crosover
worst case definition.

#### uniform crosover
Again worst case, assume if its touched its destroyed.
so either touch them all or don't touch them all.

order should be small to survive in the schema

## Schema theorom
m(h,t+1), the next generation

if the net growth factor is bigger then 1, it'll grow
selection = important for growth factor
destruction is also important, ie mutation

# Holland
proposed genetical algoritms.
Schematas are S curves, after a while the strongest 'genes'/bits will converge
and compete with each other, so you'll have a domino effect.

## Bandit problem
two normal distribution in which you can sample but you want to reduce
the sample size of the wrong distribution.

GA will figure out the best normal, but it doesn't maximise the sum
so there is some debate going on about holland's method.

# Building blocks
any search algorithm has a bias.
small population size with high selection presure won't work.

in a GA you don't want to rely on mutation
# loose & thight
thight things that are close are importent ie (bit form an integer)
loose the order of the bits doesn't matter (I imagine some sort of bit mask)

# deceptive vs non deciptive
has to do with the schema's. If you look of the schema averages of all the
schama's of the blocks. and ignore the optimum, you'll drift to the wrong
side, with non-deceptive trap functions you won't.

with non-deceptive you'll drift towards the optimum instead of the local
optimum.

# practical

## Why do we do this?

Understand about what things we're talking about. get an understanding of
the basic concepts.

## assignments

### uniformly scaled counting ones function
Count the number of ones. Its easy to model
### linearly scaled counting ones function
first has fitness of 1, the second of 2 times 1 etc...

### trap functions
it uses blocks to get trapped in a local optimum.

#### Deceptive
it will probably get stuck in a local optimum

#### non deciptive
should drift towards the optimum

### randomly linked
randomly linked that the bits are placed at random position on the binary
string.

random, at the beginning. Not for the contest, but just for the creation of
new ones?
two possible methods:
	* For the creation of all subject create 1 global random mapping
	* For each instance create his own random mapping, this has the
	problem however of deciding which random mapping to use for the child:
		Father, mother or another new random.
This wasn't discussed in class SO I asume a global mapping has to be used.
But here I assume so maybe an email to the teacher could help, or just do
both methods and discuss them in the report.

### selection
You can play with it, crank it up untill you cancel all the deciptiveness is
fine.

### 2015-11-16 inteligentagents/notes3.tex
notes of intelligent agents lecture 3

\documentclass{article}
\begin{document}
\author{Jappie Klooster}
\title{Cohen \& Levesque}
\maketitle
\section{Langauge}
DONE = just done
; = sequence
? = test if its true (groovy truth)
+ = is a non detrmenistic choice, collect them in an autcome set.
* = is done a finite but arbitrary number of times, including 0

if $\phi$ then $\alpha_1$ else
$\alpha_2$ = ($\phi? ; \alpha_1) + (\neg \phi?;\alpha)$

while $\phi$ do $\alpha$ = $(\phi?;\alpha)^*;\neg\phi?$

Pz = z has property P, a predicate attom

\subsection{Semantics}
$\omega$ = variables assginemnts
$P$ = is the set of omega $\in \omega$
$E$ = is a set of primitive events
$Agt = Agt \in [E \to p]$ specifies the agent of an event $[E \to P]$ is a
function that takes an E that gives a P.
$T \subset [Z \to E];$ a set of possible worlds, give a time number and get

i = agent
n = time

\subsubsection{Predicate}
c constants
f functions
Q predicate smbols

$\phi$ interperts predicate symbols

\[\phi(c) = d \in \Theta\]
\[\phi(f) = f D_1x\dots D_l \to D_h\]
\[\phi(Q^{n,x,ne}) = R \subseteq D_1x\dots D_l \]

\subsubsection{satisfaction}

\[M,\sigma,v,n[\alpha] m\]
In the model `$M,\sigma,v,n$', $\alpha$ lies betweeen n and m.

B = `normal` modal accesibility relation indexed by the time $n$
and the agent $v(i)$. they write it like:

\[<\sigma, n>B[v(i)] \sigma\]

Better would be:
\[\sigma B_{v(i)}^n \sigma\]

\subsection{Constraints}
\subsubsection{Believes/Goals}
KD45.

\[G\subseteq B\]
means $\vdash {Bel}_i \phi \to {Goal}_i\phi$, but not the otherway around

\subsection{Achievement goal}
Doesn't believe $\phi$ is true yet, but believes it could be true later.
There exists some action to make $\phi$ true.

\subsubsection{Asumption}
Goals will end some time in the future

\subsection{Persistent goals}
Drop the goal if its true, or its impossible.

Competence means that if he believes it then its true.

\section{Rao Georgeff}
\subsection{Time logics}
Brancing vs linear, points vs intervals,
discrete (real numbers) vs continues (complex number)
We'll only look into the future.
\end{document}

### 2015-11-16 evolutionaryComputation/notes3.tex
notes from evolutionary computing lecture 3

\documentclass{article}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\usepackage{qtree}
\usepackage{tikz}
\usepackage{listings}
\usepackage{hyperref}
\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

\newcommand{\brcell}[2][l]{%
	\begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
\begin{document}
\author{Jappie Klooster}
\title{Mutable gene possibilities}
\maketitle
#Permutation problems

Clasical example is traveling salesman.
non-banary strings, will generate many illegal soltions.
possible solutions:
	Alternative search space representation.

# Mutation operators
best one depends on the problem.

\begin{itemize}
	\item absolute position
	\item adjencency
	\item relative ordering
\end{itemize}

## Swap and insert
ABCDEFGH
	insert: ABDEFCGH
	swap:	ABGDEFCH

## Two opt
Randomly select two points along the sequance and invort one of the
subsequences.

\begin{prooftree}
	\AxiomC{AB|CDE|FG}
	\UnariInfC{AB|EDC|FG}
\end{prooftree}

Adjencency is importent

## Recombination
Capital and lower laters are the same elements.

### Order crossover
Select 2 random crosspoints
Copy the subseqent croproints from p1
after the second crosspoints try to fill in the missing elements
form p2 in relative order

### Partially mapped crossofer
Select 2 random crosspoints
Copy p2 to child
Copy elements between random points to p2, for each element
copied between random points replace the crossed out point
with the newly coppied element

### Position crossover

### Maximal preservative crossover
Use a lot of adjencency information, fallback to mapped crosver if it doesn't
work??

### Cycle crossover
Check which elements are connected to each other like a binary tree in a
int array

ABCDEFGHI
fcdaebhig
111121333

1: A->f->F->b->B->c->C->d->A->a->A
2: E->e
3: G->h->H->i->I->g

cross the full cycles, so choose between any of the 1,2 or 3

## Edge recombination
Parent tours...

City	Edges
A		B F C E
B		A C D F
C 	 	B D A
D 	 	C E B
E 	 	D F A
F 	 	A E B

Random choice => B
next candiadates: ACDF

count remaining options, pick the lowest C
Next choose D because only 1
E -> A -> F

child: B -> C -> D -> E -> A -> F

But the child has the problem that it didn't inherit D->F

# Fitness correlation coefficients
(S_1, F_1) \to (S_1, F_1)
(S_2, F_2) \to (S_2, F_2)
\vdots
(S_{n-2}, F_{n-2}) \to (S_{n-2}, F_{n-2})
(S_{n-1}, F_{n-1}) \to (S_{n-1}, F_{n-1})

##crossover
(S_1, F_1) \cross (S_1, F_1)
(S_2, F_2) \cross (S_2, F_2)
\vdots
(S_{n-2}, F_{n-2}) \cross (S_{n-2}, F_{n-2})
(S_{n-1}, F_{n-1}) \cross (S_{n-1}, F_{n-1})

##Corolotion coefition $\rho$
0 = a cloud
-1 = anti evolution, natural deselection
1 = evolution, natural selection

$-1 \steq \rho \bteq 1$

### Traveling salesman TSP
Edge crossover was made for TSP
it has .9, which is very high.

### Alternative
Just run the thing and see which one is faster.

# Evolutionary strategies (ES)
ES use an advanced mutation operator which controls its own mutability
self adaptation.

This works by using gausion distrobutions, it should adapt to the topology
of the landscape. If its to big then its like random search. If its to
small it can get stuck at a local optimum.

This gausian uses more information about the problem, you know information
so you can reduce the search space.

## ES representation

\end{document}

### 2015-11-16 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala
added slf4j logging lib

@@ -1,15 +1,22 @@
 package nl.jappieklooster.hw.ec
 
+import org.slf4j.LoggerFactory
+
 import scala.util.Random
 
 object Main{
-
+	val log = LoggerFactory.getLogger(Main.getClass)
 	val random = Random
 	def main(args:Array[String]){
-		println("scala!")
+		log.info("starting ea")
 
 		val length = 20
-		val population = Population.createOneZeros(MemberFactories.globalRandomizedGenes(random.shuffle(0.to(length))), length,40)
+		val population = Population.createOneZeros(
+			MemberFactories.globalRandomizedGenes(
+				random.shuffle(0.to(length))),
+				length,
+				40
+		)
 		val evolution = new Evolution(
 			Valuation.uniformlyScaledCountOnes,
 			MateSelection.tournamentWinIsParent,
@@ -23,7 +30,7 @@ object Main{
 		for((pop, index) <- results.zipWithIndex){
 			println(s"generation $index: $pop")
 		}
-		println(s"time in ms: $runtime")
+		log.info(s"time in ms: $runtime")
 	}
 	def randomGene = s"${random.nextInt(2)}"
 }

### 2015-11-17 evolutionaryComputation/practicals/report.tex
beginning the report

--- 2015-11-10 inteligentagents/kripke.tex
@@
 \documentclass{article}
-% I just added all the things, if you have problems build, try deleting the
-% naggin packages, fat chance I don't even use it.
-\usepackage{amsmath}
 \usepackage{txfonts}
@@
 \usepackage{tikz}
-\usepackage{amssymb}
-\usetikzlibrary{automata,arrows,positioning,calc}
 \usepackage{listings}
@@
 	\begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
-
 \begin{document}
-\lstset{language=Java}
 \author{Jappie Klooster}
-\title{Logic, Kripke and Chaos}
+\title{A praktical look at genetic algorithms}
 \maketitle
 
-\section{Introduction}
-This will be an introduction to modal logic coming from someone who followed
-` hbo informatica' as pre study.
-
-In here I will write down the things I think are important for understanding
-modal logic and how to proof things in that system.
-
-% Before starting I just want to lay out the idea here, everytime I hear
-% kripke, I imagine someone waving his arm and shouting KRIPKE in anger.
-% Not that anyone will ever read these comments but its a funny idea.
-
-This course was thought by John-Jules Meyer. The `theoretic' part focused
-on the ideas behind the logics, while the `practical' focussed on writing
-propper proofs. I guess I'll treat both in here.
-
-There are 2 main sources of information I'll use, the slides by John Jules
-and the modal logic paper for artificial intelligence by rosja mastop to help
-me with the practical part. 
-% In truth I didn't understand much of the practicals, just giving a bunch
-% of examples is not a good way of teaching how to write proofs.
-% This paper seems to explain it in a more structural way.
-
-\section{Basic modal logic}
-Express intensional (context/situation-sensitve) notations such as: 
-Knowledge, Belief, Obligation, Action and Time.
-
-\subsection{Modal langauge}
-
-Imports all the things from propositional logic and is extended
-with the modal operators ($\Box\phi,\Diamond\phi$) which can be read as:
-
-\begin{tabular}{ll}
-	Symbol & Definition \\ \toprule
-	$\Box\phi$ & It is \emph{necesarry} that $\phi$ \\
-	$\Diamond\phi$ & it is \emph{possible} that $\phi$
-
-\end{tabular}
-
-\noindent
-Also $\neg\Box\neg\phi = \Diamond\phi$ and
-$\neg\Diamond\neg\phi = \Box\phi$. So you could theoratically
-drop one to create the other.
-% where $\phi$ is an object I guess?
-
-\paragraph{Semantics} Kripke %KRIPKE, *waves arm*.
-model/frame $M=\langle S,\pi,R \rangle$
-
-\begin{tabular}{ll}
-	Symbol & Definition \\ \toprule
-	% worlds implies mutability, states, well a state is by definition 
-	% imutable. I'm pretty sure we want immutability in these kind 
-	% off logics
-	$S$ & Is a set of \emph{states (or worlds)} \\
-	$\pi$ & Is a truth assignment function $\pi : S \times AT \to {tt, ff}$ \\
-	$R$	& $R \subseteq S \times S$ is an \emph{accesibility relation}\\
-\end{tabular}
-
-\subsection{Interpertation}
-
-\begin{center}
-\begin{tikzpicture}[auto, >=stealth', auto, semithick, node distance=3cm]
-\tikzstyle{every state}=[fill=white,draw=black,thick,text=black,scale=1]
-\node[state]    (S)            {$\neg p, q$};
-\node[state]    (A)[below of=S]{$p, \neg q$};
-\node[state]    (B)[right of=S]{$\neg p, \neg q$};
-\node[state]    (C)[right of=B]{$p, q$};
-\node[state]    (E)[below of=B]{$\neg p, q$};
-\path
-(S)   edge[above]     		(A)
-(S)   edge[above]     		(E)
-(S)   edge[above]     		(B)
-(B)    edge[above]     		(E)
-(E)    edge[above]     		(C);
-\end{tikzpicture}
-\end{center}
-
-Like classical propositional logic, but now realtive to a model and a state:
-
-\[ M,s \vDash \phi \]
-
-thus:
-
-\begin{tabular}{lll}
-	Modal & $\Leftrightarrow$ & Interpertation \\ \toprule
-$M,s \vDash p$ & $\Leftrightarrow$ & $\pi(s,p) = tt$ \\
-$M,s \vDash \phi \wedge \varphi $ & $\Leftrightarrow$ & $ M,s \vDash \phi
-\mbox{ and }M,s\vDash \varphi$ \\
-$M,s \vDash \Box \phi $ & $\Leftrightarrow$ & $ M,t \vDash \phi $ for 
-\emph{every} $t$ such that $R(s,t)$ \\
-$M,s \vDash \Diamond \phi $ & $\Leftrightarrow$ & $ M,t \vDash \phi $ for 
-\emph{some} $t$ such that $R(s,t)$ \\\bottomrule
-\end{tabular}
-
-\subsection{Validity}
-\begin{tabular}{llll}
-	Definition & Denotation & $\Leftrightarrow$ & Interpertation \\ \toprule
-
-$\phi$ is \emph{valid} in a model $M= \langle S, \pi, R \rangle$ &
-$M\vDash\phi$ & $\Leftrightarrow$ & $M,s \vDash \phi$ for all $s \in S$ \\
-
-$\phi$ is \emph{valid}  &
-$\vDash\phi$ & $\Leftrightarrow$ & $M\vDash\phi$ for all Kripke models $M$ \\
-
-$\phi$ is \emph{valid} wrt class $\zeta$ &
-$\zeta\vDash\phi$ & $\Leftrightarrow$ & $M\vDash\phi$ for all Kripke models
-$M \in \zeta$ \\\bottomrule
-\end{tabular}
-
-The last one doesn't occur often.
-
-\section{System K}
-We try to  axiomatize validities.
-
-\paragraph{Axioms} We import all (or enough) propositional tautologies.
-The K axiom ($(\Box\phi \wedge \Box (\phi \to \varphi))\to\Box\varphi$) is added.
-
-\paragraph{Rules} Modes ponens
-\begin{prooftree}
-	\AxiomC{$\phi$}
-	\AxiomC{$\phi \to \varphi$}
-	\BinaryInfC{$\varphi$}
-\end{prooftree}
-
-Necisitation rule:
-\begin{prooftree}
-	\AxiomC{$\phi$}
-	\UnaryInfC{$\Box\phi$}
-\end{prooftree}
-
-\subparagraph{Caution} the necisitation is different from the invalid
-assertion: $\phi \to \Box \phi$.
-
-\subsection{Derivability in K}
-A \emph{derivation} of a formula $\phi$ is a finite sequence of formulas
-$\phi_1, \dots \phi_n = \phi$. where each $\phi_i$ for $1 \ge i \ge n$
-is either an instance of the acioms (or rather axiom schemes), or the
-conclusion of one of the rules of which the premises have
-been derived already, i.e{.} appear as $\phi_j$ in the sequance with
-$j < i$.
-
-% sad thing is I understand this crap now.
-When we can derive an epistemic formula $\phi$ by using the axioms
-and rules of $K_{(m)}$ we write $K_{(m)} \vdash \phi$
-
-System $K_{(m)}$ is \emph{sound and complete}. This means
-that exactly ll valid modal assertions can be obtained by derivations in
-system $K_{(m)}$.
-
-\[\vDash \phi \Leftrightarrow K_{(m)} \vdash \phi\]
-
-% I leave out the (non) theoroms, it seems overkill.
-
-\section{Dynamic logic}
-An example of an (indexed) version of system $K$ is dynamic logic,
-wehre the $\Box$ modality is associated with the execution results of
-a program / action.
-
-$\Box_\alpha$, normally written $[\alpha]$ 
-with reading $[\alpha]\phi$: after exuction of $\alpha$ it holds that $\phi$
-
-\[<\alpha>\phi = \neg [\alpha] \neg \phi \]
-
-\subsection{Semantics}
-Accessibility relation $R_\alpha$ for every action $\alpha$
-
-\begin{tabular}{llll}
-	% I added the natural collumn, because otherwise I won't remember.
-	% Translating something in a different notation is stupid.
-	notation & = & definition & natural \\ \toprule
-	$R_{\alpha;\beta}$ & = & $R_\alpha \circ R_\beta$ &
-	execute from $R_\alpha$ until including $R_\beta$\\
-	$R_{\alpha+\beta}$ & = & $R_\alpha \cup R_\beta$ &
-	execute $R_\alpha$ or $R_\beta$\\
-	$R_{\alpha*}$ & = & $R_\alpha*$ &
-	Execute an arbitrary times \\ \bottomrule
-\end{tabular}
-
-\subsection{Interpertation}
-\begin{tabular}{lll}
-	Modal & $\Leftrightarrow$ & Interpertation \\ \toprule
-	$M,s \vDash [\alpha]\phi$ & $\Leftrightarrow$ &
-	for all $s'$ with $R_\alpha(s,s'): M,s' \vDash \phi$ \\
-	$M,s \vDash <\alpha>\phi$ & $\Leftrightarrow$ &
-	for some $s'$ with $R_\alpha(s,s'): M,s' \vDash \phi$ \\\bottomrule
-\end{tabular}
-
-\section{Special properties of accessibility relations}
-\begin{tabular}{lll}
-	R is \dots& if\dots & natural language \\ \toprule
-	reflexive & $\forall s \in S(s,s) \in R$ & all s are linked with 
-	themselves\\
-
-	transitive & $\forall s,t,u \in S:(s,t) \in R \& (t,u) \in R \Rightarrow
-	(s,u) \in R$ & \brcell{if s is related to t, and t is related to u,  \\
-		then s has a relation with u, \\ in other words, you can make shortcuts}\\
-
-	symmetrical &$\forall s,t \in S:(s,t) \in R \Rightarrow (t,s) \in R$ &
-	undirected, can go back and forth\\
-	euclidiean & $\forall s,t,u \in S:(s,t) \in R \& (s,u) \in R \Rightarrow
-	(t,u) \in R$ & \brcell{if s has a relation with t and u, then u and t \\
-	have a relation to, you can make triangles.} \\
-	serial & $\forall s \in S \exists t \in S (s,t) \in R$&
-	\brcell{All states in S are connected to \\ at least one other state in S}\\
-	an equivalance relation& R is \emph{reflexive, transitive and 
-	symetrical}& \brcell{the states R links to have something in \\ 
-	common (not equal?)} \\ \bottomrule
-\end{tabular}
 \end{document}

### 2015-11-17 evolutionaryComputation/practicals/report.tex
added some flavour text, langauge choice etc.

@@
 \author{Jappie Klooster}
-\title{A praktical look at genetic algorithms}
+\title{A practical look at genetic algorithms}
 \maketitle
 
+\section{Introduction}
+This is a report of the evolutionary practical assigment 1.
+In here I will discuss what this assignment contained, which
+methodology I used to execute the assignemnt and how the execution
+went.
+
+The intended audiance is people familiar with programming.
+
+\section{Assignment}
+The assignment was to implement the following functions and algorithms and
+then do experimentation with help of these.
+
+\subsection{Functions}
+\subsubsection{Counting functions}
+\[x_i \in \{0,1\}:CO(x_1\dots x_l)=\sum^l_{i=1}x_i\]
+\[x_i \in \{0,1\}:CO(x_1\dots x_l)=\sum^l_{i=1}i x_i\]
+\subsubsection{Trap function}
+\[TF(x_1\dots x_l) = \sum^{\frac{l}{k}-1}_{j=0}B(x_{jk_1}\dots x_{jk+k})\]
+
+with:
+
+\[B(x_1\dots x_k)=\left\{
+	\begin{matrix}
+		k & \mbox{ if } &   CO(x_1\dots x_k) = k \\
+		k - d - \frac{k-d}{k-1} CO(x) & \mbox{ if } & CO(x_1\dots x_k) < k \\
+	\end{matrix}
+\right. \]
+\section{Methodology}
+\subsection{Build tool}
+I use Gradle for building the project. Its the most advanced build platform
+available and makes it easy to add dependencies to the project.
+For example I use slf4j for logging.
+
+Its also handy in that it comes with a self installing script that can be run
+from any platform, Gradlew. So running the code I wrote on a different machine
+becomes a trivial task, Gradle makes sure the right dependencies are fetched.
+
+\subsubsection{Run guide}
+As said before its a trivial task to run the code. Downloading and installing
+of dependencies and even gradle happens automatically with help of gradlew.
+So the first run is slower, depending on your internet speed.
+
+\subsubsection{Libraries used}
+
+\subsection{Language choice}
+I chose to use Scala to implement this project. The main reason being that I've
+played with Scala before but never have done anything constructive with it.
+
+Scala is both a functional and an object orientated program that runs on the JVM
+It compiles to byte-code and is statically typed.
+
+It also can use any Java library and integrates seamlessly with Java, (although
+calling methods the methods of a Scala project from a Java project can be
+a little clumsy sometimes, because Scala supports operator overloading in
+a strange way).
+
+Other options I considered were:
+\begin{itemize}
+	\item Java, which I'm very familiar with but I find quite boring to use.
+		Its very verbose.
+	\item C\#, the language is good but setting it up is difficult since I
+		use Linux, it also has no Gradle support which makes it clunky,
+		you'd have to rely on .net libraries only which is limited. It
+		also prefers object orientated  rather than functional programming,
+		Scala prefers functional over object orientated.
+		Scala's syntax is also more concise.
+	\item Rust/Go, I'm curious about both these languages, but I'm not at
+		all familiar with them, so I probably should do first a little
+		hobby project with them before starting doing something serious like
+		this, I've already done this with scala before.
+\end{itemize}
+\subsection{Architecture}
+The architecture was created incrementally, I just implemented the functions
+without thinking much about how to combine them.  I first started with the
+fittest functions. Then I went on with selection, and finally I did the
+crossover.
+I just implemented one or 2 options of these and then created an evolution
+class that accepted functions with the right prototype. So now you can
+create an evolution by giving it functions.
+
+\section{Execution}
+
+\section{Conclusion}
+%What did I learn from my observation?
 \end{document}

### 2015-11-17 evolutionaryComputation/practicals/report.tex
added about unit tests

@@
 		hobby project with them before starting doing something serious like
-		this, I've already done this with scala before.
+		this, I've already done this with Scala before.
 \end{itemize}
@@
 
+\subsubsection{Unit tests}
+Because the assignment has very well defined behavior. I thought it was
+a very wise Idea to use unit tests. To do this I use a library called
+Scala test.
+
 \section{Execution}

### 2015-11-17 evolutionaryComputation/practicals/report.tex
added a todo list

@@
 
+\paragraph{Bugs found}
+Linearly scaled reported the wrong value because it used indeci, so it
+gave the first 1 a value of 0, and the second a value of 1 etc.
+
+Trap functions were simply wrongly implemented, instead of counting the
+ones they gave positional values of the ones in the block and used the
+block as a lookup value for that.
 \section{Execution}
+%TODO:
+% 1. Tournement selection is now implemented wrongly
+%	* group(2) to select the first half off the parents
+%	* randomly shuffle and group(2) again to select the second half
+% 2. Mutation, should be trivial with help of the members factory, decide how
+	% many bits will be mutated and then mutate those.
+% 3. Offspring generation should create 2 instead of one.
+% 4. Experiments
+%	Create tables % including:
+%		* min pop size for 29 out of 30 runs successful.
+%		* The average number of generations
+%		* The average CPU time required for running the GA with minimal pop size
+%		* average should be over 30 runs
+%	Draw a graph from the experiments
+%		* pop size on x, # succes on Y
 

### 2015-11-17 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Fitness.scala
fixed the issues indicated by the unit tests

@@ -13,16 +13,14 @@ object Fitness {
 			// select the elements where 1
 			filter(_._1 == '1').
 			// throw away the elements and keep the indeci
-			map(_._2).
+			map(_._2+1).
 			// result
 			sum
 	def blockValuation(block:Seq[Float])(s:IHasFitness) = s.getFitness.
-			// make a map like thing of it where the index is the key
-			zipWithIndex.
-			// select the elements where 1
-			filter(_._1 == '1').
-			// throw away the elements and keep the indeci
-			map(x => block.apply(x._2 % block.size)).
-			// result
-			sum
+			// make blocks of the string
+			grouped(block.size-1).
+			map(
+				// get the value specified in the block
+				str => block.apply(str.count(c=>c=='1'))
+			).sum
 }

### 2015-11-19 inteligentagents/notes4.tex
notes of lecture 4

\documentclass{article}
\begin{document}
\author{Jappie Klooster}
\title{Temporal magick or logic}
\maketitle
\section{Linear temporal magick}

$\phi_1\cup\phi_2$ = $\phi_1$ holds until $\phi_2$ becomes true.
$\Diamond\phi$ = sometimes
$O\phi$ = next time

$M=(S,\sigma, \pi)$ where $S$ is a set of states, $\sigma: N \to S$ is an
infinte seqence of states

\subsection{Semantics}
$\vdots$
$M,\sigma\vDash O\phi \LeftRightArrow M, \sigma^1 \vDash \phi$ next time
$M,\sigma\vDash \Diamond\phi \LeftRightArrow M, \sigma^n \vDash \phi$
invention, there is some time where $\phi$ is gonna hold.

\subsubsection{untill}
if $\phi_2$ holds somewhere it does not mean it holds everywhere after that
point

$\phi_1$ holds until $\phi_2$ holds.

weak untill = where $\phi_1$ holds everywhere because there is no
$\phi_2$, for example $\phi_1^{\bot}$

\section{Branching temproal magick}

Two kinds of formulas, state formulas, path formulas.
Any state formula can also act as a path formula.

\subsection{Models}
serial, any point has a continuation, time never stops.
$S \times S$ = cartesian product of S.

Fullpath is a ninfinite seq of states more or less true the tree.
These are more or less the lines of linear logic.
Fullpaths can start anywhere, not necisarly at the root.

$\sigma_1$ = already a choice of the branch, so no need for no trees.

\section{Combination of BDI}
E = optional
A = inevitable

\section{Operators magick}
$\Diamond\Diamond\phi=\Diamond\phi$
$\Box\Box\phi=\Box\phi$
$\Box\Diamond\phi$=Infenitly often there is a point where $\phi$ holds.
$\Diamond\Box\phi=$Some point in the future $\phi$ will hold indevinitly

\subsection{BDI}
Optionally $\Diamond p$ = there is a choice where p is gonna hold
Optionally $\Box r$= there is a timline where r is gonna hold
Inevitiably $\Diamond s$=  ???
Inevitiably $\Box b$=  b always holds

\section{BDI sementics}
s.t. = such that.

From t1 to t2, the agent has a memory, but he can doubt about the future.
There is the hidden asumption that agents have memory.

success = there exist some period in wehre the even was succesfull

\subsection{Axiom}
written like $GOAL(\alpha) \subset BEL(\alpha)$, but russel notation :s so
$\subset = \to$
$GOAL(\alpha) \to BEL(\alpha)$ = believe-goal compatibility
different from cohon and laveque, $BEL \alpha \to GOAL \alpha$, because
its a different framework, but the reason behind it is the same strangely
enough.

$INTEND(does(e))\to does(e)$ intention leads to action, which is controversial

\subsection{compatibility}
Sinning = ???, something with cutting of branches.

\section{CTD}
\subsection{Axioms}
$done(e) \to BEL(done(e))$,  you're awere that you've done something

\section{Praktical}
The general strategy is to write down what you have to proof and then to
read with help of the definitions.

Don't be sloppy and take an arbitrary M with for all's
\subsction{3.16}
\[\vDash \alpha \Rightarrow \vDash BEL x \alpha\]
Assume $\Vdash \alpha$, so $M,\sigma,v,n\vDash \alpha$ for all $M,\sigma,v,n$(*)
To prove $\vDash BEL x \alpha$, 
	ie $M,\sigma,v,n\vDash BEL x \alpha$ for all $M,\sigma,v,n\vDash$
	ie $M,\sigma,v,n\vDash\alpha$ for all $\sigma'$ with (\sigma,n)B[v(x)]
	\sigma' done by (*).

\subsection{3.26}
	To prove: $M,\sigma,v,n\vDash BEL x \alpha \to Goal x \alpha$ for all 
	$M,\sigma,v,n$.
	ie $M,\sigma,v,n\vDash BEL x \alpha \Rightarrow M,\sigma,v,n\vDash$ Goal
		$x \alpha$ for all $M,\sigma,v,n$
		ie $M,\sigma',v,n\vDash \alpha$ for all $\sigma'$ with $\sigma,n)
		B[v(x)]\sigam'\RIghtarrow M,\sigma'',v,n\vDash \alpha$ for all
		$\sigma''$ with $(\sigma,n)G[v(x)]\sigma''$

		Given:
		$(\sigma,n)G[v(x)]\sigma''\Rightarrow(\sigma,n)B[v(x)]\sigma''$
		$\vdots$

		Take arb $\sigma''$ with $(\sigma,n)G[v(s)]\sigam''$ then also 
		$(\sigma,n)B[v(x)]\sigma''$ so $M,\sigma,v,n\vDash \alpha$
\subsection(3.11)
Proof
\begin{prooftree}
	\AxiomC{}
	\UnaryinfC{}
\[ \vDash \neg LATER(\Diamond p) \neg LATER(\Diamond p) \LeftRightarrow 
\neg(\neg\Diamond p) \Leftrightarrow \neg(\neg\Diamond p \wedge 
\Diamond\Diamond p) \Leftrightarrow \neg(\neg\Diamond p \wedge 
\Diamond p)\]

\end{prooftree}

\subsection{3.13}
\begin{prooftree}
	\AxiomC{$\vDash \Diamond q \wedge BEFORE pq \to \Diamond p$}
	\UnaryInfC{To prove $M,\sigma,v,n\vDash \Diamond q \wedge BEFORE pq \to \Diamond p$ for all $M,\sigma,v,n$}
	\UnaryInfC{Take arbitrary $M,\sigma,v,n$ to prove $M,\sigma,v,n\vDash \vDash \Diamond q \wedge BEFORE pq \to \Diamond p$ for all $M,\sigma,v,n$}
	\Rightlabel{ie}
	\UnaryInfC{To prove if $M,\sigma,v,n\vDash \Diamond q$ and 
	$M,\sigma,v,n\vDash BEFORE pq$ then $M,\sigma,v,n\vDash Diamond p$ for all $M,\sigma,v,n$}
	\UnaryInfC{Supose $M,\sigma,v,n\vDash \vDash \Diamond q$ and 
	$M,\sigma,v,n\vDash BEFORE pq$}
	\Rightlabel{ie}
	\UnaryInfC{Supose $M,\sigma,v,n\vDash \exists x HAPPENS x;q?$ and 
		$M,\sigma,v,n\vDash \exists \forall c (HAPPENS c;q{?}) \to \exists q$ 
	$q\se c \wedge HAPPENS a;p{?}$}
	\Rightlabel{ie}
	\UnaryInfC{Supose $M,\sigma,v,n\vDash \exists x HAPPENS x;q?$ and 
		$M,\sigma,v,n\vDash \exists \forall c (HAPPENS c;q{?}) \to \exists q$ 
	$q\se c \wedge HAPPENS a;p{?}$}
\end{prooftree}

\end{document}

### 2015-11-19 evolutionaryComputation/notes4.tex
notes of lecture 4

\documentclass{article}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\usepackage{qtree}
\usepackage{tikz}
\usepackage{listings}
\usepackage{hyperref}
\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

\newcommand{\brcell}[2][l]{%
	\begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
\begin{document}
\author{Jappie Klooster}
\title{Evolutionary stretegies}
\maketitle

\section{difference from GA}
GA's are local searches, while ES are global searches.
\section{Comenly used settings}
Coveriance = the angle, no coveriance means semetrical

The closer to the optimum the smaller your variance should be. This reduces
the chance of jumping over the optimum.

Normally distributed = gausian

\section{mutation}

\section{reccombination}
the parents is the entire population

\section{selection}
 methods $(u,\lambda)$ and $(u+\lambda)$, $(u+\lambda)$ appears better from
 an optimization standpoint but self adoptation works worse with this one.

feet4, each edge = a weight, every neuron is a function.

\section{Combinatorial}
is a finite set of solutions and you have to optimize it.

\subsection{Local search}
always try to improve the current solution by applying small changes.

Meta heuristics, for example taboo list

\end{document}

### 2015-11-20 evolutionaryComputation/practicals/report.tex
updated the todo list
having the initial expiriment setup, but I should do some more debugging tomorrow...

@@
 block as a lookup value for that.
+\subsection{Uncertainties}
+Should the elitism selection use tournement selection?
+When precisly do we find the optimal solution?
+	Should I wait for convergence or just say a string
+	of all ones is good enough?
+	Is 1 solution good enough or should all members have
+	that solution
+	I'm assuming 1 solution and that I can give a string
+	of just ones.
+How many generations to consider. I chose arbitrarly 10.
+
+About randomly linked:
+	There is another way to do it: for every member randomly link
+	differently, and randomly pick 1 randomly linked instructionset
+	for offspring generation (or do the strongest).
+	Currently I do the global randomly linked, every run will however
+	have the same randomly linked value, I'm aware of this.
 \section{Execution}
 %TODO:
-% 1. Tournement selection is now implemented wrongly
-%	* group(2) to select the first half off the parents
-%	* randomly shuffle and group(2) again to select the second half
-% 2. Mutation, should be trivial with help of the members factory, decide how
-	% many bits will be mutated and then mutate those.
-% 3. Offspring generation should create 2 instead of one.
-% 4. Experiments
+% 1. Experiments
 %	Create tables % including:
@@
 %		* pop size on x, # succes on Y
+% maybe some more testing of the crossovers?
+% also try to implement a threading scheme??
 

### 2015-11-20 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Evolution.scala

@@ -7,7 +7,8 @@ class Evolution(
 	/**
 	 * fittest filter, (valuation, parents, children) => fittestpop
 	 */
-	fittestFilter: (IHasFitness => Int, Population, Population) => Population
+	fittestFilter: (IHasFitness => Int, Population, Population) => Population,
+	hasGoodEnoughSolution: Population => Boolean
 ) {
 	val mateSelector = mateSelectorFactory(valuation)
 	private def step(parents:Population):Population = fittestFilter(
@@ -19,9 +20,17 @@ class Evolution(
 				)
 			)
 		)
-	def startGenetic(seed:Population, times:Int): Array[Population]= 0.to(times)
-		.foldLeft(Array[Population]()){
-			case (Array(), i) => Array(seed)
-			case(e, i) => e :+ step(e.last)
+	def startGenetic(seed:Population, times:Int): Array[Population]= genetic(
+		Array(seed), Math.abs(times)
+	)
+	private def genetic(prev:Array[Population], times:Int): Array[Population]={
+		if (times <= 0) prev else{
+			val last = prev.last
+			if(hasGoodEnoughSolution(last)){
+				return prev
+			}
+			return genetic(prev :+ step(last),times-1)
 		}
+	}
 }
+

### 2015-11-20 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Experiment.scala

@@ -1,0 +1,77 @@
+package nl.jappieklooster.hw.ec
+
+import scala.util.Random
+/**
+ * This expirement tries to find the optemal number of population for a specific
+ * evolution.
+ * @param evolution
+ */
+class Experiment(evolution: Evolution, memberFactory:String => IMember) {
+	import Experiment._
+	private def createPopulation(size:Int) = {
+		Population.createOneZeros(memberFactory, geneLength, size)
+	}
+	def run():Seq[RunResult]= {
+		def findOptimumPopsize(consideringSize:Int, difference:Int): Seq[RunResult] = {
+			def expiriment(): Seq[RunResult]={
+				val population = createPopulation(consideringSize)
+
+				val startTime = System.currentTimeMillis()
+				val evolutionResult = evolution.startGenetic(population, generationCount)
+				val runtime = System.currentTimeMillis() - startTime
+
+				// sucesfull if early termination or the last population is good enough
+				val success = evolutionResult.length != generationCount || hasGoodEnoughSolution(evolutionResult.last)
+				Seq(RunResult(consideringSize, success, runtime))
+			}
+			if(consideringSize > maxPopSize){
+				// we went to deep so we failed
+				return Seq(RunResult(consideringSize, false, 0))
+			}
+
+			val result = expiriment()
+			// 10, 20, 40, 80, 160, 320, 240, 280, 260, 250.
+			val hadSuccess = consideringSize != difference*2
+
+			val newDiff = if(hadSuccess) difference/2 else difference*2
+
+			if(newDiff < minPopSize){
+				return result
+			}
+			if(!result.head.success){
+				return result ++ findOptimumPopsize(consideringSize+newDiff,newDiff)
+			}
+			return result ++ findOptimumPopsize(consideringSize-difference/2, difference/2)
+		}
+		findOptimumPopsize(minPopSize,minPopSize/2)
+	}
+}
+case class RunResult(popSize:Int, success:Boolean, runtime:Long)
+
+object Experiment{
+
+	val generationCount = 10
+	val minPopSize = 10
+	val maxPopSize = 1280
+	val geneLength = 100
+
+	def create(
+		random:Random,
+		valuationFunction:IHasFitness => Int,
+		variationOperators:Seq[(PairedPopulation => Population, String => IMember)]
+	):Seq[Experiment] = variationOperators.map(
+		variation => new Experiment(new Evolution(
+			valuationFunction,
[…]

### 2015-11-20 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Fitness.scala

@@ -24,5 +24,5 @@ object Fitness {
 			map(
 				// get the value specified in the block
 				str => block.apply(str.count(whereCharIsOne))
-			).sum
+			).sum.toInt
 }

### 2015-11-20 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -3,6 +3,9 @@ package nl.jappieklooster.hw.ec
 import org.slf4j.LoggerFactory
 
 import scala.util.Random
+import OffspringGenerator._
+import MemberFactories._
+import Fitness._
 
 object Main{
 	case class Configurations(strLength:Int,popSize:Int,genCount:Int, evolution:Evolution)
@@ -11,34 +14,39 @@ object Main{
 	def main(args:Array[String]){
 		log.info("starting ea")
 
-		val stringLength = 20
-		val populationSize = 20
-		val population = Population.createOneZeros(
-			MemberFactories.withCoinfliptimesMutation(
-				MemberFactories.simpleMember,
-				random
-			),
-			stringLength,
-			populationSize
+		
+		val crossMethodsTight:Seq[(PairedPopulation => Population, String => IMember)] = Seq(
+			(twoPointCross _, tightlyLinked _),
+			(uniformCross _, tightlyLinked _),
+			(uniformCross _, withCoinfliptimesMutation(tightlyLinked, random) _)
 		)
-		val evolution = new Evolution(
-			Fitness.uniformlyScaledCountOnes,
-			MateSelection.createCompeteWithRandomTournement(random),
-			OffspringGenerator.uniformCross,
-			FittestFilter.tournementElitism
+		val crossMethodsRandom:Seq[(PairedPopulation => Population, String => IMember)] = Seq(
+			(twoPointCross _, createRandomlyLinked),
+			(uniformCross _, createRandomlyLinked),
+			(uniformCross _, withCoinfliptimesMutation(createRandomlyLinked, random) _)
 		)
+		val deciptive:Seq[Float] = Seq(4,0,1,2,3)
+		val nonDeciptive:Seq[Float] = Seq(4,0,0.5f,1,1.5f)
+		val expirements =
+		Experiment.create(random, uniformlyScaledCountOnes, crossMethodsTight) ++
+		Experiment.create(random, linearlyScaledCountOnes, crossMethodsTight) ++
+		Experiment.create(random, blockValuation(deciptive), crossMethodsTight) ++
+		Experiment.create(random, blockValuation(nonDeciptive), crossMethodsTight) ++
+		Experiment.create(random, blockValuation(deciptive), crossMethodsRandom) ++
+		Experiment.create(random, blockValuation(nonDeciptive), crossMethodsRandom)
 
-		val startTime = System.currentTimeMillis()
[…]

### 2015-11-20 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Member.scala

@@ -16,8 +16,8 @@ case class ValGeneDecoupledMember(valuation:String, genes:String) extends IMembe
 }
 
 object MemberFactories{
-	def simpleMember(str:String):IMember = SinglyMember(str)
-	def globalRandomizedGenes(instructions:TraversableOnce[Int])(str:String):IMember = {
+	def tightlyLinked(str:String):IMember = SinglyMember(str)
+	def randomlyLinked(instructions:TraversableOnce[Int])(str:String):IMember = {
 		val genes = instructions.map(inx => str.charAt(inx)).mkString
 		ValGeneDecoupledMember(str,genes)
 	}

### 2015-11-20 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Population.scala

@@ -10,12 +10,12 @@ case class Population(members:Seq[IMember], memberFactory:String => IMember) ext
 object Population{
 	def createOneZeros(
 			memberFactory:String => IMember,
-			length:Int,
-			count:Int):Population =  Population(
-		0.to(count).map(
+			geneLength:Int,
+			popCount:Int):Population =  Population(
+		0.to(popCount).map(
 			x => memberFactory(
 				// it says, from 1 to length concat 1 or 0.
-				0.to(length).foldLeft(""){(b, x) => b + Main.randomGene}
+				0.to(geneLength).foldLeft(""){(b, x) => b + Main.randomGene}
 			)
 		),
 		memberFactory

### 2015-11-20 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Experiment.scala
this build runs but it doesn't find no solutions...

@@ -1,5 +1,7 @@
 package nl.jappieklooster.hw.ec
 
+import com.itextpdf.text.log.LoggerFactory
+
 import scala.util.Random
 /**
  * This expirement tries to find the optemal number of population for a specific
@@ -8,41 +10,42 @@ import scala.util.Random
  */
 class Experiment(evolution: Evolution, memberFactory:String => IMember) {
 	import Experiment._
+	val log = LoggerFactory.getLogger(this.getClass)
 	private def createPopulation(size:Int) = {
 		Population.createOneZeros(memberFactory, geneLength, size)
 	}
-	def run():Seq[RunResult]= {
-		def findOptimumPopsize(consideringSize:Int, difference:Int): Seq[RunResult] = {
-			def expiriment(): Seq[RunResult]={
-				val population = createPopulation(consideringSize)
+	private def findOptimumPopsize(consideringSize:Int, difference:Int): Seq[RunResult] = {
+		def expiriment(): Seq[RunResult]={
+			val population = createPopulation(consideringSize)
 
-				val startTime = System.currentTimeMillis()
-				val evolutionResult = evolution.startGenetic(population, generationCount)
-				val runtime = System.currentTimeMillis() - startTime
+			val startTime = System.currentTimeMillis()
+			val evolutionResult = evolution.startGenetic(population, generationCount)
+			val runtime = System.currentTimeMillis() - startTime
 
-				// sucesfull if early termination or the last population is good enough
-				val success = evolutionResult.length != generationCount || hasGoodEnoughSolution(evolutionResult.last)
-				Seq(RunResult(consideringSize, success, runtime))
-			}
-			if(consideringSize > maxPopSize){
-				// we went to deep so we failed
-				return Seq(RunResult(consideringSize, false, 0))
-			}
-
-			val result = expiriment()
-			// 10, 20, 40, 80, 160, 320, 240, 280, 260, 250.
-			val hadSuccess = consideringSize != difference*2
-
-			val newDiff = if(hadSuccess) difference/2 else difference*2
-
-			if(newDiff < minPopSize){
-				return result
-			}
-			if(!result.head.success){
-				return result ++ findOptimumPopsize(consideringSize+newDiff,newDiff)
-			}
-			return result ++ findOptimumPopsize(consideringSize-difference/2, difference/2)
+			// sucesfull if early termination or the last population is good enough
[…]

### 2015-11-20 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -35,12 +35,11 @@ object Main{
 		Experiment.create(random, blockValuation(deciptive), crossMethodsRandom) ++
 		Experiment.create(random, blockValuation(nonDeciptive), crossMethodsRandom)
 
-		val results = expirements.map(exp => 0.to(30).map(x=>{
-			log.info(s"running $x")
-			val result = exp.run
-			log.info(s"result: $result")
-			(x, result)
-		} ))
+		val results = expirements.par.flatMap(exp => 0.to(30).par.map(x=> (x,exp.run)))
+		for(result <- results){
+			log.info(s"result: ${result._1}")
+			log.info(s"${result._2}")
+		}
 
 		import org.sameersingh.scalaplot.Implicits._
 

### 2015-11-20 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Population.scala

@@ -12,7 +12,7 @@ object Population{
 			memberFactory:String => IMember,
 			geneLength:Int,
 			popCount:Int):Population =  Population(
-		0.to(popCount).map(
+		0.to(popCount-1).map(
 			x => memberFactory(
 				// it says, from 1 to length concat 1 or 0.
 				0.to(geneLength).foldLeft(""){(b, x) => b + Main.randomGene}

### 2015-11-20 evolutionaryComputation/practicals/code/src/test/scala/nl/jappieklooster/hw/ec/tests/ExpirimentTests.scala

@@ -1,0 +1,17 @@
+package nl.jappieklooster.hw.ec.tests
+
+import nl.jappieklooster.hw.ec.{SinglyMember, Population}
+import org.scalatest.{Matchers, FlatSpec}
+import nl.jappieklooster.hw.ec.Experiment._
+
+class ExpirimentTests extends FlatSpec with Matchers{
+
+	def createMember(str:Seq[String]) = Population(str.map(SinglyMember(_)), null)
+	"has good enough result " should " not be true " in {
+		hasGoodEnoughSolution(createMember(Seq("0111", "1010", "1010", "0000", "1000"))) should be (false)
+	}
+	"has good enough result " should " be true " in {
+		hasGoodEnoughSolution(createMember(Seq("0111", "1010", "111111", "1010", "0000", "1000"))) should be (true)
+	}
+
+}

### 2015-11-20 evolutionaryComputation/practicals/code/src/test/scala/nl/jappieklooster/hw/ec/tests/MateSelectionTests.scala

@@ -19,7 +19,7 @@ class MateSelectionTests extends FlatSpec with Matchers{
 		val two = SinglyMember("2")
 		val three = SinglyMember("3")
 		val four = SinglyMember("4")
-		subject.tournement(one,two,three,four) should be (Pair(two,four))
+		subject.tournement(one,two) should be (two)
 	}
 	"the random turnement when rigged " should "produce the following sequent " in {
 		val subject = createSubject

### 2015-11-21 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Evolution.scala
tried doing a stocastic one in this way, but realized its wrong

@@ -20,10 +20,10 @@ class Evolution(
 				)
 			)
 		)
-	def startGenetic(seed:Population, times:Int): Array[Population]= genetic(
-		Array(seed), Math.abs(times)
+	def startGenetic(seed:Population, times:Int): Seq[Population]= genetic(
+		Seq(seed), Math.abs(times)
 	)
-	private def genetic(prev:Array[Population], times:Int): Array[Population]={
+	private def genetic(prev:Seq[Population], times:Int): Seq[Population]={
 		if (times <= 0) prev else{
 			val last = prev.last
 			if(hasGoodEnoughSolution(last)){

### 2015-11-21 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Experiment.scala

@@ -14,47 +14,54 @@ class Experiment(evolution: Evolution, memberFactory:String => IMember) {
 	private def createPopulation(size:Int) = {
 		Population.createOneZeros(memberFactory, geneLength, size)
 	}
+	def run(consideringSize:Int): RunResult={
+		val population = createPopulation(consideringSize)
+
+		val startTime = System.currentTimeMillis()
+		val evolutionResult = evolution.startGenetic(population, generationCount)
+		val runtime = System.currentTimeMillis() - startTime
+
+		// sucesfull if early termination or the last population is good enough
+		val success = hasGoodEnoughSolution(evolutionResult.last)
+		RunResult(consideringSize, success, runtime, evolutionResult.length)
+	}
+
 	private def findOptimumPopsize(consideringSize:Int, difference:Int): Seq[RunResult] = {
-		def expiriment(): Seq[RunResult]={
-			val population = createPopulation(consideringSize)
-
-			val startTime = System.currentTimeMillis()
-			val evolutionResult = evolution.startGenetic(population, generationCount)
-			val runtime = System.currentTimeMillis() - startTime
-
-			// sucesfull if early termination or the last population is good enough
-			val success = hasGoodEnoughSolution(evolutionResult.last)
-			Seq(RunResult(consideringSize, success, runtime))
-		}
 		if(consideringSize > maxPopSize){
 			// we went to deep so we failed
-			return Seq(RunResult(consideringSize, false, 0))
+			return Seq(RunResult(consideringSize, false, 0,0))
 		}
 
-		val result = expiriment()
+		val result = run(consideringSize)
 		// 10, 20, 40, 80, 160, 320, 240, 280, 260, 250.
 		val hadSuccess = consideringSize != difference*2
 
 		val newDiff = if(hadSuccess) difference/2 else difference*2
 
-		if(newDiff < minPopSize){
-			return result
+		if(newDiff < popUnit){
+			return Seq(result)
 		}
-		if(!result.head.success){
+		if(!result.success){
 			return result ++ findOptimumPopsize(consideringSize+newDiff,newDiff)
 		}
 		return result ++ findOptimumPopsize(consideringSize-difference/2, difference/2)
 	}
-	def run():Seq[RunResult]= {
-		findOptimumPopsize(minPopSize,minPopSize/2)
+	def bisectionalSearch() = findOptimumPopsize(popUnit ,popUnit/2)
+}
+
[…]

### 2015-11-21 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -28,14 +28,17 @@ object Main{
 		val deciptive:Seq[Float] = Seq(4,0,1,2,3)
 		val nonDeciptive:Seq[Float] = Seq(4,0,0.5f,1,1.5f)
 		val expirements =
-		Experiment.create(random, uniformlyScaledCountOnes, crossMethodsTight) ++
-		Experiment.create(random, linearlyScaledCountOnes, crossMethodsTight) ++
-		Experiment.create(random, blockValuation(deciptive), crossMethodsTight) ++
-		Experiment.create(random, blockValuation(nonDeciptive), crossMethodsTight) ++
-		Experiment.create(random, blockValuation(deciptive), crossMethodsRandom) ++
-		Experiment.create(random, blockValuation(nonDeciptive), crossMethodsRandom)
+		Experiment.create(random, uniformlyScaledCountOnes, Seq((uniformCross _, tightlyLinked _))) ++
+		//Experiment.create(random, uniformlyScaledCountOnes, crossMethodsTight) ++
+		//Experiment.create(random, linearlyScaledCountOnes, crossMethodsTight) ++
+		//Experiment.create(random, blockValuation(deciptive), crossMethodsTight) ++
+		//Experiment.create(random, blockValuation(nonDeciptive), crossMethodsTight) ++
+		//Experiment.create(random, blockValuation(deciptive), crossMethodsRandom) ++
+		//Experiment.create(random, blockValuation(nonDeciptive), crossMethodsRandom) ++
+		Nil
 
-		val results = expirements.par.flatMap(exp => 0.to(30).par.map(x=> (x,exp.run)))
+		val runCount = 4
+		val results = expirements.flatMap(exp => 1.to(runCount).map(x=> (x,exp.run(10))))
 		for(result <- results){
 			log.info(s"result: ${result._1}")
 			log.info(s"${result._2}")

### 2015-11-21 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/StoasticExperiment.scala

@@ -1,0 +1,43 @@
+package nl.jappieklooster.hw.ec
+
+import Experiment._
+class StoasticExperiment(experiment: Experiment) {
+
+	private def geussLowest={
+		// run it four times
+		val peekResult = 0.to(3).flatMap(x=>experiment.bisectionalSearch)
+		peekResult.foldLeft(RunResult(Int.MaxValue,false,Int.MaxValue,Int.MinValue))(
+			(cur,elem)=> cur match{
+				case RunResult(_,false,_) => elem
+				case s if elem.success => if(s.popSize < elem.popSize) s else elem
+				case s => s
+			}
+		)
+	}
+	def stoasticEnsurance(required:Int, tolerance:Int):Seq[RunResult] ={
+		val smallest = geussLowest
+		if(!smallest.success){
+			return Seq(smallest)
+		}
+		def verifyLowest(currentPop:Int, faults:Int, index:Int) : Seq[RunResult] = {
+			if(index >= required){
+				return Nil
+			}
+			// to deep gaurd
+			if(currentPop > maxPopSize){
+				return Seq(RunResult(currentPop, false, 0,0))
+			}
+			if(faults > tolerance){
+				return verifyLowest(currentPop+popUnit,0,0)
+			}
+			val exp = experiment.run(currentPop)
+			val newFaults = if (!exp.success) faults else faults+1
+			Seq(exp) ++ verifyLowest(currentPop, newFaults,index+1)
+		}
+		verifyLowest(smallest.popSize,0,0)
+	}
+}
+
+case class StoasticRun(runs:Seq[RunResult]) extends EmulateSeq[RunResult]{
+	override def baseSeq: Seq[RunResult] = runs
+}

### 2015-11-21 evolutionaryComputation/practicals/report.tex
this build runs but it doesn't find no solutions...
tried doing a stocastic one in this way, but realized its wrong
implmeneted the stoastic thingy

@@
 	have the same randomly linked value, I'm aware of this.
+
+I'm not quite sure about the graphs. I have to show the number of successes
+out of the 30 runs, but this is already set in stone, because its either
+29 or 30, or its no successes.
+
+The stocastic ensursance wants to make sure that 29 out of 30 runs are
+'good enough', but combining this with the bidisectional search is difficult.
+because the bidisectional search may happen to end up just higher than possible.
+
+I considered to combine the bidisectional search with the stocastic ensurance.
+But that would be way slower, and it won't be much better (or even worse depending
+on how you do it) then my current implementation.
+
 \section{Execution}
@@
 % 1. Experiments
+% Make the actuall expiriment work\dots
 %	Create tables % including:

### 2015-11-21 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Experiment.scala

@@ -42,9 +42,9 @@ class Experiment(evolution: Evolution, memberFactory:String => IMember) {
 			return Seq(result)
 		}
 		if(!result.success){
-			return result ++ findOptimumPopsize(consideringSize+newDiff,newDiff)
+			return result +: findOptimumPopsize(consideringSize+newDiff,newDiff)
 		}
-		return result ++ findOptimumPopsize(consideringSize-difference/2, difference/2)
+		return result +: findOptimumPopsize(consideringSize-difference/2, difference/2)
 	}
 	def bisectionalSearch() = findOptimumPopsize(popUnit ,popUnit/2)
 }

### 2015-11-21 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -37,11 +37,10 @@ object Main{
 		//Experiment.create(random, blockValuation(nonDeciptive), crossMethodsRandom) ++
 		Nil
 
-		val runCount = 4
-		val results = expirements.flatMap(exp => 1.to(runCount).map(x=> (x,exp.run(10))))
+
+		val results = expirements.map(ex => new StoasticExperiment(ex)).map(ex => ex.stoasticEnsurance(30,1))
 		for(result <- results){
-			log.info(s"result: ${result._1}")
-			log.info(s"${result._2}")
+			log.info(s"result: $result")
 		}
 
 		import org.sameersingh.scalaplot.Implicits._

### 2015-11-21 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/StoasticExperiment.scala

@@ -8,7 +8,7 @@ class StoasticExperiment(experiment: Experiment) {
 		val peekResult = 0.to(3).flatMap(x=>experiment.bisectionalSearch)
 		peekResult.foldLeft(RunResult(Int.MaxValue,false,Int.MaxValue,Int.MinValue))(
 			(cur,elem)=> cur match{
-				case RunResult(_,false,_) => elem
+				case RunResult(_,false,_,_) => elem
 				case s if elem.success => if(s.popSize < elem.popSize) s else elem
 				case s => s
 			}

### 2015-11-21 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Evolution.scala
able to count valuations, also various bug fixes

@@ -1,7 +1,7 @@
 package nl.jappieklooster.hw.ec
 
 class Evolution(
-	valuation: IHasFitness => Int,
+	val valuation: FitnessEvaluator,
 	mateSelectorFactory:(IHasFitness => Int)=>IMateSelector,
 	offSpringGenerator: PairedPopulation => Population,
 	/**
@@ -10,9 +10,9 @@ class Evolution(
 	fittestFilter: (IHasFitness => Int, Population, Population) => Population,
 	hasGoodEnoughSolution: Population => Boolean
 ) {
-	val mateSelector = mateSelectorFactory(valuation)
+	val mateSelector = mateSelectorFactory(valuation.valuate)
 	private def step(parents:Population):Population = fittestFilter(
-			valuation,
+			valuation.valuate,
 			parents,
 			offSpringGenerator(
 				mateSelector.selectFrom(
@@ -20,17 +20,19 @@ class Evolution(
 				)
 			)
 		)
-	def startGenetic(seed:Population, times:Int): Seq[Population]= genetic(
-		Seq(seed), Math.abs(times)
+	def startGenetic(seed:Population): Seq[Population]= genetic(
+		Seq(seed)
 	)
-	private def genetic(prev:Seq[Population], times:Int): Seq[Population]={
-		if (times <= 0) prev else{
-			val last = prev.last
-			if(hasGoodEnoughSolution(last)){
-				return prev
-			}
-			return genetic(prev :+ step(last),times-1)
+	private def genetic(prev:Seq[Population]): Seq[Population]={
+		val parents = prev.last
+		if(hasGoodEnoughSolution(parents)){
+			return prev
 		}
+		val children = step(parents)
+		if(children == parents){
+			return prev
+		}
+		genetic(prev :+ children)
 	}
 }
 

### 2015-11-21 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Experiment.scala

@@ -8,7 +8,7 @@ import scala.util.Random
  * evolution.
  * @param evolution
  */
-class Experiment(evolution: Evolution, memberFactory:String => IMember) {
+class Experiment(val name:String, evolution: Evolution, memberFactory:String => IMember) {
 	import Experiment._
 	val log = LoggerFactory.getLogger(this.getClass)
 	private def createPopulation(size:Int) = {
@@ -18,18 +18,18 @@ class Experiment(evolution: Evolution, memberFactory:String => IMember) {
 		val population = createPopulation(consideringSize)
 
 		val startTime = System.currentTimeMillis()
-		val evolutionResult = evolution.startGenetic(population, generationCount)
+		val evolutionResult = evolution.startGenetic(population)
 		val runtime = System.currentTimeMillis() - startTime
 
 		// sucesfull if early termination or the last population is good enough
 		val success = hasGoodEnoughSolution(evolutionResult.last)
-		RunResult(consideringSize, success, runtime, evolutionResult.length)
+		RunResult(consideringSize, success, runtime, evolutionResult.length,evolution.valuation.countCalls())
 	}
 
 	private def findOptimumPopsize(consideringSize:Int, difference:Int): Seq[RunResult] = {
 		if(consideringSize > maxPopSize){
 			// we went to deep so we failed
-			return Seq(RunResult(consideringSize, false, 0,0))
+			return Seq(RunResult.createFailed(consideringSize))
 		}
 
 		val result = run(consideringSize)
@@ -49,15 +49,17 @@ class Experiment(evolution: Evolution, memberFactory:String => IMember) {
 	def bisectionalSearch() = findOptimumPopsize(popUnit ,popUnit/2)
 }
 
-case class RunResult(popSize:Int, success:Boolean, runtime:Long, generationCount:Int){
+case class RunResult(popSize:Int, success:Boolean, runtime:Long, generationCount:Int, fitnessCallCount:Int){
 	override def toString() :String = {
 		s"Run(pop#=$popSize, success=$success, time=$runtime, gen#=$generationCount)"
 	}
 }
 
+object RunResult{
+	def createFailed(size: Int) = RunResult(size,false,0,0,0)
+}
 object Experiment{
 
-	val generationCount = 30
 	/**
 	 * Don't consider pop diferences smaller than this unit
 	 */
@@ -66,12 +68,15 @@ object Experiment{
 	val geneLength = 100
 
 	def create(
+		name:String,
 		random:Random,
[…]

### 2015-11-21 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Fitness.scala

@@ -3,6 +3,19 @@ package nl.jappieklooster.hw.ec
 trait IHasFitness{
 	def getFitness:String
 }
+trait FitnessEvaluator{
+	private var calls = 0
+	protected def getFunction:IHasFitness => Int
+	def valuate(x:IHasFitness):Int ={
+		calls += 1
+		getFunction(x)
+	}
+	def countCalls():Int = {
+		val result = calls
+		calls = 0
+		result
+	}
+}
 
 object Fitness {
 	private def whereCharIsOne(c:Char) = c == '1'
@@ -25,4 +38,7 @@ object Fitness {
 				// get the value specified in the block
 				str => block.apply(str.count(whereCharIsOne))
 			).sum.toInt
+	def createProbe(f:IHasFitness=>Int):FitnessEvaluator = new FitnessEvaluator {
+		protected override def getFunction: (IHasFitness) => Int = f
+	}
 }

### 2015-11-21 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -15,32 +15,32 @@ object Main{
 		log.info("starting ea")
 
 		
-		val crossMethodsTight:Seq[(PairedPopulation => Population, String => IMember)] = Seq(
-			(twoPointCross _, tightlyLinked _),
-			(uniformCross _, tightlyLinked _),
-			(uniformCross _, withCoinfliptimesMutation(tightlyLinked, random) _)
+		val crossMethodsTight = Seq(
+	//		(twoPointCross _, tightlyLinked _, " Tight 2x"),
+	//		(uniformCross _, tightlyLinked _, " Tight ux"),
+			(uniformCross _, withCoinfliptimesMutation(tightlyLinked, random) _, " Tight ux, R")
 		)
-		val crossMethodsRandom:Seq[(PairedPopulation => Population, String => IMember)] = Seq(
-			(twoPointCross _, createRandomlyLinked),
-			(uniformCross _, createRandomlyLinked),
-			(uniformCross _, withCoinfliptimesMutation(createRandomlyLinked, random) _)
+		val crossMethodsRandom = Seq(
+			(twoPointCross _, createRandomlyLinked, " Random 2x, "),
+			(uniformCross _, createRandomlyLinked, " Random ux"),
+			(uniformCross _, withCoinfliptimesMutation(createRandomlyLinked, random) _, " Random ux, R")
 		)
 		val deciptive:Seq[Float] = Seq(4,0,1,2,3)
 		val nonDeciptive:Seq[Float] = Seq(4,0,0.5f,1,1.5f)
 		val expirements =
-		Experiment.create(random, uniformlyScaledCountOnes, Seq((uniformCross _, tightlyLinked _))) ++
-		//Experiment.create(random, uniformlyScaledCountOnes, crossMethodsTight) ++
-		//Experiment.create(random, linearlyScaledCountOnes, crossMethodsTight) ++
+		Experiment.create("uni scaled - ", random, uniformlyScaledCountOnes, crossMethodsTight) ++
+		//Experiment.create("linearly scaled - ", random, linearlyScaledCountOnes, crossMethodsTight) ++
 		//Experiment.create(random, blockValuation(deciptive), crossMethodsTight) ++
 		//Experiment.create(random, blockValuation(nonDeciptive), crossMethodsTight) ++
 		//Experiment.create(random, blockValuation(deciptive), crossMethodsRandom) ++
[…]

### 2015-11-21 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Member.scala

@@ -29,7 +29,7 @@ object MemberFactories{
 			}
 			integer
 		}
-		val flipCount = countFlips(0)
+		val flipCount = countFlips(-1)
 		val mutated = 0.to(flipCount).foldLeft(str)((str, modficationCount) => {
 			def decideIndex(prev:Int):Int = if((prev == 0) || (prev == str.length))
 				decideIndex(random.nextInt(str.length)) else prev

### 2015-11-21 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/StoasticExperiment.scala

@@ -1,43 +1,63 @@
 package nl.jappieklooster.hw.ec
 
 import Experiment._
-class StoasticExperiment(experiment: Experiment) {
+class StoasticExperiment(val experiment: Experiment) {
 
 	private def geussLowest={
 		// run it four times
 		val peekResult = 0.to(3).flatMap(x=>experiment.bisectionalSearch)
-		peekResult.foldLeft(RunResult(Int.MaxValue,false,Int.MaxValue,Int.MinValue))(
+		peekResult.foldLeft(RunResult.createFailed(Int.MaxValue))(
 			(cur,elem)=> cur match{
-				case RunResult(_,false,_,_) => elem
+				case RunResult(_,false,_,_,_) => elem
 				case s if elem.success => if(s.popSize < elem.popSize) s else elem
 				case s => s
 			}
 		)
 	}
-	def stoasticEnsurance(required:Int, tolerance:Int):Seq[RunResult] ={
+	def stoasticEnsurance(required:Int, tolerance:Int):StoasticRun ={
 		val smallest = geussLowest
 		if(!smallest.success){
-			return Seq(smallest)
+			return StoasticRun(Seq(smallest), required)
 		}
 		def verifyLowest(currentPop:Int, faults:Int, index:Int) : Seq[RunResult] = {
+			def failed = Seq(RunResult.createFailed(currentPop))
 			if(index >= required){
 				return Nil
 			}
 			// to deep gaurd
 			if(currentPop > maxPopSize){
-				return Seq(RunResult(currentPop, false, 0,0))
+				return failed
 			}
 			if(faults > tolerance){
-				return verifyLowest(currentPop+popUnit,0,0)
+
+				val newPopCount = currentPop+popUnit
+				return verifyLowest(newPopCount,0,0)
 			}
 			val exp = experiment.run(currentPop)
-			val newFaults = if (!exp.success) faults else faults+1
+			val newFaults = if (exp.success) faults else faults+1
 			Seq(exp) ++ verifyLowest(currentPop, newFaults,index+1)
 		}
-		verifyLowest(smallest.popSize,0,0)
+		StoasticRun(verifyLowest(smallest.popSize,0,0),required)
 	}
 }
 
-case class StoasticRun(runs:Seq[RunResult]) extends EmulateSeq[RunResult]{
-	override def baseSeq: Seq[RunResult] = runs
+import util.Properties.lineSeparator
[…]

### 2015-11-22 evolutionaryComputation/practicals/report.tex
finished the table
added a pargarph of speedup atempts

@@
 
+\subsection{Speed up attempts}
+\subsubsection{Paralization}
+Since were doing string comparisons I had the idea of ussing parlization to
+speed up the algoritm by doing paralization. This can be done very easily
+in scala with par collections (and the fact that imutibility is default).
+
+However once I implemented this, I came to the conclusion it was not faster.
+The overhead for delegating and creating threads was more than the parilization
+gained.
+
+The problem is that the evaluation function, even with gene length of 100, is
+to cheap to make any reasonable paralization effecitve. However what was effictive
+is the parilization of the expiriments. This reduced total runtime 
+significantly. however this is not seen in the statistics.
+
+\subsubsection{Evaluate each member on creation}
 \section{Execution}
@@
 % 1. Experiments
-% Make the actuall expiriment work\dots
-%	Create tables % including:
-%		* min pop size for 29 out of 30 runs successful.
-%		* The average number of generations
-%		* The average CPU time required for running the GA with minimal pop size
-%		* average should be over 30 runs
+% Use stoastic within the algoritm
 %	Draw a graph from the experiments

### 2015-11-22 evolutionaryComputation/practicals/report.tex
only create children if there are new genes at play

@@
 \subsubsection{Evaluate each member on creation}
+Members are going to be evaluated always, initially I would evaluate the member
+everytime it needed to be evaluated, but in later version I just aded the
+evaluation to the case class. Its not going to change anyways.
+
+\subsubsection{Only create children if the genes are new}
+If the genes are new then create the children otherwise return the parent.
 \section{Execution}

### 2015-11-22 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -29,7 +29,7 @@ object Main{
 		val nonDeciptive:Seq[Float] = Seq(4,0,0.5f,1,1.5f)
 		val expirements =
 		Experiment.create("uni scaled - ", random, uniformlyScaledCountOnes, crossMethodsTight) ++
-		//Experiment.create("linearly scaled - ", random, linearlyScaledCountOnes, crossMethodsTight) ++
+		Experiment.create("linearly scaled - ", random, linearlyScaledCountOnes, crossMethodsTight) ++
 		//Experiment.create("block decpt", random, blockValuation(deciptive), crossMethodsTight) ++
 		//Experiment.create("block non", random, blockValuation(nonDeciptive), crossMethodsTight) ++
 		//Experiment.create("block decpt", random, blockValuation(deciptive), crossMethodsRandom) ++

### 2015-11-22 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/PairedPopulation.scala

@@ -3,7 +3,7 @@ package nl.jappieklooster.hw.ec
 trait Genetic{
 	def genes:String
 }
-case class Pair(father:Genetic, mother:Genetic)
+case class Pair(father:IMember, mother:IMember)
 case class PairedPopulation(members:Seq[Pair], memberFactory:String => IMember) extends EmulateSeq[Pair]{
 	override def baseSeq = members
 	/**
@@ -19,16 +19,28 @@ case class PairedPopulation(members:Seq[Pair], memberFactory:String => IMember) 
 		Population(
 			members.flatMap(
 				fumu => {
+					val genesF = fumu.father.genes
+					val genesM = fumu.mother.genes
 					def createGenes = how(
 						// just make key value like structure of the indiviudal genes
 						// for comparison
-						fumu.father.genes.zip(fumu.mother.genes)
+						genesF.zip(genesM)
 					// mkstring uses a builder to avoid ridiculous concatination problems
 					).mkString
+					def createIfNew(child:String):IMember ={
+						import OffspringGenerator._
+						if(strEq(child,genesF)){
+							return fumu.father
+						}
+						if(strEq(child,genesM)){
+							return fumu.mother
+						}
+						memberFactory(child)
+					}
 
 					// we need two children so we return a list, flatmap will
 					// flatten that list again later
-					List(memberFactory(createGenes), memberFactory(createGenes))
+					List(createIfNew(createGenes), createIfNew(createGenes))
 				}
 			),
 			memberFactory
@@ -36,6 +48,12 @@ case class PairedPopulation(members:Seq[Pair], memberFactory:String => IMember) 
 	}
 }
 object OffspringGenerator{
+	def strEq(one:String,two:String): Boolean={
+		if(one.hashCode != two.hashCode){
+			return false
+		}
+		one.equals(two)
+	}
 	def uniformCross(x:PairedPopulation) : Population = x.pairedToPop(
 		/* kindoff a weird implementation, but it works because
 		 1==1V0==0, then the selection is irrelevant

### 2015-11-22 evolutionaryComputation/practicals/report.tex
running all tests

@@
 If the genes are new then create the children otherwise return the parent.
+This reduces evaluations and object creations. Its becomes more effective
+at deeper generation, it even slows down the fastest GA's by a few miliseconds
+but the deeper GA are lots faster.
+
+This can be done safely because strings are immutable and so are the members.
+So I can reuse them whenever its needed.
+
+To reduce the amount of string comparisons, first the hascode function is called
+This will cache itself, I know this because I inspected the source code of the
+String class in Java.
 \section{Execution}

### 2015-11-22 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -30,10 +30,10 @@ object Main{
 		val expirements =
 		Experiment.create("uni scaled - ", random, uniformlyScaledCountOnes, crossMethodsTight) ++
 		Experiment.create("linearly scaled - ", random, linearlyScaledCountOnes, crossMethodsTight) ++
-		//Experiment.create("block decpt", random, blockValuation(deciptive), crossMethodsTight) ++
-		//Experiment.create("block non", random, blockValuation(nonDeciptive), crossMethodsTight) ++
-		//Experiment.create("block decpt", random, blockValuation(deciptive), crossMethodsRandom) ++
-		//Experiment.create("block non", random, blockValuation(nonDeciptive), crossMethodsRandom) ++
+		Experiment.create("block decpt", random, blockValuation(deciptive), crossMethodsTight) ++
+		Experiment.create("block non", random, blockValuation(nonDeciptive), crossMethodsTight) ++
+		Experiment.create("block decpt", random, blockValuation(deciptive), crossMethodsRandom) ++
+		Experiment.create("block non", random, blockValuation(nonDeciptive), crossMethodsRandom) ++
 		Nil
 
 

### 2015-11-22 evolutionaryComputation/practicals/report.tex
fixed the GA with help of 2 unit tests

@@
 block as a lookup value for that.
+
+Trap function actually reversed the output. I suspected the trap function
+was doing it wrong because all the GA's kept failing, so I added a bunch
+of very specific test for each block result which confirmend my suspicion.
+The actuall bug was a simple reverse, which meant  the population, slowly
+leaned towards 0's instead of 1's. and since my 'solution' is a string
+of only ones, the end result is that the bisectional search kept increasing
+the population size since now anwser was found and the algoritm wasn't happy.
+
+All valuation functions should return a float instead of an Int.
+I don't even understan how I could make such a stupid mistake until the
+unit test pointed it out to me. (programming for 6+ years..)
+
 \subsection{Uncertainties}

### 2015-11-22 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Experiment.scala

@@ -26,32 +26,66 @@ class Experiment(val name:String, evolution: Evolution, memberFactory:String => 
 		RunResult(consideringSize, success, runtime, evolutionResult.length,evolution.valuation.countCalls())
 	}
 
+	def verifyLowest(currentPop:Int, faults:Int, index:Int) : Seq[RunResult] = {
+		def failed = Nil
+		if(index >= requiredRuns){
+			return Nil
+		}
+		if(faults > faultTolerance){
+			return failed
+		}
+		val exp = run(currentPop)
+		val newFaults = if (exp.success) faults else faults+1
+		Seq(exp) ++ verifyLowest(currentPop, newFaults,index+1)
+	}
 	private def findOptimumPopsize(consideringSize:Int, difference:Int): Seq[RunResult] = {
 		if(consideringSize > maxPopSize){
 			// we went to deep so we failed
 			return Nil
 		}
 
-		val result = run(consideringSize)
+		val result = verifyLowest(consideringSize,0,0)
 		// 10, 20, 40, 80, 160, 320, 240, 280, 260, 250.
 		val hadSuccess = consideringSize != difference*2
 
 		val newDiff = if(hadSuccess) difference/2 else difference*2
 
-		if(newDiff < popUnit){
-			return Seq(result)
+		if(newDiff < popUnit){ // steps becoming to small, bort
+			return result
 		}
-		if(!result.success){
-			return result +: findOptimumPopsize(consideringSize+newDiff,newDiff)
+		if(result.last.success){
+			// if we had success, zoom in
+			return result ++ findOptimumPopsize(consideringSize-difference/2, difference/2)
 		}
-		return result +: findOptimumPopsize(consideringSize-difference/2, difference/2)
+		result ++ findOptimumPopsize(consideringSize+newDiff,newDiff)
 	}
-	def bisectionalSearch() = findOptimumPopsize(popUnit ,popUnit/2)
+	def bisectionalSearch() = StoasticRun(findOptimumPopsize(popUnit ,popUnit/2), requiredRuns)
 }
 
 case class RunResult(popSize:Int, success:Boolean, runtime:Long, generationCount:Int, fitnessCallCount:Int){
 	override def toString() :String = {
 		s"Run(pop#=$popSize, success=$success, time=$runtime, gen#=$generationCount)"
+	}
+}
+import util.Properties.lineSeparator
[…]

### 2015-11-22 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Fitness.scala

@@ -1,12 +1,12 @@
 package nl.jappieklooster.hw.ec
 
 trait IHasFitness{
-	def fitness:Int
+	def fitness:Float
 }
 trait FitnessEvaluator{
 	private var calls = 0
-	protected def getFunction:String => Int
-	def valuate(x:String):Int ={
+	protected def getFunction:String => Float
+	def valuate(x:String):Float ={
 		calls += 1
 		getFunction(x)
 	}
@@ -19,9 +19,9 @@ trait FitnessEvaluator{
 
 object Fitness {
 	private def whereCharIsOne(c:Char) = c == '1'
-	def uniformlyScaledCountOnes(s:String) = s.count(whereCharIsOne)
+	def uniformlyScaledCountOnes(s:String):Float = s.count(whereCharIsOne)
 
-	def linearlyScaledCountOnes(s:String) = s.
+	def linearlyScaledCountOnes(s:String):Float = s.
 			// make a map like thing of it where the index is the key
 			zipWithIndex.
 			// select the elements where 1
@@ -31,14 +31,15 @@ object Fitness {
 			map(_._2+1).
 			// result
 			sum
-	def blockValuation(block:Seq[Float])(s:String) = s.
-			// make blocks of the string
-			grouped(block.size-1).
-			map(
-				// get the value specified in the block
-				str => block.apply(str.count(whereCharIsOne))
-			).sum.toInt
-	def createProbe(f:String=>Int):FitnessEvaluator = new FitnessEvaluator {
-		protected override def getFunction: (String) => Int = f
+	def blockValuation(block:Seq[Float])(s:String):Float = {
+		val size = block.size-1
+		// make blocks of the string
+		s.grouped(size).map(
+			// get the value specified in the block
+			str => block(size - uniformlyScaledCountOnes(str).toInt)
+		).sum
+	}
+	def createProbe(f:String=>Float):FitnessEvaluator = new FitnessEvaluator {
+		protected override def getFunction: (String) => Float = f
 	}
 }

### 2015-11-22 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -25,8 +25,8 @@ object Main{
 			(uniformCross _, createRandomlyLinked, " Random ux"),
 			(uniformCross _, withCoinfliptimesMutation(createRandomlyLinked, random) _, " Random ux, R")
 		)
-		val deciptive:Seq[Float] = Seq(4,0,1,2,3)
-		val nonDeciptive:Seq[Float] = Seq(4,0,0.5f,1,1.5f)
+		val deciptive:Seq[Float] = List(4,0,1,2,3)
+		val nonDeciptive:Seq[Float] = List(4,0,0.5f,1,1.5f)
 		val expirements =
 		Experiment.create("uni scaled - ", random, uniformlyScaledCountOnes, crossMethodsTight) ++
 		Experiment.create("linearly scaled - ", random, linearlyScaledCountOnes, crossMethodsTight) ++
@@ -37,15 +37,13 @@ object Main{
 		Nil
 
 
-		val runCount = 30
 		val results = expirements.par.map(x => {
-			val ex = new StoasticExperiment(x)
-			(ex, ex.stoasticEnsurance(runCount,1))
+			(x, x.bisectionalSearch())
 		}).seq
-		log.info(s"doing stoastic with $runCount runs")
+		log.info(s"doing stoastic with ${Experiment.requiredRuns} runs")
 		for(tuple <- results){
 			import util.Properties.lineSeparator
-			log.info(s"${tuple._1.experiment.name}:$lineSeparator${tuple._2.toTable()}")
+			log.info(s"${tuple._1.name}:$lineSeparator${tuple._2.toTable()}")
 		}
 
 		import org.sameersingh.scalaplot.Implicits._

### 2015-11-22 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Member.scala

@@ -5,19 +5,19 @@ import org.slf4j.LoggerFactory
 import scala.util.Random
 
 trait IMember extends IHasFitness with Genetic
-case class Member(valuation:Int, gen:String) extends IMember{
+case class Member(valuation:Float, gen:String) extends IMember{
 	override def fitness = valuation
 	override def genes = gen
 }
 
 object MemberFactories{
-	def tightlyLinked(func:String=>Int)(str:String):IMember = Member(func(str), str)
-	def randomlyLinked(instructions:TraversableOnce[Int])(func:String=>Int)(str:String):IMember = {
+	def tightlyLinked(func:String=>Float)(str:String):IMember = Member(func(str), str)
+	def randomlyLinked(instructions:TraversableOnce[Int])(func:String=>Float)(str:String):IMember = {
 		val genes = instructions.map(inx => str.charAt(inx)).mkString
 		Member(func(str), str)
 	}
 	val log = LoggerFactory.getLogger(this.getClass)
-	def withCoinfliptimesMutation(creationFunc:(String=>Int) => String => IMember, random:Random)(func:String=>Int)(str:String):IMember = {
+	def withCoinfliptimesMutation(creationFunc:(String=>Float) => String => IMember, random:Random)(func:String=>Float)(str:String):IMember = {
 		def countFlips(integer:Integer):Integer = {
 			if(random.nextBoolean()){
 				return countFlips(integer+1)

### 2015-11-22 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/StoasticExperiment.scala

@@ -1,68 +1,0 @@
-package nl.jappieklooster.hw.ec
-
-import Experiment._
-import com.itextpdf.text.log.LoggerFactory
-
-class StoasticExperiment(val experiment: Experiment) {
-
-	val log = LoggerFactory.getLogger(getClass)
-	private def geussLowest={
-		// run it four times
-		val peekResult = 0.to(3).flatMap(x=>experiment.bisectionalSearch)
-		peekResult.foldLeft(RunResult.createFailed(Int.MaxValue))(
-			(cur,elem)=> cur match{
-				case RunResult(_,false,_,_,_) => elem
-				case s if elem.success => if(s.popSize < elem.popSize) s else elem
-				case s => s
-			}
-		)
-	}
-	def stoasticEnsurance(required:Int, tolerance:Int):StoasticRun ={
-		val smallest = geussLowest
-		log.info(s"found $smallest")
-		if(!smallest.success){
-			return StoasticRun(Seq(smallest), required)
-		}
-		def verifyLowest(currentPop:Int, faults:Int, index:Int) : Seq[RunResult] = {
-			def failed = Nil
-			if(index >= required){
-				return Nil
-			}
-			// to deep gaurd
-			if(currentPop > maxPopSize){
-				return failed
-			}
-			if(faults > tolerance){
-
-				val newPopCount = currentPop+popUnit
-				return verifyLowest(newPopCount,0,0)
-			}
-			val exp = experiment.run(currentPop)
-			val newFaults = if (exp.success) faults else faults+1
-			Seq(exp) ++ verifyLowest(currentPop, newFaults,index+1)
-		}
-		StoasticRun(verifyLowest(smallest.popSize,0,0),required)
-	}
-}
-
-import util.Properties.lineSeparator
-case class StoasticRun(runs:Seq[RunResult], required:Int) {
-	lazy val byPopSize = runs.groupBy(x=>x.popSize)
-	lazy val bestRun = byPopSize.get(bigestPopcount).getOrElse(Nil)
-	lazy val isSuccesfull = bestRun.size == required
-	override def toString(): String = {
-		byPopSize.mkString(lineSeparator)
-	}
-	lazy val bigestPopcount = runs.sortWith((a,b)=>a.popSize > b.popSize).head.popSize
-	lazy val avergeGeneration = bestRun.map(x=>x.generationCount).sum/bestRun.length
[…]

### 2015-11-22 evolutionaryComputation/practicals/code/src/test/scala/nl/jappieklooster/hw/ec/tests/FitnessTests.scala

@@ -24,17 +24,36 @@ class FitnessTests extends FlatSpec with Matchers{
 		Fitness.linearlyScaledCountOnes(subject) should be (64)
 	}
 
-	val deciptive = List(4f,0f,1f,2f,3f)
+	val deceptive = Fitness.blockValuation(List(4f,0f,1f,2f,3f)) _
 	"deceptive trap fitness" should "count subject as 10" in {
-		Fitness.blockValuation(deciptive)(subject) should be (10)
+		deceptive(subject) should be (10)
 	}
-	val nondecept = List(4f, 0f, 0.5f, 1f, 1.5f)
+	"decipte trap fitness " should "count these infividual elements correctly " in{
+		/*
+		* I suepected something was wrong with the trap function because all trap
+		* GA's failed to find solution wich didn't make much sense (wouldn't learn
+		* much from a function that doesn't work).
+		* These tests proove that
+		 */
+		deceptive("1111") should be (4)
+		deceptive("0111") should be (0)
+		deceptive("0011") should be (1f)
+		deceptive("0001") should be (2f)
+		deceptive("0000") should be (3f)
+	}
+	"trap function" should "have no trouble with wrongly block sized input " in {
+		deceptive("1111" + "0111" + "1010" + "1000" + "00") should be (10)  // because it should count the ones
+	}
+	val nondecept = Fitness.blockValuation(List(4f, 0f, 0.5f, 1f, 1.5f)) _
 	"nondeceptive trap fitness" should "count subject as 7" in {
-		Fitness.blockValuation(nondecept)(subject) should be (7)
+		nondecept(subject) should be (7)
+	}
+	"nondecipte trap fitness " should "count these infividual elements correctly " in{
+		nondecept("1111") should be (4)
+		nondecept("1110") should be (0)
+		nondecept("1100") should be (0.5f)
+		nondecept("1000") should be (1f)
+		nondecept("0000") should be (1.5f)
 	}
 
-	"trap function" should "have no trouble with wrongly block sized input " in {
-		Fitness.blockValuation(deciptive)("1111" + "0111" + "1010" + "1000" + "00"
-		) should be (10)  // because it should count the ones
-	}
 }

### 2015-11-23 inteligentagents/notes5.tex
added notes of Intelligent Agents

\documentclass{article}
\begin{document}
\author{Jappie Klooster}
\title{KARO logic}
\maketitle

There are indeed unkown unkowns. But not if you have
$\neg K \phi \to K \neg K \phi$

\section{the names are just the axioms}
jjK = ??
D = $\neg B \bot$
4 = $B\phi \to BB \phi$
5 = $\neg B \phi \to B\neg B \phi$

\section{Dynamic logics}
these actions can be non deterministic\ldots

Accessibility relation

$\alpha;\beta$ = sequentu\alphal compisition
$\alpha+\beta$ = nondeterministic choice
$\alpha*$ = execute \alphar\betaitr\alphary num\betaer of times

s' = a different variable to s, so instead of s' you can use [a-z].

Dynamic logic is complete!

does $[\alpha] ( \phi \to \varphi) \to ([\alpha] \phi \to [\alpha] \varphi)$
mean I can subtitute  $[\alpha] ( \phi \to \varphi)$ with
$([\alpha] \phi \to [\alpha] \varphi)$

$[\alpha*]\phi\to\phi$ do nothing its true
$[\alpha*]\phi\to[\alpha][\alpha*]\phi$ do once and it becomes true
\subsection{with induction}
$[\alpha*](\phi \to [\alpha]\phi)\to(\phi\to[\alpha*]\phi)$ 
$\phi\to[\alpha]\phi$
$[\alpha](\phi\to[\alpha]\phi)$
$[\alpha, \alpha](\phi\to[\alpha]\phi)$
$\vdots$

\section{Karo}
Opertunity vs capability: I have the opertunity to climb the dom tower\ldots

\subsection{KBD}
Knowledge Believe Desires.

$R_K = S5$
$R_B = KD45$
$R_B \subset \R_K = \vDash K\phi \to B\phi$
$R_D =$ no constraints, so sytem K.

infinite computation = $\nill$???

$R_\alpha(\nill)=\nill$ {divergence failure, or abort}

A = Agent is able to
$A\alpha \LeftRightarrow C(M,w)$

$<\phi?>\psi \LeftRightarrow (\psi \wedge \phi)$, if there is an accassible world
were $\psi$ holds and $\phi$ can be tested then in the world $\phi$ and $\psi$ hold.

Holds in general:

\[[\alpha]\phi \wedge <\alpha>\top\to <\alpha>\phi\]

If $\alpha$ is deterministic.

\[ <\alpha>\phi \to [\alpha]\phi \wedge <\alpha>\top\]

\end{document}

### 2015-11-23 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala
added graphs section

@@ -43,10 +43,11 @@ object Main{
 		log.info(s"doing stoastic with ${Experiment.requiredRuns} runs")
 		for(tuple <- results){
 			import util.Properties.lineSeparator
-			val e = tuple._1
-			log.info(s"${e.name} (${e.variation}}):$lineSeparator${tuple._2.toTable()}")
+			val ex = tuple._1
+			val re = tuple._2
+			log.info(s"${ex.name} (${ex.variation}}):$lineSeparator${re.toTable()}")
 		}
-
+		
 	}
 	def createRandomlyLinked = randomlyLinked(random.shuffle(0.to(Experiment.geneLength))) _
 }

### 2015-11-24 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Experiment.scala
did some graph exporting

@@ -104,7 +104,7 @@ object Experiment{
 	val popUnit = 10
 	val maxPopSize = 1280
 	val geneLength = 100
-	val requiredRuns = 4
+	val requiredRuns = 30
 	val faultTolerance = 1
 
 	def create(

### 2015-11-24 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -1,5 +1,8 @@
 package nl.jappieklooster.hw.ec
 
+import java.io.PrintWriter
+
+import org.sameersingh.scalaplot.XYData
 import org.slf4j.LoggerFactory
 
 import scala.util.Random
@@ -30,10 +33,10 @@ object Main{
 		val expirements =
 		Experiment.create("uni scaled", random, uniformlyScaledCountOnes, crossMethodsTight) ++
 		Experiment.create("linearly scaled", random, linearlyScaledCountOnes, crossMethodsTight) ++
-		Experiment.create("block decpt", random, blockValuation(deciptive), crossMethodsTight) ++
-		Experiment.create("block non", random, blockValuation(nonDeciptive), crossMethodsTight) ++
-		Experiment.create("block decpt", random, blockValuation(deciptive), crossMethodsRandom) ++
-		Experiment.create("block non", random, blockValuation(nonDeciptive), crossMethodsRandom) ++
+		//Experiment.create("block decpt", random, blockValuation(deciptive), crossMethodsTight) ++
+		//Experiment.create("block non", random, blockValuation(nonDeciptive), crossMethodsTight) ++
+		//Experiment.create("block decpt", random, blockValuation(deciptive), crossMethodsRandom) ++
+		//Experiment.create("block non", random, blockValuation(nonDeciptive), crossMethodsRandom) ++
 		Nil
 
 
@@ -47,7 +50,25 @@ object Main{
 			val re = tuple._2
 			log.info(s"${ex.name} (${ex.variation}}):$lineSeparator${re.toTable()}")
 		}
-		
+		import org.sameersingh.scalaplot.Implicits._
+
+		val values = results.groupBy(x=>x._1.name).map(_._2.map(
+			// map population size/callcount to points
+			xy => xy._2.byPopSize.map(
+				pop=> (pop._1.toDouble,pop._2.length.toDouble)
+			// I think the data points weren't sorted, which is logical considiring it was a map
+			).toSeq.sortBy(_._1)
+		))
+
+		val file = java.io.File.createTempFile("evolutionaryComputing", ""+values.hashCode())
+		file.delete()
+		file.createNewFile()
+		new PrintWriter(file.getCanonicalFile){
+			for(exp <- values){
+				write(output(SVG, xyChart(new XYData(exp(0), exp(1), exp(2)))))
+			}
+			close()
+		}
 	}
 	def createRandomlyLinked = randomlyLinked(random.shuffle(0.to(Experiment.geneLength))) _
 }

### 2015-11-24 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Experiment.scala
created proper latex tables and now also ascii graphs

@@ -83,14 +83,6 @@ case class StoasticRun(runs:Seq[RunResult], required:Int) {
 	lazy val avergeGeneration = bestRun.map(x=>x.generationCount).sum/bestRun.length
 	lazy val averageFitnessCount = bestRun.map(x=>x.fitnessCallCount).sum/bestRun.length
 	lazy val bestRunAverageTime = bestRun.map(x=>x.runtime).sum/bestRun.length
-
-	def toTable():String ={
-		s"success:			$isSuccesfull" + lineSeparator +
-		s"minimumPop: 		$bigestPopcount	" +lineSeparator+
-		s"avgGeneration:		$avergeGeneration" +lineSeparator+
-		s"avgFitness:			$averageFitnessCount" +lineSeparator+
-		s"avgTime:			$bestRunAverageTime " +lineSeparator
-	}
 }
 
 object RunResult{
@@ -98,14 +90,33 @@ object RunResult{
 }
 object Experiment{
 
+	def stoasticsToLatex(runs:Seq[(String, StoasticRun)]) = {
+		// did some html in my youth
+		val td = "&"
+		runs.foldLeft(
+			s"$lineSeparator" + // the structure becomes clearer on a new line
+			s"\\begin{tabular}{llllll}$lineSeparator" +
+			s"Variation $td Success $td Minimum population $td Average generation "+
+			s"$td Average fitness call-count $td Average time \\\\ \\toprule $lineSeparator"
+		)((a,tupple) => {
+			val r = tupple._2
+			s"$a" +
+			s"${tupple._1} $td" +
+			s"${r.isSuccesfull} $td ${r.bigestPopcount} $td" +
+			s"${r.avergeGeneration} $td ${r.averageFitnessCount} $td" +
+			s"${r.bestRunAverageTime} \\\\$lineSeparator"
+		}
+		) +
+		s"\\end{tabular}$lineSeparator"
+	}
 	/**
 	 * Don't consider pop diferences smaller than this unit
 	 */
 	val popUnit = 10
 	val maxPopSize = 1280
 	val geneLength = 100
-	val requiredRuns = 30
-	val faultTolerance = 1
+	val requiredRuns = 5
+	val faultTolerance = 0
 
 	def create(
 		name:String,

### 2015-11-24 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -1,14 +1,16 @@
 package nl.jappieklooster.hw.ec
 
-import java.io.PrintWriter
+import java.io.{File, PrintWriter}
 
 import org.sameersingh.scalaplot.XYData
 import org.slf4j.LoggerFactory
 
+import scala.reflect.io.Path
 import scala.util.Random
 import OffspringGenerator._
 import MemberFactories._
 import Fitness._
+import util.Properties.lineSeparator
 
 object Main{
 	case class Configurations(strLength:Int,popSize:Int,genCount:Int, evolution:Evolution)
@@ -19,24 +21,24 @@ object Main{
 
 		
 		val crossMethodsTight = Seq(
-			(twoPointCross _, tightlyLinked _, " Tight 2x"),
-			(uniformCross _, tightlyLinked _, " Tight ux"),
-			(uniformCross _, withCoinfliptimesMutation(tightlyLinked, random) _, " Tight ux, R")
+			(twoPointCross _, tightlyLinked _, "2X"),
+			(uniformCross _, tightlyLinked _, "UX"),
+			(uniformCross _, withCoinfliptimesMutation(tightlyLinked, random) _, " UX, Mutation")
 		)
 		val crossMethodsRandom = Seq(
-			(twoPointCross _, createRandomlyLinked, " Random 2x, "),
-			(uniformCross _, createRandomlyLinked, " Random ux"),
-			(uniformCross _, withCoinfliptimesMutation(createRandomlyLinked, random) _, " Random ux, R")
+			(twoPointCross _, createRandomlyLinked, "2X"),
+			(uniformCross _, createRandomlyLinked, "UX"),
+			(uniformCross _, withCoinfliptimesMutation(createRandomlyLinked, random) _, " UX, Mutation")
 		)
 		val deciptive:Seq[Float] = List(4,0,1,2,3)
 		val nonDeciptive:Seq[Float] = List(4,0,0.5f,1,1.5f)
 		val expirements =
-		Experiment.create("uni scaled", random, uniformlyScaledCountOnes, crossMethodsTight) ++
-		Experiment.create("linearly scaled", random, linearlyScaledCountOnes, crossMethodsTight) ++
-		//Experiment.create("block decpt", random, blockValuation(deciptive), crossMethodsTight) ++
-		//Experiment.create("block non", random, blockValuation(nonDeciptive), crossMethodsTight) ++
-		//Experiment.create("block decpt", random, blockValuation(deciptive), crossMethodsRandom) ++
[…]

### 2015-11-24 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Experiment.scala
svg graphics now automaticly created by the code

@@ -115,7 +115,7 @@ object Experiment{
 	val popUnit = 10
 	val maxPopSize = 1280
 	val geneLength = 100
-	val requiredRuns = 5
+	val requiredRuns = 4
 	val faultTolerance = 0
 
 	def create(

### 2015-11-24 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -16,6 +16,10 @@ object Main{
 	case class Configurations(strLength:Int,popSize:Int,genCount:Int, evolution:Evolution)
 	val log = LoggerFactory.getLogger(Main.getClass)
 	val random = Random
+	val br = lineSeparator
+	val folder = "../"
+	type GraphData = (String,Seq[Seq[(Double,Double)]])
+
 	def main(args:Array[String]){
 		log.info("starting ea")
 
@@ -47,36 +51,11 @@ object Main{
 		}).seq
 		log.info(s"doing stoastic with ${Experiment.requiredRuns} runs")
 
-		import org.sameersingh.scalaplot.Implicits._
-
 		val values = toPlotableStructure(results)
 
-		val br = lineSeparator
-		val latex = values.foldLeft(s"$br\\subsection{Data}")((str,duo) => {
-			var result = s"$str $br"
-			result += s"\\subsubsection{${duo._1}} $br"
-			result += s"\\begin{figure}[ht!] $br"
-			result += s"\\begin{verbatim} $br"
+		log.info(toLatexWith(asciiGraph)(values))
 
-			val xy = duo._2
-			val weirdAddedChar = 12.toChar.toString
-			result += output(ASCII, xyChart(new XYData(xy(0), xy(1), xy(2)))).replaceAll(weirdAddedChar, "")
-			result += s"\\end{verbatim} $br"
-			result += s"\\caption{${duo._1}}"
-			result += s"\\end{figure}\\clearpage $br"
-
-			result += Experiment.stoasticsToLatex(duo._3)
-			result
-		})
-		log.info(latex)
-
-		val outputFile = new File("result.tex")
-		outputFile.delete()
-		outputFile.createNewFile()
-		new PrintWriter(outputFile){
-			write(latex)
-			close()
-		}
+		write("result.tex", toLatexWith(svgGraph)(values))
 
 		val file = java.io.File.createTempFile("evolutionaryComputing", "dir")
 		file.delete()
@@ -88,6 +67,15 @@ object Main{
 			close()
 		}
 		*/
+	}
+	def write(filename:String, content:String) = {
+		val outputFile = new File(folder+filename)
+		outputFile.delete()
+		outputFile.createNewFile()
+		new PrintWriter(outputFile){
+			write(content)
+			close()
+		}
 	}
 
 	def toPlotableStructure(results:Seq[(Experiment, StoasticRun)]) =
[…]

### 2015-11-25 inteligentagents/notes.tex
put all notes into the same file

@@
 ##Agent architectures
+#Bratman
+
+Commitment, if you have a real intention, you want to go for it.
+intentions should be consistent, otherwise you do something which could be
+impossible
+But this is debatable...
+
+## Intentions
+Screen of admissibility, filter maybe later intentions that could be
+conflicitng
+
+# Cohen& levesque
+tried to formalise Bratman.
+
+## Intention = choice + commitment
+try to stretch it to formal logic.
+
+##
+agent beleives P is possible : a sort of self conficience of the agent
+
+## side effects
+agnets need not intend all the expected side effects of their intentions.
+Going to the dentist implies pain.
+then going to the dentist should be that you intend to have pain.
+
+$\vDash \phi \to \psi \to \vDash \Box \phi \to \Box \psi$
+
+#Rao & Georgeff BDI theory
+Not only proposed a logic, but also the first implementation of an agent and
+also the first application.
+
+#Rodney Brooks anti-cogn.robotics
+subsumption architecture: no cognition, just a reactive system.
+Should be emergent behaviour rather than modeling.
+
+# Aaron Sloman: How to build emotional agents
+Looks at properties of humans, such as emotions, and tries to put it into an
+AI.
+
+20 years ago, there are patiens that don't have emotions, expect: in problem
+solving such as puzzels. They go depth first, so they go on, and on and fail
+misserably.
+
+Archicterus of him are not the most enlighted, "messy".
+
+# Towards social agents/robots
+also research goes to sociology.
+
+Oppertunisitic behavior, sellar buying a cup, but cup is broken,
+the seller wants to sell it, but the buyer wants to use it. How
+to moddel this in Agents
+
+## Normative systems
+the way of societies. you shouldn't violate the norm, but sometimes it can
+be usefull to.
+
+### Asimov
+Laws of him. 4th one is stupid
+
+# Theory of mind
+The psycology one.
+The false believe test.
+
+# Logial foundations of actions
+
+The terminology is not stabels.
+desires and goals may be differetn or mean the same.
+
+Agenda -> latin things that should happen.
+
+# Modal logic
+Logic of $\Box$ and $\Diamond$
+
+## K axiom
+true for any modal logic
+
+## R
+means its accesible. No restriction on R in arbitrary Kripke
+
+### Constraints
+you can add constraints on R, but with restrictions things certain
+things become true.
+
+serial you can go on.
+
+## Frames
+Models where the true assignment functions are abstracted from.
+
+#naive BDI logic
+seperate blocks, it misses connections.
+No dymancis, changes of BDI in time.
+
+# Cohen & levesque : langauge
+$\alpha$ are complex actions and a's are primitive actions
+We can't say multiple agents do the same action at the same time.
+
+a is interpeted as a sequance of primitive events even though a is
+atomic.
+
+atomic action is a special class, which depends on the domain.
+
+; = sequence
+? = test if its true (groovy truth)
++ = is a non detrmenistic choice, collect them in an autcome set.
+* = is done a finite but arbitrary number of times, including 0
+
+if $\phi$ then $\alpha_1$ else
+$\alpha_2$ = ($\phi? ; \alpha_1) + (\neg \phi?;\alpha)$
+
+while $\phi$ do $\alpha$ = $(\phi?;\alpha)^*;\neg\phi?$
+
+Pz = z has property P, a predicate attom
+
+# Semantics
+\omega = variables assginemnts
+P = is the set of omega \in \omega
+E = is a set of primitive events
+Agt = Agt \in [E \to p] specifies the agent of an event [E \to P] is a
+function that takes an E that gives a P.
+T \subset [Z \to E]; a set of possible worlds, give a time number and get
+
+\documentclass{article}
+\begin{document}
+\author{Jappie Klooster}
+\title{Cohen \& Levesque}
+\maketitle
+\section{Langauge}
+DONE = just done
+; = sequence
+? = test if its true (groovy truth)
++ = is a non detrmenistic choice, collect them in an autcome set.
+* = is done a finite but arbitrary number of times, including 0
+
+if $\phi$ then $\alpha_1$ else
+$\alpha_2$ = ($\phi? ; \alpha_1) + (\neg \phi?;\alpha)$
+
+while $\phi$ do $\alpha$ = $(\phi?;\alpha)^*;\neg\phi?$
+
+Pz = z has property P, a predicate attom
+
+\subsection{Semantics}
+$\omega$ = variables assginemnts
+$P$ = is the set of omega $\in \omega$
+$E$ = is a set of primitive events
+$Agt = Agt \in [E \to p]$ specifies the agent of an event $[E \to P]$ is a
+function that takes an E that gives a P.
+$T \subset [Z \to E];$ a set of possible worlds, give a time number and get
+
+i = agent
+n = time
+
+\subsubsection{Predicate}
+c constants
+f functions
+Q predicate smbols
+
+$\phi$ interperts predicate symbols
+
+\[\phi(c) = d \in \Theta\]
+\[\phi(f) = f D_1x\dots D_l \to D_h\]
+\[\phi(Q^{n,x,ne}) = R \subseteq D_1x\dots D_l \]
+
+\subsubsection{satisfaction}
+
+\[M,\sigma,v,n[\alpha] m\]
+In the model `$M,\sigma,v,n$', $\alpha$ lies betweeen n and m.
+
+B = `normal` modal accesibility relation indexed by the time $n$
+and the agent $v(i)$. they write it like:
+
+\[<\sigma, n>B[v(i)] \sigma\]
+
+Better would be:
+\[\sigma B_{v(i)}^n \sigma\]
+
+\subsection{Constraints}
+\subsubsection{Believes/Goals}
+KD45.
+
+\[G\subseteq B\]
+means $\vdash {Bel}_i \phi \to {Goal}_i\phi$, but not the otherway around
+
+\subsection{Achievement goal}
+Doesn't believe $\phi$ is true yet, but believes it could be true later.
+There exists some action to make $\phi$ true.
+
+\subsubsection{Asumption}
+Goals will end some time in the future
+
+\subsection{Persistent goals}
+Drop the goal if its true, or its impossible.
+
+Competence means that if he believes it then its true.
+
+\section{Rao Georgeff}
+\subsection{Time logics}
+Brancing vs linear, points vs intervals,
+discrete (real numbers) vs continues (complex number)
+We'll only look into the future.
+\end{document}
+
+\documentclass{article}
+\begin{document}
+\author{Jappie Klooster}
+\title{Temporal magick or logic}
+\maketitle
+\section{Linear temporal magick}
+
+$\phi_1\cup\phi_2$ = $\phi_1$ holds until $\phi_2$ becomes true.
+$\Diamond\phi$ = sometimes
+$O\phi$ = next time
+
+$M=(S,\sigma, \pi)$ where $S$ is a set of states, $\sigma: N \to S$ is an
+infinte seqence of states
+
+\subsection{Semantics}
+$\vdots$
+$M,\sigma\vDash O\phi \LeftRightArrow M, \sigma^1 \vDash \phi$ next time
+$M,\sigma\vDash \Diamond\phi \LeftRightArrow M, \sigma^n \vDash \phi$
+invention, there is some time where $\phi$ is gonna hold.
+
+\subsubsection{untill}
+if $\phi_2$ holds somewhere it does not mean it holds everywhere after that
+point
+
+$\phi_1$ holds until $\phi_2$ holds.
+
+weak untill = where $\phi_1$ holds everywhere because there is no
+$\phi_2$, for example $\phi_1^{\bot}$
+
+\section{Branching temproal magick}
+
+Two kinds of formulas, state formulas, path formulas.
+Any state formula can also act as a path formula.
+
+\subsection{Models}
+serial, any point has a continuation, time never stops.
+$S \times S$ = cartesian product of S.
+
+Fullpath is a ninfinite seq of states more or less true the tree.
+These are more or less the lines of linear logic.
+Fullpaths can start anywhere, not necisarly at the root.
+
+$\sigma_1$ = already a choice of the branch, so no need for no trees.
+
+\section{Combination of BDI}
+E = optional
+A = inevitable
+
+\section{Operators magick}
+$\Diamond\Diamond\phi=\Diamond\phi$
+$\Box\Box\phi=\Box\phi$
+$\Box\Diamond\phi$=Infenitly often there is a point where $\phi$ holds.
+$\Diamond\Box\phi=$Some point in the future $\phi$ will hold indevinitly
+
+\subsection{BDI}
+Optionally $\Diamond p$ = there is a choice where p is gonna hold
+Optionally $\Box r$= there is a timline where r is gonna hold
+Inevitiably $\Diamond s$=  ???
+Inevitiably $\Box b$=  b always holds
+
+\section{BDI sementics}
+s.t. = such that.
+
+From t1 to t2, the agent has a memory, but he can doubt about the future.
+There is the hidden asumption that agents have memory.
+
+success = there exist some period in wehre the even was succesfull
+
+\subsection{Axiom}
+written like $GOAL(\alpha) \subset BEL(\alpha)$, but russel notation :s so
+$\subset = \to$
+$GOAL(\alpha) \to BEL(\alpha)$ = believe-goal compatibility
+different from cohon and laveque, $BEL \alpha \to GOAL \alpha$, because
+its a different framework, but the reason behind it is the same strangely
+enough.
+
+$INTEND(does(e))\to does(e)$ intention leads to action, which is controversial
+
+\subsection{compatibility}
+Sinning = ???, something with cutting of branches.
+
+\section{CTD}
+\subsection{Axioms}
+$done(e) \to BEL(done(e))$,  you're awere that you've done something
+
+\section{Praktical}
+The general strategy is to write down what you have to proof and then to
+read with help of the definitions.
+
+Don't be sloppy and take an arbitrary M with for all's
+\subsction{3.16}
+\[\vDash \alpha \Rightarrow \vDash BEL x \alpha\]
+Assume $\Vdash \alpha$, so $M,\sigma,v,n\vDash \alpha$ for all $M,\sigma,v,n$(*)
+To prove $\vDash BEL x \alpha$, 
+	ie $M,\sigma,v,n\vDash BEL x \alpha$ for all $M,\sigma,v,n\vDash$
+	ie $M,\sigma,v,n\vDash\alpha$ for all $\sigma'$ with (\sigma,n)B[v(x)]
+	\sigma' done by (*).
+
+\subsection{3.26}
+	To prove: $M,\sigma,v,n\vDash BEL x \alpha \to Goal x \alpha$ for all 
+	$M,\sigma,v,n$.
+	ie $M,\sigma,v,n\vDash BEL x \alpha \Rightarrow M,\sigma,v,n\vDash$ Goal
+		$x \alpha$ for all $M,\sigma,v,n$
+		ie $M,\sigma',v,n\vDash \alpha$ for all $\sigma'$ with $\sigma,n)
+		B[v(x)]\sigam'\RIghtarrow M,\sigma'',v,n\vDash \alpha$ for all
+		$\sigma''$ with $(\sigma,n)G[v(x)]\sigma''$
+
+		Given:
+		$(\sigma,n)G[v(x)]\sigma''\Rightarrow(\sigma,n)B[v(x)]\sigma''$
+		$\vdots$
+
+		Take arb $\sigma''$ with $(\sigma,n)G[v(s)]\sigam''$ then also 
+		$(\sigma,n)B[v(x)]\sigma''$ so $M,\sigma,v,n\vDash \alpha$
+\subsection(3.11)
+Proof
+\begin{prooftree}
+	\AxiomC{}
+	\UnaryinfC{}
+\[ \vDash \neg LATER(\Diamond p) \neg LATER(\Diamond p) \LeftRightarrow 
+\neg(\neg\Diamond p) \Leftrightarrow \neg(\neg\Diamond p \wedge 
+\Diamond\Diamond p) \Leftrightarrow \neg(\neg\Diamond p \wedge 
+\Diamond p)\]
+
+\end{prooftree}
+
+\subsection{3.13}
+\begin{prooftree}
+	\AxiomC{$\vDash \Diamond q \wedge BEFORE pq \to \Diamond p$}
+	\UnaryInfC{To prove $M,\sigma,v,n\vDash \Diamond q \wedge BEFORE pq \to \Diamond p$ for all $M,\sigma,v,n$}
+	\UnaryInfC{Take arbitrary $M,\sigma,v,n$ to prove $M,\sigma,v,n\vDash \vDash \Diamond q \wedge BEFORE pq \to \Diamond p$ for all $M,\sigma,v,n$}
+	\Rightlabel{ie}
+	\UnaryInfC{To prove if $M,\sigma,v,n\vDash \Diamond q$ and 
+	$M,\sigma,v,n\vDash BEFORE pq$ then $M,\sigma,v,n\vDash Diamond p$ for all $M,\sigma,v,n$}
+	\UnaryInfC{Supose $M,\sigma,v,n\vDash \vDash \Diamond q$ and 
+	$M,\sigma,v,n\vDash BEFORE pq$}
+	\Rightlabel{ie}
+	\UnaryInfC{Supose $M,\sigma,v,n\vDash \exists x HAPPENS x;q?$ and 
+		$M,\sigma,v,n\vDash \exists \forall c (HAPPENS c;q{?}) \to \exists q$ 
+	$q\se c \wedge HAPPENS a;p{?}$}
+	\Rightlabel{ie}
+	\UnaryInfC{Supose $M,\sigma,v,n\vDash \exists x HAPPENS x;q?$ and 
+		$M,\sigma,v,n\vDash \exists \forall c (HAPPENS c;q{?}) \to \exists q$ 
+	$q\se c \wedge HAPPENS a;p{?}$}
+\end{prooftree}
+
+\end{document}
+\documentclass{article}
+\begin{document}
+\author{Jappie Klooster}
+\title{KARO logic}
+\maketitle
+
+There are indeed unkown unkowns. But not if you have
+$\neg K \phi \to K \neg K \phi$
+
+\section{the names are just the axioms}
+jjK = ??
+D = $\neg B \bot$
+4 = $B\phi \to BB \phi$
+5 = $\neg B \phi \to B\neg B \phi$
+
+\section{Dynamic logics}
+these actions can be non deterministic\ldots
+
+Accessibility relation
+
+$\alpha;\beta$ = sequentu\alphal compisition
+$\alpha+\beta$ = nondeterministic choice
+$\alpha*$ = execute \alphar\betaitr\alphary num\betaer of times
+
+s' = a different variable to s, so instead of s' you can use [a-z].
+
+Dynamic logic is complete!
+
+does $[\alpha] ( \phi \to \varphi) \to ([\alpha] \phi \to [\alpha] \varphi)$
+mean I can subtitute  $[\alpha] ( \phi \to \varphi)$ with
+$([\alpha] \phi \to [\alpha] \varphi)$
+
+$[\alpha*]\phi\to\phi$ do nothing its true
+$[\alpha*]\phi\to[\alpha][\alpha*]\phi$ do once and it becomes true
+\subsection{with induction}
+$[\alpha*](\phi \to [\alpha]\phi)\to(\phi\to[\alpha*]\phi)$ 
+$\phi\to[\alpha]\phi$
+$[\alpha](\phi\to[\alpha]\phi)$
+$[\alpha, \alpha](\phi\to[\alpha]\phi)$
+$\vdots$
+
+\section{Karo}
+Opertunity vs capability: I have the opertunity to climb the dom tower\ldots
+
+\subsection{KBD}
+Knowledge Believe Desires.
+
+$R_K = S5$
+$R_B = KD45$
+$R_B \subset \R_K = \vDash K\phi \to B\phi$
+$R_D =$ no constraints, so sytem K.
+
+infinite computation = $\nill$???
+
+$R_\alpha(\nill)=\nill$ {divergence failure, or abort}
+
+A = Agent is able to
+$A\alpha \LeftRightarrow C(M,w)$
+
+$<\phi?>\psi \LeftRightarrow (\psi \wedge \phi)$, if there is an accassible world
+were $\psi$ holds and $\phi$ can be tested then in the world $\phi$ and $\psi$ hold.
+
+Holds in general:
+
+\[[\alpha]\phi \wedge <\alpha>\top\to <\alpha>\phi\]
+
+If $\alpha$ is deterministic.
+
+\[ <\alpha>\phi \to [\alpha]\phi \wedge <\alpha>\top\]
+
+\section{KARO}
+\subsection{Default}
+
+\subsection{Goal}
+implementable = a posibility to reach $\phi$
+
+A goal is a achievement goal, something which is not true
+has to become true.
+
+Maintanence goals want to keep something true.
+
+PossIntend = a very important one. An agent can
+transform a possible intent into an commitment.
+
+\subsection{Update agenda}
+\[Ag(M',w') = Ag(M,w)\cup\alpha\]
+for all $m',w'$ with $R(w',w'')$ 
+\[Ag(M',w'') = Ag(M',w')\cup\alpha\]
+\vdots
+
+The agent is awere off what is going on. It knows
+about the commitments.
+
+you can only commit possible intents
+
+You can only uncommit commited actions.
+\end{document}

### 2015-11-26 evolutionaryComputation/practicals/report.tex
added graphs section
did some graph exporting
created proper latex tables and now also ascii graphs
svg graphics now automaticly created by the code
fixed sentence
added extra text, removed a bunch of things I'm no longer uncertain about and added a begin to the uml diagram which I can make in latex aperantly.

@@
 \usepackage{qtree}
-\usepackage{tikz}
+\usepackage{tikz-uml}
 \usepackage{listings}
 \usepackage{hyperref}
+\usepackage{svg}
 \newenvironment{scprooftree}[1]%
@@
 \section{Methodology}
-\subsection{Build tool}
-I use Gradle for building the project. Its the most advanced build platform
-available and makes it easy to add dependencies to the project.
-For example I use slf4j for logging.
-
-Its also handy in that it comes with a self installing script that can be run
-from any platform, Gradlew. So running the code I wrote on a different machine
-becomes a trivial task, Gradle makes sure the right dependencies are fetched.
-
-\subsubsection{Run guide}
-As said before its a trivial task to run the code. Downloading and installing
-of dependencies and even gradle happens automatically with help of gradlew.
-So the first run is slower, depending on your internet speed.
-
-\subsubsection{Libraries used}
-
 \subsection{Language choice}
@@
 \end{itemize}
+
 \subsection{Architecture}
@@
 crossover.
+
+\begin{tikzpicture}
+\begin{umlpackage}[x=0,y=0]{ec}
+	\umlemptyclass[x=3]{Member}
+\umlemptyclass{Population}
+\umlemptyclass[x=6]{PairedPopulation}
+%\umluniassoc[geometry=-|,arg1=x,mult1=1,pos1=1.9,arg2=y,mult2=*,pos2=0.2]{Population}{Member}
+\end{umlpackage}
+\end{tikzpicture}
+
 I just implemented one or 2 options of these and then created an evolution
@@
 create an evolution by giving it functions.
+
+Most of the GA architecture was already in place before the assingment became
+available. I did this based on the notes I took in class. When the assignment
+became available I was almost completly done with the GA. There were some bugs
+in my code however and I interperted some things wrong. I also didn't write
+anything for an experiment. So I still needed to do that.
 
@@
 
+\subsubsection{Build tool}
+I use Gradle for building the project. Its the most advanced build platform
+available and makes it easy to add dependencies to the project.
+For example I use slf4j for logging.
+
+Its also handy in that it comes with a self installing script that can be run
+from any platform, Gradlew. So running the code I wrote on a different machine
+becomes easy, Gradle makes sure the right dependencies are fetched.
+
+\paragraph{Running it}
+
+Open a shell and move to the project folder then type:
+
+\lstset{language=Bash}
+\begin{lstlisting}[frame=single]
+	./gradlew run
+\end{lstlisting}
+
+\subsubsection{Libraries used}
+In here I will describe why I chose to use these libraries
+The details of the libraries can be found in the gradle file (which
+version and what maven group):
+
+\begin{lstlisting}[frame=single]
+	cat code/build.gradle
+\end{lstlisting}
+
+\begin{itemize}
+	\item The scala standart library, because I have to when using scala.
+	\item slf4j, a simple logging facade, this allows runtime switching of
+		logging schemes, for example if you want to log into a database or
+		a file instead of stdout you can do this relativly easy with slf4j
+		This is not necisary right now but the api is simple enough to use
+		everywhere.
+	\item scalaplot. A graph plotting library for scala. I didn't want to
+		do this myself, and this library supports svg plotting and ascii
+		plotting.
+	\item scalatest. A testing library for scala, it has some pretty advanced
+		features but I use it mostly like JUnit.
+\end{itemize}
 \subsection{Uncertainties}
-Should the elitism selection use tournement selection?
-When precisly do we find the optimal solution?
-	Should I wait for convergence or just say a string
-	of all ones is good enough?
-	Is 1 solution good enough or should all members have
-	that solution
-	I'm assuming 1 solution and that I can give a string
-	of just ones.
-How many generations to consider. I chose arbitrarly 10.
-
 About randomly linked:
@@
 	have the same randomly linked value, I'm aware of this.
-
-I'm not quite sure about the graphs. I have to show the number of successes
-out of the 30 runs, but this is already set in stone, because its either
-29 or 30, or its no successes.
 
@@
 The overhead for delegating and creating threads was more than the parilization
-gained.
+gained.  After 10 minutes of waiting I decided to drop this idea 
+since without parilization it took about 2 minutes (I tried this
+on just one run, so it shouldn't take that long).
 
@@
 
+So now all expiriments are run at the same time. This probably increases the
+average time of each 'run' somewhat (not that much because the concurent
+collection does \href{http://docs.scala-lang.org/overviews/parallel-collections/configuration.html}{load balancing})
+It alo drasticly decreases the amount of time I have to wait. Which is quite
+important, especially in the development phase. (although I also used
+other tricks, such as instead of 30, just doing 3 runs)
+
+The paralization of experiments can be removed by deleting `par' on line 64
+of Main.scala
+
 \subsubsection{Evaluate each member on creation}
-Members are going to be evaluated always, initially I would evaluate the member
-everytime it needed to be evaluated, but in later version I just aded the
-evaluation to the case class. Its not going to change anyways.
+Initially my members just contained the string with the `genes'. This I passed
+around and would eveulate when needed. However I realized that I could do this
+evaluation eagerly on creation, to reduce the evaluation calls. A member has
+to be eveluated anyways, so doing this on creation and just putting in the
+evaluation saves some heavy evaluation operations.
 
@@
 String class in Java.
+
+I was very proud for coming up with  this idea, but when I ran the expirment with
+and without hascode I came to the conclusion that first doing a hascode operation
+significantly *slows* execution. This becomes worse as population size increases.
+
 \section{Execution}
-%TODO:
-% 1. Experiments
-% Use stoastic within the algoritm
-%	Draw a graph from the experiments
-%		* pop size on x, # succes on Y
-% maybe some more testing of the crossovers?
-% also try to implement a threading scheme??
+So I can execute in parralel, I can easily add whatever function I want, so
+I decided to just run a whole bunch of experiments and see what comes out.
+
+\input{result.tex}
 

### 2015-11-28 evolutionaryComputation/practicals/report.tex
extended the report, now I want to make the uml, tried using tikz-uml
But it didn't work out very well, its to detailed, so now I'll switch to
emp, which seems easier in use, this is a safety commit

@@
 crossover.
-
-\begin{tikzpicture}
-\begin{umlpackage}[x=0,y=0]{ec}
-	\umlemptyclass[x=3]{Member}
-\umlemptyclass{Population}
-\umlemptyclass[x=6]{PairedPopulation}
-%\umluniassoc[geometry=-|,arg1=x,mult1=1,pos1=1.9,arg2=y,mult2=*,pos2=0.2]{Population}{Member}
-\end{umlpackage}
-\end{tikzpicture}
 
@@
 \subsection{Speed up attempts}
+I wanted my implementation to be fast, preferably faster than any other
+implementation, now I think I didn't succeed in this because I used a
+lot of slow mechanisms, such as functional combinators, and although
+they don't increase the computational complexity, they do cerainly increase
+the minimum required runtime, I'm quite sure that every combinator in sequance
+is implemented as just another loop. However because I
+run all expirments in paralel the total time may be faster.
 \subsubsection{Paralization}
@@
 
+Something to note is that my made up tournement elitism is actually in most
+cases better, or at least equal to the truncate elitism. So when I saw
+this I tried using my other initially used algorithm where I would completly
+ignore the parents (kill the parents). However it appears this does not 
+perform better.
+
+I also added two point cross with mutation and 1 point cross to see how 
+those fair, running the code now takes about 40 minutes in total with the
+entire thing diveded over 8 threads (quad core + hyperthreading). The only
+other improvement I could think of is running it on the GPU, but I think
+the algortim is to complex for that.
+
+Two point cross and single cross both seem to find results in deciptive 
+random tournement elitism.
+
+\newpage
 \input{result.tex}

### 2015-11-28 evolutionaryComputation/practicals/report.tex
forgot to add the original uml
Added a bunch of uml

@@
 \usepackage{bussproofs}
-\usepackage{graphicx}
+\usepackage[pdftex]{graphicx}
+\DeclareGraphicsRule{*}{mps}{*}{}
 \usepackage{pifont}
 \usepackage{qtree}
-\usepackage{tikz-uml}
+\usepackage{emp}
 \usepackage{listings}
@@
 \begin{document}
+\begin{empfile}
+\begin{empcmds}
+input metauml;
+\end{empcmds}
 \author{Jappie Klooster}
@@
 \subsection{Architecture}
+\begin{figure}[ht!]
+\centering
+\begin{emp}[classdiag](20,20)
+Class.mem("Member")("+Fitness:Int", "+genes:String")();
+Class.pop("Population")("+members:Seq[Member]", "+factory:String=>Member")();
+Class.pairedPop("PairedPopulation")("+members:Seq[Pair]","+factory:String=>Member")(
+"+pairedToPop(how):Population");
+Class.pai("Pair")("+father:Member","+mother:Member")();
+
+leftToRight.top(20)(pop,pairedPop);
+topToBottom.left(60)(pop,mem);
+leftToRight.top(80)(mem, pai);
+
+drawObjects(mem,pop,pairedPop,pai);
+link(association)(pai.n -- pairedPop.s);
+item(iAssoc)("*")(obj.se = pai.n);
+link(association)(mem.e -- pai.w);
+item(iAssoc)("2")(obj.sw = mem.e);
+
+link(association)(mem.n -- pop.s);
+item(iAssoc)("*")(obj.se = mem.n);
+\end{emp}
+\caption{Case classes}
+\end{figure}
+
+\begin{figure}[ht!]
+\centering
+\begin{emp}[classdiag](20,20)
+Class.evo("Evolution")(
+"+valuation: FitnessEvaluator",
+"mateSelector:IMateSelector",
+"offSpringGenerator: PairedPopulation => Population",
+"fittestFilter: (Population, Population) => Population",
+"hasGoodEnoughSolution: Population => Boolean")(
+"startGenetic(seed:Population): Seq[Population]",
+"genetic(prev:Seq[Population]): Seq[Population]",
+"step(parents:Population): Population",
+);
+Instance.fit("Fitness")(
+	"uniformlyScaled()",
+	"linearlyScaled()",
+	"blockValuation(block:Seq[Float])"
+);
+Instance.filt("FittestFilter")(
+	"killParents()",
+	"truncateElitism()",
+	"tournementElitism()"
+	);
+Instance.select("MateSelection")(
+	"tournementSelect()"
+);
+Instance.factor("MemberFactories")(
+	"tightlyLinkded()",
+	"randomlyLinked()",
+	"withCoinflipMutation(function):function",
+);
+Instance.off("OffspringGenerator")(
+	"uniformCross()",
+	"singlePointCross()",
+	"twoPointCross()"
+);
+
+topToBottom.left(20)(evo,fit,filt,factor);
+leftToRight.top(20)(fit,off);
+leftToRight.top(20)(filt,select);
+
+drawObjects(evo,fit,factor,filt,select,off);
+\end{emp}
+\caption{Algoritm}
+\end{figure}
+
+\begin{figure}[ht!]
+\centering
+\begin{emp}[classdiag](20,20)
+Class.ex("Experiment")(
+	"+name:String",
+	"+variation:String",
+	"evolution: Evolution",
+	"factory:String => IMember"
+)();
+Class.result("RunResult")(
+	"popSize:Int",
+	"success:Boolean",
+	"runtime:Long",
+	"generationCount:Int",
+	"fitnessCallCount:Int"
+	)();
+Class.stoastic("StoasticRun")(
+	"runs:Seq[RunResult]",
+	"required:Int",
+	)();
+topToBottom.left(20)(ex,result,stoastic);
+drawObjects(ex,result,stoastic);
+\end{emp}
+\caption{Experiment}
+\end{figure}
+
+\clearpage
 The architecture was created incrementally, I just implemented the functions
@@
 %What did I learn from my observation?
+\end{empfile}
 \end{document}

### 2015-11-29 evolutionaryComputation/practicals/report.tex
added a bunch of text going trough the entire architecture.

@@
 \subsection{Architecture}
+The code is split up into 4 parts I'll go trough them in the rough order
+they were implemented. First there are the case classes. These are
+imutable classes that hold the domain information (Java pojos).
+Then there are the algoritm classes and objects. These are pretty straight
+forward logical structures. After that comes the expirment and the result
+classes. And finally there is the presentation object Main. the only thing it does
+is interpit the experiment in this case into a latex structure.
+
 \begin{figure}[ht!]
@@
 \end{figure}
+
+The case classses add meaning to the structures. They also provide an API.
+A case class is a scala concept were you just define the atributes and
+equals, hascode and tostring methods are generated for you. They're also
+by default imutable, which is something I kept intact.
+
+The member is something with a gene.
+These are generated in batches into a population,
+a population will also have a function with it to generate new members.
+The population class doesn't use it itself, but when its members are paired
+up into \emph{Pair}'s. A \emph{PairedPopulation} can use the creation function
+of the original population to create new members. The PairedPopulation has
+also a utitity function pairedToPop, this receives a how, in which it can
+insert zipped genes (char by char paired up strings) and from which comes
+a charsequence. This how is usually implemented by one of the crossing methods.
+But the code around this 'how' was always the same, so paired population
+abstracted it away. The resulting string is then inserted into a member
+factory to create a new member (the member factory then allows mutation
+or obfuscation of the genes).
 
@@
 
+The evolution class is the general stucture of a GA{.} It's very simple and
+basically follows the steps from the slides:
+\begin{enumerate}
+	\item Select mates.
+	\item Generate offspring.
+	\item Select the fittests.
+\end{enumerate}
+
+To construct an Evolution you need to pass a valuation function, a mateselector.
+An offspring gnerator, a fittestfilter. And a function that determines if there
+is a good enough solution in the current population.
+
+The step calls almost all given functions in above order except for has good
+enough solution. This function is called in genitic, which will continue
+calling steps until the population is good enough, or until parents
+and children are equal. The population is a case class, so
+the generated comperison method is used to check if the children and parents
+are equal. Startgenetic makes a seq of population and calls genetic.
+
+The objects defined above all have functions with the types same named evolution
+atributes. So the functions are just basically options you can use to create an
+evolution from. The only exception is the withCoinFlipMutation function in
+memberFactories. This function acts as a decorating function. So you can wrap
+it around another function to create a new function.
+Objects are just singletons, or Java static space. Allmost all the functions
+in these objects are `pure', this means they get some input and create some
+output without any side effects. This is not true for the factories, which
+use random, and the offspring generators, which also use random. It's also
+not true for tournement elitism, this suffles randomly.
+
 \begin{figure}[ht!]
@@
 	"factory:String => IMember"
-)();
+)(
+	"+bisectionalSearch():StoasticRun",
+	"verifyLowest(currentPop:Int,faults:Int,index:Int):Seq[RunResult]",
+	"findOptimum(consideringSize:Int,difference:Int):Seq[RunResult]",
+	"run(consideringSize:Int):RunResult"
+);
 Class.result("RunResult")(
-	"popSize:Int",
-	"success:Boolean",
-	"runtime:Long",
-	"generationCount:Int",
-	"fitnessCallCount:Int"
+	"+popSize:Int",
+	"+success:Boolean",
+	"+runtime:Long",
+	"+generationCount:Int",
+	"+fitnessCallCount:Int"
 	)();
 Class.stoastic("StoasticRun")(
-	"runs:Seq[RunResult]",
-	"required:Int",
+	"+runs:Seq[RunResult]",
+	"+required:Int",
 	)();
-topToBottom.left(20)(ex,result,stoastic);
+topToBottom.left(20)(ex,stoastic);
+leftToRight.top(20)(stoastic,result);
 drawObjects(ex,result,stoastic);
+
+link(association)(stoastic.e -- result.w);
+item(iAssoc)("*")(obj.ne = result.w);
 \end{emp}

### 2015-12-02 evolutionaryComputation/notes4.tex
notes from fitfth lecture

@@
 
+\section{Second practical}
+
+MLS, ILS and GLS for graph Bipartioning.
+
+the easiest thing to do is how long and see on waht
+percentage they from the total??
+
+Graph biportitioning, cut the graph into 2 groups
+of 250
+
+if the hamming distance is bigger than 2, just take
+the complement of the 2?
+
+\section{Knapsack}
+Adaptive pursuit, pick one random with probability x, change
+probability while running.
+
+\section{Adaptive pursuit}
+probability matching is used most often in literature, but its not
+really that good.
+
+Markov!
+
+\subsection{Strategy}
+the learning pase is a prarameter you have to play with.
+
 \end{document}

### 2015-12-02 evolutionaryComputation/practicals/report.tex
worked on the report for a bit, fixed a bunch of spelling errors

--- 2015-11-28 evolutionaryComputation/practicals/report.tex
@@
 \section{Introduction}
-This is a report of the evolutionary practical assigment 1.
-In here I will discuss what this assignment contained, which
-methodology I used to execute the assignemnt and how the execution
-went.
+This is a report of the evolutionary practical assignment 1.
+In here I will discuss the execution of the assignment, what it was, which
+methodology was used to execute the assignment.
 
-The intended audiance is people familiar with programming.
+The intended audience is people familiar with programming.
 
@@
 \right. \]
+
+\subsection{Experiment}
+After implementing these functions we were tasked to make an experiment of it
+and make some graphs of the results and draw your conclusions.
+
 \section{Methodology}
+The developing method I chose to use is what I like to call the headless
+chicken approach. In here you just start implementing functionalities in
+hope the result will be good enough. I think this is preferable to other
+methodologies such as PRINCE2 or SCRUM because it has very little
+methodology overhead.  With a precise defined assignment like this one I
+think reducing overhead is one of the main concerns.
+Besides most professional methodologies are more
+aimed at making teams works rather than individuals.
+
 \subsection{Language choice}
@@
 	\item C\#, the language is good but setting it up is difficult since I
-		use Linux, it also has no Gradle support which makes it clunky,
-		you'd have to rely on .net libraries only which is limited. It
-		also prefers object orientated  rather than functional programming,
-		Scala prefers functional over object orientated.
-		Scala's syntax is also more concise.
+		use Linux, it also has no Gradle support which makes difficult to
+		use separate libraries or you have to rely on {.}net libraries only
+		which is limited. It also prefers object orientated rather than
+		functional programming, Scala prefers functional over object
+		orientated. I'm off the opinion that this assignment is quite good
+		to try functional programming on. Scala's syntax is also more concise.
+		In fact C\# feels pretty dated which it flamboyant use of `;' and the
+		strange semantic difference that exists between class brackets and
+		method brackets. Scala's type system is also more advanced than
+		C#'s, so its in a way more object orientated.
 	\item Rust/Go, I'm curious about both these languages, but I'm not at
@@
 		this, I've already done this with Scala before.
+	\item Python, I think the project is in that grey area where weakly
+		typed langauges become less powerful in use. Especially considering
+		the fact how quickly these algoritms blow up in runtime. Static type
+		checking then becomes a gift form heaven.
+	\item C++, I don't want to deal with C++.
 \end{itemize}
@@
 \subsection{Architecture}
-The architecture was created incrementally, I just implemented the functions
-without thinking much about how to combine them.  I first started with the
-fittest functions. Then I went on with selection, and finally I did the
-crossover.
+%TODO: check spell from here
+The code is split up into 4 parts, first there are the case classes. These are
+immutable classes that hold the domain information (Java beans).
+Then there are the algorithm classes and objects. These are pretty straight
+forward logical structures. They implement the atoms of possible GA's, such
+as for example uniform crossover. After that comes the experiment and the result
+classes, the experiment just uses the GA in a specific way.
+Finally there is the presentation object \emph{Main}. the only thing it does
+is interpret the experiment into this case into a latex structure. There is no
+diagram discussing Main, because there is little architecture involved.
 
-I just implemented one or 2 options of these and then created an evolution
-class that accepted functions with the right prototype. So now you can
-create an evolution by giving it functions.
+\begin{figure}[ht!]
+\centering
+\begin{emp}[classdiag](20,20)
+Class.mem("Member")("+Fitness:Int", "+genes:String")();
+Class.pop("Population")("+members:Seq[Member]", "+factory:String=>Member")();
+Class.pairedPop("PairedPopulation")("+members:Seq[Pair]","+factory:String=>Member")(
+"+pairedToPop(how):Population");
+Class.pai("Pair")("+father:Member","+mother:Member")();
 
-Most of the GA architecture was already in place before the assingment became
+Instance.population("Population")(
+"createOnesAndZeros(factory:String=>Member,popCount:Int,geneLength:Int):Population"
+);
+
+leftToRight.top(20)(pop,pairedPop);
+topToBottom.left(20)(population, pop);
+topToBottom.left(60)(pop,mem);
+leftToRight.top(80)(mem, pai);
+
+drawObjects(mem,pop,pairedPop,pai,population);
+link(association)(pai.n -- pairedPop.s);
+item(iAssoc)("*")(obj.se = pai.n);
+link(association)(mem.e -- pai.w);
+item(iAssoc)("2")(obj.sw = mem.e);
+
+link(association)(mem.n -- pop.s);
+item(iAssoc)("*")(obj.se = mem.n);
+\end{emp}
+\caption{Case classes}
+\end{figure}
+
+The case classes add meaning to the structures.
+A case class is a Scala concept were you just define the attributes. Then
+Scala generates the `equals', `hashCode' and `toString' methods for you.
+They're also by default immutable, which is something I kept intact. You
+can make mutations by using the generated copy function, in which you can
+specify what is changed.
+
+A \emph{Member} is something with a gene. The fitness value gets pre calculated by
+the factory, so that every members' fitness is only calculated once. This can
+be done reliably because the member is immutable.
+
+The members are generated in batches into a population, either by the paired
+population or by the Population companion object which provides a
+`createOnesAndZeros' function.
+A population will also have a function with it to generate new members.
+The population class doesn't use it itself, but when its members are paired
+up into \emph{Pair}'s. A \emph{PairedPopulation} can use the creation function
+of the original population to create new members.
+
+The \emph{PairedPopulation} has also a utility function pairedToPop,
+this receives a how, in which it can
+insert zipped genes (char by char paired up strings) and from which comes
+a char sequence. This how is implemented by one of the crossing methods.
+This abstraction was added because the code around the cross function was
+always the same.
+The resulting string from `how' is then inserted into a member
+factory to create a new member (the member factory then allows mutation
+or obfuscation of the new genes).
+
+\begin{figure}[ht!]
+\centering
+\begin{emp}[classdiag](20,20)
+Class.evo("Evolution")(
+"+valuation: FitnessEvaluator",
+"mateSelector:IMateSelector",
+"offSpringGenerator: PairedPopulation => Population",
+"fittestFilter: (Population, Population) => Population",
+"hasGoodEnoughSolution: Population => Boolean")(
+"startGenetic(seed:Population): Seq[Population]",
+"genetic(prev:Seq[Population]): Seq[Population]",
+"step(parents:Population): Population",
+);
+Instance.fit("Fitness")(
+	"uniformlyScaled()",
+	"linearlyScaled()",
+	"blockValuation(block:Seq[Float])",
+	"createProbe(function):FitnessEvaluator"
+);
+Instance.filt("FittestFilter")(
+	"killParents()",
+	"truncateElitism()",
+	"tournementElitism()"
+	);
+Instance.select("MateSelection")(
+	"tournementSelect()"
+);
+Instance.factor("MemberFactories")(
+	"tightlyLinkded()",
+	"randomlyLinked()",
+	"withCoinflipMutation(function):function",
+);
+Instance.off("OffspringGenerator")(
+	"uniformCross()",
+	"singlePointCross()",
+	"twoPointCross()"
+);
+
+topToBottom.left(20)(evo,fit,filt,factor);
+leftToRight.top(20)(fit,off);
+leftToRight.top(20)(filt,select);
+
+drawObjects(evo,fit,factor,filt,select,off);
+\end{emp}
+\caption{Algorithm}
+\end{figure}
+
+The evolution class is the common structure of a GA{.} It's very simple and
+basically follows these steps:
+\begin{enumerate}
+	\item Select mates.
+	\item Generate offspring.
+	\item Select the fittest.
+\end{enumerate}
+
+To construct an Evolution you need to pass a \emph{valuation function},
+a \emph{mateselector}, An \emph{offspring generator}, 
+a \emph{fittestFilter}. And a function that determines if there
+is a good enough solution in the current population. The last one was added
+because I wasn't quite sure when the GA needed to stop in the beginning.
+
+The step method calls almost all given functions in above order except for has good
+enough solution. This function is called in genetic, which will continue
+calling steps until the population is good enough, or until parents
+and children are equal. The population is a case class, so
+the generated comparison method is used to check if the children and parents
+are equal. \emph{Start genetic} makes a seq of population and calls genetic.
+
+The objects defined above all have functions with the types same named evolution
+attributes. So the functions are just basically options you can use to create an
+evolution from. There are some exceptions which allow wrapping is the
+withCoinFlipMutation function in
+memberFactories. This function acts as a decorating function. So you can wrap
+it around another function to create a new function.
+
+Objects are just singletons, or Java static space. I use the object keyword in
+Scala instead of class So I chose also to use an object in the UML
+representation. Scala seems to stick pretty close to UML notation anyways.
+The objects don't have types however, because they are the types. They're
+language enforced singletons.
+
+Fitness contains the evaluation functions. The assignment defined four. However
+the block evaluation (deceptive/nondeceptive) can be implemented as the
+same function with different values for the indices (which blockvaluation 
+does). Interesting is that the blockvaluation in code is shorter than the
+mathematical definition.
+The create probe function is a utility function that allows the counting of
+the fitness calls. The initial design didn't have a \emph{FitnessEvaluator},
+but the assignment desired the counting of calls so I just wrapped the
+function in a object that increases the count on every call.
+
+The offspringGenerator implements the 'how' part of the pairedPopulations.
+The mutation isn't part of the offspringGenerator because this is done by
+the member factory.
+
+The fittest filter decides how which of both the parents and children 
+continues. The most simple implementation is the Killparents, which
+returns the children. Truncate elitism selects the best of both parents
+and children. Tournament matches child next to each parent
+and decide which one is best. There is no kill children implementation
+because nothing would ever change.
+
+MateSelection implements tournament select. This is the only mate selection
+algorithm implemented currently.
+
+The member factories are able to create members from strings. These are either
+the tightly linked or randomly linked. Then the tightly or randomly linked
+functions can also be wrapped in a mutation function,
+which will mutate the string before passing it to the factory.
+Its an decorator idea, only with functions.
+
+\begin{figure}[ht!]
+\centering
+\begin{emp}[classdiag](20,20)
+Class.ex("Experiment")(
+	"+name:String",
+	"+variation:String",
+	"evolution: Evolution",
+	"factory:String => IMember"
+)(
+	"+bisectionalSearch():StochasticRun",
+	"verifyLowest(currentPop:Int,faults:Int,index:Int):Seq[RunResult]",
+	"findOptimum(consideringSize:Int,difference:Int):Seq[RunResult]",
+	"run(consideringSize:Int):RunResult"
+);
+Class.result("RunResult")(
+	"+popSize:Int",
+	"+success:Boolean",
+	"+runtime:Long",
+	"+generationCount:Int",
+	"+fitnessCallCount:Int"
+	)();
+Class.stoch("StochasticRun")(
+	"+runs:Seq[RunResult]",
+	"+required:Int",
+	)();
+topToBottom.left(20)(ex,stoch);
+leftToRight.top(20)(stoch,result);
+drawObjects(ex,result,stoch);
+
+link(association)(stoch.e -- result.w);
+item(iAssoc)("*")(obj.ne = result.w);
+\end{emp}
+\caption{Experiment}
+\end{figure}
+
+The experiment is setup to do the bisectional search. The bisectional search
+function will wrap the result of \emph{findOptimum} in a \emph{StochasticRun}.
+\emph{findOptimum} is a recursive function that calls verify lowest to get a
+collection of runResults. This is used to determine if the population size
+should be increased or decreased like a bisection search.
+\emph{verifyLowest} will try to run a population size 30 times, if the faults
+limit is exceeded or the index exceeds the required runs limit it will quit.
+In the latter case the run is considered to be successful.
+The run will execute the GA search according to the configuration of the 
+experiment, so this function will call startGenetic, and put the results
+of the run into a RunResult. The RunResult will actually truncate a lot of
+information. Not the entire resulting population is stored. It wasn't
+necessary for the experiment and I was concerned about memory. Or rather
+I didn't want to be too wasteful, especially considering that the JVM
+reserves big blocks of memory.
+
+\clearpage
+
+Most of the GA architecture was already in place before the assignment became
 available. I did this based on the notes I took in class. When the assignment
-became available I was almost completly done with the GA. There were some bugs
+became available I was almost completly done with the GA{.} There were some bugs
 in my code however and I interperted some things wrong. I also didn't write
@@
 
-\paragraph{Bugs found}
-Linearly scaled reported the wrong value because it used indeci, so it
-gave the first 1 a value of 0, and the second a value of 1 etc.
-
-Trap functions were simply wrongly implemented, instead of counting the
-ones they gave positional values of the ones in the block and used the
-block as a lookup value for that.
-
-Trap function actually reversed the output. I suspected the trap function
-was doing it wrong because all the GA's kept failing, so I added a bunch
-of very specific test for each block result which confirmend my suspicion.
-The actuall bug was a simple reverse, which meant  the population, slowly
-leaned towards 0's instead of 1's. and since my 'solution' is a string
-of only ones, the end result is that the bisectional search kept increasing
-the population size since now anwser was found and the algoritm wasn't happy.
-
-All valuation functions should return a float instead of an Int.
-I don't even understan how I could make such a stupid mistake until the
-unit test pointed it out to me. (programming for 6+ years..)
-
+I found about 6 different bugs alone in the valuation functions by using 
+unit-testing, so I think this saved me considerable time.
 \subsubsection{Build tool}
@@
 \begin{itemize}
-	\item The scala standart library, because I have to when using scala.
+	\item The Scala standard library, because I have to when using Scala.
 	\item slf4j, a simple logging facade, this allows runtime switching of
 		logging schemes, for example if you want to log into a database or
-		a file instead of stdout you can do this relativly easy with slf4j
-		This is not necisary right now but the api is simple enough to use
+		a file instead of stdout you can do this relatively easy with slf4j
+		This is not necessary right now but the API is simple enough to use
 		everywhere.
-	\item scalaplot. A graph plotting library for scala. I didn't want to
-		do this myself, and this library supports svg plotting and ascii
+	\item Scalaplot. A graph plotting library for Scala. I didn't want to
+		do this myself, and this library supports SVG plotting and ASCII
 		plotting.
-	\item scalatest. A testing library for scala, it has some pretty advanced
+	\item Scalatest. A testing library for Scala, it has some pretty advanced
 		features but I use it mostly like JUnit.
 \end{itemize}
-\subsection{Uncertainties}
-About randomly linked:
-	There is another way to do it: for every member randomly link
-	differently, and randomly pick 1 randomly linked instructionset
-	for offspring generation (or do the strongest).
-	Currently I do the global randomly linked, every run will however
-	have the same randomly linked value, I'm aware of this.
-
-The stocastic ensursance wants to make sure that 29 out of 30 runs are
-'good enough', but combining this with the bidisectional search is difficult.
-because the bidisectional search may happen to end up just higher than possible.
-
-I considered to combine the bidisectional search with the stocastic ensurance.
-But that would be way slower, and it won't be much better (or even worse depending
-on how you do it) then my current implementation.
 
@@
 lot of slow mechanisms, such as functional combinators, and although
-they don't increase the computational complexity, they do cerainly increase
-the minimum required runtime, I'm quite sure that every combinator in sequance
+they don't increase the computational complexity, they do certainly increase
+the minimum required runtime, I'm quite sure that every combinator in sequence
 is implemented as just another loop. However because I
-run all expirments in paralel the total time may be faster.
-\subsubsection{Paralization}
-Since were doing string comparisons I had the idea of ussing parlization to
-speed up the algoritm by doing paralization. This can be done very easily
-in scala with par collections (and the fact that imutibility is default).
+run all experiments in parallel the total time may be faster.
+\subsubsection{Parallelization}
+Since were doing string comparisons I had the idea of using parallelization to
+speed up the algorithm by doing parallelization. This can be done very easily
+in Scala with par collections (and the fact that immutability is default).
 
 However once I implemented this, I came to the conclusion it was not faster.
-The overhead for delegating and creating threads was more than the parilization
+The overhead for delegating and creating threads was more than the parallelization
 gained.  After 10 minutes of waiting I decided to drop this idea
-since without parilization it took about 2 minutes (I tried this
+since without parallelization it took about 2 minutes (I tried this
 on just one run, so it shouldn't take that long).
@@
 The problem is that the evaluation function, even with gene length of 100, is
-to cheap to make any reasonable paralization effecitve. However what was effictive
-is the parilization of the expiriments. This reduced total runtime 
-significantly. however this is not seen in the statistics.
+to cheap to make any reasonable parallelization effective. However what was effictive
+is the parallelization of the experiments. This reduced total runtime
+significantly. However this is not seen in the statistics.
 
-So now all expiriments are run at the same time. This probably increases the
-average time of each 'run' somewhat (not that much because the concurent
-collection does \href{http://docs.scala-lang.org/overviews/parallel-collections/configuration.html}{load balancing})
-It alo drasticly decreases the amount of time I have to wait. Which is quite
+So now all experiments are run at the same time. This probably increases the
+average time of each `run' somewhat (not that much because the concurrent
+collection does \href{http://docs.scala-lang.org/overviews/parallel-collections/configuration.html}{load balancing}).
+It also drastically decreases the amount of time I have to wait. Which is quite
 important, especially in the development phase. (although I also used
@@
 
-The paralization of experiments can be removed by deleting `par' on line 64
+The parallelization of experiments can be removed by deleting `par' on line 64
 of Main.scala
@@
 Initially my members just contained the string with the `genes'. This I passed
-around and would eveulate when needed. However I realized that I could do this
+around and would evaluate when needed. However I realized that I could do this
 evaluation eagerly on creation, to reduce the evaluation calls. A member has
-to be eveluated anyways, so doing this on creation and just putting in the
+to be evaluated anyways, so doing this on creation and just putting in the
 evaluation saves some heavy evaluation operations.
@@
 This reduces evaluations and object creations. Its becomes more effective
-at deeper generation, it even slows down the fastest GA's by a few miliseconds
+at deeper generation, it even slows down the fastest GA's by a few milliseconds
 but the deeper GA are lots faster.
@@
 
-To reduce the amount of string comparisons, first the hascode function is called
+To reduce the amount of string comparisons, first the hash code function is called
 This will cache itself, I know this because I inspected the source code of the
@@
 
-I was very proud for coming up with  this idea, but when I ran the expirment with
-and without hascode I came to the conclusion that first doing a hascode operation
+I was very proud for coming up with  this idea, but when I ran the experiment with
+and without hash code I came to the conclusion that first doing a hash code operation
 significantly *slows* execution. This becomes worse as population size increases.
 
-\section{Execution}
-So I can execute in parralel, I can easily add whatever function I want, so
-I decided to just run a whole bunch of experiments and see what comes out.
-
-Something to note is that my made up tournement elitism is actually in most
+\section{Experiment}
+Something to note is that my made up tournament elitism is actually in most
 cases better, or at least equal to the truncate elitism. So when I saw
-this I tried using my other initially used algorithm where I would completly
+this I tried using my other initially used algorithm where I would completely
 ignore the parents (kill the parents). However it appears this does not
@@
 those fair, running the code now takes about 40 minutes in total with the
-entire thing diveded over 8 threads (quad core + hyperthreading). The only
+entire thing divided over 8 threads (quad core + hyper-threading). The only
 other improvement I could think of is running it on the GPU, but I think
-the algortim is to complex for that.
+the algorithm is to complex for that and I don't know how to do that,
+especially considering I'm using the JVM.
 
-Two point cross and single cross both seem to find results in deciptive 
-random tournement elitism.
+Two point cross and single cross both seem to find results in deceptive
+random tournament elitism.
+
+\subsection{Hypothesis}
+The deceptive trap function will not be solved by any algorithm.
+non-deceptive may be more successful. The others will be solved.
+I also think mutation will have an important effect and may even
+allow deceptive to be solved.
+
+I also think that the provided selection pressure would be about
+average. It will probably always find a result.
+
+\subsection{Execution}
+Note that I added some extra selection methods and crossing strategies to
+the final experiment. I did this because I was curious how they would
+compete with the original assignment.
+
+After doing some sample experiments I found for example that 2x was going
+up with number of successes in the biggest population size. So I wondered 
+if doing my original interpretation of the select would work better.
+I found that in almost every case this original interpretation is at least as
+good or better. On both runtime and successes required population for success.
 
@@
 %What did I learn from my observation?
+My conclusion is that using both tournament selection and truncate elitism
+is to harsh. It will slow down the GA and often not find a solution at all
+(in case of deceptive trap, 2x should've come up with something). the 
+time vs population found ratio is also much better for just using tournament
+select.
+
+Whats also interesting is that if you just throw away the parents and use 
+the children (kill parents), then you still will often perform better than
+using tournament selection and truncate elitism, although not as 
+consistently as tournament elitism.
+
+To my surprise deceptive trap was solved with decreased selection pressure
+by both 2x and 1x, although 1x did worse. Also adding mutation doesn't do
+the algorithm much good. 2XM and UXM consistently find worse minimum populations
+than respectively 2X and UX{.}
+This probably means the mutation is too disruptive.
+
+Mutation turned out to have no significant effect other than just slowing
+things down.
+
+\end{empfile}
 \end{document}

### 2015-12-02 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Experiment.scala

@@ -67,13 +67,13 @@ class Experiment(val name:String, val variation:String, evolution: Evolution, me
 		if(newDiff < popUnit){ // steps becoming to small, bort return result
 			return result
 		}
-		if(StoasticRun.isSucces(result)){
+		if(StochasticRun.isSucces(result)){
 			// if we had success, zoom in
 			return result ++ findOptimumPopsize(consideringSize-difference/2, difference/2)
 		}
 		result ++ findOptimumPopsize(consideringSize+newDiff,newDiff)
 	}
-	def bisectionalSearch() = StoasticRun(findOptimumPopsize(popUnit ,popUnit/2), requiredRuns)
+	def bisectionalSearch() = StochasticRun(findOptimumPopsize(popUnit ,popUnit/2), requiredRuns)
 }
 
 case class RunResult(popSize:Int, success:Boolean, runtime:Long, generationCount:Int, fitnessCallCount:Int){
@@ -82,7 +82,7 @@ case class RunResult(popSize:Int, success:Boolean, runtime:Long, generationCount
 	}
 }
 import util.Properties.lineSeparator
-case class StoasticRun(runs:Seq[RunResult], required:Int) {
+case class StochasticRun(runs:Seq[RunResult], required:Int) {
 	lazy val byPopSize = runs.groupBy(x=>x.popSize)
 	lazy val bestRun = byPopSize.getOrElse(bigestPopcount, Nil)
 	lazy val isSuccesfull = bestRun.size == required
@@ -91,7 +91,7 @@ case class StoasticRun(runs:Seq[RunResult], required:Int) {
 	}
 	lazy val bigestPopcount = {
 	 	val successes = byPopSize.filter(p =>
-			StoasticRun.isSucces(p._2)
+			StochasticRun.isSucces(p._2)
 		)
 		// do the highest so its obvious its a lot
 		if(successes.isEmpty) runs.last.popSize else successes.minBy(x=>x._1)._1
@@ -101,7 +101,7 @@ case class StoasticRun(runs:Seq[RunResult], required:Int) {
 	lazy val bestRunAverageTime = bestRun.map(x=>x.runtime).sum/bestRun.length
 }
 
-object StoasticRun{
+object StochasticRun{
 	def isSucces(results:Seq[RunResult]) =  results.count(r=> r.success) >= (
 		Experiment.requiredRuns - Experiment.faultTolerance
 	)
@@ -116,6 +116,11 @@ object Experiment{
 	val requiredRuns = 30
[…]

### 2015-12-02 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -63,18 +63,18 @@ object Main{
 		val deciptive:Seq[Float] = List(4,0,1,2,3)
 		val nonDeciptive:Seq[Float] = List(4,0,0.5f,1,1.5f)
 		val expirements =
-		Experiment.create("Uniformly scaled truncate elitism", random, uniformlyScaledCountOnes, crossMethodsTight) ++
-		Experiment.create("Linearly scaled truncate elitism", random, linearlyScaledCountOnes, crossMethodsTight) ++
-		Experiment.create("Block deceptive tight truncate elitism", random, blockValuation(deciptive), crossMethodsTight) ++
-		Experiment.create("Block non-deceptive tight truncate elitism", random, blockValuation(nonDeciptive), crossMethodsTight) ++
-		Experiment.create("Block deceptive random truncate elitism", random, blockValuation(deciptive), crossMethodsRandom) ++
-		Experiment.create("Block non-deceptive random truncate elitism", random, blockValuation(nonDeciptive), crossMethodsRandom) ++
-		Experiment.create("Uniformly scaled, tournement elitism", random, uniformlyScaledCountOnes, crossMethodsTight, FittestFilter.tournementElitism) ++
-		Experiment.create("Linearly scaled, tournement elitism", random, linearlyScaledCountOnes, crossMethodsTight, FittestFilter.tournementElitism) ++
-		Experiment.create("Block deceptive tight, tournement elitism", random, blockValuation(deciptive), crossMethodsTight, FittestFilter.tournementElitism) ++
-		Experiment.create("Block non-deceptive tight, tournement elitism", random, blockValuation(nonDeciptive), crossMethodsTight, FittestFilter.tournementElitism) ++
-		Experiment.create("Block deceptive random, tournement elitism", random, blockValuation(deciptive), crossMethodsRandom, FittestFilter.tournementElitism) ++
-		Experiment.create("Block non-deceptive random, tournement elitism", random, blockValuation(nonDeciptive), crossMethodsRandom, FittestFilter.tournementElitism) ++
+		Experiment.create("Uniformly scaled, truncate elitism", random, uniformlyScaledCountOnes, crossMethodsTight) ++
+		Experiment.create("Linearly scaled, truncate elitism", random, linearlyScaledCountOnes, crossMethodsTight) ++
[…]

### 2015-12-02 inteligentagents/notes.tex
added proof deleted inbetween doc statementst s
added notes from the lecture and the practicum.

@@
+\documentclass{article}
+\begin{document}
+\author{Jappie Klooster}
+\title{Temporal magick or logic}
+\maketitle
+
 #Exam
@@
 
-\documentclass{article}
-\begin{document}
-\author{Jappie Klooster}
-\title{Cohen \& Levesque}
-\maketitle
 \section{Langauge}
@@
 We'll only look into the future.
-\end{document}
-
-\documentclass{article}
-\begin{document}
-\author{Jappie Klooster}
-\title{Temporal magick or logic}
-\maketitle
 \section{Linear temporal magick}
@@
 
-\end{document}
-\documentclass{article}
-\begin{document}
-\author{Jappie Klooster}
-\title{KARO logic}
-\maketitle
-
 There are indeed unkown unkowns. But not if you have
@@
 You can only uncommit commited actions.
+
+\section{practical}
+\begin{prooftree}
+	\AxiomC{$\vDash \Diamond (p\vee q)\wedge \Box \neg q)\to \Diamond p$}
+	\UnaryInfC{To proof $m,\sigma \vDash \Diamond (p\vee q)\wedge \Box \neg q)\to \Diamond p$}
+	\UnaryInfC{To proof $m,\sigma \vDash \Diamond (p\vee q)\wedge \Box \neg q)\Rightarrow m,\sigma \vDash \Diamond p$}
+	\UnaryInfC{To proof $m,\sigma \vDash \Diamond (p\vee q)$ and $\Box \neg q)\Rightarrow m,\sigma \vDash \Diamond p$}
+	\UnaryInfC{To proof exists $m,\sigma^n \vDash p$ or $m,\sigma^n \vDash q$
+	and for all $m,\sigma^m \vDash \neg q$ then exists $m,\sigma^k \vDash p$}
+\end{prooftree}
+
+\begin{prooftree}
+	\AxiomC{$\vDash A \phi \to \phi$}
+	\UnariyInfC{To prove: for all $M,v,w_t \vDash A \phi \to \phi$}
+	\UnariyInfC{Take arbitrary $M,v,w_t$ to prove: for all $M,v,w_t \vDash A \phi \to \vDash \phi$}
+	\UnariyInfC{To prove: for all $M,v,w_t \vDash A \phi \Rightarrow m,\sigma \vDashM,v,w_t \vDashM,v,w_tM,v,w_t \vDash \phi$}
+	\UnariyInfC{Assume $M,v,w_t \vDash A \phi$ ie  
+	$M,v,w_1, w_2, \ldots \vDash \phi$ for all 
+	$w_1, w_2, \cdots$ with $t_1 =t$
+	}
+\end{prooftree}
+\begin{prooftree}
+\end{prooftree}
+	\AxiomC{$\neg\vDash A\Diamond \phi \to E\Box \phi$}
+
+\section{Persistance problem}
+Closed world assumption = In databases, everything that is not in the
+database is not true.
+
+If I describe an action I want a minimal description of what change
+and what does not.
+
+\section{Situational calculus Macarty}
+calculus framework, not a logic proper.
+
+$Result(A,s)$ or do. Normally its written like $s\vDash p$
+
+\subsection{Sidenote}
+There is a big difference between european AI and american.
+American AI is much more applied. And McCarthy is more like a `god'.
+In europe modal logic is much more used.
+
+\subsection{Calculus}
+\subsubsection{Herbrand situation}
+A situation is always a sequance of actions performed on the initial situation
+
+positive effects starts to hold after the action, and negative it doesn't.
+
+Also describe the things that don't change. These are the 'axioms'.
+
+\subsubsection{The frame problem}
+You need a pair of axioms is needed for every fluent and action. This will
+result in way to many axioms.
+
+This is solved with default reasoning. Altough writer gave it up eventually.
+You need to do a consisty check, this is in modal logic already NP-hard.
+
+\section{Planning}
+GPS = General planning system
+this has to do with program syntesis
+
+STRIPS, depth first.
+
+\subsection{The susman anomaly}
+This can go into a loop, because focusing on one of these goals can
+make the other one false.
+
+\subsection{start backwards}
+Breadtfirst is then more feasible because there are fewer brances.
+\[[\alpha] = wp(\alpha,\gamma)\]
+
+\section{practical X}
+i.e = $\RightLeftarrow$
+\subsection{9A}
+\begin{prooftree}
+	\AxiomC{To prove $(m,s)$ $\vDash [\alpha_1;\alpha_2]\phi\LeftRightarrow
+[\alpha_1][\alpha_2]\phi$ for all $m,s$}
+
+\UnaryInfC{Suppose $(m,s)$ $\vDash [\alpha_1;\alpha_2]\phi$}
+\UnaryInfC{ie for all s': $R_{\alpha_1;\alpha_2}(s,s')\Rightarrow m,s$}
+\UnaryInfC{ie for all s': $(\exists t: R_{\alpha_1}(s,t)$ and
+$R_{\alpha_2}(t,s') \Rightarrow m,s' \vDash\phi$}
+\UnaryInfC{ie for all s': $\forall t[ : R_{\alpha_1}(s,t)$ and
+$R_{\alpha_2}(t,s') \Rightarrow m,s' \vDash\phi]$}
+
+\UnaryInfC{$\forall s': \forall t[ : R_{\alpha_1}(s,t)$ and
+$R_{\alpha_2}(t,s') \Rightarrow m,s' \vDash\phi]$}
+\UnaryInfC{$\forall t: \forall s[ : R_{\alpha_1}(s,t)$ and
+$R_{\alpha_2}(t,s') \Rightarrow m,s' \vDash\phi]$}
+
+\UnaryInfC{$\forall t: R_{\alpha_1}(s,t) \Rightarrow [\forall S'
+R_{\alpha_2}(t,s') \Rightarrow m,s' \vDash\phi]$}
+
+\UnaryInfC{$\forall t: R_{\alpha_1}(s,t) \Rightarrow m,t\vDash[\alpha_2]
+\vDash\phi]$}
+\UnaryInfC{m,s$\vDash[\alpha_1][\alpha_2]\phi$}
+\end{prooftree}
+
+\begin{prooftree}
+	\AxiomC{To prove $m,s\vDash <\alpha>(\phi\wedge
+	\varphi)\LeftRightarrow(<\alpha>\phi\wedge<\alpha>\varphi)$, for all m,s}
+	\UnaryInfC{$m,s\vDash<\alpha>(\phi\wedge\varphi)\LeftRightarrow$}
+	\UnaryInfC{$\exists: R_\alpha(s,t)&m,t\vDash\phi\wedge\varphi\LeftRightarrow$}
+\end{prooftree}
+\begin{prooftree}
+	\AxiomC{To proof: $m,s \vDash A\alpha;\alpha_2\LeftRightarrow 
+		(A_{\alpha_1}\wedge[\alpha_1]A_{\alpha_2})$}
+	\UnaryInfC{Assume arb. $m,s \vDash A\alpha_1;\alpha_2\LeftRightarrow$}
+	\UnaryInfC{$\alpha_1;\alpha_2 \exists C (m,s) \LeftRightarrow$}
+	\UnaryInfC{$\alpha_1 \exists C (m,s) \& \alpha_2 \exists C (m,s)\LeftRightarrow$}
+	\UnaryInfC{$m,s \vDash A\alpha_1 \exists C (m,s) \& R_{\alpha_1}(mcs)
+	\vDash A\alpha_2 \LeftRightarrow$}
+	\UnaryInfC{$m,s \vDash A\alpha_1 \& \forall(m',s'):R_\alpha_1
+	(m,s)(m',s')\rightarrow m',s')\vDash A\alpha_2 \LeftRightarrow$}
+	\UnaryInfC{$m,s \vDash A\alpha_1 \& m,s\vDash [\alpha_1]A\alpha_2 \LeftRightarrow$}
+	\UnaryInfC{$m,s \vDash A\alpha_1 \& \wedge [\alpha_1]A\alpha_2 \LeftRightarrow$}
+\end{prooftree}
 \end{document}

### 2015-12-02 methods/agents/summery.tex
added context logic

@@
 \end{itemize}
+
+\subsection{ContextLogic: $S5P,K45^{ij}$}
+
+\paragraph{Universal modality $\Box$} Based on S5(equivalence relation).
+\paragraph{Cluster modality [i]} referring to subrelations ('clusters') of
+universal relation. Its Transative, euclidean $K45$
+
+\subsubsection{S5P Models}
+\[M (S,\pi, S_1, S_2, \dots, S_n\]
+\begin{tikzpicture}
+    \draw (0,0) ellipse (5cm and 3cm);
+	\draw[color=red] (1,1) ellipse (2cm and 1cm);
+    \draw[color=red] (0,1) ellipse (2cm and 1cm);
+    \draw[color=red] (0,0) ellipse (2cm and 1cm);
+    \draw[color=red] (-1,-2) ellipse (2cm and 0.5cm);
+\end{tikzpicture}
+
+\subsubsection{Semantics}
+\begin{tabular}{lll}
+	Modal & $\Leftrightarrow$ & Interpertation \\ \toprule
+	$M,s \vDash \Box\phi$ & $\Leftrightarrow$ &
+	$M,s' \vDash \phi$ for all $s'$ $\in S$ \\
+
+	$M,s \vDash [i]\phi$ & $\Leftrightarrow$ &
+	$M,s' \vDash \phi $ for all $s'$ $\in S_i$ \\
+
+	$M,s \vDash \Box\phi$ & $\Leftrightarrow$ &
+	$M,s' \vDash \phi $ for all $s'$ $\in S$ such that $R(s,s')$ \\
+
+	$M,s \vDash [i]\phi$ & $\Leftrightarrow$ &
+	$M,s' \vDash \phi $ for all $s'$ $\in S_i$ such that $R_i(s,s')$ \\
+\end{tabular}
+
+\paragraph{Accessibility relations}
+\[R(s,s') \mbox{ for all } s,s' \in S\]
+\[R_i(s,s') \mbox{ for all } s,s' \in S_i\]
+
+$R$ is equivalance and $R_i$ is transative and euclidean.
+
+\[S5P = K45^{ij}_{(m)} + \Box\]
+
 \section{APL}

### 2015-12-06 evolutionaryComputation/practicals/report.tex
fixed a bunch of spelling mistakes also added more structere and text

@@
 After implementing these functions we were tasked to make an experiment of it
-and make some graphs of the results and draw your conclusions.
+and make some graphs of the results and make a conclusion based on the results.
 
 \section{Methodology}
-The developing method I chose to use is what I like to call the headless
-chicken approach. In here you just start implementing functionalities in
-hope the result will be good enough. I think this is preferable to other
-methodologies such as PRINCE2 or SCRUM because it has very little
-methodology overhead.  With a precise defined assignment like this one I
-think reducing overhead is one of the main concerns.
-Besides most professional methodologies are more
-aimed at making teams works rather than individuals.
 
@@
 a little clumsy sometimes, because Scala supports operator overloading in
-a strange way).
+a strange way, but its possible).
 
@@
 \begin{itemize}
-	\item Java, which I'm very familiar with but I find quite boring to use.
-		Its very verbose.
-	\item C\#, the language is good but setting it up is difficult since I
-		use Linux, it also has no Gradle support which makes difficult to
-		use separate libraries or you have to rely on {.}net libraries only
-		which is limited. It also prefers object orientated rather than
-		functional programming, Scala prefers functional over object
-		orientated. I'm off the opinion that this assignment is quite good
-		to try functional programming on. Scala's syntax is also more concise.
-		In fact C\# feels pretty dated which it flamboyant use of `;' and the
-		strange semantic difference that exists between class brackets and
-		method brackets. Scala's type system is also more advanced than
-		C#'s, so its in a way more object orientated.
+	\item Java, Java is an industry standard language, but the syntax of
+		Java is very verbose. So when I have the choice I try to avoid it.
+		I've also used a lot of Java when I was younger so I won't learn
+		much by using it.
+	\item C\#, the language is good, a lot less verbose then Java
+		but setting it up is difficult since I use Linux.
+		It also prefers object orientated rather than
+		functional programming. Scala's type system is also more advanced than
+		C\#'s, so its in a way more object orientated.
 	\item Rust/Go, I'm curious about both these languages, but I'm not at
@@
 		hobby project with them before starting doing something serious like
-		this, I've already done this with Scala before.
+		this, this is in contrast to Scala, with which I already played before.
 	\item Python, I think the project is in that grey area where weakly
-		typed langauges become less powerful in use. Especially considering
-		the fact how quickly these algoritms blow up in runtime. Static type
-		checking then becomes a gift form heaven.
-	\item C++, I don't want to deal with C++.
+		typed languages become less powerful in use. Especially considering
+		the fact how quickly these algorithms blow up in runtime. Static type
+		checking then helps solving a big part of the bugs before execution
+		and since execution will be slow, this saves a lot of time.
+	\item C++, C++ is extremely fast in execution times. However implementation
+		time is quite long compared to almost everything else.
+		Errors are for example reported poorly and
+		the syntax is inconsistent. Since execution speed isn't a top priority
+		I decided to not use C++.
 \end{itemize}
@@
 \subsection{Architecture}
-%TODO: check spell from here
 The code is split up into 4 parts, first there are the case classes. These are
@@
 classes, the experiment just uses the GA in a specific way.
-Finally there is the presentation object \emph{Main}. the only thing it does
-is interpret the experiment into this case into a latex structure. There is no
+Finally there is the presentation object \emph{Main}. The only thing it does
+is interpret the experiment into a latex structure. There is no
 diagram discussing Main, because there is little architecture involved.
+
+\subsubsection{Case classes}
+The case classes add meaning to the structures.
+A case class is a Scala concept were you just define the attributes. Then
+Scala generates the `equals', `hashCode' and `toString' methods for you.
+They're also by default immutable, which is something I like using. You
+can make mutations by using the generated copy function, in which you can
+specify what is changed to create a new copy with the change.
 
@@
 
-The case classes add meaning to the structures.
-A case class is a Scala concept were you just define the attributes. Then
-Scala generates the `equals', `hashCode' and `toString' methods for you.
-They're also by default immutable, which is something I kept intact. You
-can make mutations by using the generated copy function, in which you can
-specify what is changed.
+\paragraph{Member} A \emph{Member} is something with a gene. The fitness value gets pre calculated by
+the factory, so that every members' fitness is only calculated once. This can
+be done reliably because the member is immutable. There is no way to modify
+the genes so the fintness value does not have to change.
 
-A \emph{Member} is something with a gene. The fitness value gets pre calculated by
-the factory, so that every members' fitness is only calculated once. This can
-be done reliably because the member is immutable.
-
-The members are generated in batches into a population, either by the paired
+\paragraph{Population} The members are generated in batches and stored into a
+\emph{Population}, either by the paired
 population or by the Population companion object which provides a
 `createOnesAndZeros' function.
-A population will also have a function with it to generate new members.
+A population will also have a function with it to generate new members as
+data in this case, it drags the specified function along, its like a
+function pointer.
 The population class doesn't use it itself, but when its members are paired
-up into \emph{Pair}'s. A \emph{PairedPopulation} can use the creation function
+up into \emph{Pair}s. A \emph{PairedPopulation} can use the creation function
 of the original population to create new members.
 
+\paragraph{PairedPopulation}
 The \emph{PairedPopulation} has also a utility function pairedToPop,
@@
 
+I've thought about merging PairedPopulation and population into
+one class. However doing that will probably not make things more obvious.
+Although it does makes sense in that the population would actually use
+the function it drags along.
+
+\newpage
+\subsubsection{Algorithm}
+The objects defined above all have functions with the types
+same named evolution attributes. So the functions are just basically
+options you can use to create an
+evolution from. There are some exceptions which allow wrapping is the
+withCoinFlipMutation function in
+memberFactories. This function acts as a decorating function. So you can wrap
+it around another function to create a new function.
+
+Objects are just singletons, or Java static space. I use the object keyword in
+Scala instead of class So I chose also to use an object in the UML
+representation. Scala seems to stick pretty close to UML notation anyways.
+The objects don't have types however, because they are the types. They're
+language enforced singletons.
+
 \begin{figure}[ht!]
@@
 Class.evo("Evolution")(
-"+valuation: FitnessEvaluator",
+"+evaluation: FitnessEvaluator",
 "mateSelector:IMateSelector",
@@
 );
-Instance.fit("Fitness")(
+Instance.fit("Evaluation")(
 	"uniformlyScaled()",
@@
 
-The evolution class is the common structure of a GA{.} It's very simple and
-basically follows these steps:
-\begin{enumerate}
-	\item Select mates.
-	\item Generate offspring.
-	\item Select the fittest.
-\end{enumerate}
+\paragraph{Evolution}The evolution class is the common structure of a GA(Gentic algorithm){.} 
+So creating an instance of Evolution is like creating a new GA{.}
 
@@
 
-The step method calls almost all given functions in above order except for has good
-enough solution. This function is called in genetic, which will continue
-calling steps until the population is good enough, or until parents
-and children are equal. The population is a case class, so
-the generated comparison method is used to check if the children and parents
-are equal. \emph{Start genetic} makes a seq of population and calls genetic.
+Some of these functions are abstracted away by traits, but still can be seen
+as just fucntions.
+\subparagraph{Step method}
+The step method calls almost all given functions in the follwoing order:
 
-The objects defined above all have functions with the types same named evolution
-attributes. So the functions are just basically options you can use to create an
-evolution from. There are some exceptions which allow wrapping is the
-withCoinFlipMutation function in
-memberFactories. This function acts as a decorating function. So you can wrap
-it around another function to create a new function.
+\begin{enumerate}
+	\item Select mates.
+	\item Generate offspring.
+	\item Select the fittest.
+\end{enumerate}
 
-Objects are just singletons, or Java static space. I use the object keyword in
-Scala instead of class So I chose also to use an object in the UML
-representation. Scala seems to stick pretty close to UML notation anyways.
-The objects don't have types however, because they are the types. They're
-language enforced singletons.
+\subparagraph{Genetic} hasGoodEnoughSolution is called in genetic,
+which will continue calling steps until the population has a good enough
+solution, or until parents and children are equal.
+The population is a case class, so the generated comparison method is used
+to check if the children and parents are equal.
+The result of this function is a sequance of population. The first element
+is te first generation and the last is the last generation.
 
-Fitness contains the evaluation functions. The assignment defined four. However
-the block evaluation (deceptive/nondeceptive) can be implemented as the
-same function with different values for the indices (which blockvaluation 
-does). Interesting is that the blockvaluation in code is shorter than the
-mathematical definition.
+\subparagraph{Start genetic} Start genetic makes a seq of the seed population and
+calls genetic.
+
+\paragraph{Evaluation} The evaluation object
+contains the evaluation functions. The assignment asked to define four.
+However the block evaluation (trap functions) can be implemented as the
+same function with different values for the indices. This is done by
+blockvaluation.
+
 The create probe function is a utility function that allows the counting of
@@
 but the assignment desired the counting of calls so I just wrapped the
-function in a object that increases the count on every call.
+function in a object that increases the count on every call. I had to
+replace the function prototype in evolution with a trait because coutning
+the calls of a function requires some form of state (an increasing variable)
+I didn't know how to do that properly with just functions in scala yet,
+so I just used a trait instead (trait = interface + state).
 
-The offspringGenerator implements the 'how' part of the pairedPopulations.
-The mutation isn't part of the offspringGenerator because this is done by
-the member factory.
+\paragraph{Offspring generator} The offspringGenerator implements the `how'
+part of the pairedPopulations.  In there are three crossover functions
+uniform, single point and two point crossover.
 
-The fittest filter decides how which of both the parents and children 
-continues. The most simple implementation is the Killparents, which
-returns the children. Truncate elitism selects the best of both parents
-and children. Tournament matches child next to each parent
+\subparagraph{Uniform crossover}
+Uniform crossover checks character by character if they're equal
+If not then a random one or zero is used. This is a strange
+implementation, but its valid in this case because the only choice is
+between 1 and 0, unequal means that there exist in both the father and
+mother a one or zero. But not both. So we can just choose one.
+
+\subparagraph{1 point crossover}
+Generates a random number between zero and gene length. On the left side 
+the father is used and on the right side the mother.
+
+\subparagraph{2 point crossover}
+Two point crossover was initially implemented
+as first selecting a random value from 0 to gene length. The second would
+be from first to gene length. However this is a biased method in that
+on of the parents has more chance to be inherited. So
+later I would just generate two random values. Between 0 and gene length
+Then if the first that was generated is smaller then the second we use
+the father as a base string and the mother as the other sequence. Otherwise
+the other way around.
+
+However the unbiased method performs structurally worse as can bee seen
+in the experiment.
+
+\subparagraph{Mutation}
+Mutation isn't part of the offspringGenerator
+because this is done by the member factory. It was easier to implement
+in the member factory due to the architecture choices.
+
+\paragraph{Fittest filter} The fittest filter decides how which of both
+the parents and children continues.
+The most simple implementation is the \emph{Killparents}, which
+returns the children. \emph{Truncate elitism} selects the best of both parents
+and children. \emph{Tournament elitism} matches child next to each parent
 and decide which one is best. There is no kill children implementation
@@
 
-MateSelection implements tournament select. This is the only mate selection
-algorithm implemented currently.
+\paragraph{Mate selection} MateSelection implements tournament select.
+This is the only mate selection algorithm implemented currently.
 
-The member factories are able to create members from strings. These are either
-the tightly linked or randomly linked. Then the tightly or randomly linked
+\paragraph{Member factories} The member factories are able to create
+members from strings. These are either the tightly linked or randomly
+linked. Then the tightly or randomly linked
 functions can also be wrapped in a mutation function,
@@
 reserves big blocks of memory.
-
-\clearpage
-
-Most of the GA architecture was already in place before the assignment became
-available. I did this based on the notes I took in class. When the assignment
-became available I was almost completly done with the GA{.} There were some bugs
-in my code however and I interperted some things wrong. I also didn't write
-anything for an experiment. So I still needed to do that.
 
@@
 random tournament elitism.
-
 \subsection{Hypothesis}
-The deceptive trap function will not be solved by any algorithm.
-non-deceptive may be more successful. The others will be solved.
+The deceptive trap function will not be solved by any algorithm for
+the population sizes. I think this because its to desirable to have
+at least one 0000 block in the 100 string length.
+Non-deceptive will be more successful. The others will be solved.
 I also think mutation will have an important effect and may even
@@
 average. It will probably always find a result.
-
 \subsection{Execution}
-Note that I added some extra selection methods and crossing strategies to
+I added extra selection methods and crossing strategies to
 the final experiment. I did this because I was curious how they would
@@
 good or better. On both runtime and successes required population for success.
+
+\paragraph{Abreviations}
+\begin{tabular}{ll}
+	Abrivation & Full name \\ \toprule
+	2X & 2 Point cross \\
+	2XB & 2 Point cross biased \\
+	2XM & 2 Point cross mutated \\
+	2XM & 2 Point cross mutated \\
+	UX & uniform cross \\
+	UXM & uniform cross mutated \\
+	1X & 1 Point cross \\ \bottomrule
+\end{tabular}
+
+The biased two point cross will make 2 random values and use the lowest as
+the base crossing for the same parent. The ` normal' 2 point cross will make
+2 random values. If the first one generated is higher than the other one it
+will make the father the mother.
+
+It turns out however that the biased uniform cross is much more effective.
+As it solves the deciptive trap function and leaves the `normal' one in
+the dust. I actually excpected an improvement when changing the biased
+variant into the current vairant. To my shock it did major damage to the
+algoritm. To show this I decided to include both.
+
+\subsubsection{Graph observations}
+The major drops are caused by an optimization I implemented. After 2 failed
+run the stocastic ensurance function will stop. Often this happens when
+zooming in, this makes it appear as if there is a local gap, but its just
+random chanche and the fact that the bisectional search is zooming in.
 

### 2015-12-06 evolutionaryComputation/practicals/report.tex
added pargraph about mutation

@@
 \paragraph{Mate selection} MateSelection implements tournament select.
-This is the only mate selection algorithm implemented currently.
+This is the only mate selection algorithm implemented.
 
@@
 members from strings. These are either the tightly linked or randomly
-linked. Then the tightly or randomly linked
+linked. 
+\subparagraph{Mutation} Then the tightly or randomly linked
 functions can also be wrapped in a mutation function,

### 2015-12-07 evolutionaryComputation/notes4.tex
notes of lesson 5
notes from lecture 7?

@@
 
+\section{Time complexity}
+Convergence speed: easy problems easier to calculate
+
+population should be large enough, reasonably.
+
+Gentic programming uses extreme high selection pressure
+	creating lisp with GA's.
+
+Using this allows much better scalimg of the algortihm.
+
+\section{Population sizing}
+\subsection{Random walks}
+\[ P_N(x) = Probablity of reaching state N in postition x?\]
+Gamblers ruin
+
+\subsection{Exam}
+How does it work. How does it scale up?
+
+\section{Map labeling}
+Has a bunch of little local optimum features.
+
+all the little conditions were pulled from the fitness function
+and put in the local optimizer. This meant they didn't have to
+invent a bunch of weights.
+
+Don't put everything in a fitness function. Especially if you have
+to invent weights.
+
+Excact is better, but GA's tend to scale better.
+
+Try to keep the fitness function simple.
 \end{document}

### 2015-12-07 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Evolution.scala
finished assignment 1

@@ -42,9 +42,10 @@ class Evolution(
 			return prev
 		}
 		val children = step(parents)
+
 		if (children == parents) {
 			return prev
-			}
+		}
 		genetic(prev :+ children)
 	}
 }

### 2015-12-07 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -38,7 +38,6 @@ object Main{
 	val crossMethodsTight = Seq(
 		(twoPointCross(random,inverResult) _, tightlyLinked _, "2X"),
 		(twoPointCross(random,keepSelected) _, tightlyLinked _, "2XB"),
-		(twoPointCross(random,inverResult) _, WithCoinFlipTimesMutation(random,tightlyLinked), "2XM"),
 		(uniformCross _, tightlyLinked _, "UX"),
 		(uniformCross _, WithCoinFlipTimesMutation(random, tightlyLinked), " UXM"),
 		(singlePointCross(random) _, tightlyLinked _, "1X")
@@ -100,10 +99,6 @@ object Main{
 		log.info(toLatexWith(asciiGraph)(values))
 
 		write("result.tex", toLatexWith(svgGraph)(values))
-
-		val file = java.io.File.createTempFile("evolutionaryComputing", "dir")
-		file.delete()
-		file.createNewFile()
 	}
 	/**
 	 * write to file

### 2015-12-07 evolutionaryComputation/practicals/code/src/main/scala/nl/jappieklooster/hw/ec/PairedPopulation.scala

@@ -51,7 +51,7 @@ case class PairedPopulation(members:Seq[Pair], memberFactory:String => IMember) 
 						if(child == genesF){
 							return fumu.father
 						}
-						if(child == genesM){
+						if (child == genesM) {
 							return fumu.mother
 						}
 						memberFactory(child)

### 2015-12-08 evolutionaryComputation/1-practical/report.tex
split up the practical in 2 folders

--- 2015-12-06 evolutionaryComputation/practicals/report.tex
@@
 
+\label{paragraph:Fittest filter}
 \paragraph{Fittest filter} The fittest filter decides how which of both
@@
 linked.
+
 \subparagraph{Mutation} Then the tightly or randomly linked
@@
 Its an decorator idea, only with functions.
+\newpage
+
+\subsubsection{Experiment}
 
@@
 \emph{verifyLowest} will try to run a population size 30 times, if the faults
-limit is exceeded or the index exceeds the required runs limit it will quit.
-In the latter case the run is considered to be successful.
-The run will execute the GA search according to the configuration of the 
-experiment, so this function will call startGenetic, and put the results
-of the run into a RunResult. The RunResult will actually truncate a lot of
-information. Not the entire resulting population is stored. It wasn't
-necessary for the experiment and I was concerned about memory. Or rather
-I didn't want to be too wasteful, especially considering that the JVM
-reserves big blocks of memory.
+limit is exceeded or the index exceeds the required run limit it will quit.
+The run is considered to be successful if the count of successful run >= 
+run limit - faults.
 
-\subsubsection{Unit tests}
+The run will execute the \emph{Evolution} that has been passed into this
+experiment.  so this function will call startGenetic on the Evolution object, 
+and put the results of the run into a \emph{RunResult}.
+
+\subsection{Unit tests}
 Because the assignment has very well defined behavior. I thought it was
-a very wise Idea to use unit tests. To do this I use a library called
+a good Idea to use unit tests. To do this I use a library called
 Scala test.
@@
 I found about 6 different bugs alone in the valuation functions by using
-unit-testing, so I think this saved me considerable time.
-\subsubsection{Build tool}
+unit-testing, so I think this saved me considerable time. I also used this
+to precisely define the behavior of 2 point cross.
+
+By using unit tests I became very certain about the execution of my code.
+It does precisely what its supposed to do even though I use functional
+programming and it may look somewhat foreign.
+
+\subsection{Build tool}
 I use Gradle for building the project. Its the most advanced build platform
@@
 
-\subsubsection{Libraries used}
+\subsection{Libraries used}
 In here I will describe why I chose to use these libraries
-The details of the libraries can be found in the gradle file (which
+The details of the libraries can be found in the Gradle file (which
 version and what maven group):
@@
 I wanted my implementation to be fast, preferably faster than any other
-implementation, now I think I didn't succeed in this because I used a
-lot of slow mechanisms, such as functional combinators, and although
-they don't increase the computational complexity, they do certainly increase
-the minimum required runtime, I'm quite sure that every combinator in sequence
+implementation. I think I didn't succeed in this for the algorithm part
+because I used a lot of slow mechanisms, such as functional combinators.
+Although they don't increase the computational complexity,
+they do certainly increase
+the minimum required runtime. I'm quite sure that every combinator in sequence
 is implemented as just another loop. However because I
-run all experiments in parallel the total time may be faster.
+run all experiments in parallel the total time may be faster. So in a way
+I did succeed.
+
 \subsubsection{Parallelization}
@@
 gained.  After 10 minutes of waiting I decided to drop this idea
-since without parallelization it took about 2 minutes (I tried this
-on just one run, so it shouldn't take that long).
-
+since without parallelization it took about 2 minutes.
 The problem is that the evaluation function, even with gene length of 100, is
-to cheap to make any reasonable parallelization effective. However what was effictive
-is the parallelization of the experiments. This reduced total runtime
-significantly. However this is not seen in the statistics.
+to cheap to make any reasonable parallelization effective.
 
@@
 It also drastically decreases the amount of time I have to wait. Which is quite
-important, especially in the development phase. (although I also used
-other tricks, such as instead of 30, just doing 3 runs)
+important, especially in the development phase.
 
@@
 \subsubsection{Only create children if the genes are new}
-If the genes are new then create the children otherwise return the parent.
-This reduces evaluations and object creations. Its becomes more effective
-at deeper generation, it even slows down the fastest GA's by a few milliseconds
-but the deeper GA are lots faster.
+This optimization reduces the amount of evaluations required.  It becomes 
+more effective at deeper generations, however it slows down the fastest 
+GA's by a few milliseconds but the slower GA's  are quite a bit faster.
 
@@
 
-To reduce the amount of string comparisons, first the hash code function is called
-This will cache itself, I know this because I inspected the source code of the
-String class in Java.
-
-I was very proud for coming up with  this idea, but when I ran the experiment with
-and without hash code I came to the conclusion that first doing a hash code operation
-significantly *slows* execution. This becomes worse as population size increases.
+I also had an idea of using hash-codes to reduce the amount of comparisons.
+But it turns out that this increases the running time, so I quickly removed
+this optimization again.
 
 \section{Experiment}
-Something to note is that my made up tournament elitism is actually in most
-cases better, or at least equal to the truncate elitism. So when I saw
-this I tried using my other initially used algorithm where I would completely
-ignore the parents (kill the parents). However it appears this does not
-perform better.
+In here I will discuss how the experiments ran and the thoughts I
+above about the results.
+First the setup will be discussed, since I added to the
+original assignment out of curiosity. Then the hypothesis
+will be discussed. After that the results will be shown
+and finally I will discuss the conclusions I draw from the experiments.
 
-I also added two point cross with mutation and 1 point cross to see how
-those fair, running the code now takes about 40 minutes in total with the
-entire thing divided over 8 threads (quad core + hyper-threading). The only
-other improvement I could think of is running it on the GPU, but I think
-the algorithm is to complex for that and I don't know how to do that,
-especially considering I'm using the JVM.
+\subsection{Experiment titles}
+The experiment titles all follow the same structure:
 
-Two point cross and single cross both seem to find results in deceptive
-random tournament elitism.
-\subsection{Hypothesis}
-The deceptive trap function will not be solved by any algorithm for
-the population sizes. I think this because its to desirable to have
-at least one 0000 block in the 100 string length.
-Non-deceptive will be more successful. The others will be solved.
-I also think mutation will have an important effect and may even
-allow deceptive to be solved.
+\[F\mbox{ }L, E\]
+Where F is the fitness function, L is the linkage and E is the elitism type.
+The experiment results are grouped by this structure alphabetically.
+The elitism types are discussed in the fittest fileter paragraph \ref{paragraph:Fittest filter}
 
-I also think that the provided selection pressure would be about
-average. It will probably always find a result.
-\subsection{Execution}
-I added extra selection methods and crossing strategies to
-the final experiment. I did this because I was curious how they would
-compete with the original assignment.
+Block is the trap function.
 
-After doing some sample experiments I found for example that 2x was going
-up with number of successes in the biggest population size. So I wondered 
-if doing my original interpretation of the select would work better.
-I found that in almost every case this original interpretation is at least as
-good or better. On both runtime and successes required population for success.
-
-\paragraph{Abreviations}
+\subsection{Abbreviations}
 \begin{tabular}{ll}
@@
 
+\subsection{Setup}
+I added extra selection methods and crossing strategies to
+the final experiment. I did this because I was curious how they would
+compete with the original assignment.
+
+After doing some sample experiments I found for example that 2x was going
+up with number of successes in the biggest population size. So I wondered 
+if doing my original interpretation of the select would work better.
+I found that in almost every case this original interpretation is at least as
+good or better. On both runtime and successes required population for success.
+
 The biased two point cross will make 2 random values and use the lowest as
@@
 It turns out however that the biased uniform cross is much more effective.
-As it solves the deciptive trap function and leaves the `normal' one in
-the dust. I actually excpected an improvement when changing the biased
-variant into the current vairant. To my shock it did major damage to the
-algoritm. To show this I decided to include both.
+As it solves the deceptive trap function and leaves the `normal' one in
+the dust. I actually expected an improvement when changing the biased
+variant into the current variant. It was significantly worse however.
+To show this I decided to include both.
 
-\subsubsection{Graph observations}
-The major drops are caused by an optimization I implemented. After 2 failed
-run the stocastic ensurance function will stop. Often this happens when
-zooming in, this makes it appear as if there is a local gap, but its just
-random chanche and the fact that the bisectional search is zooming in.
+1 point cross was added to see how that competes with 2 point cross.
+I also used to have 2 point cross with
+mutation but it wasn't significantly different then the normal 2X
+and the graphs started to get cluttered so I decided to drop that in
+favour of discussing the biased crossover.
+
+Running the code now takes about 26 minutes in total with the
+entire thing divided over 8 threads (quad core + hyper-threading).
+
+\subsection{Hypothesis}
+The deceptive trap function will not be solved by any algorithm for
+the population sizes. I think this because its to desirable to have
+at least one 0000 block in the 100 string length.
+Non-deceptive will be more successful. The others will definitively be solved.
+I also think mutation will have an important effect and may even
+allow deceptive to be solved.
+
+The provided selection pressure would be about average.
+It will probably always find a result.
 
@@
 
+\subsubsection{Graph observations}
+The major drops are caused by an optimization I implemented. After 2 failed
+run the stochastic insurance function will stop.
+
+Often this happens when the bisectional search hones in on the optimum
+This makes it appear as if there is a local gap, but its just
+random chance.
+
 \section{Conclusion}
-%What did I learn from my observation?
 My conclusion is that using both tournament selection and truncate elitism
-is to harsh. It will slow down the GA and often not find a solution at all
-(in case of deceptive trap, 2x should've come up with something). the 
-time vs population found ratio is also much better for just using tournament
-select.
+is to harsh. It will slow down the GA and often not find a solution at all.
+This happens for example with the deceptive trap function, randomly and
+tightly linked.
+
+The time vs population found ratio is also much better for using 
+tournament select on both the elitism and selection.
 
@@
 the children (kill parents), then you still will often perform better than
-using tournament selection and truncate elitism, although not as 
-consistently as tournament elitism.
+using tournament selection and truncate elitism (this is how the original
+assignment was), although not as consistently as tournament elitism.
 
 To my surprise deceptive trap was solved with decreased selection pressure
-by both 2x and 1x, although 1x did worse. Also adding mutation doesn't do
-the algorithm much good. 2XM and UXM consistently find worse minimum populations
-than respectively 2X and UX{.}
-This probably means the mutation is too disruptive.
+by both 2x and 1x, although 1x did worse.
 
-Mutation turned out to have no significant effect other than just slowing
-things down.
+Mutation doesn't seem to have much effect UXM sometimes finds a better
+solution but usually this is not the case. I initially thought mutation
+would be very important but its just sort off random.  Mutation turned out
+to have no significant effect other than just slowing things down.
+
+Something to note is that my made up tournament elitism is actually in most
+cases better, or at least equal to the truncate elitism. So when I saw
+this I tried using my other initially used algorithm where I would completely
+ignore the parents (kill the parents). However it appears this does not
+perform better.
+
+Two point cross and single cross both seem to find results in deceptive
+random tournament elitism.
+
+Linearly scaled and uniformly scaled are very easy to solve for almost all
+algorithms. Although the 2x and 1x are at a big disadvantage compared to UX.
 

### 2015-12-08 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/EmulateSeq.scala

@@ -1,0 +1,27 @@
+// Evolutionary Computing Analyzer, it analyzes various genetic algoritms
+// Copyright (C) 2015 Jappie Klooster
+
+// This program is free software: you can redistribute it and/or modify
+// it under the terms of the GNU General Public License as published by
+// the Free Software Foundation, either version 3 of the License, or
+// (at your option) any later version.
+
+// This program is distributed in the hope that it will be useful,
+// but WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+// GNU General Public License for more details.
+
+// You should have received a copy of the GNU General Public License
+// along with this program.  If not, see <http://www.gnu.org/licenses/\>.
+package nl.jappieklooster.hw.ec
+
+/**
+ * a trait that passes all behavior onto an attribute (baseSeq)
+ * @tparam T
+ */
+trait EmulateSeq[+T] extends Seq[T] {
+	def baseSeq:Seq[T]
+	override def length: Int = baseSeq.length
+	override def iterator: Iterator[T] = baseSeq.iterator
+	override def apply(indx:Int):T = baseSeq.apply(indx)
+}

### 2015-12-08 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/Evaluation.scala

@@ -1,0 +1,61 @@
+// Evolutionary Computing Analyzer, it analyzes various genetic algoritms
+// Copyright (C) 2015 Jappie Klooster
+
+// This program is free software: you can redistribute it and/or modify
+// it under the terms of the GNU General Public License as published by
+// the Free Software Foundation, either version 3 of the License, or
+// (at your option) any later version.
+
+// This program is distributed in the hope that it will be useful,
+// but WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+// GNU General Public License for more details.
+
+// You should have received a copy of the GNU General Public License
+// along with this program.  If not, see <http://www.gnu.org/licenses/\>.
+package nl.jappieklooster.hw.ec
+
+trait IHasFitness{
+	def fitness:Float
+}
+trait FitnessEvaluator{
+	private var calls = 0
+	protected def getFunction:String => Float
+	def valuate(x:String):Float ={
+		calls += 1
+		getFunction(x)
+	}
+	def countCalls():Int = {
+		val result = calls
+		calls = 0
+		result
+	}
+}
+
+object Evaluation{
+
+	private def whereCharIsOne(c:Char) = c == '1'
+	def uniformlyScaledCountOnes(s:String):Float = s.count(whereCharIsOne)
+
+	def linearlyScaledCountOnes(s:String):Float = s.
+			// make a map like thing of it where the index is the key
+			zipWithIndex.
+			// select the elements where 1
+			filter(_._1 == '1').
+			// throw away the elements and keep the indeci
+			// also increase the index value by one
+			map(_._2+1).
+			// result
+			sum
+	def blockValuation(block:Seq[Float])(s:String):Float = {
+		val size = block.size-1
+		// make blocks of the string
+		s.grouped(size).map(
+			// get the value specified in the block
+			str => block(size - uniformlyScaledCountOnes(str).toInt)
+		).sum
+	}
+	def createProbe(f:String=>Float):FitnessEvaluator = new FitnessEvaluator {
+		protected override def getFunction: (String) => Float = f
+	}
+}

### 2015-12-08 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/Experiment.scala

@@ -1,0 +1,147 @@
+// Evolutionary Computing Analyzer, it analyzes various genetic algoritms
+// Copyright (C) 2015 Jappie Klooster
+
+// This program is free software: you can redistribute it and/or modify
+// it under the terms of the GNU General Public License as published by
+// the Free Software Foundation, either version 3 of the License, or
+// (at your option) any later version.
+
+// This program is distributed in the hope that it will be useful,
+// but WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+// GNU General Public License for more details.
+
+// You should have received a copy of the GNU General Public License
+// along with this program.  If not, see <http://www.gnu.org/licenses/\>.
+package nl.jappieklooster.hw.ec
+
+import com.itextpdf.text.log.LoggerFactory
+
+import scala.util.Random
+/**
+ * This expirement tries to find the optemal number of population for a specific
+ * evolution.
+ * @param evolution
+ */
+class Experiment(val name:String, val variation:String, evolution: Evolution, memberFactory:String => IMember) {
+	import Experiment._
+	val log = LoggerFactory.getLogger(this.getClass)
+	private def createPopulation(size:Int) = {
+		Population.createOneZeros(memberFactory, geneLength, size)
+	}
+	def run(consideringSize:Int): RunResult={
+		val population = createPopulation(consideringSize)
+
+		val startTime = System.currentTimeMillis()
+		val evolutionResult = evolution.startGenetic(population)
+		val runtime = System.currentTimeMillis() - startTime
+
+		// sucesfull if early termination or the last population is good enough
+		val success = hasGoodEnoughSolution(evolutionResult.last)
+		RunResult(consideringSize, success, runtime, evolutionResult.length,evolution.evaluation.countCalls())
+	}
+
+	def verifyLowest(currentPop:Int, faults:Int, index:Int) : Seq[RunResult] = {
+		if(index >= requiredRuns){
+			return Nil
+		}
+		if(faults > faultTolerance){
+			return Nil
+		}
+		val exp = run(currentPop)
+		val newFaults = if (exp.success) faults else faults+1
[…]

### 2015-12-08 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/FittestFilter.scala

@@ -1,0 +1,49 @@
+// Evolutionary Computing Analyzer, it analyzes various genetic algoritms
+// Copyright (C) 2015 Jappie Klooster
+
+// This program is free software: you can redistribute it and/or modify
+// it under the terms of the GNU General Public License as published by
+// the Free Software Foundation, either version 3 of the License, or
+// (at your option) any later version.
+
+// This program is distributed in the hope that it will be useful,
+// but WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+// GNU General Public License for more details.
+
+// You should have received a copy of the GNU General Public License
+// along with this program.  If not, see <http://www.gnu.org/licenses/\>.
+package nl.jappieklooster.hw.ec
+
+object FittestFilter {
+	/** drop the parents */
+	def killParents(
+		parents : Population,
+		children : Population
+	) : Population = {
+		children
+	}
+
+	def truncateElitism(
+			parents : Population,
+			children : Population
+			) : Population = {
+		val combined = parents ++ children
+		// end up sorting anyways, no idea why you would do tournement selection
+		val sorted = combined.sortWith((x, y) => x.fitness > y.fitness)
+		parents.copy(members = sorted.take(parents.length))
+	}
+
+	def tournementElitism(
+		parents : Population,
+		children : Population
+	): Population={
+		val combined = parents.members.zip(children.members)
+		Population(
+			combined.map(x=> MateSelection.pickBest(x._1, x._2)),
+			parents.memberFactory
+		)
+	}
+
+
+}

### 2015-12-08 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -1,0 +1,211 @@
+// Evolutionary Computing Analyzer, it analyzes various genetic algoritms
+// Copyright (C) 2015 Jappie Klooster
+
+// This program is free software: you can redistribute it and/or modify
+// it under the terms of the GNU General Public License as published by
+// the Free Software Foundation, either version 3 of the License, or
+// (at your option) any later version.
+
+// This program is distributed in the hope that it will be useful,
+// but WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+// GNU General Public License for more details.
+
+// You should have received a copy of the GNU General Public License
+// along with this program.  If not, see <http://www.gnu.org/licenses/\>.
+package nl.jappieklooster.hw.ec
+
+import java.io.{File, PrintWriter}
+
+import org.sameersingh.scalaplot._
+import org.slf4j.LoggerFactory
+
+import scala.reflect.io.Path
+import scala.util.Random
+import OffspringGenerator._
+import MemberFactories._
+import Evaluation._
+import scala.util.Properties.lineSeparator
+
+object Main{
+	case class Configurations(strLength:Int,popSize:Int,genCount:Int, evolution:Evolution)
+	val log = LoggerFactory.getLogger(Main.getClass)
+	val random = Random
+	val br = lineSeparator
+	val folder = "../"
+	type GraphData = (String,Seq[Seq[(Double,Double)]])
+
+	val crossMethodsTight = Seq(
+		(twoPointCross(random,inverResult) _, tightlyLinked _, "2X"),
+		(twoPointCross(random,keepSelected) _, tightlyLinked _, "2XB"),
+		(uniformCross _, tightlyLinked _, "UX"),
+		(uniformCross _, WithCoinFlipTimesMutation(random, tightlyLinked), " UXM"),
+		(singlePointCross(random) _, tightlyLinked _, "1X")
+	)
+	def main(args:Array[String]){
+		log.info("starting ea")
+
+		
+		val crossMethodsRandom = crossMethodsTight.map(a=>{
+				val method = a._2 match{
+					case WithCoinFlipTimesMutation(_,_) => {
+						log.trace(s"${a._3} is flipped")
+						WithCoinFlipTimesMutation(random, createRandomlyLinked)
+					}
[…]

### 2015-12-08 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/Population.scala

@@ -1,0 +1,39 @@
+// Evolutionary Computing Analyzer, it analyzes various genetic algoritms
+// Copyright (C) 2015 Jappie Klooster
+
+// This program is free software: you can redistribute it and/or modify
+// it under the terms of the GNU General Public License as published by
+// the Free Software Foundation, either version 3 of the License, or
+// (at your option) any later version.
+
+// This program is distributed in the hope that it will be useful,
+// but WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+// GNU General Public License for more details.
+
+// You should have received a copy of the GNU General Public License
+// along with this program.  If not, see <http://www.gnu.org/licenses/\>.
+package nl.jappieklooster.hw.ec
+
+import scala.collection.{SeqLike, GenSeq}
+import scala.collection.parallel.{SeqSplitter, ParSeq}
+import scala.util.Random
+
+case class Population(members:Seq[IMember], memberFactory:String => IMember) extends EmulateSeq[IMember]{
+	override def baseSeq: Seq[IMember] = members
+
+}
+object Population{
+	def createOneZeros(
+			memberFactory:String => IMember,
+			geneLength:Int,
+			popCount:Int):Population =  Population(
+		0.to(popCount-1).par.map(
+			x => memberFactory(
+				// it says, from 1 to length concat 1 or 0.
+				0.to(geneLength).foldLeft(""){(b, x) => b + OffspringGenerator.randomGene}
+			)
+		).seq,
+		memberFactory
+	)
+}

### 2015-12-08 evolutionaryComputation/2-practical/code/src/test/scala/nl/jappieklooster/hw/ec/tests/CrossoverTests.scala

@@ -1,0 +1,34 @@
+package nl.jappieklooster.hw.ec.tests
+
+import nl.jappieklooster.hw.ec.OffspringGenerator._
+import org.scalatest.{Matchers, FlatSpec}
+
+import scala.util.Random
+
+class CrossoverTests extends FlatSpec with Matchers{
+	def invertCross(a:IndexedSeq[(Char,Char)],b:Int,c:Int) = twoPointCross(inverResult, a,b,c).mkString
+	"2 point cross " should " work unbiased" in{
+		val parents = "abcdefg".zip("1234567")
+		// normal crosses...
+		invertCross(parents,1,3)should be("a23defg")
+		invertCross(parents,0,3)should be("123defg")
+
+		// when the random generates the first number higher then the second
+		// we excpect the thing to become reversered.
+		invertCross(parents,3,1)should be("1bc4567")
+		invertCross(parents,3,0)should be("abc4567")
+	}
+
+	def biasedCross(a:IndexedSeq[(Char,Char)],b:Int,c:Int) = twoPointCross(keepSelected, a,b,c).mkString
+	"2 point cross " should " work biased" in{
+		val parents = "abcdefg".zip("1234567")
+		// normal crosses...
+		biasedCross(parents,1,3)should be("a23defg")
+		biasedCross(parents,0,3)should be("123defg")
+
+		// when the random generates the first number higher then the second
+		// we excpect the thing to become reversered.
+		biasedCross(parents,3,1)should be("a23defg")
+		biasedCross(parents,3,0)should be("123defg")
+	}
+}

### 2015-12-08 evolutionaryComputation/2-practical/code/src/test/scala/nl/jappieklooster/hw/ec/tests/ExpirimentTests.scala

@@ -1,0 +1,17 @@
+package nl.jappieklooster.hw.ec.tests
+
+import nl.jappieklooster.hw.ec.{Member, Population}
+import org.scalatest.{Matchers, FlatSpec}
+import nl.jappieklooster.hw.ec.Experiment._
+
+class ExpirimentTests extends FlatSpec with Matchers{
+
+	def createMember(str:Seq[String]) = Population(str.map(x=>Member(1,x)), null)
+	"has good enough result " should " not be true " in {
+		hasGoodEnoughSolution(createMember(Seq("0111", "1010", "1010", "0000", "1000"))) should be (false)
+	}
+	"has good enough result " should " be true " in {
+		hasGoodEnoughSolution(createMember(Seq("0111", "1010", "111111", "1010", "0000", "1000"))) should be (true)
+	}
+
+}

### 2015-12-08 evolutionaryComputation/2-practical/code/src/test/scala/nl/jappieklooster/hw/ec/tests/FitnessTests.scala

@@ -1,0 +1,59 @@
+package nl.jappieklooster.hw.ec.tests
+
+import nl.jappieklooster.hw.ec.IHasFitness
+import nl.jappieklooster.hw.ec.Evaluation._
+import org.scalatest.{Matchers, FlatSpec}
+
+class FitnessTests extends FlatSpec with Matchers{
+
+	/**
+	 * create a known subject which we can count manually
+	 */
+	val subject =
+		"1111" + "0111" + "1010" + "1000" + "0000"
+		//		   1234     5678     9012     3
+		// so linearly scaled = 1+2+3+4 + 6+7+8 + 9+ 11+ 13 = 64
+		// decept: 4	+	0	+ 	1	+	2	+ 3 = 10
+		// non: 	4	+	0	+ 	0.5	+	1	+ 1.5 = 7
+
+
+	"Linearly counting fitness " should "count subject as 10" in {
+		uniformlyScaledCountOnes(subject) should be (10)
+	}
+	"Linearly scaled fitness " should "count subject as 64" in {
+		linearlyScaledCountOnes(subject) should be (64)
+	}
+
+	val deceptive = blockValuation(List(4f,0f,1f,2f,3f)) _
+	"deceptive trap fitness" should "count subject as 10" in {
+		deceptive(subject) should be (10)
+	}
+	"decipte trap fitness " should "count these infividual elements correctly " in{
+		/*
+		* I suepected something was wrong with the trap function because all trap
+		* GA's failed to find solution wich didn't make much sense (wouldn't learn
+		* much from a function that doesn't work).
+		* These tests proove that
+		 */
+		deceptive("1111") should be (4)
+		deceptive("0111") should be (0)
+		deceptive("0011") should be (1f)
+		deceptive("0001") should be (2f)
+		deceptive("0000") should be (3f)
+	}
+	"trap function" should "have no trouble with wrongly block sized input " in {
+		deceptive("1111" + "0111" + "1010" + "1000" + "00") should be (10)  // because it should count the ones
+	}
+	val nondecept = blockValuation(List(4f, 0f, 0.5f, 1f, 1.5f)) _
+	"nondeceptive trap fitness" should "count subject as 7" in {
+		nondecept(subject) should be (7)
+	}
+	"nondecipte trap fitness " should "count these infividual elements correctly " in{
+		nondecept("1111") should be (4)
+		nondecept("1110") should be (0)
[…]

### 2015-12-08 evolutionaryComputation/2-practical/code/src/test/scala/nl/jappieklooster/hw/ec/tests/MateSelectionTests.scala

@@ -1,0 +1,47 @@
+package nl.jappieklooster.hw.ec.tests
+
+import nl.jappieklooster.hw.ec._
+import org.scalatest.{Matchers, FlatSpec}
+import org.slf4j.LoggerFactory
+import util.Random
+
+class MateSelectionTests extends FlatSpec with Matchers{
+	val log = LoggerFactory.getLogger(this.getClass)
+	// rig random so we know what will happen
+	def rand = new Random(5)
+	// again we rig the show to make sure the right things happen
+
+	def createMember(s:String) = Member(5,s)
+	def createMember(i:Int, s:String) = Member(i,s)
+	def createSubject = new CompeteWithRandomTournement(rand)
+	"the tournament function when rigged" should "always let the rihgt side win " in {
+		val one = createMember("1")
+		val two = createMember("2")
+		MateSelection.pickBest(one,two) should be (two)
+	}
+	"the random turnement when rigged " should "produce the following sequent " in {
+		val subject = createSubject
+		val one = createMember("1")
+		val two = createMember("2")
+		val three = createMember("3")
+		val four = createMember("4")
+		// with the rigging of this rand, the shuffle will produce List(4,1,2,3)
+		// as can been seen from the other test the right side will always win.
+		// so expected is:
+		val expected = PairedPopulation(List(Pair(two, one), Pair(four, three)), null)
+		subject.selectFrom(Population(List(one,two,three,four), null)) should be (expected)
+	}
+	"the random turnement when rigged " should "produce the best two as winners " in {
+		val subject = createSubject
+		val one = createMember(4, "a")
+		val two = createMember(5, "b")
+		val three = createMember(8, "c")
+		val four = createMember(3, "d")
+		// with the rigging of this rand, the shuffle will produce List(4,1,2,3)
+		// as can been seen from the other test the right side will always win.
+		// so expected is:
+		val expected = PairedPopulation(List(Pair(two, one), Pair(three, three)), null)
+		subject.selectFrom(Population(List(one,two,three,four), null)) should be (expected)
+	}
+
+}

### 2015-12-08 evolutionaryComputation/2-practical/report.tex
name of the report

--- 2015-12-08 evolutionaryComputation/1-practical/report.tex
@@
 \author{Jappie Klooster}
-\title{A practical look at genetic algorithms}
+\title{Graph bipartitioning and you}
 \maketitle
@@
 \section{Introduction}
-This is a report of the evolutionary practical assignment 1.
+This is a report of the evolutionary practical assignment 2.
 In here I will discuss the execution of the assignment, what it was, which
@@
 \section{Assignment}
-The assignment was to implement the following functions and algorithms and
-then do experimentation with help of these.
-
-\subsection{Functions}
-\subsubsection{Counting functions}
-\[x_i \in \{0,1\}:CO(x_1\dots x_l)=\sum^l_{i=1}x_i\]
-\[x_i \in \{0,1\}:CO(x_1\dots x_l)=\sum^l_{i=1}i x_i\]
-\subsubsection{Trap function}
-\[TF(x_1\dots x_l) = \sum^{\frac{l}{k}-1}_{j=0}B(x_{jk_1}\dots x_{jk+k})\]
-
-with:
-
-\[B(x_1\dots x_k)=\left\{
-	\begin{matrix}
-		k & \mbox{ if } &   CO(x_1\dots x_k) = k \\
-		k - d - \frac{k-d}{k-1} CO(x) & \mbox{ if } & CO(x_1\dots x_k) < k \\
-	\end{matrix}
-\right. \]
-
-\subsection{Experiment}
-After implementing these functions we were tasked to make an experiment of it
-and make some graphs of the results and make a conclusion based on the results.
+Graphbipartitioning is dividing a graph into 2, with as little edges between
+both graphs as possible.
 

### 2015-12-08 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/Experiment.scala
deleted all unused code

@@ -23,22 +23,11 @@ import scala.util.Random
  * evolution.
  * @param evolution
  */
-class Experiment(val name:String, val variation:String, evolution: Evolution, memberFactory:String => IMember) {
+class Experiment(val name:String) {
 	import Experiment._
 	val log = LoggerFactory.getLogger(this.getClass)
-	private def createPopulation(size:Int) = {
-		Population.createOneZeros(memberFactory, geneLength, size)
-	}
 	def run(consideringSize:Int): RunResult={
-		val population = createPopulation(consideringSize)
-
-		val startTime = System.currentTimeMillis()
-		val evolutionResult = evolution.startGenetic(population)
-		val runtime = System.currentTimeMillis() - startTime
-
-		// sucesfull if early termination or the last population is good enough
-		val success = hasGoodEnoughSolution(evolutionResult.last)
-		RunResult(consideringSize, success, runtime, evolutionResult.length,evolution.evaluation.countCalls())
+		RunResult(consideringSize,false, 4,4,4)
 	}
 
 	def verifyLowest(currentPop:Int, faults:Int, index:Int) : Seq[RunResult] = {
@@ -117,31 +106,6 @@ object Experiment{
 	val faultTolerance = 1
 
 	def create(
-		name:String,
-		random:Random,
-		valuationFunction:String=>Float,
-		variationOperators:Seq[(PairedPopulation => Population, (String=>Float) => String => IMember, String)],
-		filter:(Population, Population) => Population = FittestFilter.truncateElitism
-	):Seq[Experiment] = variationOperators.map(
-		variation => {
-			val probe = Evaluation.createProbe(valuationFunction)
-		 new Experiment(
-			name,
-			variation._3,
-			new Evolution(
-				probe,
-				MateSelection.createCompeteWithRandomTournement(random),
-				variation._1,
-				filter,
-				hasGoodEnoughSolution
-			),
-			variation._2(probe.valuate)
-			)
-		}
-	)
-	def hasGoodEnoughSolution(population: Population):Boolean = {
-		// if there exists a member in a population that only contains 1's.
-		population.exists(member => !member.genes.exists(c => c == '0'))
-	}
-
+		name:String
+	):Seq[Experiment] = Seq(new Experiment(name))
 }

### 2015-12-08 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/FittestFilter.scala

@@ -1,49 +1,0 @@
-// Evolutionary Computing Analyzer, it analyzes various genetic algoritms
-// Copyright (C) 2015 Jappie Klooster
-
-// This program is free software: you can redistribute it and/or modify
-// it under the terms of the GNU General Public License as published by
-// the Free Software Foundation, either version 3 of the License, or
-// (at your option) any later version.
-
-// This program is distributed in the hope that it will be useful,
-// but WITHOUT ANY WARRANTY; without even the implied warranty of
-// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-// GNU General Public License for more details.
-
-// You should have received a copy of the GNU General Public License
-// along with this program.  If not, see <http://www.gnu.org/licenses/\>.
-package nl.jappieklooster.hw.ec
-
-object FittestFilter {
-	/** drop the parents */
-	def killParents(
-		parents : Population,
-		children : Population
-	) : Population = {
-		children
-	}
-
-	def truncateElitism(
-			parents : Population,
-			children : Population
-			) : Population = {
-		val combined = parents ++ children
-		// end up sorting anyways, no idea why you would do tournement selection
-		val sorted = combined.sortWith((x, y) => x.fitness > y.fitness)
-		parents.copy(members = sorted.take(parents.length))
-	}
-
-	def tournementElitism(
-		parents : Population,
-		children : Population
-	): Population={
-		val combined = parents.members.zip(children.members)
-		Population(
-			combined.map(x=> MateSelection.pickBest(x._1, x._2)),
-			parents.memberFactory
-		)
-	}
-
-
-}

### 2015-12-08 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala
now can read the file properly

@@ -35,8 +35,9 @@ object Main{
 		log.info("starting graph bipartitioning")
 		val verteci = Source.fromFile("src/main/resources/data.txt").getLines().zipWithIndex.map((elem) => {
 			val id = elem._2 + 1
-			val line = elem._1
-			Vertex(id, xy._1,xy._2,Nil)
+
+			val connections:Seq[Int] = if(elem._1.length > 0) elem._1.split(" ").map(_.toInt) else Nil
+			Vertex(id, connections)
 		})
 		for (vertex <- verteci){
 			log.info(vertex.toString)

### 2015-12-08 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/Vertex.scala

@@ -1,7 +1,6 @@
 package nl.jappieklooster.hw.ec
 
 case class Vertex(id:Int, connections:Seq[Int]) {
-
 }
 
 case class Graph(verteci:Array[Vertex]){

### 2015-12-08 evolutionaryComputation/2-practical/report.tex
removed a lot of text from the previous

@@
 \section{Methodology}
+I coppied all the code from the previous practical into this new folder.
+Then I went onto a delete rage. Most code was deleted except the Expirment
+class, this will probably still be usefull.
+
+I use git, so it doens't really matter to accidently delete to much. After
+that I changed the data, I removed everything which was useless. Now the file
+only contains the connection numbers. This makes parsing the file a lot
+easier. I don't want to do advanced pattern matching when I can quickly modfy
+the file with a few vim macros. (in this case I just did a `30x').
 
@@
 
-\subsection{Speed up attempts}
-I wanted my implementation to be fast, preferably faster than any other
-implementation. I think I didn't succeed in this for the algorithm part
-because I used a lot of slow mechanisms, such as functional combinators.
-Although they don't increase the computational complexity,
-they do certainly increase
-the minimum required runtime. I'm quite sure that every combinator in sequence
-is implemented as just another loop. However because I
-run all experiments in parallel the total time may be faster. So in a way
-I did succeed.
-
-\subsubsection{Parallelization}
-Since were doing string comparisons I had the idea of using parallelization to
-speed up the algorithm by doing parallelization. This can be done very easily
-in Scala with par collections (and the fact that immutability is default).
-
-However once I implemented this, I came to the conclusion it was not faster.
-The overhead for delegating and creating threads was more than the parallelization
-gained.  After 10 minutes of waiting I decided to drop this idea
-since without parallelization it took about 2 minutes.
-The problem is that the evaluation function, even with gene length of 100, is
-to cheap to make any reasonable parallelization effective.
-
-So now all experiments are run at the same time. This probably increases the
-average time of each `run' somewhat (not that much because the concurrent
-collection does \href{http://docs.scala-lang.org/overviews/parallel-collections/configuration.html}{load balancing}).
-It also drastically decreases the amount of time I have to wait. Which is quite
-important, especially in the development phase.
-
-The parallelization of experiments can be removed by deleting `par' on line 64
-of Main.scala
-
-\subsubsection{Evaluate each member on creation}
-Initially my members just contained the string with the `genes'. This I passed
-around and would evaluate when needed. However I realized that I could do this
-evaluation eagerly on creation, to reduce the evaluation calls. A member has
-to be evaluated anyways, so doing this on creation and just putting in the
-evaluation saves some heavy evaluation operations.
-
-\subsubsection{Only create children if the genes are new}
-This optimization reduces the amount of evaluations required.  It becomes 
-more effective at deeper generations, however it slows down the fastest 
-GA's by a few milliseconds but the slower GA's  are quite a bit faster.
-
-This can be done safely because strings are immutable and so are the members.
-So I can reuse them whenever its needed.
-
-I also had an idea of using hash-codes to reduce the amount of comparisons.
-But it turns out that this increases the running time, so I quickly removed
-this optimization again.
-
 \section{Experiment}

### 2015-12-17 evolutionaryComputation/notes4.tex
added some lecture notes

@@
 Try to keep the fitness function simple.
+
+\section{Model GA}
+Learn structure from good solutions.
+Is very broad applicable because few assumptions.
+
+Probability model is a vector(1,0,0.5)
+Where 0.5 = random,
+0 = 100 percent certainty of 0
+1 = 100 percent certainty of 1
+
+Overfitting, you don't learn anything because you
+made it perfect.
+
+deception is the problem, not non lineararity.
+
+\section{PMGA}
+Modelling baysion networks you should look at the fitting of data *and* model
+complexity
+
+Foss can be anything, for example:
+  3456
+0110110
+1234
+
+with the first and alst row being indeci of subsets.
+
+univariate is just one big set.
+marginal product = Mutex set
+Hiercial tree, multiple level of linkage, the bottom no linkage the top 
+uniform and marginal product in between in varying levels.
+
+mutual information is super important concept\ldots
+
 \end{document}

### 2016-01-07 evolutionaryComputation/2-practical/report.tex
implemented mls, and treid ILS, but I think I'm doing ti wrong since the results are quite a lot worse then MLS

@@
 
-\subsection{Language choice}
-I chose to use Scala to implement this project. The main reason being that I've
-played with Scala before but never have done anything constructive with it.
-
-Scala is both a functional and an object orientated program that runs on the JVM
-It compiles to byte-code and is statically typed.
-
-It also can use any Java library and integrates seamlessly with Java, (although
-calling methods the methods of a Scala project from a Java project can be
-a little clumsy sometimes, because Scala supports operator overloading in
-a strange way, but its possible).
-
-Other options I considered were:
-\begin{itemize}
-	\item Java, Java is an industry standard language, but the syntax of
-		Java is very verbose. So when I have the choice I try to avoid it.
-		I've also used a lot of Java when I was younger so I won't learn
-		much by using it.
-	\item C\#, the language is good, a lot less verbose then Java
-		but setting it up is difficult since I use Linux.
-		It also prefers object orientated rather than
-		functional programming. Scala's type system is also more advanced than
-		C\#'s, so its in a way more object orientated.
-	\item Rust/Go, I'm curious about both these languages, but I'm not at
-		all familiar with them, so I probably should do first a little
-		hobby project with them before starting doing something serious like
-		this, this is in contrast to Scala, with which I already played before.
-	\item Python, I think the project is in that grey area where weakly
-		typed languages become less powerful in use. Especially considering
-		the fact how quickly these algorithms blow up in runtime. Static type
-		checking then helps solving a big part of the bugs before execution
-		and since execution will be slow, this saves a lot of time.
-	\item C++, C++ is extremely fast in execution times. However implementation
-		time is quite long compared to almost everything else.
-		Errors are for example reported poorly and
-		the syntax is inconsistent. Since execution speed isn't a top priority
-		I decided to not use C++.
-\end{itemize}
-
-\subsection{Architecture}
-The code is split up into 4 parts, first there are the case classes. These are
-immutable classes that hold the domain information (Java beans).
-Then there are the algorithm classes and objects. These are pretty straight
-forward logical structures. They implement the atoms of possible GA's, such
-as for example uniform crossover. After that comes the experiment and the result
-classes, the experiment just uses the GA in a specific way.
-Finally there is the presentation object \emph{Main}. The only thing it does
-is interpret the experiment into a latex structure. There is no
-diagram discussing Main, because there is little architecture involved.
-
-\subsubsection{Case classes}
-The case classes add meaning to the structures.
-A case class is a Scala concept were you just define the attributes. Then
-Scala generates the `equals', `hashCode' and `toString' methods for you.
-They're also by default immutable, which is something I like using. You
-can make mutations by using the generated copy function, in which you can
-specify what is changed to create a new copy with the change.
-
-\begin{figure}[ht!]
-\centering
-\begin{emp}[classdiag](20,20)
-Class.mem("Member")("+Fitness:Int", "+genes:String")();
-Class.pop("Population")("+members:Seq[Member]", "+factory:String=>Member")();
-Class.pairedPop("PairedPopulation")("+members:Seq[Pair]","+factory:String=>Member")(
-"+pairedToPop(how):Population");
-Class.pai("Pair")("+father:Member","+mother:Member")();
-
-Instance.population("Population")(
-"createOnesAndZeros(factory:String=>Member,popCount:Int,geneLength:Int):Population"
-);
-
-leftToRight.top(20)(pop,pairedPop);
-topToBottom.left(20)(population, pop);
-topToBottom.left(60)(pop,mem);
-leftToRight.top(80)(mem, pai);
-
-drawObjects(mem,pop,pairedPop,pai,population);
-link(association)(pai.n -- pairedPop.s);
-item(iAssoc)("*")(obj.se = pai.n);
-link(association)(mem.e -- pai.w);
-item(iAssoc)("2")(obj.sw = mem.e);
-
-link(association)(mem.n -- pop.s);
-item(iAssoc)("*")(obj.se = mem.n);
-\end{emp}
-\caption{Case classes}
-\end{figure}
-
-\paragraph{Member} A \emph{Member} is something with a gene. The fitness value gets pre calculated by
-the factory, so that every members' fitness is only calculated once. This can
-be done reliably because the member is immutable. There is no way to modify
-the genes so the fintness value does not have to change.
-
-\paragraph{Population} The members are generated in batches and stored into a
-\emph{Population}, either by the paired
-population or by the Population companion object which provides a
-`createOnesAndZeros' function.
-A population will also have a function with it to generate new members as
-data in this case, it drags the specified function along, its like a
-function pointer.
-The population class doesn't use it itself, but when its members are paired
-up into \emph{Pair}s. A \emph{PairedPopulation} can use the creation function
-of the original population to create new members.
-
-\paragraph{PairedPopulation}
-The \emph{PairedPopulation} has also a utility function pairedToPop,
-this receives a how, in which it can
-insert zipped genes (char by char paired up strings) and from which comes
-a char sequence. This how is implemented by one of the crossing methods.
-This abstraction was added because the code around the cross function was
-always the same.
-The resulting string from `how' is then inserted into a member
-factory to create a new member (the member factory then allows mutation
-or obfuscation of the new genes).
-
-I've thought about merging PairedPopulation and population into
-one class. However doing that will probably not make things more obvious.
-Although it does makes sense in that the population would actually use
-the function it drags along.
-
-\newpage
-\subsubsection{Algorithm}
-The objects defined above all have functions with the types
-same named evolution attributes. So the functions are just basically
-options you can use to create an
-evolution from. There are some exceptions which allow wrapping is the
-withCoinFlipMutation function in
-memberFactories. This function acts as a decorating function. So you can wrap
-it around another function to create a new function.
-
-Objects are just singletons, or Java static space. I use the object keyword in
-Scala instead of class So I chose also to use an object in the UML
-representation. Scala seems to stick pretty close to UML notation anyways.
-The objects don't have types however, because they are the types. They're
-language enforced singletons.
-
-\begin{figure}[ht!]
-\centering
-\begin{emp}[classdiag](20,20)
-Class.evo("Evolution")(
-"+evaluation: FitnessEvaluator",
-"mateSelector:IMateSelector",
-"offSpringGenerator: PairedPopulation => Population",
-"fittestFilter: (Population, Population) => Population",
-"hasGoodEnoughSolution: Population => Boolean")(
-"startGenetic(seed:Population): Seq[Population]",
-"genetic(prev:Seq[Population]): Seq[Population]",
-"step(parents:Population): Population",
-);
-Instance.fit("Evaluation")(
-	"uniformlyScaled()",
-	"linearlyScaled()",
-	"blockValuation(block:Seq[Float])",
-	"createProbe(function):FitnessEvaluator"
-);
-Instance.filt("FittestFilter")(
-	"killParents()",
-	"truncateElitism()",
-	"tournementElitism()"
-	);
-Instance.select("MateSelection")(
-	"tournementSelect()"
-);
-Instance.factor("MemberFactories")(
-	"tightlyLinkded()",
-	"randomlyLinked()",
-	"withCoinflipMutation(function):function",
-);
-Instance.off("OffspringGenerator")(
-	"uniformCross()",
-	"singlePointCross()",
-	"twoPointCross()"
-);
-
-topToBottom.left(20)(evo,fit,filt,factor);
-leftToRight.top(20)(fit,off);
-leftToRight.top(20)(filt,select);
-
-drawObjects(evo,fit,factor,filt,select,off);
-\end{emp}
-\caption{Algorithm}
-\end{figure}
-
-\paragraph{Evolution}The evolution class is the common structure of a GA(Gentic algorithm){.} 
-So creating an instance of Evolution is like creating a new GA{.}
-
-To construct an Evolution you need to pass a \emph{valuation function},
-a \emph{mateselector}, An \emph{offspring generator},
-a \emph{fittestFilter}. And a function that determines if there
-is a good enough solution in the current population. The last one was added
-because I wasn't quite sure when the GA needed to stop in the beginning.
-
-Some of these functions are abstracted away by traits, but still can be seen
-as just fucntions.
-\subparagraph{Step method}
-The step method calls almost all given functions in the follwoing order:
-
-\begin{enumerate}
-	\item Select mates.
-	\item Generate offspring.
-	\item Select the fittest.
-\end{enumerate}
-
-\subparagraph{Genetic} hasGoodEnoughSolution is called in genetic,
-which will continue calling steps until the population has a good enough
-solution, or until parents and children are equal.
-The population is a case class, so the generated comparison method is used
-to check if the children and parents are equal.
-The result of this function is a sequance of population. The first element
-is te first generation and the last is the last generation.
-
-\subparagraph{Start genetic} Start genetic makes a seq of the seed population and
-calls genetic.
-
-\paragraph{Evaluation} The evaluation object
-contains the evaluation functions. The assignment asked to define four.
-However the block evaluation (trap functions) can be implemented as the
-same function with different values for the indices. This is done by
-blockvaluation.
-
-The create probe function is a utility function that allows the counting of
-the fitness calls. The initial design didn't have a \emph{FitnessEvaluator},
-but the assignment desired the counting of calls so I just wrapped the
-function in a object that increases the count on every call. I had to
-replace the function prototype in evolution with a trait because coutning
-the calls of a function requires some form of state (an increasing variable)
-I didn't know how to do that properly with just functions in scala yet,
-so I just used a trait instead (trait = interface + state).
-
-\paragraph{Offspring generator} The offspringGenerator implements the `how'
-part of the pairedPopulations.  In there are three crossover functions
-uniform, single point and two point crossover.
-
-\subparagraph{Uniform crossover}
-Uniform crossover checks character by character if they're equal
-If not then a random one or zero is used. This is a strange
-implementation, but its valid in this case because the only choice is
-between 1 and 0, unequal means that there exist in both the father and
-mother a one or zero. But not both. So we can just choose one.
-
-\subparagraph{1 point crossover}
-Generates a random number between zero and gene length. On the left side 
-the father is used and on the right side the mother.
-
-\subparagraph{2 point crossover}
-Two point crossover was initially implemented
-as first selecting a random value from 0 to gene length. The second would
-be from first to gene length. However this is a biased method in that
-on of the parents has more chance to be inherited. So
-later I would just generate two random values. Between 0 and gene length
-Then if the first that was generated is smaller then the second we use
-the father as a base string and the mother as the other sequence. Otherwise
-the other way around.
-
-However the unbiased method performs structurally worse as can bee seen
-in the experiment.
-
-\subparagraph{Mutation}
-Mutation isn't part of the offspringGenerator
-because this is done by the member factory. It was easier to implement
-in the member factory due to the architecture choices.
-
-\label{paragraph:Fittest filter}
-\paragraph{Fittest filter} The fittest filter decides how which of both
-the parents and children continues.
-The most simple implementation is the \emph{Killparents}, which
-returns the children. \emph{Truncate elitism} selects the best of both parents
-and children. \emph{Tournament elitism} matches child next to each parent
-and decide which one is best. There is no kill children implementation
-because nothing would ever change.
-
-\paragraph{Mate selection} MateSelection implements tournament select.
-This is the only mate selection algorithm implemented.
-
-\paragraph{Member factories} The member factories are able to create
-members from strings. These are either the tightly linked or randomly
-linked.
-
-\subparagraph{Mutation} Then the tightly or randomly linked
-functions can also be wrapped in a mutation function,
-which will mutate the string before passing it to the factory.
-Its an decorator idea, only with functions.
-\newpage
-
-\subsubsection{Experiment}
-
-\begin{figure}[ht!]
-\centering
-\begin{emp}[classdiag](20,20)
-Class.ex("Experiment")(
-	"+name:String",
-	"+variation:String",
-	"evolution: Evolution",
-	"factory:String => IMember"
-)(
-	"+bisectionalSearch():StochasticRun",
-	"verifyLowest(currentPop:Int,faults:Int,index:Int):Seq[RunResult]",
-	"findOptimum(consideringSize:Int,difference:Int):Seq[RunResult]",
-	"run(consideringSize:Int):RunResult"
-);
-Class.result("RunResult")(
-	"+popSize:Int",
-	"+success:Boolean",
-	"+runtime:Long",
-	"+generationCount:Int",
-	"+fitnessCallCount:Int"
-	)();
-Class.stoch("StochasticRun")(
-	"+runs:Seq[RunResult]",
-	"+required:Int",
-	)();
-topToBottom.left(20)(ex,stoch);
-leftToRight.top(20)(stoch,result);
-drawObjects(ex,result,stoch);
-
-link(association)(stoch.e -- result.w);
-item(iAssoc)("*")(obj.ne = result.w);
-\end{emp}
-\caption{Experiment}
-\end{figure}
-
-The experiment is setup to do the bisectional search. The bisectional search
-function will wrap the result of \emph{findOptimum} in a \emph{StochasticRun}.
-\emph{findOptimum} is a recursive function that calls verify lowest to get a
-collection of runResults. This is used to determine if the population size
-should be increased or decreased like a bisection search.
-\emph{verifyLowest} will try to run a population size 30 times, if the faults
-limit is exceeded or the index exceeds the required run limit it will quit.
-The run is considered to be successful if the count of successful run >= 
-run limit - faults.
-
-The run will execute the \emph{Evolution} that has been passed into this
-experiment.  so this function will call startGenetic on the Evolution object, 
-and put the results of the run into a \emph{RunResult}.
-
-\subsection{Unit tests}
-Because the assignment has very well defined behavior. I thought it was
-a good Idea to use unit tests. To do this I use a library called
-Scala test.
-
-I found about 6 different bugs alone in the valuation functions by using
-unit-testing, so I think this saved me considerable time. I also used this
-to precisely define the behavior of 2 point cross.
-
-By using unit tests I became very certain about the execution of my code.
-It does precisely what its supposed to do even though I use functional
-programming and it may look somewhat foreign.
-
-\subsection{Build tool}
-I use Gradle for building the project. Its the most advanced build platform
-available and makes it easy to add dependencies to the project.
-For example I use slf4j for logging.
-
-Its also handy in that it comes with a self installing script that can be run
-from any platform, Gradlew. So running the code I wrote on a different machine
-becomes easy, Gradle makes sure the right dependencies are fetched.
-
-\paragraph{Running it}
-
-Open a shell and move to the project folder then type:
-
-\lstset{language=Bash}
-\begin{lstlisting}[frame=single]
-	./gradlew run
-\end{lstlisting}
-
-\subsection{Libraries used}
-In here I will describe why I chose to use these libraries
-The details of the libraries can be found in the Gradle file (which
-version and what maven group):
-
-\begin{lstlisting}[frame=single]
-	cat code/build.gradle
-\end{lstlisting}
-
-\begin{itemize}
-	\item The Scala standard library, because I have to when using Scala.
-	\item slf4j, a simple logging facade, this allows runtime switching of
-		logging schemes, for example if you want to log into a database or
-		a file instead of stdout you can do this relatively easy with slf4j
-		This is not necessary right now but the API is simple enough to use
-		everywhere.
-	\item Scalaplot. A graph plotting library for Scala. I didn't want to
-		do this myself, and this library supports SVG plotting and ASCII
-		plotting.
-	\item Scalatest. A testing library for Scala, it has some pretty advanced
-		features but I use it mostly like JUnit.
-\end{itemize}
-
-\section{Experiment}
-In here I will discuss how the experiments ran and the thoughts I
-above about the results.
-First the setup will be discussed, since I added to the
-original assignment out of curiosity. Then the hypothesis
-will be discussed. After that the results will be shown
-and finally I will discuss the conclusions I draw from the experiments.
-
-\subsection{Experiment titles}
-The experiment titles all follow the same structure:
-
-\[F\mbox{ }L, E\]
-Where F is the fitness function, L is the linkage and E is the elitism type.
-The experiment results are grouped by this structure alphabetically.
-The elitism types are discussed in the fittest fileter paragraph \ref{paragraph:Fittest filter}
-
-Block is the trap function.
-
-\subsection{Abbreviations}
-\begin{tabular}{ll}
-	Abrivation & Full name \\ \toprule
-	2X & 2 Point cross \\
-	2XB & 2 Point cross biased \\
-	2XM & 2 Point cross mutated \\
-	2XM & 2 Point cross mutated \\
-	UX & uniform cross \\
-	UXM & uniform cross mutated \\
-	1X & 1 Point cross \\ \bottomrule
-\end{tabular}
-
-\subsection{Setup}
-I added extra selection methods and crossing strategies to
-the final experiment. I did this because I was curious how they would
-compete with the original assignment.
-
-After doing some sample experiments I found for example that 2x was going
-up with number of successes in the biggest population size. So I wondered 
-if doing my original interpretation of the select would work better.
-I found that in almost every case this original interpretation is at least as
-good or better. On both runtime and successes required population for success.
-
-The biased two point cross will make 2 random values and use the lowest as
-the base crossing for the same parent. The ` normal' 2 point cross will make
-2 random values. If the first one generated is higher than the other one it
-will make the father the mother.
-
-It turns out however that the biased uniform cross is much more effective.
-As it solves the deceptive trap function and leaves the `normal' one in
-the dust. I actually expected an improvement when changing the biased
-variant into the current variant. It was significantly worse however.
-To show this I decided to include both.
-
-1 point cross was added to see how that competes with 2 point cross.
-I also used to have 2 point cross with
-mutation but it wasn't significantly different then the normal 2X
-and the graphs started to get cluttered so I decided to drop that in
-favour of discussing the biased crossover.
-
-Running the code now takes about 26 minutes in total with the
-entire thing divided over 8 threads (quad core + hyper-threading).
-
-\subsection{Hypothesis}
-The deceptive trap function will not be solved by any algorithm for
-the population sizes. I think this because its to desirable to have
-at least one 0000 block in the 100 string length.
-Non-deceptive will be more successful. The others will definitively be solved.
-I also think mutation will have an important effect and may even
-allow deceptive to be solved.
-
-The provided selection pressure would be about average.
-It will probably always find a result.
-
-\newpage
-\input{result.tex}
-
-\subsubsection{Graph observations}
-The major drops are caused by an optimization I implemented. After 2 failed
-run the stochastic insurance function will stop.
-
-Often this happens when the bisectional search hones in on the optimum
-This makes it appear as if there is a local gap, but its just
-random chance.
-
-\section{Conclusion}
-My conclusion is that using both tournament selection and truncate elitism
-is to harsh. It will slow down the GA and often not find a solution at all.
-This happens for example with the deceptive trap function, randomly and
-tightly linked.
-
-The time vs population found ratio is also much better for using 
-tournament select on both the elitism and selection.
-
-Whats also interesting is that if you just throw away the parents and use
-the children (kill parents), then you still will often perform better than
-using tournament selection and truncate elitism (this is how the original
-assignment was), although not as consistently as tournament elitism.
-
-To my surprise deceptive trap was solved with decreased selection pressure
-by both 2x and 1x, although 1x did worse.
-
-Mutation doesn't seem to have much effect UXM sometimes finds a better
-solution but usually this is not the case. I initially thought mutation
-would be very important but its just sort off random.  Mutation turned out
-to have no significant effect other than just slowing things down.
-
-Something to note is that my made up tournament elitism is actually in most
-cases better, or at least equal to the truncate elitism. So when I saw
-this I tried using my other initially used algorithm where I would completely
-ignore the parents (kill the parents). However it appears this does not
-perform better.
-
-Two point cross and single cross both seem to find results in deceptive
-random tournament elitism.
-
-Linearly scaled and uniformly scaled are very easy to solve for almost all
-algorithms. Although the 2x and 1x are at a big disadvantage compared to UX.
-
-\end{empfile}
+\section{TODO}
+MLS with several results
 \end{document}

### 2016-01-07 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -17,13 +17,16 @@ package nl.jappieklooster.hw.ec
 
 import java.io.{File, PrintWriter}
 
-import nl.jappieklooster.hw.ec.model.Vertex
+import nl.jappieklooster.hw.ec.algorithm.{LocalSearch, Evaluation}
+import nl.jappieklooster.hw.ec.experiment.Timer
+import nl.jappieklooster.hw.ec.model._
 import org.sameersingh.scalaplot._
 import org.slf4j.LoggerFactory
 
 import scala.io.Source
 import scala.util.Random
 import scala.util.Properties.lineSeparator
+
 
 object Main{
 	val log = LoggerFactory.getLogger(Main.getClass)
@@ -34,15 +37,49 @@ object Main{
 
 	def main(args:Array[String]){
 		log.info("starting graph bipartitioning")
-		val verteci = Source.fromFile("src/main/resources/data.txt").getLines().zipWithIndex.map((elem) => {
-			val id = elem._2 + 1
 
-			val connections:Seq[Int] = if(elem._1.length > 0) elem._1.split(" ").map(_.toInt) else Nil
-			Vertex(id, connections)
-		})
-		for (vertex <- verteci){
-			log.info(vertex.toString)
+		def strToInts(str:String):Seq[Int] = if(str.length > 0) str.split(" ").map(
+			// they have 1 based index
+			_.toInt-1
+		) else Nil
+		val graph = Graph.create(Source.fromFile("src/main/resources/data.txt").getLines().map(strToInts).toSeq:_*)
+		log.info(graph.toString)
+
+		/*
+		Implement MLS with the VSN local search. Use the first improvement version of
+		the local search. Generate 1000 local optima for a single run of MLS. Measure the
+		computational time required.
+		 */
+		val factory = MemberFactories.tightlyLinked(Evaluation.graphValuation(graph)) _
+		val localSearch = new LocalSearch(LocalSearch.vertexSwapFirstImprovement(graph, factory))
+		val starts = Population.createEqualOnesZeros(random, factory, graph.verteci.length, 1000)
+		var timer = Timer()
+		//val localOptima = starts.par.map(localSearch.search).toArray.sortBy(-_.fitness)
+		//log.info(s"runtime: ${timer.timeSinceCreation}")
+		//log.info("{"+localOptima.foldLeft("")((a,b) => s"$a, ${b.fitness}") + "}")
+
+		//  1862.0, 1842.0, 1798.0, 1792.0
+		/*
[…]

### 2016-01-07 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/algorithm/LocalSearch.scala

@@ -4,19 +4,49 @@ import nl.jappieklooster.hw.ec.model.{Population, IMember, Graph}
 import scala.annotation.tailrec
 import scala.util.Random
 
-class LocalSearch(method:IMember => IMember, stopCondition:(IMember, IMember)=>Boolean = (a,b)=> a==b){
+import LocalSearch._
+class LocalSearch(
+		method:IMember => IMember,
+		stopCondition:(IMember, IMember)=>Boolean = StopConditions.onEqual,
+		decideResult:(IMember, IMember) => IMember = ResultDecision.current
+	){
 	@tailrec
-	final def search(member: IMember):IMember = {
-		val result = method(member)
-		if(stopCondition(result,member)){
-			return result
+	final def search(previous: IMember):IMember = {
+		val result = method(previous)
+		if(stopCondition(result,previous)){
+			return decideResult(result, previous)
 		}else{
 			return search(result)
 		}
 	}
 }
 object LocalSearch {
+	object StopConditions{
+		def onEqual(a:IMember, b:IMember):Boolean = a==b
+		def isWorse(a:IMember, b:IMember):Boolean = a.fitness < b.fitness
+	}
+	object ResultDecision{
+		def current(cur:IMember, prev:IMember) = cur
+		def previous(cur:IMember, prev:IMember) = prev
+	}
 
+	object Iterative{
+		def swapMutation(times:Int, random:Random, memberFactory:String => IMember)(member:IMember):IMember ={
+			val gen = member.genes
+			def randomGenePoint = random.nextInt(gen.length)
+			def create(prev:String, x:Int):String = {
+				val one = randomGenePoint
+				val two = randomGenePoint
+				if(prev(one) == prev(two)){
+					create(prev,x)
+				}else{
+					swapChar(prev)(one, two)
+				}
+			}
+
+			return memberFactory(1.to(times).foldLeft(gen)(create))
+		}
+	}
 	/**
 	* Swap characters at specified positions.
 	*/
@@ -38,7 +68,7 @@ object LocalSearch {
 		val swapVertex = swapChar(member.genes) _
 
 		// I just don't know a good way to do this with recursion or combbinators
-		for(index <- 0.to(graph.verteci.length)){
+		for(index <- 0.to(graph.verteci.length-1)){
 
 			val vert = graph.verteci(index)
 			val options = vert.connections.filter(conindx => partitioning(index) != partitioning(conindx))

### 2016-01-07 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/experiment/Experiment.scala

@@ -13,10 +13,10 @@
 
 // You should have received a copy of the GNU General Public License
 // along with this program.  If not, see <http://www.gnu.org/licenses/\>.
-package nl.jappieklooster.hw.ec
+package nl.jappieklooster.hw.ec.experiment
 
 import com.itextpdf.text.log.LoggerFactory
-import nl.jappieklooster.hw.ec.algorithm.{FittestFilter, MateSelection, Evolution, Evaluation}
+import nl.jappieklooster.hw.ec.algorithm.{Evaluation, Evolution, FittestFilter, MateSelection}
 import nl.jappieklooster.hw.ec.model.{IMember, PairedPopulation, Population}
 
 import scala.util.Random
@@ -83,7 +83,7 @@ case class RunResult(popSize:Int, success:Boolean, runtime:Long, generationCount
 		s"Run(pop#=$popSize, success=$success, time=$runtime, gen#=$generationCount)"
 	}
 }
-import util.Properties.lineSeparator
+import scala.util.Properties.lineSeparator
 case class StochasticRun(runs:Seq[RunResult], required:Int) {
 	lazy val byPopSize = runs.groupBy(x=>x.popSize)
 	lazy val bestRun = byPopSize.getOrElse(bigestPopcount, Nil)

### 2016-01-07 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/experiment/Timer.scala

@@ -1,0 +1,9 @@
+package nl.jappieklooster.hw.ec.experiment
+
+class Timer {
+	val creationTime = System.currentTimeMillis()
+	def timeSinceCreation:Float = (System.currentTimeMillis() - creationTime)/1000
+}
+object Timer{
+	def apply() = new Timer()
+}

### 2016-01-07 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/model/Population.scala

@@ -53,7 +53,7 @@ object Population{
 		if(geneLength % 2 == 1){
 			throw new Exception("Doesn't work with odd genelength")
 		}
-		val initial = createOneZeros(memberFactory, geneLength/2, popCount)
+		val initial = createOneZeros((s) => Member(0,s), geneLength/2, popCount)
 		Population(
 			initial.map(x=> memberFactory(
 				// we need to shuffle the inverse mirror, because otherwise

### 2016-01-07 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/model/Vertex.scala

@@ -4,6 +4,7 @@ case class Vertex(id:Int, connections:Seq[Int]) {
 }
 
 case class Graph(verteci:Array[Vertex]){
+	override def toString = "Graph{"+ verteci.foldLeft("")((p, v) => p + s"(${v.id}:{${v.connections.mkString(",")}}) ") +"}"
 }
 object Graph{
 	def create(elements:Seq[Int]*) :Graph = Graph(

### 2016-01-07 evolutionaryComputation/2-practical/code/src/test/scala/nl/jappieklooster/hw/ec/tests/ExpirimentTests.scala

@@ -1,9 +1,9 @@
 package nl.jappieklooster.hw.ec.tests
 
+import nl.jappieklooster.hw.ec.experiment.Experiment
 import nl.jappieklooster.hw.ec.model.Member
 import nl.jappieklooster.hw.ec.model.{Member, Population}
 import org.scalatest.{Matchers, FlatSpec}
-import nl.jappieklooster.hw.ec.Experiment
 
 class ExpirimentTests extends FlatSpec with Matchers{
 

### 2016-01-09 evolutionaryComputation/2-practical/report.tex
Implemented ILS and GLS

@@
 
+While implementing ILS, I had it turn out results much worse than MLS.
+The initial implmentation would just stop if it found a worse local optimum.
+The description sais it should go back to the previous optimum. So I guess
+that you should try again after that? But then I wonder how often you should
+try again.
+
+I got much better results with a system where the local search got several chances.
+If it failed it would fallback, and then try again. 
+With chances around 30 I managed to get up and around 1900.
+When I reset the chance counter on a succesfull find
+I managed to get consistently 2200, one result was as high as 2310.
+
+I don't think there is a way to do it any better without drastically making
+the algorithm more complex.
+This current search for improvement was just based on the fact that ILS was
+underperforming to MLS, which shouldn't be the case according to my memory
+from the classes. Now its overperforming quite drastically so I'm happy.
+
+I implemented GLS with the evolution class. The only thing I had to do really
+was write the crossover method, the rest was already functional.
 \section{TODO}

### 2016-01-09 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -17,7 +17,8 @@ package nl.jappieklooster.hw.ec
 
 import java.io.{File, PrintWriter}
 
-import nl.jappieklooster.hw.ec.algorithm.{LocalSearch, Evaluation}
+import nl.jappieklooster.hw.ec.algorithm.LocalSearch.{RetryOnResult, Probe}
+import nl.jappieklooster.hw.ec.algorithm._
 import nl.jappieklooster.hw.ec.experiment.Timer
 import nl.jappieklooster.hw.ec.model._
 import org.sameersingh.scalaplot._
@@ -52,11 +53,14 @@ object Main{
 		 */
 		val factory = MemberFactories.tightlyLinked(Evaluation.graphValuation(graph)) _
 		val localSearch = new LocalSearch(LocalSearch.vertexSwapFirstImprovement(graph, factory))
-		val starts = Population.createEqualOnesZeros(random, factory, graph.verteci.length, 1000)
-		var timer = Timer()
-		//val localOptima = starts.par.map(localSearch.search).toArray.sortBy(-_.fitness)
-		//log.info(s"runtime: ${timer.timeSinceCreation}")
-		//log.info("{"+localOptima.foldLeft("")((a,b) => s"$a, ${b.fitness}") + "}")
+		val starts = Population.createEqualOnesZeros(factory, graph.verteci.length, 1000)
+		val timer = Timer()
+		/*val localOptima = timer.time{
+			starts.par.map(localSearch.search).toArray.sortBy(-_.fitness)
+		}
+		log.info(s"runtime: ${timer.seconds}")
+		log.info("{"+localOptima.foldLeft("")((a,b) => s"$a, ${b.fitness}") + "}")
+		*/
 
 		//  1862.0, 1842.0, 1798.0, 1792.0
 		/*
@@ -68,17 +72,59 @@ object Main{
 		 Are the results obtained with the different mutation/perturbation sizes statistically
 		 different from each other ?
 		 */
-		log.info(starts.head + "")
-		for(size <- 1.to(20)){
+		//  1 fitness {2310.0,
+		/*log.info(starts.head + "")
+		for(size <- 1.to(5).par){
+			val probe = new Probe
+			val resultplexer = new RetryOnResult(30, null)
+			import LocalSearch._
 			val iterativeLocalsearch = new LocalSearch(
-				(localSearch.search _ ).compose(LocalSearch.Iterative.swapMutation(size,random,factory)),
-				stopCondition = LocalSearch.StopConditions.isWorse,
[…]

### 2016-01-09 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/algorithm/Evolution.scala

@@ -17,21 +17,24 @@ package nl.jappieklooster.hw.ec.algorithm
 
 import nl.jappieklooster.hw.ec.model.{PairedPopulation, Population}
 
+import Evolution._
+
+import scala.annotation.tailrec
+
 class Evolution(
-	val evaluation: FitnessEvaluator,
 	mateSelector:IMateSelector,
-	offSpringGenerator: PairedPopulation => Population,
+	offSpringGenerator: OffspringGenerator,
 	/**
 	 * fittest filter, (valuation, parents, children) => fittestpop
 	 */
-	fittestFilter: (Population, Population) => Population,
+	fittestFilter: FittestFilter,
 
 	/**
 	 * If there already exists a really good solution you're
 	 * happy with you can use this so you don't have to wait untill
 	 * convergence
 	 */
-	hasGoodEnoughSolution: Population => Boolean
+	hasGoodEnoughSolution: SolutionJudge = SolutionJudge.badSolution
 ) {
 	private def step(parents:Population):Population = fittestFilter(
 			parents,
@@ -44,6 +47,7 @@ class Evolution(
 	def startGenetic(seed:Population): Seq[Population]= genetic(
 		Seq(seed)
 	)
+	@tailrec
 	private def genetic(prev:Seq[Population]): Seq[Population]={
 		val parents = prev.last
 		if(hasGoodEnoughSolution(parents)){
@@ -58,3 +62,11 @@ class Evolution(
 	}
 }
 
+object Evolution{
+	type OffspringGenerator = PairedPopulation => Population
+	type FittestFilter = (Population, Population) => Population
+	type SolutionJudge = Population => Boolean
+	object SolutionJudge{
+		def badSolution(x:Population) = false
+	}
+}

### 2016-01-09 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/algorithm/LocalSearch.scala

@@ -6,9 +6,9 @@ import scala.util.Random
 
 import LocalSearch._
 class LocalSearch(
-		method:IMember => IMember,
-		stopCondition:(IMember, IMember)=>Boolean = StopConditions.onEqual,
-		decideResult:(IMember, IMember) => IMember = ResultDecision.current
+		method:SearchMethod,
+		stopCondition:StopCondition = StopCondition.onEqual,
+		decideResult:ResultDecision = ResultDecision.current
 	){
 	@tailrec
 	final def search(previous: IMember):IMember = {
@@ -21,10 +21,20 @@ class LocalSearch(
 	}
 }
 object LocalSearch {
-	object StopConditions{
+	type SearchMethod = IMember => IMember
+	type StopCondition = (IMember, IMember)=>Boolean
+	object StopCondition{
 		def onEqual(a:IMember, b:IMember):Boolean = a==b
 		def isWorse(a:IMember, b:IMember):Boolean = a.fitness < b.fitness
+		def resetRetryOnBetter(which:RetryOnResult, stopCondition: StopCondition)(current: IMember, previous: IMember): Boolean = {
+			val result = stopCondition(current,previous)
+			if(!result){
+				which.reset
+			}
+			result
+		}
 	}
+	type ResultDecision = (IMember, IMember) => IMember
 	object ResultDecision{
 		def current(cur:IMember, prev:IMember) = cur
 		def previous(cur:IMember, prev:IMember) = prev
@@ -37,10 +47,13 @@ object LocalSearch {
 			def create(prev:String, x:Int):String = {
 				val one = randomGenePoint
 				val two = randomGenePoint
-				if(prev(one) == prev(two)){
+				val acmp = prev(one)
+				val bcmp = prev(two)
+				if(acmp == bcmp){
 					create(prev,x)
 				}else{
-					swapChar(prev)(one, two)
+					val result = swapChar(prev)(one, two)
+					result
 				}
 			}
 
@@ -85,5 +98,25 @@ object LocalSearch {
 			}
 		}
 		member
+	}
+
+	class Probe extends SearchMethod{
+		var tracked = List[IMember]()
+		override def apply(member: IMember): IMember = {
+			tracked = tracked :+ member
+			member
+		}
+	}
+	class RetryOnResult(repitions:Int, var method: SearchMethod) extends ResultDecision{
+		val resetvalue = 0
+		private var current = resetvalue
[…]

### 2016-01-09 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/algorithm/OffspringGenerator.scala

@@ -1,0 +1,104 @@
+package nl.jappieklooster.hw.ec.algorithm
+
+import com.itextpdf.text.log.LoggerFactory
+import nl.jappieklooster.hw.ec.model.{PairedPopulation, Population}
+
+import scala.util.Random
+
+object OffspringGenerator{
+	def uniformCross(x:PairedPopulation) : Population = x.pairedToPop(
+		/* kindoff a weird implementation, but it works because
+		 1==1V0==0, then the selection is irrelevant
+		 but iff 1==0 or 0==1, then the randomGene will just
+		 create a 1 or 0 without looking at the parents.
+		*/
+		genes =>
+			// where the genes are equal will keep the same, else random
+			genes.map(fm => if (fm._1 == fm._2) fm._1 else randomGene)
+	)
+
+	def balancedUniformCross(x:PairedPopulation) :Population = x.pairedToPop(genes =>{
+		val hammingDistance = genes.count(a => a._1 != a._2)
+
+		// if the hamming distance is to big, crossover will have little effect,
+		// so inverting one of the parents will result in the same score but a
+		// closer hamming distance
+		val hammingClosed = (if(hammingDistance > genes.length/2){
+			genes.map(a => (flipBit(a._1), a._2))
+		}else{
+			genes
+		}).zipWithIndex
+
+		// filter is a little confusing, it selects the predicate that matches
+		val notMatching = hammingClosed.filter(cur => cur._1._1 != cur._1._2)
+
+		// Can there be an uneven amount of notMatching bits in two balanced solutions?
+		// No, this is impossible. So we just let the program crash if this occurs
+		val newGenesInstructions = Map(randomBalancedGenes(notMatching.length).zipWithIndex.map(
+			pair => notMatching(pair._2)._2 -> pair._1
+		):_*)
+
+		hammingClosed.map(fm => if(fm._1._1 == fm._1._2) fm._1._1 else newGenesInstructions(fm._2) )
+	})
+	def singlePointCross(rand:Random)(x:PairedPopulation) : Population = x.pairedToPop(
+		genes => {
+			val point = rand.nextInt(genes.length)
+			// where the genes are equal will keep the same, else random
+			genes.zipWithIndex.map(fm => {
+				if (point < fm._2) fm._1._1 else fm._1._2
+			})
+		}
+	)
+	val log = LoggerFactory.getLogger("offspringernerator")
[…]

### 2016-01-09 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/experiment/Experiment.scala

@@ -40,7 +40,7 @@ class Experiment(val name:String, val variation:String, evolution: Evolution, me
 
 		// sucesfull if early termination or the last population is good enough
 		val success = hasGoodEnoughSolution(evolutionResult.last)
-		RunResult(consideringSize, success, runtime, evolutionResult.length,evolution.evaluation.countCalls())
+		RunResult(consideringSize, success, runtime, evolutionResult.length,0)
 	}
 
 	def verifyLowest(currentPop:Int, faults:Int, index:Int) : Seq[RunResult] = {
@@ -131,7 +131,6 @@ object Experiment{
 			name,
 			variation._3,
 			new Evolution(
-				probe,
 				MateSelection.createCompeteWithRandomTournement(random),
 				variation._1,
 				filter,

### 2016-01-09 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/experiment/Timer.scala

@@ -1,8 +1,15 @@
 package nl.jappieklooster.hw.ec.experiment
 
 class Timer {
-	val creationTime = System.currentTimeMillis()
-	def timeSinceCreation:Float = (System.currentTimeMillis() - creationTime)/1000
+	var last:Long = 0
+	def time[R](block: => R): R = {
+		val t0 = System.nanoTime()
+		val result = block    // call-by-name
+		val t1 = System.nanoTime()
+		last = t1 - t0
+		result
+	}
+	def seconds:Double = last / 1000000000.0
 }
 object Timer{
 	def apply() = new Timer()

### 2016-01-09 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/model/Member.scala

@@ -18,7 +18,7 @@ package nl.jappieklooster.hw.ec.model
 import org.slf4j.LoggerFactory
 
 import scala.util.Random
-import nl.jappieklooster.hw.ec.algorithm.IHasFitness
+import nl.jappieklooster.hw.ec.algorithm.{LocalSearch, IHasFitness}
 
 trait IMember extends IHasFitness with Genetic
 case class Member(valuation:Float, gen:String) extends IMember{
@@ -28,6 +28,7 @@ case class Member(valuation:Float, gen:String) extends IMember{
 
 
 object MemberFactories{
+	type MemberFactory = (String=>Float) => String => IMember
 	def tightlyLinked(func:String=>Float)(str:String):IMember =
 		Member(func(str), str)
 	def randomlyLinked
@@ -40,8 +41,8 @@ object MemberFactories{
 	val log = LoggerFactory.getLogger(this.getClass)
 	case class WithCoinFlipTimesMutation(
 		random:Random,
-		creationFunc:(String=>Float) => String => IMember
-	) extends ((String=>Float)=>(String) => IMember){
+		creationFunc:MemberFactory
+	) extends MemberFactory {
 
 		def apply(func:String=>Float) = (str:String) => {
 			val flipCount = countFlips(-1)
@@ -63,5 +64,10 @@ object MemberFactories{
 			}
 			integer
 		}
+	}
+
+	def withLocalSearch(localSearch:LocalSearch, creationFunction: MemberFactory)(func:String=>Float)(str:String):IMember ={
+		val result = creationFunction(func)(str)
+		localSearch.search(result)
 	}
 }

### 2016-01-09 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/model/PairedPopulation.scala

@@ -65,56 +65,4 @@ case class PairedPopulation(members:Seq[Pair], memberFactory:String => IMember) 
 		)
 	}
 }
-object OffspringGenerator{
-	def uniformCross(x:PairedPopulation) : Population = x.pairedToPop(
-		/* kindoff a weird implementation, but it works because
-		 1==1V0==0, then the selection is irrelevant
-		 but iff 1==0 or 0==1, then the randomGene will just
-		 create a 1 or 0 without looking at the parents.
-		*/
-		genes =>
-			// where the genes are equal will keep the same, else random
-			genes.map(fm => if (fm._1 == fm._2) fm._1 else randomGene)
-	)
-	def singlePointCross(rand:Random)(x:PairedPopulation) : Population = x.pairedToPop(
-		genes => {
-			val point = rand.nextInt(genes.length)
-			// where the genes are equal will keep the same, else random
-			genes.zipWithIndex.map(fm => {
-				if (point < fm._2) fm._1._1 else fm._1._2
-			})
-		}
-	)
-	val log = LoggerFactory.getLogger("offspringernerator")
-	def inverResult(selected:Char, one:Char, two:Char) = if(selected == one) two else one
-	def keepSelected(selected:Char, one:Char, two:Char) = selected
-	def twoPointCross(onFirstBig:(Char,Char,Char)=>Char, genes:IndexedSeq[(Char,Char)], first:Int, second:Int):IndexedSeq[Char] = {
-		val isFirstSmall = first < second
-		val result = genes.zipWithIndex.map(fm => {
-			def select(start:Int, end:Int):Char = {
-				if(fm._2 < start) fm._1._1 else{
-					if(fm._2 < end){
-						fm._1._2
-					}else fm._1._1
-				}
-			}
-			if(isFirstSmall){
-				select(first,second)
-			}else{
-				onFirstBig(select(second,first),fm._1._1,fm._1._2)
-			}
-		})
-		result
-	}
-	def twoPointCross(rand:Random, onFirstBig:(Char,Char,Char)=>Char)(x:PairedPopulation) : Population = x.pairedToPop(
-		genes => {
-			val first = rand.nextInt(genes.length)
-			val second = rand.nextInt(genes.length)
-			log.debug(s"crossing with $first and $second")
-			// where the genes are equal will keep the same, else random
-			twoPointCross(onFirstBig, genes,first,second)
-		}
-	)
-	val random = Random
-	def randomGene = s"${random.nextInt(2)}"
-}
+

### 2016-01-09 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/model/Population.scala

@@ -15,6 +15,8 @@
 // along with this program.  If not, see <http://www.gnu.org/licenses/\>.
 package nl.jappieklooster.hw.ec.model
 
+import nl.jappieklooster.hw.ec.algorithm.OffspringGenerator
+
 import scala.util.Random
 
 case class Population(members:Seq[IMember], memberFactory:String => IMember) extends EmulateSeq[IMember] {
@@ -22,50 +24,38 @@ case class Population(members:Seq[IMember], memberFactory:String => IMember) ext
 
 }
 object Population{
-	def createOneZeros(
+	type SeedMethod = Int=>String
+	def createWithMethod(
 			memberFactory:String => IMember,
 			geneLength:Int,
-			popCount:Int):Population =  Population(
-		0.to(popCount-1).map(
+			popCount:Int,
+			seedMethod:SeedMethod
+			):Population = Population(
+		1.to(popCount).map(
 			x => memberFactory(
 				// it says, from 1 to length concat 1 or 0.
-				0.to(geneLength).foldLeft(""){(b, x) => b + OffspringGenerator.randomGene}
+				seedMethod(geneLength)
 			)
 		),
 		memberFactory
 	)
+	def createOneZeros(
+			memberFactory:String => IMember,
+			geneLength:Int,
+			popCount:Int):Population = createWithMethod(
+		memberFactory,
+		geneLength,
+		popCount,
+		OffspringGenerator.randomGenes
+	)
 
-	def flipBit(c:Char):Char = if (c == '0') '1' else '0'
-	/**
-	 * return the original and flip every other bit, this basically makes an
-	 * equal amount of ones and zeros.
-	 * @param input
-	 * @return
-	 */
-	def inverseMirror(input:String):String ={
-		input + input.map(flipBit)
-	}
 	def createEqualOnesZeros(
-		random:Random,
 		memberFactory:String => IMember,
 		geneLength:Int,
-		popCount:Int):Population =  {
-		if(geneLength % 2 == 1){
-			throw new Exception("Doesn't work with odd genelength")
-		}
-		val initial = createOneZeros((s) => Member(0,s), geneLength/2, popCount)
-		Population(
-			initial.map(x=> memberFactory(
-				// we need to shuffle the inverse mirror, because otherwise
-				// it'll lose some entropy (the posibility space becomes less
-				// without shuffeling because the first part is only randomly
-				// generated, the second part is its mirror image)
[…]

### 2016-01-09 evolutionaryComputation/2-practical/code/src/test/scala/nl/jappieklooster/hw/ec/tests/CrossoverTests.scala

@@ -1,13 +1,14 @@
 package nl.jappieklooster.hw.ec.tests
 
-import nl.jappieklooster.hw.ec.model.OffspringGenerator
+import nl.jappieklooster.hw.ec.algorithm.OffspringGenerator
 import OffspringGenerator._
+import nl.jappieklooster.hw.ec.algorithm.OffspringGenerator
 import org.scalatest.{Matchers, FlatSpec}
 
 import scala.util.Random
 
 class CrossoverTests extends FlatSpec with Matchers{
-	def invertCross(a:IndexedSeq[(Char,Char)],b:Int,c:Int) = twoPointCross(inverResult, a,b,c).mkString
+	def invertCross(a:IndexedSeq[(Char,Char)],b:Int,c:Int) = twoPointCross(inverseResult, a,b,c).mkString
 	"2 point cross " should " work unbiased" in{
 		val parents = "abcdefg".zip("1234567")
 		// normal crosses...

### 2016-01-14 evolutionaryComputation/2-practical/report.tex
using the original data

@@
 was write the crossover method, the rest was already functional.
+
+I wanted to see if the GLS popsize and selection pressure would have any
+effect on the average fitness found. I suspect that a bigger population size
+would find better results and that weaker selection would find better
+average scores.
+
+I chose to investigate with population sizes of 25, 50 and 75. I didn't want
+to wait to long on this and the runtime for 50 was approaching ~10 minutes
+for the slowest. Luckily I have a powerfull laptop and can use paralization.
+
+Doing this 30 times to cancel the staostic effect would result into ~5 hours.
 \section{TODO}
 MLS with several results
+GLS -> I think the localsearch will recursively unecesrly decend into its members
+because the local search is packed in the member factory. Maybe remove the
+local search from the memberfactory?
+
+Use original file by doing a substr(30)
 \end{document}

### 2016-01-14 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -17,9 +17,9 @@ package nl.jappieklooster.hw.ec
 
 import java.io.{File, PrintWriter}
 
-import nl.jappieklooster.hw.ec.algorithm.LocalSearch.{RetryOnResult, Probe}
+import nl.jappieklooster.hw.ec.algorithm.Search.{RetryOnResult, Probe}
 import nl.jappieklooster.hw.ec.algorithm._
-import nl.jappieklooster.hw.ec.experiment.Timer
+import nl.jappieklooster.hw.ec.experiment.{Experiment, Timer}
 import nl.jappieklooster.hw.ec.model._
 import org.sameersingh.scalaplot._
 import org.slf4j.LoggerFactory
@@ -39,11 +39,21 @@ object Main{
 	def main(args:Array[String]){
 		log.info("starting graph bipartitioning")
 
-		def strToInts(str:String):Seq[Int] = if(str.length > 0) str.split(" ").map(
-			// they have 1 based index
-			_.toInt-1
-		) else Nil
+		def strToInts(str:String):Seq[Int] = {
+			if(str.length <= 30){
+				return Nil
+			}
+			val sub = str.substring(30)
+			if(sub.length > 0) sub.split(" ").map(
+				// they have 1 based index
+				_.toInt-1
+			) else Nil
+		}
 		val graph = Graph.create(Source.fromFile("src/main/resources/data.txt").getLines().map(strToInts).toSeq:_*)
+		def printResults(members:Seq[IMember]) = {
+			val results = members.foldLeft("")((a,b) => s"$a, ${graph.edgeCount-b.fitness}").substring(2)
+			log.info(s"{$results}")
+		}
 		log.info(graph.toString)
 
 		/*
@@ -52,15 +62,18 @@ object Main{
 		computational time required.
 		 */
 		val factory = MemberFactories.tightlyLinked(Evaluation.graphValuation(graph)) _
-		val localSearch = new LocalSearch(LocalSearch.vertexSwapFirstImprovement(graph, factory))
-		val starts = Population.createEqualOnesZeros(factory, graph.verteci.length, 1000)
+		val localSearch = new Search(Search.vertexSwapFirstImprovement(graph, factory))
+		val starts = Population.createEqualOnesZeros(factory, graph.verteci.length, 100)
 		val timer = Timer()
-		/*val localOptima = timer.time{
+		val localOptima = timer.time{
[…]

### 2016-01-14 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/algorithm/FittestFilter.scala

@@ -18,6 +18,7 @@ package nl.jappieklooster.hw.ec.algorithm
 import nl.jappieklooster.hw.ec.model.Population
 
 object FittestFilter {
+	type FittestFilter = (Population, Population) => Population
 	/** drop the parents */
 	def killParents(
 		parents : Population,

### 2016-01-14 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/algorithm/Search.scala

@@ -1,11 +1,21 @@
 package nl.jappieklooster.hw.ec.algorithm
 
 import nl.jappieklooster.hw.ec.model.{Population, IMember, Graph}
+
 import scala.annotation.tailrec
 import scala.util.Random
 
-import LocalSearch._
-class LocalSearch(
+import Search._
+
+/**
+ * Search, it depends on how you configure it if its global or local.
+ * it executes the search method, then checks if its happy, if not it'll
+ * try again, if so the decideResult is called
+ * @param method
+ * @param stopCondition
+ * @param decideResult
+ */
+class Search(
 		method:SearchMethod,
 		stopCondition:StopCondition = StopCondition.onEqual,
 		decideResult:ResultDecision = ResultDecision.current
@@ -20,7 +30,7 @@ class LocalSearch(
 		}
 	}
 }
-object LocalSearch {
+object Search {
 	type SearchMethod = IMember => IMember
 	type StopCondition = (IMember, IMember)=>Boolean
 	object StopCondition{

### 2016-01-14 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/experiment/Experiment.scala

@@ -1,148 +1,21 @@
-// Evolutionary Computing Analyzer, it analyzes various genetic algoritms
-// Copyright (C) 2015 Jappie Klooster
-
-// This program is free software: you can redistribute it and/or modify
-// it under the terms of the GNU General Public License as published by
-// the Free Software Foundation, either version 3 of the License, or
-// (at your option) any later version.
-
-// This program is distributed in the hope that it will be useful,
-// but WITHOUT ANY WARRANTY; without even the implied warranty of
-// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-// GNU General Public License for more details.
-
-// You should have received a copy of the GNU General Public License
-// along with this program.  If not, see <http://www.gnu.org/licenses/\>.
 package nl.jappieklooster.hw.ec.experiment
 
-import com.itextpdf.text.log.LoggerFactory
-import nl.jappieklooster.hw.ec.algorithm.{Evaluation, Evolution, FittestFilter, MateSelection}
-import nl.jappieklooster.hw.ec.model.{IMember, PairedPopulation, Population}
+import scala.collection.GenSeq
 
-import scala.util.Random
-/**
- * This expirement tries to find the optemal number of population for a specific
- * evolution.
- * @param evolution
- */
-class Experiment(val name:String, val variation:String, evolution: Evolution, memberFactory:String => IMember) {
-	import Experiment._
-	val log = LoggerFactory.getLogger(this.getClass)
-	private def createPopulation(size:Int) = {
-		Population.createOneZeros(memberFactory, geneLength, size)
-	}
-	def run(consideringSize:Int): RunResult={
-		val population = createPopulation(consideringSize)
-
-		val startTime = System.currentTimeMillis()
-		val evolutionResult = evolution.startGenetic(population)
-		val runtime = System.currentTimeMillis() - startTime
-
-		// sucesfull if early termination or the last population is good enough
-		val success = hasGoodEnoughSolution(evolutionResult.last)
-		RunResult(consideringSize, success, runtime, evolutionResult.length,0)
+class Experiment[R](code: => R) {
+	val totalTimer = Timer()
+	def run:RunResult[R] = {
+		val timer = Timer()
[…]

### 2016-01-14 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/experiment/Timer.scala

@@ -1,16 +1,18 @@
 package nl.jappieklooster.hw.ec.experiment
 
 class Timer {
-	var last:Long = 0
+	var measurement:Option[Moment] = None
 	def time[R](block: => R): R = {
 		val t0 = System.nanoTime()
 		val result = block    // call-by-name
 		val t1 = System.nanoTime()
-		last = t1 - t0
+		measurement = Some(Moment(t1 - t0))
 		result
 	}
-	def seconds:Double = last / 1000000000.0
 }
 object Timer{
 	def apply() = new Timer()
+}
+case class Moment(duration:Long){
+	lazy val seconds:Double = duration / 1000000000.0
 }

### 2016-01-14 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/model/Member.scala

@@ -18,7 +18,7 @@ package nl.jappieklooster.hw.ec.model
 import org.slf4j.LoggerFactory
 
 import scala.util.Random
-import nl.jappieklooster.hw.ec.algorithm.{LocalSearch, IHasFitness}
+import nl.jappieklooster.hw.ec.algorithm.{Search, Search$, IHasFitness}
 
 trait IMember extends IHasFitness with Genetic
 case class Member(valuation:Float, gen:String) extends IMember{
@@ -66,7 +66,7 @@ object MemberFactories{
 		}
 	}
 
-	def withLocalSearch(localSearch:LocalSearch, creationFunction: MemberFactory)(func:String=>Float)(str:String):IMember ={
+	def withLocalSearch(localSearch:Search, creationFunction: MemberFactory)(func:String=>Float)(str:String):IMember ={
 		val result = creationFunction(func)(str)
 		localSearch.search(result)
 	}

### 2016-01-14 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/model/Vertex.scala

@@ -4,7 +4,10 @@ case class Vertex(id:Int, connections:Seq[Int]) {
 }
 
 case class Graph(verteci:Array[Vertex]){
-	override def toString = "Graph{"+ verteci.foldLeft("")((p, v) => p + s"(${v.id}:{${v.connections.mkString(",")}}) ") +"}"
+	override def toString = s"Graph(edges:$edgeCount){${
+		verteci.foldLeft("")((p, v) => p + s"(${v.id}:{${v.connections.mkString(",")}}) ")
+	}}"
+	lazy val edgeCount = verteci.map(_.connections.length).sum
 }
 object Graph{
 	def create(elements:Seq[Int]*) :Graph = Graph(

### 2016-01-14 evolutionaryComputation/2-practical/code/src/test/scala/nl/jappieklooster/hw/ec/tests/SearchTest.scala

@@ -3,9 +3,9 @@ package nl.jappieklooster.hw.ec.tests
 import nl.jappieklooster.hw.ec.algorithm.Evaluation
 import org.scalatest.{Matchers, FlatSpec}
 import nl.jappieklooster.hw.ec.model.{Member, Graph}
-import nl.jappieklooster.hw.ec.algorithm.LocalSearch._
+import nl.jappieklooster.hw.ec.algorithm.Search._
 
-class LocalSearchTest extends FlatSpec with Matchers{
+class SearchTest extends FlatSpec with Matchers{
 
 	val graph = Graph.create(
 		Seq(2),

### 2016-01-16 evolutionaryComputation/2-practical/report.tex
found the right graph

@@
 \usepackage{svg}
+\usepackage{pgfplots}
+\usepackage{pgfplotstable}
+
+\pgfplotsset{
+    compat=1.9,
+    compat/bar nodes=1.8
+}
+
+\begin{filecontents*}{dataTable.txt}
+zero 	xpos 		firsty 		secondy		errbotstart		errbotend 		errtopstart			errtopend
+0 		1           0         	1			0.1				0.2				0.1	               0.2
+0 		2           1         	2			0.3				0.1             1.5               0.4
+0 		3           2         	3.5			0.5				0.2             0.3               0.4
+\end{filecontents*}
+
 \newenvironment{scprooftree}[1]%
@@
 
-Use original file by doing a substr(30)
+\pgfplotstableread{dataTable.txt}\dataTable
+\begin{tikzpicture}
+\pgfplotsset{set layers}
+\begin{axis}[ybar stacked]
+
+% Layers needed to get error bars on top of lower blue rectangle
+\addplot[color=white]
+table[on layer=axis grid, x=xpos, y=firsty]{\dataTable};
+
+\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
+table[x=xpos, y=zero, y error minus=errbotstart, y error plus=errbotend]{\dataTable};
+
+\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
+table[x=xpos, y=secondy, y error minus=errtopstart, y error plus=errtopend]{\dataTable};
+
+\end{axis}
+
+\end{tikzpicture}
+
+\end{empfile}
 \end{document}

### 2016-01-17 evolutionaryComputation/2-practical/report.tex
this has a sane plot thing that works

--- 2016-01-14 evolutionaryComputation/2-practical/report.tex
@@
 \usepackage{svg}
+\usepackage{pgfplots}
+\usepackage{pgfplotstable}
+
+\pgfplotsset{
+    compat=1.9,
+    compat/bar nodes=1.8
+}
+
+\begin{filecontents*}{dataTable.txt}
+zero 	name 		firsty 		secondy		errbotstart		errbotend 		errtopstart			errtopend
+0 		MLS           0         	1			0.1				0.2				0.1	               0.2
+0 		ILS           1         	2			0.3				0.1             1.5               0.4
+0 		GLS           2         	3.5			0.5				0.2             0.3               0.4
+\end{filecontents*}
+
 \newenvironment{scprooftree}[1]%
@@
 
-Use original file by doing a substr(30)
+\pgfplotstableread{dataTable.txt}\dataTable
+\begin{tikzpicture}
+\pgfplotsset{set layers}
+\begin{axis}[
+	ybar stacked,
+	xlabel={Algorithm},
+	xticklabels from table={dataTable.txt}{name},
+	xtick=data,
+	ylabel={Time}
+]
+
+% Layers needed to get error bars on top of lower blue rectangle
+\addplot[color=white]
+table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};
+
+\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
+table[x expr=\coordindex, y=zero, y error minus=errbotstart, y error plus=errbotend]{\dataTable};
+
+\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
+table[x expr=\coordindex, y=secondy, y error minus=errtopstart, y error plus=errtopend]{\dataTable};
+
+\end{axis}
+
+\end{tikzpicture}
+
+\end{empfile}
 \end{document}

### 2016-01-22 evolutionaryComputation/summery.tex
cheatsheet

\documentclass[10pt,landscape]{article}
\usepackage{multicol}
\usepackage{calc}
\usepackage{ifthen}
\usepackage[landscape]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage{color,graphicx,overpic}
\usepackage{hyperref}

\pdfinfo{
  /Title (example.pdf)
  /Creator (TeX)
  /Producer (pdfTeX 1.40.0)
  /Author (Seamus)
  /Subject (Example)
  /Keywords (pdflatex, latex,pdftex,tex)}

% This sets page margins to .5 inch if using letter paper, and to 1cm
% if using A4 paper. (This probably isn't strictly necessary.)
% If using another size paper, use default 1cm margins.
\ifthenelse{\lengthtest { \paperwidth = 11in}}
    { \geometry{top=.5in,left=.5in,right=.5in,bottom=.5in} }
    {\ifthenelse{ \lengthtest{ \paperwidth = 297mm}}
        {\geometry{top=1cm,left=1cm,right=1cm,bottom=1cm} }
        {\geometry{top=1cm,left=1cm,right=1cm,bottom=1cm} }
    }

% Turn off header and footer
\pagestyle{empty}

% Redefine section commands to use less space
\makeatletter
\renewcommand{\section}{\@startsection{section}{1}{0mm}%
                                {-1ex plus -.5ex minus -.2ex}%
                                {0.5ex plus .2ex}%x
                                {\normalfont\large\bfseries}}
\renewcommand{\subsection}{\@startsection{subsection}{2}{0mm}%
                                {-1explus -.5ex minus -.2ex}%
                                {0.5ex plus .2ex}%
                                {\normalfont\normalsize\bfseries}}
\renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{0mm}%
                                {-1ex plus -.5ex minus -.2ex}%
                                {1ex plus .2ex}%
                                {\normalfont\small\bfseries}}
\makeatother

% Define BibTeX command
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

% Don't print section numbers
\setcounter{secnumdepth}{0}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0pt plus 0.5ex}

%My Environments
\newtheorem{example}[section]{Example}
% -----------------------------------------------------------------------

\begin{document}
\raggedright
\footnotesize
\begin{multicols}{3}

% multicol parameters
% These lengths are set only within the two main columns
%\setlength{\columnseprule}{0.25pt}
\setlength{\premulticols}{1pt}
\setlength{\postmulticols}{1pt}
\setlength{\multicolsep}{1pt}
\setlength{\columnsep}{2pt}

\begin{center}
     \Large{\underline{Evolutionary Computing}} \\
\end{center}

\section{Genetic algorithms}
\subsection{Schema}
$o(h)$: schema order $o(11\#\#0)$ ie nonwildcard \# elements.
$\delta(h)$: schema defining length, maximum distance between two defining
symbols. $m(h,t)$: number of schema h instances at generation t.
$f(h,t)=\sum_{i\in P}f_i$ schema fitness is average fitness of members.
\subsubsection{Growth by selection}
reproduction ratio $\phi(h,t)=\frac{m(h,t^s)}{m(h,t)}$.
\paragraph{Proportionate selection}
$f_i$ fitness ind i. Probability individiual i selected: $\frac{f_i}{\sum f_i}$.
$N$ pop size. Expected number of copies
$\frac{f_i}{\sum f_i} \cdot N = \frac{f_i}{f(t)}$.
Expected number of copies of schema $h$ members:
$m(h,t^s) = m(h,t)\phi(h,t)=m(h,t)\frac{f(h,t)}{f(t)}$.
\paragraph{Tournament selction}
tournament size $s: 0 \leq \phi(h,t) \leq s$.

\subsubsection{Distruption by mutation}
probability bit flipped: $p_m$.
schema $h$ survives iff all the bit values are \emph{not} mutated
$p_{survival}=(1-P_m)^{o(h)}$. for small values $p_m << 1$,
$(1-p_m)^{o(h)}\approx 1 - o(h)\cdot p_m$.
disruption factor $\epsilon(h,t)$ by mutation: $\epsilon(h,t)=o(h)\cdot P_m$.

\subsubsection{Distruption by mutation}
Probablity crossover applied $P_c$. Bit swap probability: $p_x$.
\paragraph{1-point crossover} schema $h$ survives iff cutpoint \emph{not}
withing defining length $\delta$: $p_{survival} = 1-\frac{\delta(h,t)}{l-1}$
\paragraph{Uniform crossover} Schema $h$ survives iff none or all bits
swapped together. $p_{survival}=p^{o(h)}_x + (1-p_x)^{o(h)}$.

\paragraph{Distruption factor} $\epsilon(h,t)$ by recombination:
$\epsilon(h,t)=P_c\cdot(1-p_{survival})$

\subsubsection{Theorom}
Selection mutation and recombination combined:
$m(h,t+1)\leq m(h,t)\phi(h,t)[1-\epsilon(h,t)]$.
Net growth factor: $\gamma(h,t)=\frac{m(h,t+1)}{m(h,t)}$.

\subsection{Fitness Correlation coefficents}
$F_p$: mean fitness of the parents. $F_c$: mean fitness of the children.
$\sigma(F_p) =$ standard deviation of fitness parents
$\sigma(F_c) =$ std. dev. of children. coveriance fitness between parents and children:
$cov(F_p,F_c)=\sum^n_{i=1}\frac{(f(p_{gi})-F_p)(f(c_{gi})-f_c)}{n}$.
Operator fitness correlation coeficient
$\rho_{op}=\frac{cov(F_p,F_c)}{sigma(F_p)\sigma(F_c)}$.

\section{ES representation}
Fitness function: $f(x_1,\dots,x_n):\mathbb{R}^n\to \mathbb{R}$.
Genotype representation of an individual solution:
$(x_1,\dots,x_n,\sigma^2_1,\dots,\sigma^2_n, c_{12},\dots,c_{n-1,n})$,
paremeters $(x_1, \dots,x_n)$ need to be optimized.
Individiual solution consists of 3 parts: $\vec{x}$: problem vairables
$\Rightarrow$ Fitness $f(\vec{x})$. $\vec{sigma}$: standard deviations
$\Rightarrow$ variances. $\vec{\alpha}$: rotation angles $\Rightarrow$
covariances.

\end{multicols}
\end{document}

### 2016-01-23 evolutionaryComputation/summery.tex
space!
sumerizing....

@@
 $\Rightarrow$ variances. $\vec{\alpha}$: rotation angles $\Rightarrow$
-covariances.
+covariances. The $n$-dimensional normal probability density function:
+$p(X=x_1,\dots,x_n)=\frac{exp(-\frac{1}{2}X^TC^{-1}X)}{\sqrt{(2\pi)^n|C|}}$.
+$C$: correlation matrix ($c_{ij}$; $|C|$ determinant $\Rightarrow$ rotation 
+angles $\alpha_{ij}$: tan $2\alpha_{ij}=2c_{ij}/(\sigma^2_i-\sigma^2_j)$.
+1-dimensional gaussian function:
+$f(x)=\frac{1}{\sqrt{2\pi \sigma^2}}exp\frac{(x-\mu)^2}{2\sigma^2}$
+\subsection{Mutation}
+Complete covairance matrix: $\sigma_1,\dots,\sigma_n$; $c_{ij}\neq 0 (i\neq j)$.
+The strategy paremeters are mutated:
+$\sigma'_i=\sigma_ie^{
+	\frac{\mathbb{N}(0,1)}{\sqrt{2n}}+
+	\frac{\mathbb{N}_i(0,1)}{\sqrt{2\sqrt{n}}}
+}$
+$\alpha'_j=\alpha_j+\beta\mathbb{N}_j(0,1)$.
+$\beta = 5$ degree in radians. $\mathbb{N}(0,1)$: standard normal
+distribution. Then the problem paremeters are mutated with the new
+strategy: $\vec{x}'=\vec{x}+\vec{N}(\vec{0},\vec{\sigma},\vec{\alpha}')$.
+$\vec{N}$: n-dimensional normal distribution.
+Often this is simplified by not doing angles or even just having a single
+standard deviation (ie no index at $\sigma$ and nothing after the + 
+in the power $e$).
+\subsection{Recombination}
+create one offspring from several parents selected randomly. Problem parameters
+are averaged. The std devation is selected from one of the parents. Rotation
+angles are not recombined
+\subsection{Selection}
+ES applies a high selection pressure: from $\mu$ parents $\lambda$ offspring
+are generated with $\lambda >> \mu$ (~5 to 10 times). The best $\mu$ solutions
+of the $\lambda$ offspring are selected, (or from both offspring and parents,
+but experiments show just offspring has better selfadaptation).
 

### 2016-01-23 evolutionaryComputation/summery.tex
added about ada pursuit and some about modaling the behavior

@@
 \section{Genetic algorithms}
+%TODO: crossover methods I don't understand
+%TODO: schemas, buidling block theory
 \subsection{Schema}
@@
 
-\subsubsection{Distruption by mutation}
+\subsubsection{Distruption by recombination}
 Probablity crossover applied $P_c$. Bit swap probability: $p_x$.
-\paragraph{1-point crossover} schema $h$ survives iff cutpoint \emph{not}
+
+\emph{1-point crossover} schema $h$ survives iff cutpoint \emph{not}
 withing defining length $\delta$: $p_{survival} = 1-\frac{\delta(h,t)}{l-1}$
-\paragraph{Uniform crossover} Schema $h$ survives iff none or all bits
+
+\emph{Uniform crossover} Schema $h$ survives iff none or all bits
 swapped together. $p_{survival}=p^{o(h)}_x + (1-p_x)^{o(h)}$.
 
-\paragraph{Distruption factor} $\epsilon(h,t)$ by recombination:
+\emph{Distruption factor} $\epsilon(h,t)$ by recombination:
 $\epsilon(h,t)=P_c\cdot(1-p_{survival})$
@@
 but experiments show just offspring has better selfadaptation).
+\section{Ada pursuit}
+Given: Set of $K$ \emph{operators} $A=\{a_1,\dots,a_k\}$. 
+\emph{Probability vector} $P(t) = \{P_1(t),\dots,P_k(t)\}$. operator $a_i$ 
+applied at time t in proportion to probability $P_i(t)$. env. returns rewards
+$R_i(t) \geq 0$. Goal: Adapt $P(t)$ so that expected value of cumalitive
+reward $\varepsilon[R]=\sum^T_{t=1}R_i(t)$ is maximized.
+
+\subsection{Probabilty matching}
+update $P(t)$ so probability of applying operator $a_i$ matches
+the proportion of the estimated reward $Q_i(t)$ to the sum of all reward 
+estimates $\sum^K_{a=1} Q_a(t)$
+\subsubsection{PM: Reward estimate}
+compute an estimate of the rewards, and older rewards should be less
+influantial in non-stationary envs:
+$Q_a(t+1)=Q_a(t)+\alpha[R_a(t)-Q_a(t)]$
+
+\subsubsection{PM: Probability adaptation}
+In non-stationary envs the probabilty of applying
+any operator should never be less than some minmal \emph{thershold} $P_{min} > 0$.
+For $K$ operators maximal probablity $P_{max} = 1-(K-1)P_{min}$. updating rule
+for $P(t)$: $P_a(t+1)=P_{min}+(1-K\cdot P_{min})\frac{Q_a(t)}{\sum^K_{i=1}Q_i(t)}$.
+\subsubsection{PM: Problem}
+If one operator is consistently better probablity will still converge to average.
+
+\subsection{Adaptive Pursuit strategy}
+Update $P(t)$ so \emph{operator} $a^*$ that currently has the max estimated
+reward $Q_{a^*}(t)$ is \emph{persued}. To achieve this the pursuit method
+\emph{increases} selection probability $P_{a^*}$ and decreases all other
+probabilities $P_a(t),\forall a\neq a^*$. So here it can deal with non-static envs.
+\subsubsection{APS: Adaptive method}
+Similiar to probability matching:
+$Q_a(t+1)=Q_a(t)+\alpha[R_a(t)-Q_a(t)]$. Different from probablity matching:
+Selection probability $P(t)$ is adapted in a greedy way.
+\subsubsection{APS: Probability adoptation}
+The selection probability of the current best operator
+$a^*=argmax_a[Q_a(t+1)]$ is \emph{increased} $(0<\beta<1)$:
+$P_{a^*}(t+1)=P_{a^*}(t)+\beta[P_{max}-P_{a^*}(t)]$.
+The others are \emph{decreased}: 
+$\forall a\neq a^*: P_a(t+1)=P_a(t)+\beta[P_{min}-P_a(t)]$
+
+%TODO: the psuedo code??
+
+\section{Convergence models}
+Number of fitness function evaluations is proportional to the
+number of generations run and pop size $FitFunc.count = Gen.count * Popsize$.
+Selection algorithm and variation operators can be ignored.
+
+\emph{Convergence speed:}
+High selection presure is faster covnergence. Bigger popsize give better solutions.
+Minimal fitness function evals is a tradeoff between selection presure and popsize.
+\subsection{Selection Intensity}
+Quantifies the speed of convergence. $I$ selection intensity. Response to
+selection $R(t)$ is difference between mean fitness of generation $t+1$ and fitness
+of generation $t$. Selection differential $S(t)$ is the diff between parent pop
+at generation $t$ and mean fitness of generation $t$. 
+$S(t)=\overline{f(t^s)}-\overline{f(t)}$. Assuming the pop fitness is normally
+distributed $N(\overline(f),\sigma^2)$ we can scale the selection differential
+by thestandard deiation $\sigma(t)$. This scaled selection differential is 
+called selection intensity $I(t)$. This is a dimensionless number since 
+the std deiation has the same units as the selection response:
+$I(t)=\frac{S(t)}{\sigma(t)}=\frac{\overline{f(t^s)}-\overline{f(t)}}{\sigma(t)}$.
+Standerdizing the normal distribution $(\overline(f)=0,\sigma=1)$ shows that
+selection intensity $I$ is simply the epected average fitness
+of the population after applying the selection scheme to a population
+with standardized normal distibuted fitness $(N(0,1))$.
+The relation between the response to selection $R$ and the selection
+differential $S$ is given by the heritability $h^2$: $R(t)=h^2S(t)$ or
+$R(t)=h^2\sigma(t)I(t)$.
+
+\subsection{Proportionate selection}
+$P_i(t)$ the proportion of occurences of individual $i$ at generation $t$.
+$i$ has fitness $f_i$ and the mean fitness of generation $t$ is $\overline{f(t)}$.
+$P_i(t^s)$ the proportion of individual $i$ in the parent pool
+after applying proportionate selection: $P_(t^s)=P_i(t)\frac{f_i}{\overline{f(t)}}$.
+The selection differential of proportiaonate selection is: 
+$S(t)=\overline{f(t^s)}-\overline{f(t)}=\frac{\sigma^2(t)}{\overline{f(t)}}$
+The selection intensity $I(t)=\frac{S(t)}{\sigma(t)}$ is tus eaual
+to the ratio of the std devation of the fitness and the pop mean fitness:
+$I(t)=\frac{\sigma(t)}{\overline{f(t)}}$
 

### 2016-01-23 evolutionaryComputation/summery.tex
finished page 1 of the topics (excluding self knowledge)

@@
     {\ifthenelse{ \lengthtest{ \paperwidth = 297mm}}
-        {\geometry{top=1cm,left=1cm,right=1cm,bottom=1cm} }
-        {\geometry{top=1cm,left=1cm,right=1cm,bottom=1cm} }
+        {\geometry{top=0.5cm,left=0.5cm,right=0.5cm,bottom=0.5cm} }
+        {\geometry{top=0.5cm,left=0.5cm,right=0.5cm,bottom=0.5cm} }
     }
@@
 $f(h,t)=\sum_{i\in P}f_i$ schema fitness is average fitness of members.
+
+\emph{Buidling blocks} low order high performance schemate receive exponentiaonally
+(geometrically) increasing trials.
+
 \subsubsection{Growth by selection}
 reproduction ratio $\phi(h,t)=\frac{m(h,t^s)}{m(h,t)}$.
-\paragraph{Proportionate selection}
+
+\emph{Proportionate selection}
 $f_i$ fitness ind i. Probability individiual i selected: $\frac{f_i}{\sum f_i}$.
@@
 $m(h,t^s) = m(h,t)\phi(h,t)=m(h,t)\frac{f(h,t)}{f(t)}$.
-\paragraph{Tournament selction}
+
+\emph{Tournament selction}
 tournament size $s: 0 \leq \phi(h,t) \leq s$.
@@
 
-\subsubsection{Theorom}
+\subsubsection{Theorem}
 Selection mutation and recombination combined:
@@
 Net growth factor: $\gamma(h,t)=\frac{m(h,t+1)}{m(h,t)}$.
+\subsection{Permutation problems}
+2-opt m, select 2 points and reverse a subsequance.
+adjacency of elements is important, 2-opt has minimal change.
+relative ordering is important: 2-opt large change.
+\subsubsection{Crossover}
+std. crosover won't work cause they can make invalid solution.
+ie : $abcd+CBDA\to Cbcd$. Order crossover: rnd select 2 points,
+copy subseq between crosspoints from p1, starting at 2nd crosspoint
+fill in missing elements retaining order from p2. partially
+mapped crosover: rnd select 2 points, copy p2 to child, copy
+elements between crospoint from p1 to childe while placing
+the replaced element from p2 at the location where the replacer is.
+Position crossover: rnd select k postitions, copy unmarked elements from
+p1 to child, scan p2 from lft to rght and fill in missing elemnts. Max 
+preservative cross: rnd select two points, copy subseq between crospoints
+from p1, add succesivily an adjecent element from p2 starting at
+last element in child, if already placed take and adjecent elemnt form
+p1. Cycle crossover: mark cycles, cross full cycles.
 
 \subsection{Fitness Correlation coefficents}
+Genetic operators should perserve usefull fitness characaristics.
+Calcualte the fitness coorlation coeficient to quantify this.
 $F_p$: mean fitness of the parents. $F_c$: mean fitness of the children.
@@
 
+\section{Metaheursitics}
+Crossover can have mutliple search biases:
+1. Random sampling witin a specific subspace. 2. place together for contrast
+partial solutions from two parent solutions.
+
+\subsection{PBMS}
+Probablistic Modal Building: Probability distributions model dependencies 
+between variables present in good solutions. sSelection makes these fitness 
+based dependencies stand out. Estimating a probablity modal over the selected
+solutions identifies these dependencies. Drawing new samples form the probability
+modal will respect the dependencies.
+PMBLS = PBM + LS. Bivariate probalbistic modal learns the pairwise dependencies
+between problem dependencies. Buildign a dependency tree=maximum spanning tree
+over the dependency graph. New solutions obtained by sampling from the dependency
+tree.
+For graphbipartitioning. Probablistic model: count the frequnceis that two
+verticies are in the same parition. Dep tree build over most extreme frequency
+values. reduce computational complexity by only considering the
+pairwise interactions between connected vertices
 \section{ES representation}

### 2016-01-23 evolutionaryComputation/summery.tex
added about pop sizing

@@
 
+\subsection{Rank based selection}
+Proporitionate selection has not a constant selection pressure.
+
+\section{Population sizing}
+Selection viewed as descision making proscess withing partitions:
+schama competition. When best schema looses comptetition we have a 
+selection descision error. The amount of decision errors that
+can be afforded at a given pop size can be explained by random walks.
+\subsection{Selection error}
+Probablity selection error is equal to the probablity that the best schema is represented
+by a string with fitness less than the representative worst schema. wich is also equal
+to the probability that the fitness difference is nagative.
+%TODO: add fancy pants maths.
+
+\subsection{Random walks}
+One dim space of size $N + 1$. A particle in above
+space in postition $x\in \{0,\dots,N\}$. The particle can move
+left with probability $p$ and right with probability $1-p$.
+when the particle reaches the boundries the walk ends.
+$P_N(x)$ the probability that the particle is absorbed
+by boundry $N$. When it currently is at pos $x$.
+Diff equation:
+$P_N(x)=pP_N(x+1)+(1-p)P_N(x-1)$
+
+\subsection{Population sizing \& gambler's ruin}
+The number of optimal bit values `1' in the population at a certain
+position. - this is, for the partition considered - corresponds
+to the position $x$ in the gambler's ruin model.
+The boundries correspond to all bit vlaues in the population being
+equal to a bit value. WHen the boudaries are reached the random walk
+ends. The proba bility that the amount of optoimal string values in the
+population at the partition is increased by one is equal to the probability
+that the selection decision making is correct for that partition. This
+corresponds to the probability that the particle moves to the right(p).
+Desired convergence corresponds to the particle reaching the $x=N$ boundary.
+Premature convergence - this is, loosing the optimal bit value in the
+population - correspons the the particle reaching the $x=0$ boundary.
+$P[OptConv]=\frac{1-(\frac{P[SellErr]}{1-P[SellErr]})^{N/2}}{1-(\frac{P[SellErr]}{1-P[SellErr]})^N}
+\approx 1-(\frac{\sqrt{\pi l}-2}{\sqrt{\pi l}+2})^{N/2}$
+Critical pop size: $N \approx ln(1-P[OptConv]\frac{\sqrt{pi l}}{-2}$
+
 \end{multicols}

### 2016-01-23 evolutionaryComputation/summery.tex
Reduced the text size significantly and continued steamrolling trough the slides. Also added my proper name to the text instead of the stackoverflow thing.

@@
-\documentclass[10pt,landscape]{article}
+\documentclass[8pt,landscape]{extarticle}
 \usepackage{multicol}
@@
 \pdfinfo{
-  /Title (example.pdf)
+  /Title (EC cheatsheet)
   /Creator (TeX)
   /Producer (pdfTeX 1.40.0)
-  /Author (Seamus)
-  /Subject (Example)
-  /Keywords (pdflatex, latex,pdftex,tex)}
+  /Author (Jappie Klooster)
+  /Subject (Evulotionary computing)
+  /Keywords (Evolutionary, Computing, Gentic, Algorithm, Heuristics, Strategy)}
 
@@
 
+\section{Multi objective EA's}
+\subsection{Pareto}
+\emph{Dominance} A solution $x^0$ is said to (Paretor) \emph{dominate} a
+solution $x^1$ (denoted $x^0\succ x^1$) iff
+$(\forall i \in \{0,1,\dots,m-1\}:f_i(x^0)\leq f_i(x^1))\wedge
+ (\exists i \in \{0,1,\dots,m-1\}:f_i(x^0)< f_i(x^1))$.
+A solution is pareto optimal iff $\neg\exists x^1:x^1\succ x^0$.
+The parote optimall set: $P_s=\{x^0|\neg\exists x^1:x^1\succ x^0\}$.
+The paroto optimal front is the $P_f$ of all objective function values
+correspoinding to the solutions in $P_s$: $P_F=\{f(x)=(f_0(x),f_1(x),\dots,
+	f_{m-1}(x))|x\in P_s$
+\subsection{Weighted aggregation}
+The graphs with the linerider lines. It doesn't work with bowl
+like objectives because only the edges are found.
+\subsection{Evolutionary Multi-objective Optimization}
+Desired result: representative subset of $P_s$.
+\emph{Domination count}
+For each solution count how many times
+its dominatad. Then use one-dimensional selection (truncate or tournament).
+\emph{Domination rank} solutions of rank 0 have domination count 0, solutions
+of rank i are solutions of rank 0 when solutions of lower ranks are disregarded.
+
+\subsection{Elitism}
+What if number of rank 0 solutions is bigger than the next generation?
+Must diregard some of the rank 0 solutions. The space dominated by
+elitist solutions changes, this can cause the elitist solutions/front to
+become \emph{worse}. This isn't true eltism!
+For this reason an elitistarchive is often used. Current best solutions
+are stored in the archive if they dominate at least one solution in the
+archive or if they are not dominated by any solution in the archive.
+Dominated solutions are removed from the archive. It can grow exponentially
+large. Pruning the archive causes the same problems again. One solution:
+discretize objective space, store at most one solution per
+discretization box. Typically for better convergence also select
+solutions from archive.
+\subsection{Performance indicators}
+Performance indicator is a function: Input approximation set, output: single
+value representing a certaing goodnes. However no single or combination of
+performance indiciators taht compare approximation sets can always tell wether
+one approximation set is better than another.
+
+\section{Model-based EAs: Est. of dist. algorithms}
+Model based evolutionary algorthms: Population based stochastic search
+algorithms. Exploitation: selection, Exploration: Learn model from selected
+solutions, generate new solutions from the model (\&pop).
+\subsection{Univariate PMBGA}
+No statistical interaction between variables considered.
+Model: probability vector $[p_1,\dots,P_l]$($l$:string length).
+$p_i$: probability of value 1 at string position $i$.
+$p(x)=\Pi^l_{i=1}p(x_i) \quad$ ($p(x_i)$: univariate marginal distribution).
+Learn modal: count proportions of 1 in selected population. Sample model:
+Generate new solutions with specified probabilities.
+
+%TODO: lack of space remove variants
+\subsubsection{Variants}
+PBIL: Prob. vector incrementally updated over successive generations.
+UMDA: No incremental updates. Compact GA: Models steady-state GA with
+tournament selection. DEUM: Uses markov random field modeling.
+
+\subsection{Bivariate PMBGA}
+Need more than just probabilities of bit values, model pairwise interactions:
+conditional probablities.
+
+\subsubsection{Ntropy}
+Random variable $X$ with probability istrubtion function $p(X)$. Entropy
+$H(X$ is a measure of unceartainty about a random variable $X$:
+$H(X) =\sum_{x\in X}-p(x)\mbox{log }p(x)$. Conditionaly entropy $H(Y|X)$ is
+a measure of uncertainty remianing about $Y$ after $X$ is known (what $X$
+does not say about $Y$):
+$H(Y|X) =\sum\limits_{x\in X}\sum\limits_{y\in Y}p(x,y)\mbox{log }
+\frac{p(x)}{p(x,y)}$. The mutual information $I(X,Y)$ of two random variables
+is a measure of the variables' mutual dependence. Mutual information determines
+how similiar the joint distribution $P(X,Y)$ is to the products of factored
+marginal distribution $p(X)p(Y)$:
+$I(Y,X) =\sum\limits_{x\in X}\sum\limits_{y\in Y}p(x,y)\mbox{log }
+\frac{p(x,y)}{p(x)p(y)}$. Mutual information in relation to Ntropy:
+$I(X,Y) = H(Y)-H(Y|X)=H(X)+H(Y)-H(X,Y)$.
+
 \end{multicols}

### 2016-01-24 evolutionaryComputation/summery.tex
CHEATSHEET FINAL FORM

@@
 \renewcommand{\subsection}{\@startsection{subsection}{2}{0mm}%
-                                {-1explus -.5ex minus -.2ex}%
-                                {0.5ex plus .2ex}%
+                                {-0.8ex plus -.5ex minus -.2ex}%
+                                {0.2ex plus .2ex}%
                                 {\normalfont\normalsize\bfseries}}
 \renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{0mm}%
-                                {-1ex plus -.5ex minus -.2ex}%
-                                {1ex plus .2ex}%
+                                {-0ex plus -.5ex minus -.2ex}%
+                                {0.1ex plus .2ex}%
                                 {\normalfont\small\bfseries}}
@@
 \footnotesize
-\begin{multicols}{3}
+\begin{multicols}{4}
 
@@
 \section{Genetic algorithms}
-%TODO: crossover methods I don't understand
-%TODO: schemas, buidling block theory
+Darwinian system $\Rightarrow$ GA: 
+Information structures $\Rightarrow$ binary strings, real-valued vectors,
+programs etc. Structures are copied $\Rightarrow$ Selection algorithm.
+Copies partially vary from original $\Rightarrow$ mutation \& crosover
+operators. Structures are competing for a limited resource $\Rightarrow$
+selecting fixed size parent pool. Reproductive succes depends on env
+$\Rightarrow$ user defined fitness function.
 \subsection{Schema}
@@
 \emph{Buidling blocks} low order high performance schemate receive exponentiaonally
-(geometrically) increasing trials.
+(geometrically) increasing trials. ie there exists some schemas that when combined
+will produce a near optimal solution, the building blocks.
 
@@
 like objectives because only the edges are found.
-\subsection{Evolutionary Multi-objective Optimization}
+\subsection{Multi-objective Optimization}
 Desired result: representative subset of $P_s$.
@@
 
-\section{Model-based EAs: Est. of dist. algorithms}
+\section{Model EAs: Est. dist. algorithms}
 Model based evolutionary algorthms: Population based stochastic search
@@
 
+\subsubsection{COMIT}
+Optimal dependency tree. Compute fully connected weighted graph between problem
+variables. Weights are the mutual infromation $I(X,Y)$ between the variables.
+Commit computes the maximam spanning treee of the weighted graph. The approximating
+probability model is restricted to factorization in which the conditional
+probabilty distribtuion for any random vraiable depends on the value of at
+most one other random variable
+$p(x)=\overset{n}{\underset{i=1}{\Pi}}p(X_i|X_{parent(i)})$.
+$p(X)$ is the class of distubtions with a tree as graphical model.
+The optimal tree model: Create a fully connected weighed graph $G$,
+each vertix $V_i$ corresponds to randome variable $X_i$ each weight $W_{ij}$
+for the edge betweeen $V_i$ and $V_j$ is equal to the mutal information
+$I(X_i, X_j)$. The edges in the maximum spanning tree of $G$ determine an
+optimal set $n-1$ conditional probabilites with which to aproximate the joint
+probability distribution. Note that all ordered trees conforming the unorderd
+spanning tree model identical distributions.
+\emph{Algorithm} Calculate unconditional and conditinoal probabilties
+$p(X_i)$ and $p(X_i,X_j)$, and the mutual information $I(X_i,X_j)$.
+Select arbitrary random variable $X_r$ as root of the tree. Finda the
+pair of variables $X_{in}$ and $X_{out}$ where $X_{in}$ is already in the
+tree and $X_{out}$ is not, with the maximum mutual information
+$I(X_{in}, X_{out})$. Add $X_{out}$ to the tree with $X_{in}$ as parent,
+reapeat untill all variables are in the tree.
+
+\subsection{Multivariate PMBGA}
+MDL is a measure of complexity (Information Theroy).
+$MDL(M,D)=D_{Model}+D_{Data}$ Model complexity $D_{Model}$:
+complexity of describing the model. Compressed population complexity
+$D_{Data}$: complexity of describing the data within the model (=measure
+of goodness of the probability distribution estimation).
+Best model = the one with the lowest MDL score.
+$D_{Model}=log_2(N+1)\sum_i(2^{s_i}-1)$. Compressed pop complexity
+$D_{Data}=N\sum_i H(M_i)$. $N$ pop size. $S_i$ size of partition $i$.
+$M_i$ marginal distribution of the partition $i$. $H(M_i)$ entropy of the
+marginal distribution of the partition i.
+
+\subsubsection{Extended compact GA}
+Greedily searches for the marginal product model that minzimizes the MDL.
+$p(X)=\Pi^G_{g=1}p(x_g)$ Choose the probability distribution with the lowest
+MDL score. Start from the simplest model: the univariate factorization.
+Join two grousp that result in the largest improvemnt in the used scoring
+measure. Stop when no jioning of two groups impvoes the score further.
+\section{Model EAs: Linkage Tree GA}
+Identify groups of problem variables that together make an important
+contribution to the quality of solutions.
+These variable groups are interacting in a non-linear way and should
+be processed as a block = \emph{building block} the dependency struct
+is a set of subsets of the problem.
+\subsection{Family of subsets (dep struct)}
+Let there be $l$ problem vars $x_0,\dots,x_{l-1}$. Let $S$ be a set of all
+variable indecies $\{0,\dots,l-1\}$. A FOS $F$ is a set of subsets of the
+set $S$. FOS $F$ i a subset of the powerset of $S(F\subseteq P(S))$.
+FOS can be written more specifically as $F=\{F^0,\dots,F^{|F|-1}\}$. where
+$F^i \subseteq \{0,\dots,l-1\}, i\in\{0,\dots,|F|-1\}$. Every var is
+in at least one subset in the FOS: $\forall i \in \{0,\dots,l-1\}:(
+\exists j \in \{0,\dots,|F|-1\}:i\in F^j)$.
+\subsection{Linkage tree learning}
+start from univariate structure (ie all vars in independent sets). Build
+linkage tree using bottum-up hierarchial clustering algortim.
+Similiartiy measuer: Between individual vairables X and Y: Mutual
+informaiton I(X,Y).  Between clusture groups $X_{F^i}$ and $X_{F^j}$:
+average pairwise linkage clustering (=unweighted pair group method
+with arithmic mean).
+$I^{UPGMA}(X_{F^i}, X_{F^j} = \frac{1}{|X_{F^i}||X_{F^j}|}
+\sum\limits_{X\in X_{F^i}}\sum\limits_{Y\in X_{F^i}} I(X,Y)$
+\subsubsection{Reciprocal nearest neighbor chain clustering}
+Construct a nearsest neighbor chian: start from an arbitrary
+start cluster which is followed by its nearest neigbro, which
+again is followed by its nearest neigbor from among the remaining
+points and so on.
+When the nearest neigbor for the last chain element $n$ is the elements
+$n-1$ we stop with the generation with the NNchan, we have found
+a $RNN$ pair and can agglomerate it.
+The nearest neigbor assignment stay valid for the remaining chain members,
+which can thus be reused for the next iteration of agglomeration. If
+we do not find any more RNN pairs in the remaining NN chain we create a new one
+starting from a random start point.
+\subsection{Optimal Mixing Evolutionary Algorhms}
+OMEA is a model building EA that uses a FOS as its linkage model.
+It uses intermediate function evalutations (inside variation).
+Can be regarded as greedy improvment of exisiting solutions.
+Coined ``optimal'' mixing because better instances for substructures
+are imediatly accepted and not dependent on noise coming form other parts
+of te solution.
+\subsection{GOMEA}
+FOS liknage models to specify the linked vars. A subset of the fos is used
+as crosover mask. Crosover is greedy: only improvments (or equal) are accepted.
+Each generation a new FOS model is build from selected solutions. For each solution
+in the population all subsets of the FOS are treid with a donor solution
+randomly picked fro mthe population. Recombainative OM is GA-like: select single
+donor solution to perform OM with. GOM is EDA-like: select donor solution for each
+subset in OM.
+\subsection{Linkage Tree genetic algorithm.}
+The LTGA is an instance of GOMEA that uses a linkage tree as fos model. Each
+generation a new hierechial cluster tree is build. For each solution in
+population, traverse tree starting at the top. Node (=clusters) in the linkage
+tree used as crosover masks. Select random donor solution and its values at
+the crossover mask replace the variable values from the current solution.
+Evaluate new solution and accept if better/equal, otherwise reject.
+\section{Seminars}
+\subsection{ILS for Qaudratic assignment problem}
+Population based extension work by replacing the worst. Using evolution
+strategies. and don't do recombination?
+\subsection{GA with ILS for solving location routing}
+Assignment Configuration $A(x)$: location. Permutation vector $P(x)$: routing.
+\emph{crossover} $A(x)$ one point cross. $P(x)$ focus on relative ordering.
+\subsection{GASAT}
+\emph{Corrective clause crossover} set all the bits of the child to undefined.
+Go over all false clauses of both parents (and child), find the best to flip,
+keep the sum of parents' fitness the lowest, flip this bit in the child.
+End step: all undefined bits of the child are uniformly crossed over.
+\emph{Corrective clause and truth maintance crossover} Obvious problem
+with CC: Child may make clause false that is true for both parents. A possible
+olution: CCTM. DO CC except for end step. Go over all clauses that are true
+for parents, false for child. Look at the bits of the clause but only those
+different from the child and at least one parent. Find the best bit to flip,
+keep the sum of the parents' fitness the lowest. Flip this bit in the cild.
+End step of CC.
+\emph{Fleuent and ferland's crossover} Go over all clauses true for one parent
+and false for another. Copy all of the former's bits for that clause. End
+step of CC.
+\emph{Also uniform crossover} copy a bit of x or y with equal probablity.
+\subsection{Fitness landscapes in graphbitpart}
+Researh suggest that the optimal soution is often found near the
+center of local-optimum space. Thus a good idea for a search
+tactic is to find a way to get near this center and use a lcoal
+optimization function to try and get near the global optimum.
+\subsection{Solving bus terminal problems}
+Standar rollete wheel with linear ranking is assigning fitness to
+possible solutions. The higher the fitness the more chance of being
+selected.
+\subsection{GLS w. dist preserv recombo}
+Order crossover (see earlier). single parent cross, select with probability 0.5
+from the parent for customers not inserted pick one at random and insert at the
+best position. Convert into chromosone.
+String cross
 \end{multicols}

### 2016-01-26 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala
can do stastical significant stuff now, although I don't claim to understand it.

@@ -77,17 +77,12 @@ object Main{
 		println(s"total time ${first.seconds}")
 		val times = first.result.map( x=> x.seconds.toDouble).sorted
 		Plot.write("runtimes.table", DataTable.createTable(
-			TableRow("MLS", times.head, times.last,0.1,0.1)
+			DataTable.createRow("MLS", times)
 		))
 		val error = times.length/10
 		val top = times.take(error).sum/error
 		println(first.result.flatMap(x=>x.result).map(x=>graph.edgeCount-x.fitness).sorted)
-		/*val topError = times.take(error).head - top
-		val botError = top - times.take(error).last
-		println(Plot.asciiGraph(("blah", first.result.map( x=> x.seconds.toDouble))))
 
-		println(times)
-		*/
 		//  1862.0, 1842.0, 1798.0, 1792.0
 		/*
 		 Investigate the impact of different mutation/perturbation sizes for ILS with the VSN
@@ -99,7 +94,7 @@ object Main{
 		 different from each other ?
 		//  1 fitness {2310.0,
 		log.info(starts.head + "")
-		for(size <- 1.to(10).par){
+		for(size <- 1.to(5).par){
 			val probe = new Probe
 			val resultplexer = new RetryOnResult(30, null)
 			import Search._
@@ -109,7 +104,7 @@ object Main{
 				decideResult = resultplexer
 			)
 			resultplexer.method = iterativeLocalsearch.search
-			timer.time{
+			val runtime = Timer.measure{
 				iterativeLocalsearch.search(starts.head)
 			}
 			val results = probe.tracked.sortBy(-_.fitness)
@@ -118,9 +113,7 @@ object Main{
 		}
 
 
-		*/
-		/*
-		val popSizes = List(25,50,75)
+		val popSizes = List(5,10,20)
 		val selectMethods = List(
 			("Kill parents", FittestFilter.killParents _),
 			("Tournament", FittestFilter.tournementElitism _),
@@ -133,7 +126,7 @@ object Main{
 				OffspringGenerator.balancedUniformCross,
 				method._2
 			)
-			val result = timer.time{
+			val result = Timer.measure{
 				gls.startGenetic(
 					Population.createEqualOnesZeros(
 						MemberFactories.withLocalSearch(
@@ -147,13 +140,12 @@ object Main{
 			}
 
[…]

### 2016-01-26 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/experiment/DataTable.scala

@@ -1,13 +1,40 @@
 package nl.jappieklooster.hw.ec.experiment
 
+import org.apache.commons.math3.distribution.TDistribution
+import org.apache.commons.math3.exception.MathIllegalArgumentException
+import org.apache.commons.math3.stat.descriptive.SummaryStatistics
+
+
 object DataTable {
 
 	def createTable(tableRow: TableRow*): String ={
 		val toprow = "zero    name        firsty      secondy     errbotstart     errbotend    errtopstart    errtopend"
 		tableRow.foldLeft(toprow)((a,b)=>a+System.lineSeparator() +b.toString)
 	}
+
+	def createRow(name:String, data:Seq[Double]):TableRow ={
+		// Build summary statistics of the dataset "data"
+		val stats = new SummaryStatistics()
+		data.foreach(stats.addValue)
+
+		def calcMeanCI(confidence:Double) = {
+			// Create T Distribution with N-1 degrees of freedom
+			val tDist = new TDistribution(stats.getN - 1)
+			// Calculate critical value
+			val critVal = tDist.inverseCumulativeProbability(1.0 - (1 - confidence) / 2)
+			// Calculate confidence interval
+			critVal * stats.getStandardDeviation / Math.sqrt(stats.getN)
+		}
+
+		// Calculate 95% confidence interval
+		val ci = calcMeanCI(0.95)
+		val lower = stats.getMean - ci
+		val upper = stats.getMean + ci
+		val sorted = data.sorted
+		TableRow(name, lower, upper,sorted.head-lower,sorted.last - upper)
+	}
 }
 
 case class TableRow(name:String, firstY:Double, secondY:Double, firstErr:Double, secondErr:Double){
-	override def toString:String = s"0	$name	$firstY		$secondY	$firstErr	0	0	$secondErr"
+	override def toString:String = s"0	$name	$firstY		$secondY	0 	$firstErr	0	$secondErr"
 }

### 2016-01-29 evolutionaryComputation/2-practical/report.tex
can do stastical significant stuff now, although I don't claim to understand it.
now can generate tables properly also started working on fidducia and mathesis search

@@
 Doing this 30 times to cancel the staostic effect would result into ~5 hours.
+
+\subsection{Counting double}
+I had an issue in which my classmates would get much better results than me.
+After discussing it for a while it turned out that I was counting double. However
+This didn't make sense because I just used scala combinators to count the 
+elements. Basically this means that there would have to be a bug in the std
+Scala library which would seem highly unlikely. I knew I was counting double
+because all my results where even numbers. After thinking about it for a while
+I decided to investigate the graph file, it turns out that every connection
+is made double, so the double counting was in fact correct.
+
 \section{TODO}
@@
 
-\pgfplotstableread{dataTable.txt}\dataTable
+\section{Results}
+\subsection{Time}
+\pgfplotstableread{runtimes.table}\dataTable
 \begin{tikzpicture}
@@
 	xlabel={Algorithm},
-	xticklabels from table={dataTable.txt}{name},
+	xticklabels from table={runtimes.table}{name},
 	xtick=data,
+	x tick label style={rotate=90,anchor=east},
 	ylabel={Time}
@@
 
+\input{runtimes.textable}
+
+\subsection{Optima}
+\pgfplotstableread{results.table}\dataTable
+\begin{tikzpicture}
+\pgfplotsset{set layers}
+\begin{axis}[
+	ybar stacked,
+	xlabel={Algorithm},
+	xticklabels from table={results.table}{name},
+	xtick=data,
+	x tick label style={rotate=90,anchor=east},
+	ylabel={Fitness (less is better)}
+]
+% Layers needed to get error bars on top of lower blue rectangle
+\addplot[color=white]
+table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};
+
+\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
+table[x expr=\coordindex, y=zero, y error minus=errbotstart, y error plus=errbotend]{\dataTable};
+
+\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
+table[x expr=\coordindex, y=secondy, y error minus=errtopstart, y error plus=errtopend]{\dataTable};
+
+\end{axis}
+
+\end{tikzpicture}
+\\
+\input{results.textable}
 \end{empfile}

### 2016-01-29 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala

@@ -17,7 +17,8 @@ package nl.jappieklooster.hw.ec
 
 import java.io.{File, PrintWriter}
 
-import nl.jappieklooster.hw.ec.algorithm.Search.{RetryOnResult, Probe}
+import nl.jappieklooster.hw.ec.algorithm.search.{VertexSwapFirstImprovement, Search}
+import Search.{RetryOnResult, Probe}
 import nl.jappieklooster.hw.ec.algorithm._
 import nl.jappieklooster.hw.ec.experiment._
 import nl.jappieklooster.hw.ec.model._
@@ -39,114 +40,139 @@ object Main{
 	val random = Random
 	val folder = "../"
 
-	def main(args:Array[String]){
+	val glsPopSizes = List(5)
+	val mlsStartsize = 10
+
+	def main(args:Array[String]) {
 		log.info("starting graph bipartitioning")
 
-		def strToInts(str:String):Seq[Int] = {
-			if(str.length <= 30){
+
+		def strToInts(str: String): Seq[Int] = {
+			if (str.length <= 30) {
 				return Nil
 			}
 			val sub = str.substring(30)
-			if(sub.length > 0) sub.split(" ").map(
+			if (sub.length > 0) sub.split(" ").map(
 				// they have 1 based index
-				_.toInt-1
-			) else Nil
+				_.toInt - 1
+			)
+			else Nil
 		}
-		val graph = Graph.create(Source.fromFile("src/main/resources/data.txt").getLines().map(strToInts).toSeq:_*)
-		def printResults(members:Seq[IMember]) = {
-			val results = members.foldLeft("")((a,b) => s"$a, ${graph.edgeCount-b.fitness}").substring(2)
+		val graph = Graph.create(Source.fromFile("src/main/resources/data.txt").getLines().map(strToInts).toSeq: _*)
+		def printResults(members: Seq[IMember]) = {
+			val results = members.foldLeft("")((a, b) => s"$a, ${graph.edgeCount - b.fitness}").substring(2)
 			log.info(s"{$results}")
 		}
 		log.info(graph.toString)
 
+		def parseResults(results: Seq[Computation[Array[IMember]]]) = {
+			results.flatMap(x => x.result).map(x => (graph.edgeCount - x.fitness)/2).sorted
+		}
 		/*
 		Implement MLS with the VSN local search. Use the first improvement version of
 		the local search. Generate 1000 local optima for a single run of MLS. Measure the
 		computational time required.
[…]

### 2016-01-29 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/algorithm/Evaluation.scala

@@ -15,6 +15,7 @@
 // along with this program.  If not, see <http://www.gnu.org/licenses/\>.
 package nl.jappieklooster.hw.ec.algorithm
 
+import nl.jappieklooster.hw.ec.experiment.Timer
 import nl.jappieklooster.hw.ec.model.Graph
 
 trait IHasFitness{

### 2016-01-29 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/algorithm/search/FidduciaMathesisSearch.scala

@@ -1,0 +1,133 @@
+package nl.jappieklooster.hw.ec.algorithm.search
+
+import Search._
+import nl.jappieklooster.hw.ec.model.{IMember, Graph, Vertex}
+
+/**
+ * Used as a link list like structure in the map.
+ * Now its possible to manipulate the list in place with a reference to this cell
+ * Its impossible to use std structures because they encapsulte this away.
+ *
+ * Note that the connections are mutable on intention.
+ *
+ * @param item the vertex
+ * @param next next cell
+ * @param previous previous cell
+ */
+class Cell(val item:Vertex, var gain:Int, var next:Option[Cell]=None, var previous:Option[Cell]=None){}
+
+/**
+ * In the original paper the bucket was an array indicate the gain that could be
+ * gotten. However I take enough with a map because it is constant-ish and I
+ * can generate it with std scala methods (although I could fold the map again into
+ * an array I don't think I'll save time by doing that)
+ *
+ * This bucket class serves as an encapusaltion of several related mutable fields
+ * Its a mutable sturcture based on the imutable map, however it encapsulates this
+ * in combination of the maxgain variable and a more usefull (for this algorithm atleast) pop method
+ * @param partitioning
+ * @param cells
+ * @param as
+ * @return
+ */
+class Bucket(private var cellMap:Map[Int,Cell]) {
+	private var maxGain = cellMap.keySet.max
+	def gain = maxGain
+	def pop:Option[Cell] = {
+		if(cellMap.isEmpty){
+			return None
+		}
+		if(!cellMap.contains(maxGain)){
+			maxGain = maxGain -1
+			return pop
+		}
+		val result = cellMap(maxGain)
+		cellMap -= maxGain
+		for(next <- result.next){
+			next.previous = None
+			cellMap += (maxGain -> next)
+		}
+		Some(result)
+	}
+	def isEmpty = cellMap.isEmpty
+}
+object Bucket{
+	def apply(partitioning:String, cells:Array[Cell], as:Char):Bucket = {
+		new Bucket(
+			cells.filter(x => partitioning(x.item.id) == as).groupBy(_.gain).
+			mapValues(
+				_.foldLeft[Cell](null)((a,b)=>{
+					if(a==null){
+						b
+					}else{
+						a.next = Some(b)
+						b.previous = Some(a)
+						b
+					}
+				})
+			)
+		)
+	}
+}
+/**
[…]

### 2016-01-29 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/algorithm/search/Search.scala

@@ -1,11 +1,10 @@
-package nl.jappieklooster.hw.ec.algorithm
+package nl.jappieklooster.hw.ec.algorithm.search
 
-import nl.jappieklooster.hw.ec.model.{Population, IMember, Graph}
+import nl.jappieklooster.hw.ec.algorithm.search.Search._
+import nl.jappieklooster.hw.ec.model.{Graph, IMember, Vertex}
 
 import scala.annotation.tailrec
 import scala.util.Random
-
-import Search._
 
 /**
  * Search, it depends on how you configure it if its global or local.
@@ -57,13 +56,10 @@ object Search {
 			def create(prev:String, x:Int):String = {
 				val one = randomGenePoint
 				val two = randomGenePoint
-				val acmp = prev(one)
-				val bcmp = prev(two)
-				if(acmp == bcmp){
+				if(prev(one)== prev(two)){
 					create(prev,x)
 				}else{
-					val result = swapChar(prev)(one, two)
-					result
+					swapChar(prev)(one, two)
 				}
 			}
 
@@ -78,36 +74,6 @@ object Search {
 		builder.setCharAt(one, partitioning(two))
 		builder.setCharAt(two, partitioning(one))
 		builder.mkString
-	}
-	/**
-	* The local search algorithm we use is the vertex
-	* swap neighborhood search (VSN). In VSN a neighboring solution is generated by swapping
-	* two vertices, one from each partitioning.
-	*/
-	def vertexSwapFirstImprovement(graph:Graph, memberFactory:String => IMember)(member:IMember):IMember = {
-		val partitioning = member.genes
-
-		// swapping of vertex in partitions
-		val swapVertex = swapChar(member.genes) _
-
-		// I just don't know a good way to do this with recursion or combbinators
-		for(index <- 0.to(graph.verteci.length-1)){
-			val vert = graph.verteci(index)
-			val options = vert.connections.filter(conindx => partitioning(index) != partitioning(conindx))
-
-			var lastMember:IMember = null // optimization, prevents member creation and valuation twice
-
-			// first improvement variant
-			val result = options.find(conindx => {
-				lastMember = memberFactory(swapVertex(conindx, index))
-				lastMember.fitness > member.fitness
-			})
-			if(!result.isEmpty){
-				return lastMember
-			}
-
-		}
-		member
 	}
 
 	class Probe extends SearchMethod{

### 2016-01-29 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/algorithm/search/VertexSwapFirstImprovement.scala

@@ -1,0 +1,38 @@
+package nl.jappieklooster.hw.ec.algorithm.search
+
+import nl.jappieklooster.hw.ec.model.{Graph, IMember}
+
+import Search._
+
+/**
+ * The local search algorithm we use is the vertex
+ * swap neighborhood search (VSN). In VSN a neighboring solution is generated by swapping
+ * two vertices, one from each partitioning.
+ */
+class VertexSwapFirstImprovement(graph:Graph, memberFactory:String => IMember) extends SearchMethod{
+
+	override def apply(member: IMember): IMember = {
+		val partitioning = member.genes
+		// swapping of vertex in partitions
+		val swapVertex = swapChar(member.genes) _
+
+		// I just don't know a good way to do this with recursion or combbinators
+		for(index <- 0.to(graph.verteci.length-1)){
+			val vert = graph.verteci(index)
+			val options = vert.connections.filter(conindx => partitioning(index) != partitioning(conindx))
+
+			var lastMember:IMember = null // optimization, prevents member creation and valuation twice
+
+			// first improvement variant
+			val result = options.find(conindx => {
+				lastMember = memberFactory(swapVertex(conindx, index))
+				lastMember.fitness > member.fitness
+			})
+			if(!result.isEmpty){
+				return lastMember
+			}
+
+		}
+		member
+	}
+}

### 2016-01-29 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/experiment/DataTable.scala

@@ -7,15 +7,37 @@ import org.apache.commons.math3.stat.descriptive.SummaryStatistics
 
 object DataTable {
 
-	def createTable(tableRow: TableRow*): String ={
+	val significance = 0.95D
+	val br = System.lineSeparator()
+	def toPgfRow(row:TableRow) = s"0	${row.name}	${row.firstY}	${row.secondY-row.firstY}	${row.firstErr} 0		0	${row.secondErr}"
+	def createPgfPlotTable(tableRow: TableRow*): String ={
 		val toprow = "zero    name        firsty      secondy     errbotstart     errbotend    errtopstart    errtopend"
-		tableRow.foldLeft(toprow)((a,b)=>a+System.lineSeparator() +b.toString)
+		tableRow.foldLeft(toprow)((a,b)=>a+ br +toPgfRow(b))
 	}
 
-	def createRow(name:String, data:Seq[Double]):TableRow ={
+	def createLatexResultTable(tableRow: TableRow*): String ={
+		val toprow = s"\\begin{tabular}{lllll}$br name & low & high & error bottom & error top \\\\ \\toprule"
+
+		tableRow.foldLeft(toprow)((a,b)=> a + br +
+		  f"${b.name}%s & ${b.firstY}%4.2f & ${b.secondY}%4.2f & ${b.firstErr}%4.2f & ${b.secondErr}%4.2f \\\\") +
+		"\\\\bottomrule" + br + "\\end{tabular}"
+	}
+	def createLatexTTestTable(tableRow: TableRow*): String ={
+
+		val toprow = s"\\begin{tabular}{${
+			tableRow.foldLeft("l")((a,b)=> a+"l")
+		}}$br ttests & ${
+			val r = tableRow.foldLeft("")((a,b)=>a+b.name+" &")
+			r.substring(0,r.length-1)
+		} \\\\ \\toprule"
+
+		toprow + "\\end{tabular}"
+	}
+
+	def createRow(name:String, data:Seq[Float]):TableRow ={
 		// Build summary statistics of the dataset "data"
 		val stats = new SummaryStatistics()
-		data.foreach(stats.addValue)
+		data.foreach(stats.addValue(_))
 
 		def calcMeanCI(confidence:Double) = {
 			// Create T Distribution with N-1 degrees of freedom
@@ -26,15 +48,16 @@ object DataTable {
 			critVal * stats.getStandardDeviation / Math.sqrt(stats.getN)
 		}
 
-		// Calculate 95% confidence interval
[…]

### 2016-01-29 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/model/Member.scala

@@ -15,10 +15,11 @@
 // along with this program.  If not, see <http://www.gnu.org/licenses/\>.
 package nl.jappieklooster.hw.ec.model
 
+import nl.jappieklooster.hw.ec.algorithm.search.Search
 import org.slf4j.LoggerFactory
 
 import scala.util.Random
-import nl.jappieklooster.hw.ec.algorithm.{Search, Search$, IHasFitness}
+import nl.jappieklooster.hw.ec.algorithm.IHasFitness
 
 trait IMember extends IHasFitness with Genetic
 case class Member(valuation:Float, gen:String) extends IMember{

### 2016-01-29 evolutionaryComputation/2-practical/code/src/test/scala/nl/jappieklooster/hw/ec/tests/SearchTest.scala

@@ -1,9 +1,10 @@
 package nl.jappieklooster.hw.ec.tests
 
 import nl.jappieklooster.hw.ec.algorithm.Evaluation
+import nl.jappieklooster.hw.ec.algorithm.search.Search
 import org.scalatest.{Matchers, FlatSpec}
 import nl.jappieklooster.hw.ec.model.{Member, Graph}
-import nl.jappieklooster.hw.ec.algorithm.Search._
+import Search._
 
 class SearchTest extends FlatSpec with Matchers{
 

### 2016-01-29 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/Main.scala
added student ttest and cleared some other stuff up

@@ -42,6 +42,9 @@ object Main{
 
 	val glsPopSizes = List(5)
 	val mlsStartsize = 10
+	val ilsUntill = 3
+
+	val experimentCount = 8
 
 	def main(args:Array[String]) {
 		log.info("starting graph bipartitioning")
@@ -98,7 +101,7 @@ object Main{
 		def runILS(times: Int) = {
 			val starts = Population.createEqualOnesZeros(factory, graph.verteci.length, 1)
 			log.info(s"ils start with: ${starts.head}")
-			val results = 1.to(5).map { x =>
+			val results = 1.to(ilsUntill).map { x =>
 				val experiment = new Experiment({
 					// can't do this outside of the experiment, because the probe
 					// needs to be added per experiment.
@@ -156,17 +159,24 @@ object Main{
 				(createRow(results.map(_.seconds.toFloat)), createRow(parseResults(results)))
 			})
 		}
-		val runtimes = 8
-		val result =  List(runMLS(runtimes)) ::: runILS(runtimes).toList ::: runGLS(runtimes)
+		val result =  List(runMLS(experimentCount)) ::: runILS(experimentCount).toList ::: runGLS(experimentCount)
 
-		Plot.write("runtimes.table", DataTable.createPgfPlotTable(
+		import DataTable._
+		Plot.write("runtimes.table", createPgfPlotTable(
 			result.map(x => x._1): _*
-		))
-		Plot.write("runtimes.textable", DataTable.createLatexResultTable(result.map(x=>x._1):_*))
-		Plot.write("results.table", DataTable.createPgfPlotTable(
+		)
+		)
+		Plot.write("runtimes.textable",
+			createLatexResultTable(result.map(_._1):_*)+ br + br +
+			createLatexTTestTable(result.map(_._1):_*)
+		)
+		Plot.write("results.table", createPgfPlotTable(
 			result.map(x => x._2): _*
 		))
-		Plot.write("results.textable", DataTable.createLatexResultTable(result.map(x=>x._2):_*))
+		Plot.write("results.textable",
+			createLatexResultTable(result.map(_._2):_*)+ br+ br +
+			createLatexTTestTable(result.map(_._2):_*)
+		)
 	}
 
 }

### 2016-01-29 evolutionaryComputation/2-practical/code/src/main/scala/nl/jappieklooster/hw/ec/experiment/DataTable.scala

@@ -3,6 +3,7 @@ package nl.jappieklooster.hw.ec.experiment
 import org.apache.commons.math3.distribution.TDistribution
 import org.apache.commons.math3.exception.MathIllegalArgumentException
 import org.apache.commons.math3.stat.descriptive.SummaryStatistics
+import org.apache.commons.math3.stat.inference.TTest
 
 
 object DataTable {
@@ -20,18 +21,33 @@ object DataTable {
 
 		tableRow.foldLeft(toprow)((a,b)=> a + br +
 		  f"${b.name}%s & ${b.firstY}%4.2f & ${b.secondY}%4.2f & ${b.firstErr}%4.2f & ${b.secondErr}%4.2f \\\\") +
-		"\\\\bottomrule" + br + "\\end{tabular}"
+		"\\\\ \\bottomrule" + br + "\\end{tabular}"
 	}
 	def createLatexTTestTable(tableRow: TableRow*): String ={
 
 		val toprow = s"\\begin{tabular}{${
 			tableRow.foldLeft("l")((a,b)=> a+"l")
-		}}$br ttests & ${
-			val r = tableRow.foldLeft("")((a,b)=>a+b.name+" &")
+		}}$br tTests & ${
+			val r = tableRow.foldLeft("")((a,b)=>s"$a {\\small ${b.name}} &")
 			r.substring(0,r.length-1)
-		} \\\\ \\toprule"
+		} \\\\ \\toprule" + br
 
-		toprow + "\\end{tabular}"
+		val tester = new TTest
+		val rows = tableRow.foldLeft("")((prevRows,currentRow)=>{
+			val td = " & "
+			val varColls = tableRow.foldLeft(td)((prevCells,currentCell)=>{
+				val cell = if(currentCell == currentRow){"x"}else{
+					if(tester.tTest(currentRow.data.map(_.toDouble).toArray, currentCell.data.map(_.toDouble).toArray, 1- significance)){
+						"T"
+					}else{
+						"F"
+					}
+				}
+				prevCells + cell+  td
+			})
+			prevRows+currentRow.name+varColls.substring(0,varColls.length-td.length) + "\\\\" + br
+		})
+		toprow + rows + "\\end{tabular}"
 	}
 
 	def createRow(name:String, data:Seq[Float]):TableRow ={
@@ -53,11 +69,11 @@ object DataTable {
 		val lower = (stats.getMean - ci).toFloat
 		val upper = (stats.getMean + ci).toFloat
 		val sorted = data.sorted
[…]

### 2016-01-30 evolutionaryComputation/2-practical/report.tex
added student ttest and cleared some other stuff up
adding more text and both tables

@@
 
+\subsection{MLS}
+MLS basically involved implementing the initial vertex swap first improvement
+local search and tying it into the existing framework. I did this by just
+using the existing createPopMethods with a new solution generator. Then
+I just mapped this population with the local search (so the type is
+$IMember \to IMember$, but the solution changes).
+
+\subsection{ILS}
 While implementing ILS, I had it turn out results much worse than MLS.
@@
 underperforming to MLS, which shouldn't be the case according to my memory
-from the classes. Now its overperforming quite drastically so I'm happy.
+from class. Now its overperforming quite drastically so I'm happy.
 
+\subsection{GLS}
 I implemented GLS with the evolution class. The only thing I had to do really
@@
 average scores.
+
+Here I also started to think about my own topic to investigate, at first
+I just wanted to do varying population sizes and varying selection methods.
+But I thinks this takes a little to much time and it doesn't quite fit
+on the tTest table so I decided the varying popsizes (spoilers: it produces
+better solution at the cost of exponentially more time).
 
@@
 
+\subsubsection{Graph pruning}
+I decided to use a prune graph to speed up the algortihm. In the beginning I
+create 2 graphs. One normal and one pruned, The pruning works by selecting
+comparing all connections against the id, if lower we ignore the connection.
+To make sure we only do this on doubly linked graphs 2 graphs are created.
+If the prune graph has exactly half the score of a random solution we assume
+its a doubly linked graph so we can prune it.
+The pruning effectivally halfs the ammount of connections that need to be
+checked by each itteration (this happens in the evaluation). The
+FidduciaMathesis has even more gain because it has a lot of book keeping to
+do with the connections (although it tries to reduce this with the freecell
+array, having a smaller initial graph is even better).
+
+\subsection{Fidducia Mathesis}
+I was somewhat sceptical of my local search implementation. I had quite 
+different results from my classmates so to compare I decided to implement
+the fidducia mathesis local search to. It was quite a lot of work but it
+truly is significantly faster.
+
+Finding the paper was somewhat difficult as the original was behind a paywal.
+But somoene put a copy on github so I just used that one. The algorithm 
+involves a lot of cleaver book keeping:
+
 \section{TODO}
@@
 \section{Results}
-\subsection{Time}
-\pgfplotstableread{runtimes.table}\dataTable
+\subsection{Fiduccia Mathesis}
+\subsubsection{Time}
+\pgfplotstableread{fidRuntimes.table}\dataTable
 \begin{tikzpicture}
@@
 	xlabel={Algorithm},
-	xticklabels from table={runtimes.table}{name},
+	xticklabels from table={fidRuntimes.table}{name},
 	xtick=data,
@@
 ]
-
 % Layers needed to get error bars on top of lower blue rectangle
@@
 
-\input{runtimes.textable}
+\input{fidRuntimes.textable}
 
-\subsection{Optima}
-\pgfplotstableread{results.table}\dataTable
+\subsubsection{Optima}
+\pgfplotstableread{fidResults.table}\dataTable
 \begin{tikzpicture}
@@
 	xlabel={Algorithm},
-	xticklabels from table={results.table}{name},
+	xticklabels from table={fidResults.table}{name},
 	xtick=data,
@@
 \\
-\input{results.textable}
+\input{fidResults.textable}
+
+\subsection{Vertex swap first improvemnt}
+\subsubsection{Time}
+\pgfplotstableread{vertRuntimes.table}\dataTable
+\begin{tikzpicture}
+\pgfplotsset{set layers}
+\begin{axis}[
+	ybar stacked,
+	xlabel={Algorithm},
+	xticklabels from table={vertRuntimes.table}{name},
+	xtick=data,
+	x tick label style={rotate=90,anchor=east},
+	ylabel={Time}
+]
+% Layers needed to get error bars on top of lower blue rectangle
+\addplot[color=white]
+table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};
+
+\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
+table[x expr=\coordindex, y=zero, y error minus=errbotstart, y error plus=errbotend]{\dataTable};
+
+\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
+table[x expr=\coordindex, y=secondy, y error minus=errtopstart, y error plus=errtopend]{\dataTable};
+
+\end{axis}
+
+\end{tikzpicture}
+
+\input{vertRuntimes.textable}
+
+\subsubsection{Optima}
+\pgfplotstableread{vertResults.table}\dataTable
+\begin{tikzpicture}
+\pgfplotsset{set layers}
+\begin{axis}[
+	ybar stacked,
+	xlabel={Algorithm},
+	xticklabels from table={vertResults.table}{name},
+	xtick=data,
+	x tick label style={rotate=90,anchor=east},
+	ylabel={Fitness (less is better)}
+]
+% Layers needed to get error bars on top of lower blue rectangle
+\addplot[color=white]
+table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};
+
+\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
+table[x expr=\coordindex, y=zero, y error minus=errbotstart, y error plus=errbotend]{\dataTable};
+
+\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
+table[x expr=\coordindex, y=secondy, y error minus=errtopstart, y error plus=errtopend]{\dataTable};
+
+\end{axis}
+
+\end{tikzpicture}
+\\
+\input{vertResults.textable}
 \end{empfile}

### 2016-01-31 evolutionaryComputation/2-practical/report.tex
talking about findings

@@
 \usepackage{pgfplotstable}
+\usepackage{seqsplit}
 
@@
 But somoene put a copy on github so I just used that one. The algorithm 
-involves a lot of cleaver book keeping:
+involves a lot of cleaver book keeping. First of all the concept of gain is
+introduced, this indicates how much gain there will be from moving a vertex
+to the other partition. Then the vertexes are put into an array with as index
+their gain. If there are more verteci at a particular index then they're
+put into an linked list. To support this I introduced a container class
+Cell that holds the vertex and also has an optional previous and next
+element.
+These cells are also put in another array based on their vertex ID. This
+array is used for constant speed access when the connections of a vert their
+gain needs to be updated after moving it to the other partition.
+So you can just unty the ellemnt of the linked list (or if it has no
+previous in my case its in the fornt of the bucket and I remove it from
+there and reinsert the next element).
 
-\section{TODO}
-MLS with several results
-GLS -> I think the localsearch will recursively unecesrly decend into its members
-because the local search is packed in the member factory. Maybe remove the
-local search from the memberfactory?
+After knowing the basic datastructers the implementation is quite trivial
+(although I had some tedious bugs such as index out of bounds errors or
+next elements being wrongly updated).
+It doesn't find the local optimum in one go, 4 to 5 iterations are
+generally required, which is still much and much better than the first
+improvment vertex swap.
 
+\section{Own research topic}
+I had several ideas for my own research to investigate, first of all I wanted
+to do the really easy to implement ones, such as selection pressure and pop
+sizes for GLS. But later, when waiting for results, I got curious about the
+other local search I heard about in class. With several days left I decided
+to try and implement the Fidducia Mathesis local search heuristic.
+I think I've succeeded so my final research question is how does the 
+Fidducia Mathesis algorithm compare to first
+improvement vertex swap?
+
+As can be seen in the results section the Fidducia Mathesis algorithm is
+much faster than first improvement vertex swap. The main reason for this is
+that a single pass will climb very steeply in valuation. It also appears
+to produce better results, I think this is becuase highest gain is considered
+first.
+
+My other research topic is selection pressure, turns out that tournament is 
+the best for GLS in case of fidducia mathesis, but it also takes the mot 
+amount of time.
+
+\subsection{Statistics}
+The error marges in the graph are the maximum range of the results so these
+are round numbers when substracted from the low or added to the high. The low
+and high values are calulated with something called a confidence interval.
+It alligns pretty well with the tTest results so I guess its ok.
+
+\newpage
 \section{Results}
+Fidducia Mathesis dominates vertex swap first improvement in both runtime and
+optima found. Combined with GLS it gets such a good score that I became
+skeptical and decided to test the anwser more thoroughly.
+
+Timewise Fidducia Mathesis dominates vertex swap first improvement. This
+is because it starts with an incredibally steep hillclimb that flattens
+out quite quicly. So you just need to pass over the string 4 or 5
+times and then you are at the local optimum. Vertex swap first improvemt
+increases linearly however. So it takes a lot more string itterations
+and evaluations to get there.
+
+The improvements found by Fidducia Mathesis are also a lot better, especially
+the GLS variants. I'm not quite sure why this is the case, in fact I
+added a paragraph scepticism and put a gene in there that found a solution
+with 4 crossings on the original graph. Vertex swap first improvement gets
+noway near to that. The lowest it found is like 280 crossings on the original
+graph. I think the GLS variant doesn't find any better is because I always
+start at index 0 of the string and take any improvment I can find, this
+means the search space is biased. A random start could solve this issue but
+then the function is no longer pure. The other bias vertex swap first
+improvement has is that it either goes from left to write or from right to
+left, you can't really do both. Fidducia Mathesis doesn't suffer from both
+of these problem because it just picks whatever improvement has the maximum
+gain (while keeping balance in mind), so its biased towards maximum gain,
+which isn't a bad bias at all.
+
+Something I found interesting is that the Fiduccia Mathesis local search
+doesn't seem to respond well to ILS it just stays in the range of the lower
+error marging of MLS. To verify this I cranked up the validation number to
+100 (ie 100 times no better improvement found the local search end), and
+increased the pertubation to 7 (because initially it looked like 1 to 5
+where slowly going down).
+However when you look to vertex swap first improvement you can see ILS
+responding violently. In fact ILS-1 is tTest similiar to GLS:Killparents.
+
+Selection method seems to be statistically irrelevant for the optimum in
+Fidducia mathesis, although you pay a much bigger time price with both
+tournament and kill parents selection. Tournament also touches the best 
+result, but in other runs the other two did this to.
+
+\newpage
 \subsection{Fiduccia Mathesis}
@@
 
+\newpage
 \subsubsection{Optima}
@@
 
+\paragraph{Sceptisicm} So these results seem a little to good to be true honestly,
+especially compared to vertex swap first improvement. So I decided to catch
+one of the results of GLS and throw it trough a bunch of tests to see if I 
+could find anything wrong with it. The result is this one and it appears
+to be OK with a fitness score of 2 (or 4 if you use the doubly connected graph).
+I also checked if there was an equal amount of zeros and ones just to be sure.
+
+\seqsplit{11100000101001111000000100110111101010000000110011100110011110101110110010111110010010010011000000101100111010011100001000111000110111110101110000010011111001110110100110110101110010101011000101000111100010111100010001100100000110001001010010101001110010111110001001010000001111011111111101101000101100000100011011101111001010010111001101101111100001011000001100100001011101000101111000101011101001001010101011011000110000110111111101101000001111000010010001111101000101000010111100101011100111100101}
+
+There is a good chanche this is a global optimum.
+
+\newpage
 \subsection{Vertex swap first improvemnt}

### 2016-01-31 evolutionaryComputation/2-practical/report.tex
Spelling

--- 2016-01-30 evolutionaryComputation/2-practical/report.tex
@@
 \section{Assignment}
-Graphbipartitioning is dividing a graph into 2, with as little edges between
+Graph bi-partitioning is dividing a graph into 2, with as little edges between
 both graphs as possible.
@@
 \section{Methodology}
-I coppied all the code from the previous practical into this new folder.
-Then I went onto a delete rage. Most code was deleted except the Expirment
-class, this will probably still be usefull.
+I copied all the code from the previous practical into this new folder.
+Then I went onto a delete rage. Most code was deleted except the Experiment
+class, this will probably still be useful.
 
-I use git, so it doens't really matter to accidently delete to much. After
+I use git, so it doens't really matter to accidentally delete to much. After
 that I changed the data, I removed everything which was useless. Now the file
 only contains the connection numbers. This makes parsing the file a lot
-easier. I don't want to do advanced pattern matching when I can quickly modfy
+easier. I don't want to do advanced pattern matching when I can quickly modify
 the file with a few vim macros. (in this case I just did a `30x').
@@
 While implementing ILS, I had it turn out results much worse than MLS.
-The initial implmentation would just stop if it found a worse local optimum.
-The description sais it should go back to the previous optimum. So I guess
+The initial implementation would just stop if it found a worse local optimum.
+The description says it should go back to the previous optimum. So I guess
 that you should try again after that? But then I wonder how often you should
@@
 I got much better results with a system where the local search got several chances.
-If it failed it would fallback, and then try again. 
+If it failed it would fall back, and then try again.
 With chances around 30 I managed to get up and around 1900.
-When I reset the chance counter on a succesfull find
+When I reset the chance counter on a successful find
 I managed to get consistently 2200, one result was as high as 2310.
@@
 This current search for improvement was just based on the fact that ILS was
-underperforming to MLS, which shouldn't be the case according to my memory
-from class. Now its overperforming quite drastically so I'm happy.
+under performing to MLS, which shouldn't be the case according to my memory
+from class. Now its over-performing quite drastically.
 
@@
 
-I wanted to see if the GLS popsize and selection pressure would have any
+I wanted to see if the GLS population size and selection pressure would have any
 effect on the average fitness found. I suspect that a bigger population size
@@
 But I thinks this takes a little to much time and it doesn't quite fit
-on the tTest table so I decided the varying popsizes (spoilers: it produces
-better solution at the cost of exponentially more time).
-
-I chose to investigate with population sizes of 25, 50 and 75. I didn't want
-to wait to long on this and the runtime for 50 was approaching ~10 minutes
-for the slowest. Luckily I have a powerfull laptop and can use paralization.
-
-Doing this 30 times to cancel the staostic effect would result into ~5 hours.
+on the tTest table so I decided to drop the varying pop-sizes.
 
@@
 After discussing it for a while it turned out that I was counting double. However
-This didn't make sense because I just used scala combinators to count the 
+This didn't make sense because I just used Scala combinators to count the
 elements. Basically this means that there would have to be a bug in the std
@@
 \subsubsection{Graph pruning}
-I decided to use a prune graph to speed up the algortihm. In the beginning I
+I decided to use a prune graph to speed up the algorithm. In the beginning I
 create 2 graphs. One normal and one pruned, The pruning works by selecting
@@
 its a doubly linked graph so we can prune it.
-The pruning effectivally halfs the ammount of connections that need to be
-checked by each itteration (this happens in the evaluation). The
-FidduciaMathesis has even more gain because it has a lot of book keeping to
-do with the connections (although it tries to reduce this with the freecell
-array, having a smaller initial graph is even better).
+The pruning effectively halves the amount of connections that need to be
+checked by each iteration (this happens in the evaluation). The
+Fiduccia Mattheyses has even more gain because it has a lot of book keeping to
+do with the connections.
 
-\subsection{Fidducia Mathesis}
+\subsection{Fiduccia Mattheyses}
 I was somewhat sceptical of my local search implementation. I had quite 
 different results from my classmates so to compare I decided to implement
-the fidducia mathesis local search to. It was quite a lot of work but it
+the Fiduccia Mattheyses local search to. It was quite a lot of work but it
 truly is significantly faster.
 
-Finding the paper was somewhat difficult as the original was behind a paywal.
-But somoene put a copy on github so I just used that one. The algorithm 
-involves a lot of cleaver book keeping:
+Finding the paper was somewhat difficult as the original was behind a paywall.
+But someone put a copy on GitHub so I just used that one. The algorithm 
+involves a lot of cleaver book keeping. First of all the concept of gain is
+introduced, this indicates how much gain there will be from moving a vertex
+to the other partition. Then the vertexes are put into an array with as index
+their gain. If there are more vertices at a particular index then they're
+put into an linked list. To support this I introduced a container class
+Cell that holds the vertex and also has an optional previous and next
+element.
+These cells are also put in another array based on their vertex ID. This
+array is used for constant speed access when the connections of a vertex their
+gain needs to be updated after moving it to the other partition.
+So you can just untie the element of the linked list (or if it has no
+previous in my case its in the front of the bucket and I remove it from
+there and reinsert the next element).
 
-\section{TODO}
-MLS with several results
-GLS -> I think the localsearch will recursively unecesrly decend into its members
-because the local search is packed in the member factory. Maybe remove the
-local search from the memberfactory?
+After knowing the basic data structures the implementation is quite trivial
+(although I had some tedious bugs such as index out of bounds errors or
+next elements being wrongly updated).
+It doesn't find the local optimum in one go, 4 to 5 iterations are
+generally required, which is still much and much better than the first
+improvement vertex swap.
 
+\section{Own research topic}
+I had several ideas for my own research to investigate, first of all I wanted
+to do the really easy to implement ones, such as selection pressure and pop
+sizes for GLS. But later, when waiting for results, I got curious about the
+other local search I heard about in class. With several days left I decided
+to try and implement the Fiduccia Mattheyses local search heuristic.
+I think I've succeeded so my final research question is how does the 
+Fiduccia Mattheyses algorithm compare to first
+improvement vertex swap?
+
+As can be seen in the results section the Fiduccia Mattheyses algorithm is
+much faster than first improvement vertex swap. The main reason for this is
+that a single pass will climb very steeply in valuation. It also appears
+to produce better results, I think this is because highest gain is considered
+first.
+
+My other research topic is selection pressure, turns out that tournament is
+the best for GLS in case of Fiduccia Mattheyses, but it also takes the mot
+amount of time.
+
+\subsection{Statistics}
+The error margins in the graph are the maximum range of the results so these
+are round numbers when subtracted from the low or added to the high. The low
+and high values are calculated with something called a confidence interval.
+It aligns pretty well with the tTest results so I guess its OK.
+
+\newpage
 \section{Results}
-\subsection{Fiduccia Mathesis}
+Fiduccia Mattheyses dominates vertex swap first improvement in both runtime and
+optima found. Combined with GLS it gets such a good score that I became
+skeptical and decided to test the answer more thoroughly.
+
+Time wise Fiduccia Mattheyses dominates vertex swap first improvement. This
+is because it starts with an incredibly steep hill climb that flattens
+out quite quickly. So you just need to pass over the string 4 or 5
+times and then you are at the local optimum. Vertex swap first improvement
+increases linearly however. So it takes a lot more string iterations
+and evaluations to get there.
+
+The improvements found by Fiduccia Mattheyses are also a lot better, especially
+the GLS variants. I'm not quite sure why this is the case, in fact I
+added a paragraph scepticism and put a gene in there that found a solution
+with 4 crossings on the original graph. Vertex swap first improvement gets
+noway near to that. The lowest it found is like 280 crossings on the original
+graph. I think the GLS variant doesn't find any better is because I always
+start at index 0 of the string and take any improvement I can find, this
+means the search space is biased. A random start could solve this issue but
+then the function is no longer pure. The other bias vertex swap first
+improvement has is that it either goes from left to write or from right to
+left, you can't really do both. Fiduccia Mattheyses doesn't suffer from both
+of these problem because it just picks whatever improvement has the maximum
+gain (while keeping balance in mind), so its biased towards maximum gain,
+which isn't a bad bias at all.
+
+Something I found interesting is that the Fiduccia Mattheyses local search
+doesn't seem to respond well to ILS it just stays in the range of the lower
+error margin of MLS. To verify this I cranked up the validation number to
+100 (100 times no better improvement found the local search end), and
+increased the perturbation to 7 (because initially it looked like 1 to 5
+where slowly going down).
+However when you look to vertex swap first improvement you can see ILS
+responding violently. In fact ILS-1 is tTest similar to GLS:Killparents.
+
+Selection method seems to be statistically irrelevant for the optimum in
+Fiduccia Mattheyses, although you pay a much bigger time price with both
+tournament and kill parents selection. Tournament selection does seem to
+be best for vertex swap first improvement, but again you pay a time price
+for this.
+
+\newpage
+\subsection{Fiduccia Mattheyses}
 \subsubsection{Time}
@@
 
-\subsection{Vertex swap first improvemnt}
+\paragraph{Scepticism} So these results seem a little to good to be true honestly,
+especially compared to vertex swap first improvement. So I decided to catch
+one of the results of GLS and throw it trough a bunch of tests to see if I 
+could find anything wrong with it. The result is this one and it appears
+to be OK with a fitness score of 2 (or 4 if you use the doubly connected graph).
+I also checked if there was an equal amount of zeros and ones just to be sure.
+
+\seqsplit{11100000101001111000000100110111101010000000110011100110011110101110110010111110010010010011000000101100111010011100001000111000110111110101110000010011111001110110100110110101110010101011000101000111100010111100010001100100000110001001010010101001110010111110001001010000001111011111111101101000101100000100011011101111001010010111001101101111100001011000001100100001011101000101111000101011101001001010101011011000110000110111111101101000001111000010010001111101000101000010111100101011100111100101}
+
+There is a good chance this is a global optimum.
+
+\newpage
+\subsection{Vertex swap first improvement}
 \subsubsection{Time}

### 2016-01-31 evolutionaryComputation/2-practical/report.tex
added a bunch of text!

@@
 	\begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
+
+% "define" Scala
+\lstdefinelanguage{scala}{
+  morekeywords={abstract,case,catch,class,def,%
+    do,else,extends,false,final,finally,%
+    for,if,implicit,import,match,mixin,%
+    new,null,object,override,package,%
+    private,protected,requires,return,sealed,%
+    super,this,throw,trait,true,try,%
+    type,val,var,while,with,yield},
+  otherkeywords={=>,<-,<\%,<:,>:,\#,@},
+  sensitive=true,
+  morecomment=[l]{//},
+  morecomment=[n]{/*}{*/},
+  morestring=[b]",
+  morestring=[b]',
+  morestring=[b]"""
+}
 \begin{document}
@@
 \author{Jappie Klooster}
-\title{Graph bipartitioning and you}
+\title{Graphs, partitioning and you}
 \maketitle
@@
 \section{Assignment}
-Graph bi-partitioning is dividing a graph into 2, with as little edges between
-both graphs as possible.
+Graph bi-partitioning is dividing a graph into 2 partitions, 
+with as little crossings between both partitionings as possible.
+
+So we try to reduce the amount of crossing by modifiying which vertex resides
+in what partition. The graph itself is however not modified. The amount of
+vertexes that reside in on partition should also be even to the other partition.
 
@@
 
+\newpage
+\subsubsection{Reading the graph}
+The first thing I did was reading the graph into a memory structure. That 
+looks like this:
+
+\begin{figure}[ht!]
+\centering
+\begin{emp}[classdiag](20,20)
+Class.graph("Graph")("+verteci:Array[Vertex]", "+edgeCount:Int")();
+Class.vertex("Vertex")("+id:Int", "+connections:Seq[Int]")();
+
+leftToRight.top(30)(graph,vertex);
+
+drawObjects(graph,vertex);
+
+link(association)(graph.e -- vertex.w);
+item(iAssoc)("*")(obj.se = vertex.w);
+
+\end{emp}
+\caption{Case classes}
+\end{figure}
+
+So the file is stored in an imutable memory structure. This is possible
+because the connections are just the number that essently work as pointers
+if you have the graph at hand.
+
+\subsubsection{Solution generation}
+Then I went on doing the generation of the solutions. I did this by creating
+a random string of half the graph length, then copying this string and invert
+it. This ensured there were as many ones as zeros. Then I combined these
+two strings and randomly shuffled them:
+
+\lstset{language=scala}
+\begin{lstlisting}
+	val random = Random
+	def randomGene = random.nextInt(2).toString
+	def randomGenes(geneLength:Int) = 1.to(geneLength).
+		foldLeft(""){(b, x) => b + OffspringGenerator.randomGene}
+	def randomBalancedGenes(geneLength:Int):String = {
+		if(geneLength % 2 == 1){
+			throw new Exception("Doesn't work with odd genelength")
+		}
+		val initial = randomGenes(geneLength/2)
+		random.shuffle(
+			inverseMirror(initial).toCharArray.toSeq
+		).mkString
+	}
+	def flipBit(c:Char):Char = inverseResult(c,'1','0')
+	def inverseMirror(input:String):String ={
+		input + input.map(flipBit)
+	}
+\end{lstlisting}
+
+\subsubsection{Valuation}
+After that I did the evaluation function. Which was just folding the vertexes
+into a float and then counting the string indexes that did match. I think
+I did this wrong by mistake at first, but I realized it was advantageous to
+count the connections that are the same instead of different. Because now
+we want to maximize again as in the previous assignment, which made it
+possible to leave the selection as is (so I don't have to modify that when
+implementing GLS). To represent the internal valuation 
+(which is trying to be maximized) as the minimizing
+problem I simply do $graph.edgecount-valuation$.
+
+\subsubsection{Local search}
+Finally I designed the search class, which was first called LocalSearch but
+was later generalized to Search:
+
+\begin{figure}
+\begin{emp}[classdiag](20,20)
+Class.search("Search")(
+	"-method:IMember=>IMember", 
+	"-stopCondition:(IMember, IMember)=>Boolean",
+	"-decideResult:(IMember, IMember)=>IMember"
+	)(
+	"+search(previous:IMember):IMember"
+	);
+
+drawObjects(search);
+\end{emp}
+\caption{Search}
+\end{figure}
+
+A very compact class indeed, we execute method on previous, check the
+stopcondition if true we decide result else we call ourself (tail) recursively.
+The stopcondition and decide result are by default equal and return the result
+of the method. Why these functions are configurable will be discussed in the 
+ILS chapter.
+Then the method is either filled with vertexSwapFirstImprovement or
+FiducciaMattheysesSearch.
+
 \subsection{ILS}
@@
 When I reset the chance counter on a successful find
-I managed to get consistently 2200, one result was as high as 2310.
+I managed to get consistently 2200, one result was as high as 2310. Note:
+I was maximizing here and still counting double. For the final run I put
+the validation count up to 100. So before we decide something is at a local
+optimum it first has to return the same result 100 times.
 
-I don't think there is a way to do it any better without drastically making
-the algorithm more complex.
-This current search for improvement was just based on the fact that ILS was
-under performing to MLS, which shouldn't be the case according to my memory
-from class. Now its over-performing quite drastically.
+\subsubsection{Implementation}
+At first I started to wrote a separate ILS class for search below the then
+local search class. However I noticed the code I was producing looked 
+suspiciously similar to what I already had in local search. So I decided
+to try and combine the functionalities by making LocalSearch more configurable
+and I succeeded. The end result was the Search class (I had to rename
+the local search class because it was no longer just doing local search).
 
+To explain how it works the first thing one should do is notice how the method
+of search has the same prototype as the search method of the search class.
+This means you can nest a search method of one search class inside another
+search class. So we now have one inner search class doing the local search
+and one outer search class doing iteration and perturbation of solutions.
+
+Actually the perturbation is also nested inside a method like prototype and
+I use function composition to let the perturbation be executed just before
+the local search can be executed.
+
+as stopcondition of the outer search we have an isWorse function inside a
+resetretryOnBetter function. this function has also access to the retryOnResult
+object which counts to the validation (and can be rest on a better result).
 \subsection{GLS}
@@
 
-Here I also started to think about my own topic to investigate, at first
+So I also started to think about my own topic to investigate, at first
 I just wanted to do varying population sizes and varying selection methods.
@@
 This didn't make sense because I just used Scala combinators to count the
-elements. Basically this means that there would have to be a bug in the std
-Scala library which would seem highly unlikely. I knew I was counting double
-because all my results where even numbers. After thinking about it for a while
+elements. This would mean that there should be a bug in the std
+Scala library which seems highly unlikely. I knew I was counting double
+because all my results were even numbers. After thinking about it for a while
 I decided to investigate the graph file, it turns out that every connection
@@
 I decided to use a prune graph to speed up the algorithm. In the beginning I
-create 2 graphs. One normal and one pruned, The pruning works by selecting
+create 2 graphs. One normal and one pruned, The pruning works by
 comparing all connections against the id, if lower we ignore the connection.
 To make sure we only do this on doubly linked graphs 2 graphs are created.
-If the prune graph has exactly half the score of a random solution we assume
-its a doubly linked graph so we can prune it.
+If the pruned graph has exactly half the score of a random solution we assume
+its a doubly linked graph so we can prune it. I know its not a
+guaranteed to work.
 The pruning effectively halves the amount of connections that need to be
@@
 the Fiduccia Mattheyses local search to. It was quite a lot of work but it
-truly is significantly faster.
+truly is significantly faster. Finding the paper was somewhat difficult
+as the original was behind a paywall.  But someone put a copy on GitHub so
+I just used that one.
 
-Finding the paper was somewhat difficult as the original was behind a paywall.
-But someone put a copy on GitHub so I just used that one. The algorithm 
-involves a lot of cleaver book keeping. First of all the concept of gain is
-introduced, this indicates how much gain there will be from moving a vertex
-to the other partition. Then the vertexes are put into an array with as index
-their gain. If there are more vertices at a particular index then they're
+The algorithm involves a lot of clever book keeping.
+First of all the concept of gain is introduced, this indicates how much
+gain there will be from moving a vertex to the other partition. Then the
+vertexes are split based on partition and put into an array with as index
+their gain (the paper calls this a bucket, as do I).
+If there are more vertices at a particular index then they're
 put into an linked list. To support this I introduced a container class
 Cell that holds the vertex and also has an optional previous and next
-element.
-These cells are also put in another array based on their vertex ID. This
-array is used for constant speed access when the connections of a vertex their
-gain needs to be updated after moving it to the other partition.
+element. (Again the paper uses the name cell to).
+
+These cells are also put in another array based on their vertex ID (position
+in the solution string). This
+array is used for constant access speed when the \emph{connections of a vertex}
+their gain needs to be updated after moving the initial vertex to the other
+partition.
 So you can just untie the element of the linked list (or if it has no
-previous in my case its in the front of the bucket and I remove it from
-there and reinsert the next element).
+previous in my case, its in the front of the bucket and I remove it from
+there and reinsert the next element and itself to the same bucket at the
+proper gain index).
 
@@
 I think I've succeeded so my final research question is how does the 
-Fiduccia Mattheyses algorithm compare to first
-improvement vertex swap?
+Fiduccia Mattheyses algorithm compare to first improvement vertex swap?
 
@@
 to produce better results, I think this is because highest gain is considered
-first.
-
-My other research topic is selection pressure, turns out that tournament is
-the best for GLS in case of Fiduccia Mattheyses, but it also takes the mot
-amount of time.
+first. But more about this in the results.
 
@@
 and high values are calculated with something called a confidence interval.
-It aligns pretty well with the tTest results so I guess its OK.
+It aligns pretty well with the tTest results.
 
@@
 \end{empfile}
+
+\section{Conclusion}
+Fiduccia Mattheyses combined with GLS finds the best solution at the cost of
+quite some time, but the results are very close to the global optimum.
+
 \end{document}

### 2016-01-31 evolutionaryComputation/2-practical/report.tex
deleted the old bitchiness and added some flavour text

@@
 \section{Methodology}
-I copied all the code from the previous practical into this new folder.
-Then I went onto a delete rage. Most code was deleted except the Experiment
-class, this will probably still be useful.
-
-I use git, so it doens't really matter to accidentally delete to much. After
-that I changed the data, I removed everything which was useless. Now the file
-only contains the connection numbers. This makes parsing the file a lot
-easier. I don't want to do advanced pattern matching when I can quickly modify
-the file with a few vim macros. (in this case I just did a `30x').
+This time again I chose to use scala to solve this assignment. I was pleasently
+surprised last time by the power of scala's combinators and hope this will be
+the case again. I've become quite a fan of the langauge.
 

### 2016-01-31 evolutionaryComputation/2-practical/report.tex
spelling

@@
 \section{Methodology}
-This time again I chose to use scala to solve this assignment. I was pleasently
-surprised last time by the power of scala's combinators and hope this will be
-the case again. I've become quite a fan of the langauge.
+This time again I chose to use Scala to solve this assignment. I was pleasantly
+surprised last time by the power of Scala's combinators and hope this will be
+the case again. I've become quite a fan of the language.
 
@@
 
-So the file is stored in an imutable memory structure. This is possible
-because the connections are just the number that essently work as pointers
+So the file is stored in an immutable memory structure. This is possible
+because the connections are just the number that essentially work as pointers
 if you have the graph at hand.

### 2016-01-31 evolutionaryComputation/2-practical/report.tex
added about building

@@
 
+\subsection{Statistics}
+The error margins in the graph are the maximum range of the results so these
+are round numbers when subtracted from the low or added to the high. The low
+and high values are calculated with something called a confidence interval.
+It aligns pretty well with the tTest results.
+
+\subsection{Build tool}
+I use Gradle for building the project. Its the most advanced build platform
+available and makes it easy to add dependencies to the project.
+For example I use slf4j for logging.
+
+Its also handy in that it comes with a self installing script that can be run
+from any platform, Gradlew. So running the code I wrote on a different machine
+becomes easy, Gradle makes sure the right dependencies are fetched.
+
+\paragraph{Running it}
+
+Open a shell and move to the project folder then type:
+
+\lstset{language=Bash}
+\begin{lstlisting}[frame=single]
+	./gradlew run
+\end{lstlisting}
+
+\subsection{Libraries used}
+In here I will describe why I chose to use these libraries
+The details of the libraries can be found in the Gradle file (which
+version and what maven group):
+
+\begin{lstlisting}[frame=single]
+	cat code/build.gradle
+\end{lstlisting}
+
+\begin{itemize}
+	\item The Scala standard library, because I have to when using Scala.
+	\item slf4j, a simple logging facade, this allows runtime switching of
+		logging schemes, for example if you want to log into a database or
+		a file instead of stdout you can do this relatively easy with slf4j
+		This is not necessary right now but the API is simple enough to use
+		everywhere.
+	\item Scalaplot. A graph plotting library for Scala. I didn't want to
+		do this myself, and this library supports SVG plotting and ASCII
+		plotting.
+	\item Scalatest. A testing library for Scala, it has some pretty advanced
+		features but I use it mostly like JUnit.
+	\item appache commons math: For the statistics calculations.
+\end{itemize}
+
 \section{Own research topic}
@@
 first. But more about this in the results.
-
-\subsection{Statistics}
-The error margins in the graph are the maximum range of the results so these
-are round numbers when subtracted from the low or added to the high. The low
-and high values are calculated with something called a confidence interval.
-It aligns pretty well with the tTest results.
 

### 2016-02-15 geometricAlgorithms/exercisze/excersizes.tex
excersizes

--- 2016-01-31 evolutionaryComputation/2-practical/report.tex
@@
 }
-
-\begin{filecontents*}{dataTable.txt}
-zero 	name 		firsty 		secondy		errbotstart		errbotend 		errtopstart			errtopend
-0 		MLS           0         	1			0.1				0.2				0.1	               0.2
-0 		ILS           1         	2			0.3				0.1             1.5               0.4
-0 		GLS           2         	3.5			0.5				0.2             0.3               0.4
-\end{filecontents*}
 
@@
 \author{Jappie Klooster}
-\title{Graphs, partitioning and you}
+\title{Stay in Shape! A geometric algorithmic excersize collection}
 \maketitle
@@
 \section{Introduction}
-This is a report of the evolutionary practical assignment 2.
-In here I will discuss the execution of the assignment, what it was, which
-methodology was used to execute the assignment.
+Referencing he book Computational geometry algorithms and application.
+This collection  tries to solve a selection of the excersizes in that book.
 
-The intended audience is people familiar with programming.
+Referer to this work with care as its meant to be homework (although
+I try to make it correct).
 
-\section{Assignment}
-Graph bi-partitioning is dividing a graph into 2 partitions, 
-with as little crossings between both partitionings as possible.
+\section{The intersection detection problem}
+The intersection detection problem for a set S of n line segments is to
+determine whether there exists a pair of segments in S that intersect. Give
+a plane sweep algorithm that solves the intersection detection problem in
+$O(n \cdot  log_2(n))$ time.
 
-So we try to reduce the amount of crossing by modifiying which vertex resides
-in what partition. The graph itself is however not modified. The amount of
-vertexes that reside in on partition should also be even to the other partition.
+\begin{enumerate}
+	\item Fold $S$ into a binary tree where left and right is decided by Y
+	coordinate of the end points,
+	make sure the end points have a reference ot eachother.
+	Call this new Binary tree $E_{bin}$.
+\item Now Pop the top two events from $E_{bin}$
+\end{enumerate}
 
-\section{Methodology}
-This time again I chose to use scala to solve this assignment. I was pleasently
-surprised last time by the power of scala's combinators and hope this will be
-the case again. I've become quite a fan of the langauge.
-
-\subsection{MLS}
-MLS basically involved implementing the initial vertex swap first improvement
-local search and tying it into the existing framework. I did this by just
-using the existing createPopMethods with a new solution generator. Then
-I just mapped this population with the local search (so the type is
-$IMember \to IMember$, but the solution changes).
-
-\newpage
-\subsubsection{Reading the graph}
-The first thing I did was reading the graph into a memory structure. That 
-looks like this:
-
-\begin{figure}[ht!]
-\centering
-\begin{emp}[classdiag](20,20)
-Class.graph("Graph")("+verteci:Array[Vertex]", "+edgeCount:Int")();
-Class.vertex("Vertex")("+id:Int", "+connections:Seq[Int]")();
-
-leftToRight.top(30)(graph,vertex);
-
-drawObjects(graph,vertex);
-
-link(association)(graph.e -- vertex.w);
-item(iAssoc)("*")(obj.se = vertex.w);
-
-\end{emp}
-\caption{Case classes}
-\end{figure}
-
-So the file is stored in an imutable memory structure. This is possible
-because the connections are just the number that essently work as pointers
-if you have the graph at hand.
-
-\subsubsection{Solution generation}
-Then I went on doing the generation of the solutions. I did this by creating
-a random string of half the graph length, then copying this string and invert
-it. This ensured there were as many ones as zeros. Then I combined these
-two strings and randomly shuffled them:
-
-\lstset{language=scala}
-\begin{lstlisting}
-	val random = Random
-	def randomGene = random.nextInt(2).toString
-	def randomGenes(geneLength:Int) = 1.to(geneLength).
-		foldLeft(""){(b, x) => b + OffspringGenerator.randomGene}
-	def randomBalancedGenes(geneLength:Int):String = {
-		if(geneLength % 2 == 1){
-			throw new Exception("Doesn't work with odd genelength")
-		}
-		val initial = randomGenes(geneLength/2)
-		random.shuffle(
-			inverseMirror(initial).toCharArray.toSeq
-		).mkString
-	}
-	def flipBit(c:Char):Char = inverseResult(c,'1','0')
-	def inverseMirror(input:String):String ={
-		input + input.map(flipBit)
-	}
-\end{lstlisting}
-
-\subsubsection{Valuation}
-After that I did the evaluation function. Which was just folding the vertexes
-into a float and then counting the string indexes that did match. I think
-I did this wrong by mistake at first, but I realized it was advantageous to
-count the connections that are the same instead of different. Because now
-we want to maximize again as in the previous assignment, which made it
-possible to leave the selection as is (so I don't have to modify that when
-implementing GLS). To represent the internal valuation 
-(which is trying to be maximized) as the minimizing
-problem I simply do $graph.edgecount-valuation$.
-
-\subsubsection{Local search}
-Finally I designed the search class, which was first called LocalSearch but
-was later generalized to Search:
-
-\begin{figure}
-\begin{emp}[classdiag](20,20)
-Class.search("Search")(
-	"-method:IMember=>IMember", 
-	"-stopCondition:(IMember, IMember)=>Boolean",
-	"-decideResult:(IMember, IMember)=>IMember"
-	)(
-	"+search(previous:IMember):IMember"
-	);
-
-drawObjects(search);
-\end{emp}
-\caption{Search}
-\end{figure}
-
-A very compact class indeed, we execute method on previous, check the
-stopcondition if true we decide result else we call ourself (tail) recursively.
-The stopcondition and decide result are by default equal and return the result
-of the method. Why these functions are configurable will be discussed in the 
-ILS chapter.
-Then the method is either filled with vertexSwapFirstImprovement or
-FiducciaMattheysesSearch.
-
-\subsection{ILS}
-While implementing ILS, I had it turn out results much worse than MLS.
-The initial implementation would just stop if it found a worse local optimum.
-The description says it should go back to the previous optimum. So I guess
-that you should try again after that? But then I wonder how often you should
-try again.
-
-I got much better results with a system where the local search got several chances.
-If it failed it would fall back, and then try again.
-With chances around 30 I managed to get up and around 1900.
-When I reset the chance counter on a successful find
-I managed to get consistently 2200, one result was as high as 2310. Note:
-I was maximizing here and still counting double. For the final run I put
-the validation count up to 100. So before we decide something is at a local
-optimum it first has to return the same result 100 times.
-
-\subsubsection{Implementation}
-At first I started to wrote a separate ILS class for search below the then
-local search class. However I noticed the code I was producing looked 
-suspiciously similar to what I already had in local search. So I decided
-to try and combine the functionalities by making LocalSearch more configurable
-and I succeeded. The end result was the Search class (I had to rename
-the local search class because it was no longer just doing local search).
-
-To explain how it works the first thing one should do is notice how the method
-of search has the same prototype as the search method of the search class.
-This means you can nest a search method of one search class inside another
-search class. So we now have one inner search class doing the local search
-and one outer search class doing iteration and perturbation of solutions.
-
-Actually the perturbation is also nested inside a method like prototype and
-I use function composition to let the perturbation be executed just before
-the local search can be executed.
-
-as stopcondition of the outer search we have an isWorse function inside a
-resetretryOnBetter function. this function has also access to the retryOnResult
-object which counts to the validation (and can be rest on a better result).
-\subsection{GLS}
-I implemented GLS with the evolution class. The only thing I had to do really
-was write the crossover method, the rest was already functional.
-
-I wanted to see if the GLS population size and selection pressure would have any
-effect on the average fitness found. I suspect that a bigger population size
-would find better results and that weaker selection would find better
-average scores.
-
-So I also started to think about my own topic to investigate, at first
-I just wanted to do varying population sizes and varying selection methods.
-But I thinks this takes a little to much time and it doesn't quite fit
-on the tTest table so I decided to drop the varying pop-sizes.
-
-\subsection{Counting double}
-I had an issue in which my classmates would get much better results than me.
-After discussing it for a while it turned out that I was counting double. However
-This didn't make sense because I just used Scala combinators to count the
-elements. This would mean that there should be a bug in the std
-Scala library which seems highly unlikely. I knew I was counting double
-because all my results were even numbers. After thinking about it for a while
-I decided to investigate the graph file, it turns out that every connection
-is made double, so the double counting was in fact correct.
-
-\subsubsection{Graph pruning}
-I decided to use a prune graph to speed up the algorithm. In the beginning I
-create 2 graphs. One normal and one pruned, The pruning works by
-comparing all connections against the id, if lower we ignore the connection.
-To make sure we only do this on doubly linked graphs 2 graphs are created.
-If the pruned graph has exactly half the score of a random solution we assume
-its a doubly linked graph so we can prune it. I know its not a
-guaranteed to work.
-The pruning effectively halves the amount of connections that need to be
-checked by each iteration (this happens in the evaluation). The
-Fiduccia Mattheyses has even more gain because it has a lot of book keeping to
-do with the connections.
-
-\subsection{Fiduccia Mattheyses}
-I was somewhat sceptical of my local search implementation. I had quite
-different results from my classmates so to compare I decided to implement
-the Fiduccia Mattheyses local search to. It was quite a lot of work but it
-truly is significantly faster. Finding the paper was somewhat difficult
-as the original was behind a paywall.  But someone put a copy on GitHub so
-I just used that one.
-
-The algorithm involves a lot of clever book keeping.
-First of all the concept of gain is introduced, this indicates how much
-gain there will be from moving a vertex to the other partition. Then the
-vertexes are split based on partition and put into an array with as index
-their gain (the paper calls this a bucket, as do I).
-If there are more vertices at a particular index then they're
-put into an linked list. To support this I introduced a container class
-Cell that holds the vertex and also has an optional previous and next
-element. (Again the paper uses the name cell to).
-
-These cells are also put in another array based on their vertex ID (position
-in the solution string). This
-array is used for constant access speed when the \emph{connections of a vertex}
-their gain needs to be updated after moving the initial vertex to the other
-partition.
-So you can just untie the element of the linked list (or if it has no
-previous in my case, its in the front of the bucket and I remove it from
-there and reinsert the next element and itself to the same bucket at the
-proper gain index).
-
-After knowing the basic data structures the implementation is quite trivial
-(although I had some tedious bugs such as index out of bounds errors or
-next elements being wrongly updated).
-It doesn't find the local optimum in one go, 4 to 5 iterations are
-generally required, which is still much and much better than the first
-improvement vertex swap.
-
-\section{Own research topic}
-I had several ideas for my own research to investigate, first of all I wanted
-to do the really easy to implement ones, such as selection pressure and pop
-sizes for GLS. But later, when waiting for results, I got curious about the
-other local search I heard about in class. With several days left I decided
-to try and implement the Fiduccia Mattheyses local search heuristic.
-I think I've succeeded so my final research question is how does the 
-Fiduccia Mattheyses algorithm compare to first improvement vertex swap?
-
-As can be seen in the results section the Fiduccia Mattheyses algorithm is
-much faster than first improvement vertex swap. The main reason for this is
-that a single pass will climb very steeply in valuation. It also appears
-to produce better results, I think this is because highest gain is considered
-first. But more about this in the results.
-
-\subsection{Statistics}
-The error margins in the graph are the maximum range of the results so these
-are round numbers when subtracted from the low or added to the high. The low
-and high values are calculated with something called a confidence interval.
-It aligns pretty well with the tTest results.
-
-\newpage
-\section{Results}
-Fiduccia Mattheyses dominates vertex swap first improvement in both runtime and
-optima found. Combined with GLS it gets such a good score that I became
-skeptical and decided to test the answer more thoroughly.
-
-Time wise Fiduccia Mattheyses dominates vertex swap first improvement. This
-is because it starts with an incredibly steep hill climb that flattens
-out quite quickly. So you just need to pass over the string 4 or 5
-times and then you are at the local optimum. Vertex swap first improvement
-increases linearly however. So it takes a lot more string iterations
-and evaluations to get there.
-
-The improvements found by Fiduccia Mattheyses are also a lot better, especially
-the GLS variants. I'm not quite sure why this is the case, in fact I
-added a paragraph scepticism and put a gene in there that found a solution
-with 4 crossings on the original graph. Vertex swap first improvement gets
-noway near to that. The lowest it found is like 280 crossings on the original
-graph. I think the GLS variant doesn't find any better is because I always
-start at index 0 of the string and take any improvement I can find, this
-means the search space is biased. A random start could solve this issue but
-then the function is no longer pure. The other bias vertex swap first
-improvement has is that it either goes from left to write or from right to
-left, you can't really do both. Fiduccia Mattheyses doesn't suffer from both
-of these problem because it just picks whatever improvement has the maximum
-gain (while keeping balance in mind), so its biased towards maximum gain,
-which isn't a bad bias at all.
-
-Something I found interesting is that the Fiduccia Mattheyses local search
-doesn't seem to respond well to ILS it just stays in the range of the lower
-error margin of MLS. To verify this I cranked up the validation number to
-100 (100 times no better improvement found the local search end), and
-increased the perturbation to 7 (because initially it looked like 1 to 5
-where slowly going down).
-However when you look to vertex swap first improvement you can see ILS
-responding violently. In fact ILS-1 is tTest similar to GLS:Killparents.
-
-Selection method seems to be statistically irrelevant for the optimum in
-Fiduccia Mattheyses, although you pay a much bigger time price with both
-tournament and kill parents selection. Tournament selection does seem to
-be best for vertex swap first improvement, but again you pay a time price
-for this.
-
-\newpage
-\subsection{Fiduccia Mattheyses}
-\subsubsection{Time}
-\pgfplotstableread{fidRuntimes.table}\dataTable
-\begin{tikzpicture}
-\pgfplotsset{set layers}
-\begin{axis}[
-	ybar stacked,
-	xlabel={Algorithm},
-	xticklabels from table={fidRuntimes.table}{name},
-	xtick=data,
-	x tick label style={rotate=90,anchor=east},
-	ylabel={Time}
-]
-% Layers needed to get error bars on top of lower blue rectangle
-\addplot[color=white]
-table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};
-
-\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
-table[x expr=\coordindex, y=zero, y error minus=errbotstart, y error plus=errbotend]{\dataTable};
-
-\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
-table[x expr=\coordindex, y=secondy, y error minus=errtopstart, y error plus=errtopend]{\dataTable};
-
-\end{axis}
-
-\end{tikzpicture}
-
-\input{fidRuntimes.textable}
-
-\newpage
-\subsubsection{Optima}
-\pgfplotstableread{fidResults.table}\dataTable
-\begin{tikzpicture}
-\pgfplotsset{set layers}
-\begin{axis}[
-	ybar stacked,
-	xlabel={Algorithm},
-	xticklabels from table={fidResults.table}{name},
-	xtick=data,
-	x tick label style={rotate=90,anchor=east},
-	ylabel={Fitness (less is better)}
-]
-% Layers needed to get error bars on top of lower blue rectangle
-\addplot[color=white]
-table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};
-
-\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
-table[x expr=\coordindex, y=zero, y error minus=errbotstart, y error plus=errbotend]{\dataTable};
-
-\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
-table[x expr=\coordindex, y=secondy, y error minus=errtopstart, y error plus=errtopend]{\dataTable};
-
-\end{axis}
-
-\end{tikzpicture}
-\\
-\input{fidResults.textable}
-
-\paragraph{Scepticism} So these results seem a little to good to be true honestly,
-especially compared to vertex swap first improvement. So I decided to catch
-one of the results of GLS and throw it trough a bunch of tests to see if I 
-could find anything wrong with it. The result is this one and it appears
-to be OK with a fitness score of 2 (or 4 if you use the doubly connected graph).
-I also checked if there was an equal amount of zeros and ones just to be sure.
-
-\seqsplit{11100000101001111000000100110111101010000000110011100110011110101110110010111110010010010011000000101100111010011100001000111000110111110101110000010011111001110110100110110101110010101011000101000111100010111100010001100100000110001001010010101001110010111110001001010000001111011111111101101000101100000100011011101111001010010111001101101111100001011000001100100001011101000101111000101011101001001010101011011000110000110111111101101000001111000010010001111101000101000010111100101011100111100101}
-
-There is a good chance this is a global optimum.
-
-\newpage
-\subsection{Vertex swap first improvement}
-\subsubsection{Time}
-\pgfplotstableread{vertRuntimes.table}\dataTable
-\begin{tikzpicture}
-\pgfplotsset{set layers}
-\begin{axis}[
-	ybar stacked,
-	xlabel={Algorithm},
-	xticklabels from table={vertRuntimes.table}{name},
-	xtick=data,
-	x tick label style={rotate=90,anchor=east},
-	ylabel={Time}
-]
-% Layers needed to get error bars on top of lower blue rectangle
-\addplot[color=white]
-table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};
-
-\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
-table[x expr=\coordindex, y=zero, y error minus=errbotstart, y error plus=errbotend]{\dataTable};
-
-\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
-table[x expr=\coordindex, y=secondy, y error minus=errtopstart, y error plus=errtopend]{\dataTable};
-
-\end{axis}
-
-\end{tikzpicture}
-
-\input{vertRuntimes.textable}
-
-\newpage
-\subsubsection{Optima}
-\pgfplotstableread{vertResults.table}\dataTable
-\begin{tikzpicture}
-\pgfplotsset{set layers}
-\begin{axis}[
-	ybar stacked,
-	xlabel={Algorithm},
-	xticklabels from table={vertResults.table}{name},
-	xtick=data,
-	x tick label style={rotate=90,anchor=east},
-	ylabel={Fitness (less is better)}
-]
-% Layers needed to get error bars on top of lower blue rectangle
-\addplot[color=white]
-table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};
-
-\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
-table[x expr=\coordindex, y=zero, y error minus=errbotstart, y error plus=errbotend]{\dataTable};
-
-\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
-table[x expr=\coordindex, y=secondy, y error minus=errtopstart, y error plus=errtopend]{\dataTable};
-
-\end{axis}
-
-\end{tikzpicture}
-\\
-\input{vertResults.textable}
 \end{empfile}
-
-\section{Conclusion}
-Fiduccia Mattheyses combined with GLS finds the best solution at the cost of
-quite some time, but the results are very close to the global optimum.
-
 \end{document}

### 2016-02-15 geometricAlgorithms/exercisze/excersizes.tex
trying to write me some psuedocode

@@
 \usepackage{seqsplit}
-
+\usepackage{algorithmicx}
+\usepackage{algpseudocode}
 \pgfplotsset{
@@
 
-\begin{enumerate}
-	\item Fold $S$ into a binary tree where left and right is decided by Y
-	coordinate of the end points,
-	make sure the end points have a reference ot eachother.
-	Call this new Binary tree $E_{bin}$.
-\item Now Pop the top two events from $E_{bin}$
-\end{enumerate}
+\begin{algorithmic}
+	\State Sort all line segment $s \in S$ end points $p$ based on
+		$y$ coordinate.
+		$p$ should have a reference to its coorisponding point (ie the start
+		point has a reference to the end point and vice versa). The resulting
+		collection is called $E_{sorted}$
+	\For{point$\in E_{sorted}$}
+		\State
+		\If{begin point} insert in State. On inserted point calculate
+		if there is a crossing if so insert in $E_{sorted}$.
+		\EndIf
+		\If{end point} remove begin point from state.
+		\EndIf
+		\If{Crossing point} add to result.
+		\EndIf
+	\EndFor
+\end{algorithmic}
 

### 2016-02-19 geometricAlgorithms/homeworkExam/homeworkExam.tex
yesterdays modifications

--- 2016-02-15 geometricAlgorithms/exercisze/excersizes.tex
@@
 \author{Jappie Klooster}
-\title{Stay in Shape! A geometric algorithmic excersize collection}
+\title{Examination of shapes! (homework exam)}
 \maketitle
 
-\section{Introduction}
-Referencing he book Computational geometry algorithms and application.
-This collection  tries to solve a selection of the excersizes in that book.
+\section{Polyhedron}
 
-Referer to this work with care as its meant to be homework (although
-I try to make it correct).
+\subsection{When is the $xy$-silhouette of a convex polyhedron $P$ is unique}
+%TODO this anwser seems to complex for 0.5 points, maybe I should look in the book
+% TODO: don't give a crappy discription
+%TODO: define uniqueness?
+First we have to define what uniqueness means to anwser this question. I'm 
+assuming uniqueness in this case means an $xy$-silhouette that can only be
+generated from a polyhedron P when looking at it from a specific angle.
 
-\section{The intersection detection problem}
-The intersection detection problem for a set S of n line segments is to
-determine whether there exists a pair of segments in S that intersect. Give
-a plane sweep algorithm that solves the intersection detection problem in
-$O(n \cdot  log_2(n))$ time.
+% TODO: wtf?
+So there existions a quaternion $\exists q \in Q$ where there is an $xy$-silhoeutte
+$s$ that is not equal to any other sillhoute.
 
+So an $xy$ sillhouette is unique if there is no repitition in the outer plan that
+can be reshown trough rotation.
+
+Although it does have to be notated that you can always flip the entire thing 180
+degree and receive the exact sam sillhoutte.
+
+\subsubsection{When it is not unique}
+%TODO: \dots
+\subsection{Give an algorithm that detrmines an $xy$-silhouette of a conves polyhedron $P$ in O(n) time}
 \begin{algorithmic}
@@
 
+\subsection{Why takes this algorithm linear time}
+\section{Red and blue lines}
+\section{Polygon cut}
+\section{Diagonal polygon cut}
 \end{empfile}

### 2016-02-19 geometricAlgorithms/homeworkExam/homeworkExam.tex
question 1A

@@
 
-% TODO: wtf?
-So there existions a quaternion $\exists q \in Q$ where there is an $xy$-silhoeutte
-$s$ that is not equal to any other sillhoute.
+So when projecting a polyhedron P on an $xy$ plane (ie removing the z values).
+you need to have 3 angles. $x$ rotation, $y$ rotation and $z$ rotation.
 
-So an $xy$ sillhouette is unique if there is no repitition in the outer plan that
-can be reshown trough rotation.
+An $xy$-silhoutte is said to be unique if there is only one set of rotations
+that can generate that sillhouette.
 
-Although it does have to be notated that you can always flip the entire thing 180
-degree and receive the exact sam sillhoutte.
+I had initially some questions about this defintion, because is it not possible
+to always generate another sillouthe by just applying enough rotations.
+So I started drawing a simple sillhoutte, a square with an angle.
+
+\begin{figure}
+\begin{tikzpicture}[
+rotor/.style={inner sep=3pt,outer sep=0,fill opacity=0.3,minimum width=2cm,circle},
+crossline/.style={gray,line width=3pt,opacity=0.5,line cap=round}
+]
+	\node[rotor] (n2) at (4,2)  {\quad y};
+
+	\draw [->,line width=1pt] (n2) ++(140:5mm) arc (-220:40:5mm);
+	\draw (1,1) -- (3,1) -- (3,3) -- (1,3) -- (1,1);
+	\draw (2,2) -- (0,2) -- (0,1);
+
+	\draw (5,1) -- (7,1) -- (7,3) -- (5,3) -- (5,1);
+	\draw (6,2) -- (8,2) -- (8,1);
+
+	\draw (4,0) -- (4,4);
+\end{tikzpicture}
+\caption{Cube with arm sticking out sillhoutte}
+\end{figure}
+
+I did applied similiar rotations for the $x$ and $z$ axis. If you
+do both 180 degrees you end up looking at the same side. So for this
+particiluar figure there exists a unique $xy$-sillhoutte.
 
 \subsubsection{When it is not unique}
-%TODO: \dots
+
+It is not unique when certain parts of the figure can be mirrored. For
+example consider the follwoing figure:
+
+\begin{figure}
+\begin{tikzpicture}[
+rotor/.style={inner sep=3pt,outer sep=0,fill opacity=0.3,minimum width=2cm,circle},
+crossline/.style={gray,line width=3pt,opacity=0.5,line cap=round}
+]
+	\node[rotor] (n2) at (5,2)  {\quad y};
+
+	\draw [->,line width=1pt] (n2) ++(140:5mm) arc (-220:40:5mm);
+	\draw (1,1) -- (3,1) -- (3,3) -- (1,3) -- (1,1);
+	\draw (2,2) -- (0,2) -- (0,1);
+	\draw (2,2) -- (4,2) -- (4,3);
+
+	\draw (7,1) -- (9,1) -- (9,3) -- (7,3) -- (7,1);
+	\draw (8,2) -- (10,2) -- (10,1);
+	\draw (8,2) -- (6,2) -- (6,3);
+
+	\draw (5,0) -- (5,4);
+\end{tikzpicture}
+\caption{Cube with two arms in opposite direction sillhoutte}
+\end{figure}
+
+The issue with this figue is that if you apply 180 degrees $z$ rotation you
+end up with the same sillhoutte. So having some symetry ends up causing will
+cause having redundant sillhouttes.
+
 \subsection{Give an algorithm that detrmines an $xy$-silhouette of a conves polyhedron $P$ in O(n) time}

### 2016-02-20 geometricAlgorithms/homeworkExam/homeworkExam.tex
improved the algorithm to not be the wrong one

@@
 \subsection{Give an algorithm that detrmines an $xy$-silhouette of a conves polyhedron $P$ in O(n) time}
+I'm gonna use recursion for this.
+
+%TODO looks to simple\dots
 \begin{algorithmic}
-	\State Sort all line segment $s \in S$ end points $p$ based on
-		$y$ coordinate.
-		$p$ should have a reference to its coorisponding point (ie the start
-		point has a reference to the end point and vice versa). The resulting
-		collection is called $E_{sorted}$
-	\For{point$\in E_{sorted}$}
+	\Function{findxySilhouette}{startVert, currentVert, result}
+	\If{startVert is equal to currentVert} \State \Return result \EndIf
 	\State
-		\If{begin point} insert in State. On inserted point calculate
-		if there is a crossing if so insert in $E_{sorted}$.
-		\EndIf
-		\If{end point} remove begin point from state.
-		\EndIf
-		\If{Crossing point} add to result.
-		\EndIf
-	\EndFor
+	$nextVert \gets currentVert.incidentEdge.next.origin$
+	\State
+	\Return findxySilhoeutte(startVert, nextVert, result.append(nextVert))
+	\EndFunction
+	\Function{findxySilAlgoirthm}{startVert}
+	\State \Return
+	\Call{findxySilhoeutte}{startVert, startVert.incidentEdge.next.origin, Seq(startVert)}
+	\EndFunction
+	\State 
 \end{algorithmic}
@@
 \subsection{Why takes this algorithm linear time}
+
 \section{Red and blue lines}

### 2016-02-22 geometricAlgorithms/homeworkExam/homeworkExam.tex
I think I finished the recursive function

--- 2016-02-19 geometricAlgorithms/homeworkExam/homeworkExam.tex
@@
 \subsection{Give an algorithm that detrmines an $xy$-silhouette of a conves polyhedron $P$ in O(n) time}
+%TODO: should we find the side?
+
 \begin{algorithmic}
-	\State Sort all line segment $s \in S$ end points $p$ based on
-		$y$ coordinate.
-		$p$ should have a reference to its coorisponding point (ie the start
-		point has a reference to the end point and vice versa). The resulting
-		collection is called $E_{sorted}$
-	\For{point$\in E_{sorted}$}
+	\Function{findxySilhouette}{$v_{start}, v_{current}, S_{result}$}
+	\If{$v_{start}$ is equal to $v_{current}$} 
+		\State \Return $S_{result}$
+	\EndIf
 	\State
-		\If{begin point} insert in State. On inserted point calculate
-		if there is a crossing if so insert in $E_{sorted}$.
-		\EndIf
-		\If{end point} remove begin point from state.
-		\EndIf
-		\If{Crossing point} add to result.
-		\EndIf
-	\EndFor
+	$v_{next} \gets Origin(Next(IncidentEdge(v_{current})))$
+	\State
+	\Return \Call{findxySilhoeutte}{
+		$v_{start}, v_{next}, S_{result} \cup v_{current}$
+	}
+	\EndFunction
+	\Function{findxySilAlgoirthm}{$v_{start}$}
+	\Comment{Helper function to suffice the contract (init work)}
+	\State \Return
+	\Call{findxySilhoeutte}{
+		$v_{start}, Origin(Next(IncidentEdge(v_{start}))), Seq(v_{start})$
+	}
+	\Comment{The $Seq$ just makes an arbitrary sequence with constand append time}
+	\EndFunction
+	\State 
 \end{algorithmic}
 
+Because the Polyhedron $P$ is convex (so no weird dents), and
+the initial vertex $v_{start}$ was already on the boundry this
+alogrithm is correct.
+In class we used marker bits but its simpler to just pass the
+starting point, less chance to shoot yourself in the foot with
+mutable state.
+
 \subsection{Why takes this algorithm linear time}
+Because all operations are linear or faster. In fact the slowest
+operation, the recursive step is in itself linear. It is gauranteed to stop
+as long as it is passed a cycle.
+
 \section{Red and blue lines}
+\subsection{The status structure}
+A binary tree sorted on the x axis.
+\subsection{Initialization}
+\subsection{Distinguish the events and describe for each event how its handled}
+\subsection{Analyze the time}
 \section{Polygon cut}

### 2016-02-23 geometricAlgorithms/homeworkExam/homeworkExam.tex
improved the algorithm from assignment 1
1.b finished I think
some changes to exam?
probably final version of 2b... probably

@@
 
+\newpage
 \subsection{Give an algorithm that detrmines an $xy$-silhouette of a conves polyhedron $P$ in O(n) time}
 %TODO: should we find the side?
+%TODO: walk to the side of the silhoutte
 
 \begin{algorithmic}
-	\Function{findxySilhouette}{$v_{start}, v_{current}, S_{result}$}
-	\If{$v_{start}$ is equal to $v_{current}$} 
+	\State
+	\Comment{Helper function to suffice the contract (init work)}
+	\Function{findxySilAlgoirthm}{$v_{start}$}
+	\State
+	\Comment{Find the border of the silloutte}
+	\State
+	$\vec{e}_{init} \gets IncidentEdge(\Call{findxyBorderFirstImprovment}{v_{start}})$
+	\State
+	\Comment{The $Seq$ just makes an arbitrary sequence with constand append time}
+	\State \Return
+	\Call{findxySilhoeutte}{
+		$Origin(\vec{e}_{init}), \vec{e}_{init}, Seq(\vec{e}_{init})$
+	}
+	\EndFunction
+	\State
+	\Comment{Finds the xyBorder by maximizing X}
+	\Function{findxyBorderFirstImprovement}{$v_{start}$}
+		\State
+		\Comment{The incident is used to check when we've traversed all edges}
+		\State
+		$\vec{e}_{incident} \gets IncidentEdge(v_{start})$
+		\Repeat
+			\State
+			$\vec{e}_{next} \gets Next(Twin(\vec{e}_{incident}))$
+			\State
+			$v_{candidate} \gets Origin(Next(\vec{e}_{next}))$
+			\If{$v_{candidate}$ has bigger x than $v_{start}$}
+				\State
+				\Comment{The return statment will break the loop}
+				\State
+				\Return \Call{findxyBorderFirstImprovement}{$v_{candidate}$}
+			\EndIf
+		\Until{$\vec{e}_{next}$ is not equal to $\vec{e}_{incident}$}
+		\State
+		\Comment{If the loop finds no improvment we've maximized X and are on the border}
+		\State \Return $v_{start}$
+	\EndFunction
+	\Function{findxySilhouette}{$v_{start}, \vec{e}_{current}, S_{result}$}
+	\State
+	\Comment{Base case, passing the start all the way down doesn't require us to do marking}
+	\If{$v_{start}$ is equal to $Origin(\vec{e}_{current})$}
 		\State \Return $S_{result}$
@@
 	\State
-	$v_{next} \gets Origin(Next(IncidentEdge(v_{current})))$
+	\Comment{Figuring out which edge to do next}
+	\State
+	$\vec{e}_{next} \gets Next(\vec{e}_{current})$
+	\State
+	\Comment{We use another recursive funtion for this}
+	\State
+	$\vec{e}_{next} \gets \Call{findLeastRotatingEdge}{Twin(\vec{e}_{current}), Next(Twin(\vec{e}_{next})), \vec{e}_{next}}$
 	\State
 	\Return \Call{findxySilhoeutte}{
-		$v_{start}, v_{next}, S_{result} \cup v_{current}$
+		$v_{start}, \vec{e}_{next}, S_{result} \cup \vec{e}_{current}$
 	}
 	\EndFunction
-	\Function{findxySilAlgoirthm}{$v_{start}$}
-	\Comment{Helper function to suffice the contract (init work)}
-	\State \Return
-	\Call{findxySilhoeutte}{
-		$v_{start}, Origin(Next(IncidentEdge(v_{start}))), Seq(v_{start})$
-	}
-	\Comment{The $Seq$ just makes an arbitrary sequence with constand append time}
+	\Function{findLeastRotatingEdge}{$\vec{e}_{start}, \vec{e}_{current}, \vec{e}_{best}$}
+
+	\Comment{Trying to get an as shallow rotation as possible}
+	\If{$\vec{e}_{current}$ rotation is bigger than $\vec{e}_{best}$ (with as orientation point $Prev(\vec{e}_{start})$)}
+		\State
+		$\vec{e}_{best} \gets \vec{e}_{current}$
+	\EndIf
+	\State
+	\Comment{Base case for recursion}
+	\If{$\vec{e}_{start}$ is equal to $\vec{e}_{current}$}
+		\State
+		\Return $\vec{e}_{best}$
+	\EndIf
+	\Comment{Going deeper because there is another step}
+	\State
+	\Return \Call{findLeastRotatingEdge}{$\vec{e}_{start},Next(Twin(\vec{e}_{current})),\vec{e}_{best}$}
 	\EndFunction
@@
 \subsection{Initialization}
+
 \subsection{Distinguish the events and describe for each event how its handled}
 \subsection{Analyze the time}
+
 \section{Polygon cut}

### 2016-02-23 geometricAlgorithms/homeworkExam/homeworkExam.tex
trying to draw a custom trianlge

@@
 \usepackage{amsmath}
-\usepackage{txfonts}
-\usepackage{booktabs}
-\usepackage{color}
-\usepackage{bussproofs}
-\usepackage[pdftex]{graphicx}
-\DeclareGraphicsRule{*}{mps}{*}{}
-\usepackage{pifont}
-\usepackage{qtree}
-\usepackage{emp}
-\usepackage{listings}
-\usepackage{hyperref}
-\usepackage{svg}
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\usepackage{seqsplit}
 \usepackage{algorithmicx}
 \usepackage{algpseudocode}
-\pgfplotsset{
-    compat=1.9,
-    compat/bar nodes=1.8
-}
+\usepackage{pst-solides3d}
+\usepackage{auto-pst-pdf}
 
-\newenvironment{scprooftree}[1]%
-{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
-{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }
-
-\newcommand{\brcell}[2][l]{%
-	\begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
-
-% "define" Scala
-\lstdefinelanguage{scala}{
-  morekeywords={abstract,case,catch,class,def,%
-    do,else,extends,false,final,finally,%
-    for,if,implicit,import,match,mixin,%
-    new,null,object,override,package,%
-    private,protected,requires,return,sealed,%
-    super,this,throw,trait,true,try,%
-    type,val,var,while,with,yield},
-  otherkeywords={=>,<-,<\%,<:,>:,\#,@},
-  sensitive=true,
-  morecomment=[l]{//},
-  morecomment=[n]{/*}{*/},
-  morestring=[b]",
-  morestring=[b]',
-  morestring=[b]"""
-}
 \begin{document}
-\begin{empfile}
-\begin{empcmds}
-input metauml;
-\end{empcmds}
 \author{Jappie Klooster}
@@
 
-\begin{figure}
-\begin{tikzpicture}[
-rotor/.style={inner sep=3pt,outer sep=0,fill opacity=0.3,minimum width=2cm,circle},
-crossline/.style={gray,line width=3pt,opacity=0.5,line cap=round}
-]
-	\node[rotor] (n2) at (4,2)  {\quad y};
-
-	\draw [->,line width=1pt] (n2) ++(140:5mm) arc (-220:40:5mm);
-	\draw (1,1) -- (3,1) -- (3,3) -- (1,3) -- (1,1);
-	\draw (2,2) -- (0,2) -- (0,1);
-
-	\draw (5,1) -- (7,1) -- (7,3) -- (5,3) -- (5,1);
-	\draw (6,2) -- (8,2) -- (8,1);
-
-	\draw (4,0) -- (4,4);
-\end{tikzpicture}
-\caption{Cube with arm sticking out sillhoutte}
-\end{figure}
+\begin{pspicture}(-2,-2)(3,3)
+\psset{viewpoint=100 30 20,Decran=100}
+\axesIIID[showOrigin=false](1,1,1)(3,2,2.5)
+\psSolid[object=new,
+        action=draw,
+        sommets=
+		0 2 0
+		-2 0 -2
+		2 0 -2
+		2 0 2
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 2 3]
+       }]%
+\end{pspicture}
 
@@
 particiluar figure there exists a unique $xy$-sillhoutte.
-
-\subsubsection{When it is not unique}
-
-It is not unique when certain parts of the figure can be mirrored. For
-example consider the follwoing figure:
-
-\begin{figure}
-\begin{tikzpicture}[
-rotor/.style={inner sep=3pt,outer sep=0,fill opacity=0.3,minimum width=2cm,circle},
-crossline/.style={gray,line width=3pt,opacity=0.5,line cap=round}
-]
-	\node[rotor] (n2) at (5,2)  {\quad y};
-
-	\draw [->,line width=1pt] (n2) ++(140:5mm) arc (-220:40:5mm);
-	\draw (1,1) -- (3,1) -- (3,3) -- (1,3) -- (1,1);
-	\draw (2,2) -- (0,2) -- (0,1);
-	\draw (2,2) -- (4,2) -- (4,3);
-
-	\draw (7,1) -- (9,1) -- (9,3) -- (7,3) -- (7,1);
-	\draw (8,2) -- (10,2) -- (10,1);
-	\draw (8,2) -- (6,2) -- (6,3);
-
-	\draw (5,0) -- (5,4);
-\end{tikzpicture}
-\caption{Cube with two arms in opposite direction sillhoutte}
-\end{figure}
-
-The issue with this figue is that if you apply 180 degrees $z$ rotation you
-end up with the same sillhoutte. So having some symetry ends up causing will
-cause having redundant sillhouttes.
-
 \newpage
@@
 \section{Diagonal polygon cut}
-\end{empfile}
 \end{document}

### 2016-02-23 geometricAlgorithms/homeworkExam/homeworkExam.tex
having a nice triangle
completly overdid the first question...

@@
 
+\begin{figure}
 \begin{pspicture}(-2,-2)(3,3)
-\psset{viewpoint=100 30 20,Decran=100}
-\axesIIID[showOrigin=false](1,1,1)(3,2,2.5)
+\psset{viewpoint=100 0 30,Decran=100}
+\psSolid[object=tetrahedron,
+		r=2,
+		RotZ=65,
+		fillcolor=magenta!20,
+        action=draw**,
+		numfaces=all]%
+\psSolid[object=tetrahedron,
+		r=2,
+		RotZ=125,
+		fillcolor=magenta!20,
+        action=draw**,
+		numfaces=all](0, 4, 0)%
+\psSolid[object=tetrahedron,
+		r=2,
+		RotZ=185,
+		fillcolor=magenta!20,
+        action=draw**,
+		numfaces=all](0, 8, 0)%
+\end{pspicture}
+
+\caption{Not unique}
+\end{figure}
+\newpage
+
+\begin{pspicture}(-2,-2)(3,3)
+\psset{viewpoint=100 0 0,Decran=100}
 \psSolid[object=new,
-        action=draw,
+        action=draw**,
+		fillcolor=cyan,
+		numfaces=all,
         sommets=
-		0 2 0
-		-2 0 -2
-		2 0 -2
-		2 0 2
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 1.5 0,
    faces={
@@
        [0 3 1]
-       [1 2 3]
-       }]%
+       [1 3 2]}](0,0,0)%
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		RotZ=65,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 1.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,4,0)%
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		RotZ=125,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 1.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,8,0)%
+	   %%%%%%%%%%%%%%%% X
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 1.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,0,-4)%
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		RotX=65,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 1.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,4,-4)%
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		RotX=120,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 1.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,8,-4)%
+	   %%%%%%%%%%%%%%%% Y
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 1.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,0,-8)%
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		RotY=85,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 1.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,4,-8)%
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		RotY=125,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 1.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,8,-8)%
 \end{pspicture}
@@
 A binary tree sorted on the x axis.
+The event list is a structure of 
 \subsection{Initialization}

### 2016-02-24 geometricAlgorithms/homeworkExam/homeworkExam.tex
playing more with the tetrahedrons

--- 2016-02-23 geometricAlgorithms/homeworkExam/homeworkExam.tex
@@
 
+\begin{figure}
 \begin{pspicture}(-2,-2)(3,3)
-\psset{viewpoint=100 30 20,Decran=100}
-\axesIIID[showOrigin=false](1,1,1)(3,2,2.5)
+\psset{viewpoint=100 0 30,Decran=100}
+\psSolid[object=tetrahedron,
+		r=2,
+		RotZ=65,
+		fillcolor=magenta!20,
+        action=draw**,
+		numfaces=all]%
+\psSolid[object=tetrahedron,
+		r=2,
+		RotZ=125,
+		fillcolor=magenta!20,
+        action=draw**,
+		numfaces=all](0, 4, 0)%
+\psSolid[object=tetrahedron,
+		r=2,
+		RotZ=185,
+		fillcolor=magenta!20,
+        action=draw**,
+		numfaces=all](0, 8, 0)%
+\end{pspicture}
+
+\caption{Not unique}
+\end{figure}
+\newpage
+
+\begin{figure}
+\begin{pspicture}(-2,-2)(3,3)
+\psset{viewpoint=100 0 0,Decran=100}
 \psSolid[object=new,
-        action=draw,
+        action=draw**,
+		fillcolor=cyan,
+		numfaces=all,
         sommets=
-		0 2 0
-		-2 0 -2
-		2 0 -2
-		2 0 2
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 2.5 0,
    faces={
@@
        [0 3 1]
-       [1 2 3]
-       }]%
+       [1 3 2]}](0,0,0)%
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		RotZ=65,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 2.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,4.5,0)%
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		RotZ=125,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 2.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,9,0)%
+	   %%%%%%%%%%%%%%%% X
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 2.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,0,-4)%
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		RotX=65,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 2.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,4.5,-4)%
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		RotX=120,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 2.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,9,-4)%
+	   %%%%%%%%%%%%%%%% Y
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 2.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,0,-8)%
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		RotY=85,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 2.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,4.5,-8)%
+\psSolid[object=new,
+        action=draw**,
+		fillcolor=cyan,
+		RotY=125,
+		numfaces=all,
+        sommets=
+		0 0 2
+		-1.5 -0.1 0
+		1 -1 0
+		1.5 2.5 0,
+   faces={
+       [0 1 2]
+       [0 2 3]
+       [0 3 1]
+       [1 3 2]}](0,9,-8)%
 \end{pspicture}
+\caption{Unique}
+\end{figure}
 

### 2016-02-24 geometricAlgorithms/homeworkExam/homeworkExam.tex
started working on the second assignment

@@
 \subsection{Give an algorithm that detrmines an $xy$-silhouette of a conves polyhedron $P$ in O(n) time}
-%TODO: should we find the side?
-%TODO: walk to the side of the silhoutte
-
-\begin{algorithmic}
+\begin{algorithmic}[1]
 	\State
@@
 	\State
+	%TODO 
 	$\vec{e}_{init} \gets IncidentEdge(\Call{findxyBorderFirstImprovment}{v_{start}})$
@@
 			\EndIf
-		\Until{$\vec{e}_{next}$ is not equal to $\vec{e}_{incident}$}
+		\Until{$\vec{e}_{next}$ is equal to $\vec{e}_{incident}$}
 		\State
@@
 	\EndFunction
-	\State 
 \end{algorithmic}
@@
 \subsection{Why takes this algorithm linear time}
-Because all operations are linear or faster. In fact the slowest
-operation, the recursive step is in itself linear. It is gauranteed to stop
-as long as it is passed a cycle.
+To determin this lets go over the function and see how much time they each
+maximally can take.
+To start the algorithm you call FindXYSilAlgorithm and pass it a starting
+vertex. Then to find a boundary of the XYsillhoute the
+FindBorderFirstImprovement is used. This tries to maximize X to ensure we
+are on the border of the sillhoutte. It takes any improvement, so it may
+not take the shortest path, but at most it will take $O(N)$ time.
 
+After the border is found we pass it 
 \section{Red and blue lines}
-\subsection{The status structure}
-A binary tree sorted on the x axis.
-The event list is a structure of 
-\subsection{Initialization}
+\begin{algorithmic}[1]
+	\State
+	$P_{ends} \gets R \cup B$ (flat) Mapped to endpoints where each segment
+	produces two end points, the endpoints contain a reference to eachother
+	and they aslo contain which color they are.
+	\State
+	$P_{sorted} \gets P_{ends}$ sorted on Y coordinate.
+	\ForAll{$p_{current} \gets Next(P_{sorted})$}
+	\EndFor
+\end{algorithmic}
 
-\subsection{Distinguish the events and describe for each event how its handled}
-\subsection{Analyze the time}
+The status of the sweep line is a set of linesegments intersecting it at
+its current position. ordered from left to right (x axis).
 

### 2016-02-24 geometricAlgorithms/homeworkExam/homeworkExam.tex
now doing the initilization properly??

@@
 \subsection{Give an algorithm that detrmines an $xy$-silhouette of a conves polyhedron $P$ in O(n) time}
+
+\subsubsection{initilization}
+To initialize we consider a function, that receives the first vertex as
+described in the assignment.
+
 \begin{algorithmic}[1]
-	\State
-	\Comment{Helper function to suffice the contract (init work)}
 	\Function{findxySilAlgoirthm}{$v_{start}$}
@@
 	\State
-	\Comment{The $Seq$ just makes an arbitrary sequence with constand append time}
+	\State
+	\Comment{So now we have an edge comming out of an origin with a
+	maximized x value, however we still have to find the edge that
+	goes over the sillhoutte}
+	\State
+	$\vec{e}_{init} \gets \Call{findFirstSilhouteEdge}{
+		\vec{e}_{init}, Next(Twin(\vec{e}_{init})), \vec{e}_{init}
+	}$
+	\State
 	\State \Return
@@
 	}
+	\Comment{The $Seq$ just makes an arbitrary sequence with constand append time}
 	\EndFunction
@@
 	\EndFunction
+	\Function{findFirstSilhoutteEdge}{$\vec{e}_{start}, \vec{e}_{current}, \vec{e}_{best}$}
+	\If{$\vec{e}_{current}$ is equal to $\vec{e}_{start}$}
+		\State
+		\Return $\vec{e}_{best}$
+	\EndIf
+	\If{$\vec{e}_{current}$ angle compared to the x axis is more shallow than $\vec{e}_{best}$}
+		\State
+		$\vec{e}_{best} \gets \vec{e}_{current}$
+	\EndIf
+	\State
+	\Return \Call{findFirstSilhouteEdge}{$\vec{e}_{start}, Next(Twin(\vec{e}_{current})), \vec{e}_{best}$}
+	\EndFunction
+\end{algorithmic}
+
+\subsubsection{Algorithm}
+Ones initilization is done and we have the first edge that is on the xy
+sillhoutte the findxySilhoutte function is called.
+This is the heart of the algorithm.
+\begin{algorithmic}[1]
 	\Function{findxySilhouette}{$v_{start}, \vec{e}_{current}, S_{result}$}

### 2016-02-26 geometricAlgorithms/homeworkExam/homeworkExam.tex
some stuff about speed??
homwork exam finished 2 mostly...

@@
 
-After the border is found we pass it 
+After the border is found we pass it to the FindXYSilhouette function.
+This funtion will take at most $O(N)$ time, because the function
+istelf will always try to stay at much on the border as possible with help
+of the least rotating edge function. This function traverses all connecting
+edges. This sounds like a loop in a loop but it isn't really, because the
+first loop doesn't consider all elements (only the outer ring) and the second
+loop considers the connection in the outer rings. So some elements may be
+considered dubble but this is just one element (the least rotating edge).
+Its also worth noting that the ``inside'' edges may never be considered by
+either of these functions because they try to stay on the outside.
 \section{Red and blue lines}
@@
 	produces two end points, the endpoints contain a reference to eachother
-	and they aslo contain which color they are.
+	and they aslo contain which color they are and are also marked
+	valid.
 	\State
 	$P_{sorted} \gets P_{ends}$ sorted on Y coordinate.
+	\State
+	$S_{tatus} \gets $ empty binary tree that sorts on X axis
+	$S_{result} \gets $ empty list
 	\ForAll{$p_{current} \gets Next(P_{sorted})$}
+		\If{$p_{current}$ is a start point of a blue line}
+			Insert $p_{current}$ into $S_{tatus}$
+		\State
+		\ElsIf{$p_{current}$ is a start point of a red line}
+			\State
+			$t_{neigbours} \gets $ Insert $p_{current}$ into
+			$S_{tatus}$ which returns the left and right neigbours
+			in a tupple
+			\If{$t_{neigbours}$ are not both blue lines}
+				\State
+				$p_{current}$ gets marked invalid.
+			\EndIf
+		\State
+		\ElsIf{$p_{current}$ is an end point of a blue line}
+			\State
+			$t_{neigbours} \gets $ Insert $p_{current}$ into
+			$S_{tatus}$ which returns the left and right neigbours
+			in a tupple
+			\If{$t_{neigbours}$ a red line}
+				\State
+				mark that line invalid if $t_{neigbours}$ does
+				not contain a blue line.
+			\EndIf
+			\State
+			Remove $p_{current}$ and its twin (so the start point) from$S_{tatus}$
+		\State
+		\ElsIf{$p_{current}$ is an end point of a red line}
+			\State
+			$t_{neigbours} \gets $ Insert $p_{current}$ into
+			$S_{tatus}$ which returns the left and right neigbours
+			in a tupple
+			\If{both $t_{neigbours}$ are blue and the twin of $p_{current}$ was still marked valid}
+				Add $p_{current}$ to $S_{result}$
+			\EndIf
+			\State
+			Remove $p_{current}$ and its twin (so the start point) from$S_{tatus}$
+		\EndIf
 	\EndFor
@@
 
+The events are the beginning of either a red or blue line and the ending of
+a red or blue line.
+
+If the beginning of a blue line is encoutnered it gets inserted into the
+status structure.
+
+If the beginning of a red line is encountered it gets also
+inserted into the status structure, Once inserted we check if the line left
+and right are blue lines, if not we mark it invalid.
+
+If the ending of a blue line is encoured we check its neighbours. If either
+of them are red we make sure the other neigbor is blue, if not we mark the red line as invalid.
+Then we remove the blue line from the status structure.
+
+If the ending of a red line is encounred we check if its invalid,
+if so we remove it from the structure and continue.
+If its not invalid we insert the ending into the status structure.
+If this ending is has a blue two blue lines as neigbors we add it
+to the result set.
+Finally we remove it and continue.
 \section{Polygon cut}

### 2016-02-27 geometricAlgorithms/homeworkExam/homeworkExam.tex
running time and status structure added

@@
 \section{Red and blue lines}
+\subsection{algoirthm}
+To design the algorithm I wrote it down in psuedo code:
 \begin{algorithmic}[1]
@@
 	\State
-	$S_{tatus} \gets $ empty binary tree that sorts on X axis
+	$S_{tatus} \gets $ empty binary tree that sorts on X axis, it also
+	returns the two direct neigbours on insert.
+	\State
 	$S_{result} \gets $ empty list
-	\ForAll{$p_{current} \gets Next(P_{sorted})$}
+	\For{$p_{current} \gets Next(P_{sorted})$}
 		\If{$p_{current}$ is a start point of a blue line}
@@
 			\State
-			$t_{neigbours} \gets $ Insert $p_{current}$ into
-			$S_{tatus}$ which returns the left and right neigbours
-			in a tupple
+			$t_{neigbours} \gets $ Insert $p_{current}$ into $S_{tatus}$
+
 			\If{$t_{neigbours}$ are not both blue lines}
@@
 			$t_{neigbours} \gets $ Insert $p_{current}$ into
-			$S_{tatus}$ which returns the left and right neigbours
-			in a tupple
+			$S_{tatus}$
 			\If{$t_{neigbours}$ a red line}
@@
 			$t_{neigbours} \gets $ Insert $p_{current}$ into
-			$S_{tatus}$ which returns the left and right neigbours
-			in a tupple
+			$S_{tatus}$
 			\If{both $t_{neigbours}$ are blue and the twin of $p_{current}$ was still marked valid}
@@
 	\EndFor
+	\State
+	\Return $S_{result}$
 \end{algorithmic}
 
+\subsection{Status}
 The status of the sweep line is a set of linesegments intersecting it at
@@
 
+The status is a binary tree with a special insert operation that also
+returns the neigbours of the inserted element.
+
+\subsubsection{Structure}
+A node in the binary tree has a reference to a parent, and a left and right child.
+A node has a value, which will be used to decide if you have to go left or right.
+The initials node value is just the x coordinate.
+
+\subsubsection{Insertion}
+If another element gets
+inserted we compare its x coordinate with that of the current node. If its
+smaller we go left otherwise right. If x coordinate is equal we use the other
+endpoint's X coordinate as a tie breaker, if that is equal to we look at
+Y coordinate.
+In case we are at the bottom of the tree we create a new connection node
+and take the average of the old bottom element and the inserted element.
+The parent node gets updated to let the right child point at the new
+connection node and the new connection node has as children the old bottom
+element and the newly inserted element.
+
+\paragraph{Neigbours on insert}
+But we defined a special binary tree that will return the neigbours upon
+insertion. So at this point we already found one of the neigbours, to find
+the other one we have to go up again until we can go into the neigbour
+direction and this direction is not equal to where we cam from and then go
+all the way in the other direction.
+So if we try to find the right neigbour we go up and try to go to the right
+child, if this is the child where we came from we go up further and try again.
+Until we find the root, which means there is no right neigbor, or until
+we find a child that is not the direction where we came from, which means
+we have to go left all the way down in the tree.
+
+\subsubsection{Removal}
+Removal is done by finding the node we want to remove, then going to its
+parent parent and saying one the child you came from is equal to its neigbor
+
+\subsubsection{Balancing}
+To ensure $O(log N)$ acces time an AVL implementation
+can be used. But I think describing that is somewhat out of the scope
+of this assignment.
+
+\subsection{Events}
 The events are the beginning of either a red or blue line and the ending of
@@
 
-If the ending of a red line is encounred we check if its invalid,
-if so we remove it from the structure and continue.
-If its not invalid we insert the ending into the status structure.
-If this ending is has a blue two blue lines as neigbors we add it
+If the ending of a red line is encountered we insert the ending into the status structure.
+If this ending has two blue lines as neigbors and its twin point wasn't marked invalid.
 to the result set.
 Finally we remove it and continue.
+
+\subsection{Running time}
+To do the plane sweep we first sort the points on Y access. This means
+we have at least already a running time of:
+
+\[ O((n+m) log (n+m))\]
+
+Now the question is can we stay below this for the rest of the algorithm?
+For the algorithm we consider all events which can be expressed as:
+
+\[ 2(n+m) \]
+
+Because for every red $n$ line there are two events, namely the beginpoint and
+end point of that line.
+So we loop trough this and use a binary tree to detirmin if there are neigbours
+or not.
+let $k$ be the size of $S_{tatus}$.
+Doing the test for neigbours takes at most:
+
+\[2 log(k)\]
+
+Assuming its a balanced tree. Its times 2 because one neigbour is found for
+free on insertion, but for the other neigbor in worst case we have to
+consider the entire tree.
+
+So the final running time for the algorithm is:
+
+\[O(2(n+m) 2log(k))\]
+
+Plus the initilization:
+
+\[O((n+m)log(n+m)+2(n+m) 2log(k))\]
+
 \section{Polygon cut}

### 2016-02-27 geometricAlgorithms/homeworkExam/homeworkExam.tex
assignment I think I got the gits of it

@@
 \usepackage{auto-pst-pdf}
+\usepackage{tikz}
 
@@
 \section{Polygon cut}
+Consider the following polygon:
+
+\begin{tikzpicture}
+	\draw (0,1.25) -- (2,1.25) -- (2,0) --
+	 (1,1) -- (0,0) -- cycle;
+\end{tikzpicture}
+
+The best horizontal cut in this case is a cut that looks like so:
+
+\begin{tikzpicture}
+	\draw (0,1.25) -- (2,1.25) -- (2,0) --
+	 (1,1) -- (0,0) -- cycle;
+	 \draw [dotted] (0,1) -- (1,1);
+\end{tikzpicture}
+
+The reason for this is that the two triangles are bigger in area than
+the connecting rectangle. But if we try to cut the entire figure we
+end up with 3 different areas instead of 2. So the best cut we can
+make will end up to  the depth of the trianlge enclave.
+
+Now if you make the rectangle bigger something funny happens:
+
+\begin{tikzpicture}
+	\draw (0,2) -- (2,2) -- (2,0) --
+	 (1,1) -- (0,0) -- cycle;
+	 \draw [dotted] (0,1.25) -- (2,1.25);
+\end{tikzpicture}
+
+The best cut moves up to devide everything perfectly.
+So we get an imperfect cut that *has* to end on a vertex according to
+the lemma. Or we ge a perfect cut that may or may not end on a vertex.
+
 \section{Diagonal polygon cut}
+The most obvious solution is the brute force solution, which involves
+having two loops that compare every possible cut from every vertex.
+and just keep track of the best cut. this will take $O(n^2)$ time.
 \end{document}

### 2016-02-27 geometricAlgorithms/homeworkExam/homeworkExam.tex
A lot of spelling errors, and the foundation of 3 an also went trough 1a again.
So this should be handinable without a to bad grade??

@@
 \subsection{When is the $xy$-silhouette of a convex polyhedron $P$ is unique}
-%TODO this anwser seems to complex for 0.5 points, maybe I should look in the book
-% TODO: don't give a crappy discription
-%TODO: define uniqueness?
-First we have to define what uniqueness means to anwser this question. I'm 
-assuming uniqueness in this case means an $xy$-silhouette that can only be
+
+Assuming uniqueness in this case means an $xy$-silhouette that can only be
 generated from a polyhedron P when looking at it from a specific angle.
@@
 So when projecting a polyhedron P on an $xy$ plane (ie removing the z values).
-you need to have 3 angles. $x$ rotation, $y$ rotation and $z$ rotation.
+You need to have 3 angles. $x$ rotation, $y$ rotation and $z$ rotation.
 
-An $xy$-silhoutte is said to be unique if there is only one set of rotations
-that can generate that sillhouette.
+An $xy$-silhouette is said to be unique if there is only one set of rotations
+that can generate that silhouette.
 
-I had initially some questions about this defintion, because is it not possible
-to always generate another sillouthe by just applying enough rotations.
-So I started drawing a simple sillhoutte, a square with an angle.
+\subsubsection{Not unique}
+A good and simple example of not unique can be seen in figure
+\ref{fig:Not unique} a tetrahedron.
 
-\begin{figure}
+\begin{figure}[h]
 \begin{pspicture}(-2,-2)(3,3)
@@
 
+\label{fig:Not unique}
 \caption{Not unique}
@@
 
-\begin{figure}
+\subsubsection{Unique}
+From the unique tetrahedron I tried making a not unique one, I did
+this initially with blender but later recreated it with the pst-solides3d
+latex package as can be seen in figure \ref{fig:Unique} a malformed
+tetrahedron. The figure will always start from the same side and apply
+rotation on different axis.
+No matter how you rotate the tetrahedron the xy-silhoutte will look
+different until you finish the full rotation.
+
+\begin{figure}[h]
 \begin{pspicture}(-2,-2)(3,3)
@@
 \end{pspicture}
+\label{fig:Unique}
 \caption{Unique}
@@
 
-I did applied similiar rotations for the $x$ and $z$ axis. If you
-do both 180 degrees you end up looking at the same side. So for this
-particiluar figure there exists a unique $xy$-sillhoutte.
+So a polyhedron is xy-axis will be unique trough rotation if there
+exists symmetry in the polyhedron.
 \newpage
-\subsection{Give an algorithm that detrmines an $xy$-silhouette of a conves polyhedron $P$ in O(n) time}
+\subsection{Give an algorithm that determines an $xy$-silhouette of a convex polyhedron $P$ in O(n) time}
 
-\subsubsection{initilization}
-To initialize we consider a function, that receives the first vertex as
+\subsubsection{initialization}
+To initialize we use a function that receives the first vertex as
 described in the assignment.
@@
 \begin{algorithmic}[1]
-	\Function{findxySilAlgoirthm}{$v_{start}$}
+	\Function{findxySilAlgorithm}{$v_{start}$}
 	\State
-	\Comment{Find the border of the silloutte}
-	\State
-	%TODO 
 	$\vec{e}_{init} \gets IncidentEdge(\Call{findxyBorderFirstImprovment}{v_{start}})$
 	\State
-	\State
-	\Comment{So now we have an edge comming out of an origin with a
+	\Comment{So now we have an edge coming out of an origin with a
 	maximized x value, however we still have to find the edge that
-	goes over the sillhoutte}
+	goes over the silhouette}
 	\State
@@
 	\State \Return
-	\Call{findxySilhoeutte}{
-		$Origin(\vec{e}_{init}), \vec{e}_{init}, Seq(\vec{e}_{init})$
+	\Call{findxySilhouette}{
+		$Origin(\vec{e}_{init}), \vec{e}_{init}, Seq(Origin(\vec{e}_{init}))$
 	}
-	\Comment{The $Seq$ just makes an arbitrary sequence with constand append time}
 	\EndFunction
 	\State
-	\Comment{Finds the xyBorder by maximizing X}
+	\Comment{Finds the xy-border by maximizing X}
 	\Function{findxyBorderFirstImprovement}{$v_{start}$}
@@
 				\State
-				\Comment{The return statment will break the loop}
+				\Comment{The return statement will break the loop}
 				\State
@@
 		\State
-		\Comment{If the loop finds no improvment we've maximized X and are on the border}
+		\Comment{If the loop finds no improvement we've maximized X and are on the border}
 		\State \Return $v_{start}$
 	\EndFunction
-	\Function{findFirstSilhoutteEdge}{$\vec{e}_{start}, \vec{e}_{current}, \vec{e}_{best}$}
+	\Function{findFirstSilhouetteEdge}{$\vec{e}_{start}, \vec{e}_{current}, \vec{e}_{best}$}
 	\If{$\vec{e}_{current}$ is equal to $\vec{e}_{start}$}
@@
 	\State
-	\Return \Call{findFirstSilhouteEdge}{$\vec{e}_{start}, Next(Twin(\vec{e}_{current})), \vec{e}_{best}$}
+	\Return \Call{findFirstSilhouetteEdge}{$\vec{e}_{start}, Next(Twin(\vec{e}_{current})), \vec{e}_{best}$}
 	\EndFunction
@@
 
+\newpage
 \subsubsection{Algorithm}
-Ones initilization is done and we have the first edge that is on the xy
-sillhoutte the findxySilhoutte function is called.
+Ones initialization is done and we have the first edge that is on the xy
+silhouette the findxySilhouette function is called.
 This is the heart of the algorithm.
@@
 	\State
-	\Comment{We use another recursive funtion for this}
+	\Comment{We use another recursive function for this}
 	\State
@@
 	\State
-	\Return \Call{findxySilhoeutte}{
+	\Return \Call{findxySilhouette}{
 		$v_{start}, \vec{e}_{next}, S_{result} \cup \vec{e}_{current}$
@@
 
-Because the Polyhedron $P$ is convex (so no weird dents), and
-the initial vertex $v_{start}$ was already on the boundry this
-alogrithm is correct.
-In class we used marker bits but its simpler to just pass the
-starting point, less chance to shoot yourself in the foot with
-mutable state.
+Because the Polyhedron $P$ is convex  and
+the initial vertex $v_{start}$ was already on the boundary this
+algorithm is correct.
 
 \subsection{Why takes this algorithm linear time}
-To determin this lets go over the function and see how much time they each
+To determine this lets go over the function and see how much time they each
 maximally can take.
-To start the algorithm you call FindXYSilAlgorithm and pass it a starting
-vertex. Then to find a boundary of the XYsillhoute the
+To start the algorithm you call findxySilAlgorithm and pass it a starting
+vertex. Then to find a boundary of the xy-silhouette the
 FindBorderFirstImprovement is used. This tries to maximize X to ensure we
-are on the border of the sillhoutte. It takes any improvement, so it may
+are on the border of the silhouette. It takes any improvement, so it may
 not take the shortest path, but at most it will take $O(N)$ time.
 
-After the border is found we pass it to the FindXYSilhouette function.
-This funtion will take at most $O(N)$ time, because the function
-istelf will always try to stay at much on the border as possible with help
-of the least rotating edge function. This function traverses all connecting
-edges. This sounds like a loop in a loop but it isn't really, because the
-first loop doesn't consider all elements (only the outer ring) and the second
-loop considers the connection in the outer rings. So some elements may be
-considered dubble but this is just one element (the least rotating edge).
-Its also worth noting that the ``inside'' edges may never be considered by
-either of these functions because they try to stay on the outside.
+After the border is found we pass it to the findxySilhouette function.
+This function will take at most $O(N)$ time, this function will always
+tries to stay on the least xy rotation as possible.
+
+This function traverses all connecting edges of the current edge's endpoint,
+so it sounds like we're having an $O(N^2)$ function here, but this isn't the
+case because the outer function doesn't consider all elements
+(only the outer ring) and the second function only considers the local
+connections of the current edge. Because this is a local consideration
+and most of these edges don't overlap (in fact only 2 do per consideration,
+the starting on and the answer), it takes $O(N)$ running time.
+
+So the total running time of this algorithm is:
+
+\[ O(N+N) \]
+
+Which simplifies to:
+
+\[ O(N)\]
+
 \section{Red and blue lines}
-\subsection{algoirthm}
-To design the algorithm I wrote it down in psuedo code:
+\subsection{algorithm}
+To design the algorithm I wrote it down in pseudo code:
 \begin{algorithmic}[1]
@@
 	$P_{ends} \gets R \cup B$ (flat) Mapped to endpoints where each segment
-	produces two end points, the endpoints contain a reference to eachother
-	and they aslo contain which color they are and are also marked
+	produces two end points, the endpoints contain a reference to each other
+	and they also contain which color they are and are also marked
 	valid.
@@
 	$S_{tatus} \gets $ empty binary tree that sorts on X axis, it also
-	returns the two direct neigbours on insert.
+	returns the two direct neighbours on insert.
 	\State
@@
 \subsection{Status}
-The status of the sweep line is a set of linesegments intersecting it at
+The status of the sweep line is a set of line segments intersecting it at
 its current position. ordered from left to right (x axis).
@@
 The status is a binary tree with a special insert operation that also
-returns the neigbours of the inserted element.
+returns the neighbours of the inserted element.
 
@@
 
-\paragraph{Neigbours on insert}
-But we defined a special binary tree that will return the neigbours upon
-insertion. So at this point we already found one of the neigbours, to find
-the other one we have to go up again until we can go into the neigbour
+\paragraph{Neighbours on insert}
+But we defined a special binary tree that will return the neighbours upon
+insertion. So at this point we already found one of the neighbours, to find
+the other one we have to go up again until we can go into the neighbour
 direction and this direction is not equal to where we cam from and then go
 all the way in the other direction.
-So if we try to find the right neigbour we go up and try to go to the right
+So if we try to find the right neighbour we go up and try to go to the right
 child, if this is the child where we came from we go up further and try again.
-Until we find the root, which means there is no right neigbor, or until
+Until we find the root, which means there is no right neighbour, or until
 we find a child that is not the direction where we came from, which means
@@
 Removal is done by finding the node we want to remove, then going to its
-parent parent and saying one the child you came from is equal to its neigbor
+parent parent and saying one the child you came from is equal to its neighbour
 
 \subsubsection{Balancing}
-To ensure $O(log N)$ acces time an AVL implementation
+To ensure $O(log N)$ access time an AVL implementation
 can be used. But I think describing that is somewhat out of the scope
@@
 
-If the beginning of a blue line is encoutnered it gets inserted into the
+If the beginning of a blue line is encountered it gets inserted into the
 status structure.
@@
 
-If the ending of a blue line is encoured we check its neighbours. If either
-of them are red we make sure the other neigbor is blue, if not we mark the red line as invalid.
+If the ending of a blue line is encountered we check its neighbours. If either
+of them are red we make sure the other neighbour is blue, if not we mark the red line as invalid.
 Then we remove the blue line from the status structure.
@@
 If the ending of a red line is encountered we insert the ending into the status structure.
-If this ending has two blue lines as neigbors and its twin point wasn't marked invalid.
+If this ending has two blue lines as neighbours and its twin point wasn't marked invalid.
 to the result set.
@@
 
-Because for every red $n$ line there are two events, namely the beginpoint and
+Because for every red $n$ line there are two events, namely the begin point and
 end point of that line.
-So we loop trough this and use a binary tree to detirmin if there are neigbours
+So we loop trough this and use a binary tree to determine if there are neigbours
 or not.
 let $k$ be the size of $S_{tatus}$.
-Doing the test for neigbours takes at most:
+Doing the test for neighbours takes at most:
 
@@
 Assuming its a balanced tree. Its times 2 because one neigbour is found for
-free on insertion, but for the other neigbor in worst case we have to
+free on insertion, but for the other neighbour in worst case we have to
 consider the entire tree.
@@
 
-Plus the initilization:
+Plus the initialization:
 
@@
 end up with 3 different areas instead of 2. So the best cut we can
-make will end up to  the depth of the trianlge enclave.
+make will end up to  the depth of the triangle enclave.
 
@@
 
-The best cut moves up to devide everything perfectly.
+The best cut moves up to divide everything perfectly.
 So we get an imperfect cut that *has* to end on a vertex according to
-the lemma. Or we ge a perfect cut that may or may not end on a vertex.
+the lemma. Or we get a perfect cut that may or may not end on a vertex.
 
+In case we can do a perfect cut, do that.
+If we can't do a perfect cut we have to minimize the largest area,
+so start
+
+\subsection{Vertex types}
+Merge and split by definition, but on occasion a regular may happen to.
 \section{Diagonal polygon cut}

### 2016-02-28 geometricAlgorithms/homeworkExam/homeworkExam.tex
to proof, also excersize 4 better

@@
 
-Now if you make the rectangle bigger something funny happens:
+Now if you make the rectangle bigger something happens:
 
@@
 
-In case we can do a perfect cut, do that.
-If we can't do a perfect cut we have to minimize the largest area,
-so start
+\paragraph{To proof} Given a simple polygon $P$, the best horizontal cut
+has at least one endpoint on a vertex of $P$, or it cuts $P$ into two pieces
+with equal area (or both).
+
+\paragraph{Proof} There are two cases to be proven:
+	In case we can best horizontal cut a polygon $P$ into two equal areas
+	we're done (since in this case there are no restrictions on the best
+	horizontal cut line).
+	In case we can't best horizontal cut polygon $P$ into two equal areas,
+	the best horizontal cut has to minimise the larger area. In this case
+	the best horizontal cut has to end on a vertex of $P$ according to
+	the lemma. To prove this lets assume its not the case we end up of a
+	vertex of $P$.
+	However because the best horizontal cut has to minimize the larger
+	area it will always end up on some sort of angle of the polygon, which
+	is defined by a vertex and it can't equally divide the polygon because
+	that was already handled
+	by case one. So this shows that you end up with a contradiction.
+	This is what had to be shown.
 
@@
 The most obvious solution is the brute force solution, which involves
-having two loops that compare every possible cut from every vertex.
-and just keep track of the best cut. this will take $O(n^2)$ time.
+having two loops that compare every possible cut from every vertex,
+while keeping track of the best cut. This will take $O(n^2)$ time.
+
+A faster method is just finding all the merge and split vertexes. Which takes
+$O(n)$ time, and then comparing each of these with all other vertexes and keeping
+track of the best area. This will take at most $O(n+(m+s)n)$
+with $m$ as merge vertexes and $s$ as split vertexes. This is correct because in
+exercise 3.b we determined that the vertexes a best horizontal cut intersects
+with either a merge or split vertex, so the best diagonal cut will to.
 \end{document}

### 2016-02-29 geometricAlgorithms/homeworkExam/homeworkExam.tex
Improved the proof?

@@
 	horizontal cut line).
+
 	In case we can't best horizontal cut polygon $P$ into two equal areas,
@@
 	the best horizontal cut has to end on a vertex of $P$ according to
-	the lemma. To prove this lets assume its not the case we end up of a
-	vertex of $P$.
+	the lemma.
+
+	To prove this lets assume its not the case a best horizontal cut's endpoint
+	ends lies on a vertex of $P$ if dividing into two equal areas is impossible.
 	However because the best horizontal cut has to minimize the larger
 	area it will always end up on some sort of angle of the polygon, which
-	is defined by a vertex and it can't equally divide the polygon because
-	that was already handled
-	by case one. So this shows that you end up with a contradiction.
-	This is what had to be shown.
+	is defined by a vertex. It can't equally divide the polygon because
+	that was already handled by case one.
+	So it can't be the case that the best horizontal cut does not
+	end up on a vertex of $P$. This is what had to be shown.
 

### 2016-03-15 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
initial homework exam II

\documentclass{article}
\usepackage{amsmath}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{pst-solides3d}
\usepackage{auto-pst-pdf}
\usepackage{tikz}
\usepackage{enumerate}
\renewcommand{\thesubsection}{\thesection.\alph{subsection}}
\begin{document}
\author{Jappie Klooster}
\title{Stay in shape! Homework exam II}
\maketitle
\section{A disk and some half-planes}
\subsection{Per case, what can you say about $H$}
\begin{enumerate}[i]
	\item \emph{There is no disk that has its centor non the $x$ axis
			that lies inside all half-planes.}

		There exists some half plan whose's area never crosses the x axis
		(so is either completly above or below y=0).
	\item \emph{There are arbitrarly large disk with the specifications.}

		The half planes do not enclose the disk area.
	\item \emph{The largest disk exists and is unique.}

	\item \emph{There are several equally large largest disks}
\end{enumerate}
\subsection{Deal with case ii in linear worstcase time}
\subsection{Formulate and proof the RIC lemma}
\subsection{Give the RIC algorithm}
\subsection{Prove the expected running time}
% Don't forget to remove constants
\subsection{Eplain why you can't use RIC for points}
\section{A datastructure for horizontal line segments}
\subsection{Analyze the storage requirements}
\subsection{Argue that all horizontal line segs are found}
\subsection{Give psuedo code for the query algorithm}
\subsection{Analyze the query time}
\section{Voronoi nearest neigbour\emph{s}}
%TODO: wtf? maybe with higher order voronoi?? But it won't be log n
\end{document}

### 2016-03-15 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
1a

@@
 
-		There exists some half plan whose's area never crosses the x axis
-		(so is either completly above or below y=0).
+		The intersection of halfplanes of $H$ never crosses the x axis,
+		so it exists entirely above or below $y=0$.
+
 	\item \emph{There are arbitrarly large disk with the specifications.}
 
-		The half planes do not enclose the disk area.
+		The intersection of halfplanes of $H$ do not enclose the disk
+		area.
+		In fact the area they do enclose is smaller than $\pi$ radials.
+
 	\item \emph{The largest disk exists and is unique.}
 
+		The intersection of halfplanes of $H$ produces an area with no
+		parralel lines.
+
 	\item \emph{There are several equally large largest disks}
+
+		The intersection of halfplanes of $H$ produces an area with
+		parralel lines.
+
 \end{enumerate}
+
 \subsection{Deal with case ii in linear worstcase time}
+throw new exception(``done'');
 \subsection{Formulate and proof the RIC lemma}

### 2016-03-15 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
more about shapes!

--- 2016-03-15 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
@@
 \subsection{Per case, what can you say about $H$}
+Let $H_i$ be the intersection of halfplanes of $H$.
 \begin{enumerate}[i]
@@
 
-		There exists some half plan whose's area never crosses the x axis
-		(so is either completly above or below y=0).
+		$H_i$ never crosses the x axis,
+		so it exists entirely above or below $y=0$.
+
 	\item \emph{There are arbitrarly large disk with the specifications.}
 
-		The half planes do not enclose the disk area.
+		$H_i$ is open on one side with no line parralel to te x axis and
+		each line can at most have $\pi$ radials angle difference
+		compared to the line that is most paralel to the x axis.
+
 	\item \emph{The largest disk exists and is unique.}
 
+		$H_i$ produces an area with no x-axis parralel lines.
+
 	\item \emph{There are several equally large largest disks}
+
+		$H_i$ produces an area with x-axis parralel lines.
+
 \end{enumerate}
+
 \subsection{Deal with case ii in linear worstcase time}
+throw new exception(``done'');
 \subsection{Formulate and proof the RIC lemma}

### 2016-03-18 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
after meeting suggestions

--- 2016-03-15 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
@@
 \subsection{Per case, what can you say about $H$}
+Let $H_{int}$ be the intersection of halfplanes of $H$.
 \begin{enumerate}[i]
@@
 
-		There exists some half plan whose's area never crosses the x axis
-		(so is either completly above or below y=0).
+		$H_{int}$ never crosses the x axis,
+		so it exists entirely above or below $y=0$. Or the intersection is empty.
+
 	\item \emph{There are arbitrarly large disk with the specifications.}
 
-		The half planes do not enclose the disk area.
+		$H_{int}$ has no parallel line  to te x axis and
+		at most one line crossing the x axis.
+
 	\item \emph{The largest disk exists and is unique.}
 
+		$H_{int}$ crosses the x-axis at least 2 times and
+		if there is a line parallel to the x axis then both x-axis crossing
+		to be in range of the diameter of the disk.
+
 	\item \emph{There are several equally large largest disks}
+
+		$H_{int}$ produces an area with lines parallel to the x-axis.
+
 \end{enumerate}
+
+\newpage
 \subsection{Deal with case ii in linear worstcase time}
+\begin{algorithmic}[1]
+	\Function{hasArbitrarlylargeDisc}{$H$}
+		\State
+		$b_r, b_l \gets None$
+		\For{$i \gets 1$ to $n$ with $h_i \in H$}
+
+			\If{$h_i$ is parallel to the x-axis}
+				\State
+				\Return $False$
+			\EndIf
+			\State
+			\Comment{Check where the x axes crossings are, and write it into
+			a variable pair, if both pairs are filled its not case ii.}
+			\If{$h_i$ points left}
+			\State
+				$b_r \gets h_i$
+			\EndIf
+			\If{$h_i$ points right}
+			\State
+				$b_l \gets h_i$
+			\EndIf
+			\If{$b_r$ not equal to $None$ and $b_l$ not equal to $None$}
+				\State
+				\Return $False$
+			\EndIf
+		\EndFor
+		\State
+		\Return $True$
+	\EndFunction
+	\State 
+	\State 
+\end{algorithmic}
 \subsection{Formulate and proof the RIC lemma}
+%TODO: een beeetje overtypen van 4.5
+Let $1 \leq i \leq n$, and let $c_i$ be the
+
 \subsection{Give the RIC algorithm}
+
 \subsection{Prove the expected running time}
@@
 \section{A datastructure for horizontal line segments}
+
 \subsection{Analyze the storage requirements}
+%TODO: warning: this is x and y descisions.
 \subsection{Argue that all horizontal line segs are found}
 \subsection{Give psuedo code for the query algorithm}
+%TODO: sla lijnen op geen punten.
+
 \subsection{Analyze the query time}
+
 \section{Voronoi nearest neigbour\emph{s}}
 %TODO: wtf? maybe with higher order voronoi?? But it won't be log n
+%TODO: we hebben ook bogen. Kun je representateren als een lijn en een punt.
 \end{document}

### 2016-03-24 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
special constraints idea...

@@
 \subsection{Per case, what can you say about $H$}
-Let $H_{int}$ be the intersection of halfplanes of $H$.
+We number the half-planes of $H$: $h_1, h_2,\dots,h_i$.
+Let $H_{i}$ be the set of the first $i$ constraints with special constraints
+$x_1$ and $x_2$ both one unit away from the x axis and pointing to the x
+axis, and
+let $C_{i}$ be the feasible region defined by these constranits:
+
+\[ H_i := \{x_1, x_2, h_1, h_2,\dots, h_i\}\]
+\[ C_i := x_1 \cap x_2 \cap h_1 \cap h_2 \cap \dots \cap h_i\]
+
 \begin{enumerate}[i]
@@
 
-		$H_{int}$ never crosses the x axis,
-		so it exists entirely above or below $y=0$. Or the intersection is empty.
+	The intersection $C_i$ is empty.
 
@@
 
-		$H_{int}$ has no parallel line  to te x axis and
+		$C_i$ has no parallel line  to te x axis and
 		at most one line crossing the x axis.
@@
 
-		$H_{int}$ crosses the x-axis at least 2 times and
+		$C_i$ crosses the x-axis at least 2 times and
 		if there is a line parallel to the x axis then both x-axis crossing
@@
 
-		$H_{int}$ produces an area with lines parallel to the x-axis.
+		$C_i$ produces an area with lines parallel to the x-axis.
 
@@
 	\EndFunction
-	\State 
-	\State 
 \end{algorithmic}
@@
 %TODO: een beeetje overtypen van 4.5
-Let $1 \leq i \leq n$, and let $c_i$ be the
+Let $1 \leq i \leq n$, let $C_i$ be the defined as above and let $d_i$ be
+the largest disk in $C_i$. Then we have
+\begin{enumerate}[i]
+	\item If $d_{i-1} \in h_{i},$ $then$ $d_i = d_{i-1}$
+	\item If $d_{i-1} \in h_{i},$ $then$ $d_i = d_{i-1}$
+\end{enumerate}
 

### 2016-03-24 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
threw away special constraint idea

@@
 We number the half-planes of $H$: $h_1, h_2,\dots,h_i$.
-Let $H_{i}$ be the set of the first $i$ constraints with special constraints
-$x_1$ and $x_2$ both one unit away from the x axis and pointing to the x
-axis, and
+Let $H_{i}$ be the set of the first $i$ constraints, and
 let $C_{i}$ be the feasible region defined by these constranits:
 
-\[ H_i := \{x_1, x_2, h_1, h_2,\dots, h_i\}\]
-\[ C_i := x_1 \cap x_2 \cap h_1 \cap h_2 \cap \dots \cap h_i\]
+\[ H_i := \{h_1, h_2,\dots, h_i\}\]
+\[ C_i := h_1 \cap h_2 \cap \dots \cap h_i\]
 
@@
 
-	The intersection $C_i$ is empty.
+		$C_i$ never crosses the x axis,
+		so it exists entirely above or below $y=0$. Or the intersection is empty.
 

### 2016-03-30 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
trying to prove the lemma

@@
 
-\begin{enumerate}[i]
+\begin{enumerate}[(i)]
 	\item \emph{There is no disk that has its centor non the $x$ axis
@@
 \subsection{Formulate and proof the RIC lemma}
-%TODO: een beeetje overtypen van 4.5
+\paragraph{Lemma}
 Let $1 \leq i \leq n$, let $C_i$ be the defined as above and let $d_i$ be
 the largest disk in $C_i$. Then we have
-\begin{enumerate}[i]
-	\item If $d_{i-1} \in h_{i},$ $then$ $d_i = d_{i-1}$
-	\item If $d_{i-1} \in h_{i},$ $then$ $d_i = d_{i-1}$
+\begin{enumerate}[(i)]
+	\item If $d_{i-1} \in h_i,$ $then$ $d_i = d_{i-1}$
+%TODO: C = nill? dat klopt volgens mij niet
+	\item If $d_{i-1} \not\in h_{i},$ $then$ $C_i = \emptyset$
+		or $b_i \in l_i$ where $b_i$ is the boundary of $d_i$ and $l_i$
+		is the line bounding $h_i$.
+\end{enumerate}
+\paragraph{Proof}
+\begin{enumerate}[(i)]
+	\item Let $d_{i-1} \in h_i$. Because $C_i = C_{i-1}\cap h_i$ and
+		$d_{i-1} \in C_{i-1}$. Furtheremore,the optimal disk in $C_i$
+		cannot be better than the optimal disk in $C_{i-1}$, since
+		$C_i \subseteq C_{i-1}$. Hence, $d_{i-1}$ is the optimal disk
+		in $C_i$ as well.
+	\item Let $d_{i-1} \not\in h_{i}$. Suppose for a contradiction that
+		$C_i \not\emptyset$ and that $b_i \not\in l_i$. Now consider
+		the line segment $\overline{cx_{i-1}cx_i}$ where $cx_i$ is the
+		center on the x-axis of disk $d_i$. We have
+		$d_{i-1} \in C_{i-1}$ and, since $C_i \subset C_{i-1}$, also
+		$d_i \in C_{i-1}$. 
 \end{enumerate}

### 2016-03-31 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
added algorithm and extended the proof

--- 2016-03-24 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
@@
 
-		$C_i$ never crosses the x axis,
-		so it exists entirely above or below $y=0$. Or the intersection is empty.
+		$C_i$ does not intersect the x axis. So its either an empty set or entirly above or below the x axis.
 
@@
 \subsection{Formulate and proof the RIC lemma}
-%TODO: een beeetje overtypen van 4.5
+\paragraph{Lemma}
 Let $1 \leq i \leq n$, let $C_i$ be the defined as above and let $d_i$ be
 the largest disk in $C_i$. Then we have
-\begin{enumerate}[i]
-	\item If $d_{i-1} \in h_{i},$ $then$ $d_i = d_{i-1}$
-	\item If $d_{i-1} \in h_{i},$ $then$ $d_i = d_{i-1}$
+\begin{enumerate}[(i)]
+	\item If $d_{i-1} \subset h_i,$ $then$ $d_i = d_{i-1}$
+%TODO: C = nill? dat klopt volgens mij niet
+	\item If $d_{i-1} \not\subset h_{i},$ $then$ $C_i$ does not intersect the x-axis
+		or $b_i \cap l_i \neq \emptyset$ where $b_i$ is the boundary of $d_i$ and $l_i$
+		is the line bounding $h_i$.
 \end{enumerate}
+\paragraph{Proof}
+\newpage
+\begin{enumerate}[(i)]
+	\item Let $d_{i-1} \subset h_i$. Because $C_i = C_{i-1}\cap h_i$ and
+		$d_{i-1} \subset C_{i-1}$ this means that $d_{i-1} \subset C_i$.
+		Furtheremore,the optimal disk in $C_i$
+		cannot be better than the optimal disk in $C_{i-1}$, since
+		$C_i \subseteq C_{i-1}$. Hence, $d_{i-1}$ is the optimal disk
+		in $C_i$ as well.
+%TODO: replace with using a line between center(d_{i-1}) and
+%		center(d_i) then reason about how d_i will keep growing
+%		 until it hits h_i.
+	\item Let $d_{i-1} \not\subset h_{i}$. Suppose for a contradiction that
+		$C_i$ intersects the x-axis and that $b_i \cap l_i = \emptyset$.
+		Consider the linesegment $s$ between $center(d_i)$ and 
+		$center(d_{i-1})$.
+		Where $center$ is a function that returns the center vertex of a disk.
+		We have $d_{i-1} \subset C_{i-1}$ and
+		since $C_i \subseteq C_{i-1}$, also $d_i \subset C_{i-1}$.
+		Together with the convexity of $C_{i-1}$, this implies that
+		the line segment $s$ is contained within $C_{i-1}$.
+		Since $d_{i-1}$ is the optimal disk in $C_{i-1}$ and the optimum is to
+		maximize $r_i$ the radius of $d_i$. Consider $d_i$
+		where $s$ is maximized, now to increase $r_i$ we move $d_i$ along $s$,
+		we can keep doing this until $b_i \cap h_i$.
+		So $d_i$ needs to intersect $h_i$ otherwise $d_i$ is not the optimum
+		disk in $C_i$. This shows it cannot be the case that $C_i$ intersects
+		the x-axis and $b_i\cap l_i = \emptyset$.
 
+\end{enumerate}
 \subsection{Give the RIC algorithm}
+
+\begin{algorithmic}[1]
+	\Function{calulateBiggestDisk}{$H$}
+		\State
+		$d_0 \gets arbitrarlyLargeDisk()$
+		\If{hasArbitrarlylargeDisc($H$)}
+			\State
+			\Return $d_0$
+		\EndIf
+		\State
+		$H_r \gets$ random shuffle $H$
+		\For{$i\gets 1$ to $n$ with $h_i \in H_r$}
+			\State
+			$d_i \gets d_{i-1}$
+			\If{$d_{i-1} \not\in h_i$}
+				$d_i \gets$
+			\EndIf
+		\EndFor
+	\EndFunction
+\end{algorithmic}
 

### 2016-03-31 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
overhauled 1a and added a temporarily solution for the RIC

@@
 
+And so $C$ is $\cap H$.
+
 \begin{enumerate}[(i)]
@@
 
-		$C_i$ does not intersect the x axis. So its either an empty set or entirly above or below the x axis.
+		$C$ does not intersect the x axis. So its either an empty set or entirly above or below the x axis.
 
-	\item \emph{There are arbitrarly large disk with the specifications.}
+	\item \emph{There lare arbitrarly large disk with the specifications.}
 
-		$C_i$ has no parallel line  to te x axis and
-		at most one line crossing the x axis.
+		$C$ has no parallel bounds to te x axis and
+		at most one bounding linesegment crossing the x axis.
 
@@
 
-		$C_i$ crosses the x-axis at least 2 times and
-		if there is a line parallel to the x axis then both x-axis crossing
-		to be in range of the diameter of the disk.
+		$C$ bounding linesegments cross the x-axis 2 times and
+			the disk border is intersected at least 2 times by non x axis
+			parralel bounding linesegments.
 
@@
 
-		$C_i$ produces an area with lines parallel to the x-axis.
+		$C$ produces an area with one or two bounding linesgments
+		parallel to the x-axis and these linesegment intersect the border
+		of the possible disks.
 
@@
 	\item If $d_{i-1} \subset h_i,$ $then$ $d_i = d_{i-1}$
-%TODO: C = nill? dat klopt volgens mij niet
 	\item If $d_{i-1} \not\subset h_{i},$ $then$ $C_i$ does not intersect the x-axis
@@
 \paragraph{Proof}
-\newpage
 \begin{enumerate}[(i)]
@@
 			$d_i \gets d_{i-1}$
-			\If{$d_{i-1} \not\in h_i$}
-				$d_i \gets$
+			\If{$d_{i-1} \not\subset h_i$}
+			\State
+			$d_i \gets createLargestDiskFrom(h_i, \{h_0\dots,h_{i-1}\})$
 			\EndIf
@@
 % Don't forget to remove constants
+
 \subsection{Eplain why you can't use RIC for points}

### 2016-03-31 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
expended 1a even more

@@
 			the disk border is intersected at least 2 times by non x axis
-			parralel bounding linesegments.
+			parralel bounding linesegments, one linesegment intersects the
+			left part of the disk and another the right part.
+			An alternative defintion is not case 1,2 and 4.
 

### 2016-03-31 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
most of the psuedo code foundation for calulate the biggest disk

@@
 			\State
-			$d_i \gets createLargestDiskFrom(h_i, \{h_0\dots,h_{i-1}\})$
+			$d_i \gets calculateBiggestDisk(h_i, \{h_0\dots,h_{i-1}\})$
 			\EndIf
 		\EndFor
+	\EndFunction
+	\Function{calulateBiggestDisk}{$h_{bound}$,$H_{prev}$}
+		\State
+		$d_0 \gets arbitrarlyLargeDisk()$
+		\For{$i\gets 1$ to $n$ with $h_i \in H_{prev}$}
+			\State
+			$d_i \gets d_{i-1}$
+			\If{$h_i$ does not point in the same direction as $h_{bound}$}
+			\If{$d_{i-1} \not\subset h_i$}
+				\If{$h_i \cap h_{bound}$ does not intersect the x-axis}
+					\State
+					\Return $None$
+				\EndIf
+				\If{$h_i$ is parralel to the x axis}
+					%TODO: parralel to x axis!
+					$d_i \gets$ Calulate the largest disk by
+					creating an imaginary triangle with vertex $a$ being the
+					intersection between $l_i$ and $l_{bound}$ and vertex $b$
+				\Else
+					$d_i \gets$ Calulate the largest disk by
+					creating an imaginary triangle with vertex $a$ being the
+					intersection between $l_i$ and $l_{bound}$ and vertex $b$
+					and $c$ being the $-Ycoordinate(a)$ equal to $l_i$ and
+					$l_{bound}$ respectivily. (because you use negative Y
+					coordinate the disk center will always be on the x axis).
+				\EndIf
+				\State
+			\EndIf
+			\EndIf
+		\EndFor
+
 	\EndFunction

### 2016-04-02 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
finsihed RIC algorithm
started with the analises
finished e and f

--- 2016-03-31 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
@@
 \begin{algorithmic}[1]
-	\Function{calulateBiggestDisk}{$H$}
+	\Function{calculateBiggestDisk}{$H$}
 		\State
@@
 			\State
-			$d_i \gets createLargestDiskFrom(h_i, \{h_0\dots,h_{i-1}\})$
+			$d_i \gets calculateBiggestDiskFrom(h_i, \{h_0\dots,h_{i-1}\})$
+			\If{$d_i = None$}
+				\State
+				\Return $None$
+			\EndIf
 			\EndIf
 		\EndFor
+		\State
+		\Return $d_n$
+	\EndFunction
+	\Function{calculateBiggestDiskFrom}{$h_{bound}$,$H_{prev}$}
+		\State
+		$d_0 \gets arbitrarlyLargeDisk()$
+		\For{$i\gets 1$ to $n$ with $h_i \in H_{prev}$}
+			\State
+			$d_i \gets d_{i-1}$
+			\If{$h_i$ does not point in the same direction as $h_{bound}$}
+			\If{$d_{i-1} \not\subset h_i$}
+				\If{$h_i \cap h_{bound}$ does not intersect the x-axis}
+					\State
+					\Return $None$
+				\EndIf
+				\If{$h_i$ is parralel to the x axis}
+				\State
+				%TODO, the calculation with math?
+				$d_i \gets$ Calculate the leftmost largest disk in $h_{bound} \cap h_{i}$
+				\Else
+				\State
+				$d_i \gets$ Calculate the largest disk in $h_{bound} \cap h_{i}$
+				\EndIf
+				\State
+			\EndIf
+			\EndIf
+		\EndFor
+		\State
+		\Return $d_n$
 	\EndFunction
@@
 \subsection{Prove the expected running time}
+It should be noted that both $calculateBiggestDisk(H)$ and
+$calculateBiggestDiskFrom(h,H)$ are both linear time functions.
+And so worst case $calculateBiggestDisk(H)$ is worst case:
+
+\[O(n^2)\]
+
 % Don't forget to remove constants
+We will prove that a$calculateBiggestDisk(H)$ will run in $\Theta(n)$ expected
+time with help of backwards analises: Consider the situation after $h_i$ is
+inserted, and $d_i$ is computed (either by $d_i = d_{i-1}$ or with help
+of $calculateBiggestDiskFrom(h_i,H)$). If $h_i$ does not bound the feasible
+region, or $d_i \subset h_i$, then the addition step was cheap
+and took $\Theta(1)$ time. But if $d_i \not\subset h_i$ then the addition
+step was expensive.
+
+There are $i$ half-planes that could have been defining $d_i$, and $i-2$
+of these are in random order. Since the order was reandom eacho of the $i-2$
+half-planes has the same probability to be the last one added, and only
+$\leq$ 2 of these cause the expensive step.
+
+Excpected time for $i$-th addition:
+
+\[\frac{i-4}{i-2}\cdot\Theta(1)+\frac{2}{i-2}\cdot\Theta(i)=\Theta(1)\]
+
+Total expected running time:
+
+\[3\Theta(n) = \Theta(n)\]
 
 \subsection{Eplain why you can't use RIC for points}
+To create a RIC algorithm with linear expected time we need to have the
+following properties, lets go trough each of one to see if they get violated.
+
+\begin{itemize}
+	\item \emph{The test wether the next input object violates the current
+		optimum must be possible and fast.}
+		This is true for this problem. Checking if a point is inside a disk
+		is trivially easy.
+	\item \emph{If the next input boject violates the current optimum,
+		finding the new optimum must be an easier problem than the general
+		problem.}
+		This is the violating property. because you can still move
+		the disk having a point violate the current disk may not be a point
+		wich with the disks boundary has to intersect.
+
+		A more crude way to put it is that you can have multiple optimum
+		that don't intersect the same points.
+	\item \emph{The optimum must already be defined by O(1) of the input
+		objects.}
+		I think this is true, you can use some math to figure out a new
+		optimum.
+	\item \emph{The analysis must work out.}
+		It doesn't work out because you can't gaurantee that you've just
+		found a local optimum instead of a global optimum. I mean you
+		could use this as a heuristic but not as an anwser.
+\end{itemize}
 \section{A datastructure for horizontal line segments}

### 2016-04-03 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
dedicaded functions

@@
 \renewcommand{\thesubsection}{\thesection.\alph{subsection}}
+\algrenewcomment[1]{\(\triangleright\) #1}
 \begin{document}
@@
 			\State
-			$d_i \gets calculateBiggestDiskFrom(h_i, \{h_0\dots,h_{i-1}\})$
+			$d_i \gets \Call{calculateBiggestDiskFrom}{h_i, \{h_0\dots,h_{i-1}\}}$
 			\If{$d_i = None$}
@@
 \section{A datastructure for horizontal line segments}
+\subsection{Analyze the storage requirements}
+As can be read from lemma 10.2 (and the suceeding proof) from
+the book \cite{book}, an interval tree uses $O(n)$ storage.
 
-\subsection{Analyze the storage requirements}
-%TODO: warning: this is x and y descisions.
+Now we have the outer tree that is a binary tree based on $y$ coordinates.
+For each interval node $v$ we have a canoconical subset $S_v \subseteq S$
+of all linesegments that appear in the subtree rooted at node $v$.
+Note that $subtree(c_{left}(v))\cup subtree(c_{right}(v)) = subtree(v)$
+where $subtree$ is a function that returns the subtree of $v$ and 
+$c_{right}$ returns the right child and $c_{left}$ returns the left child.
+Every level of the outer tree has an $n$ amount of items stored in its
+subtrees. Because at deeper level the subtrees get split further.
+This results in the following order of storage:
+
+\[ O(n \cdot log(n))\]
+
 \subsection{Argue that all horizontal line segs are found}
+All line segments are found because the binary search tree will
+select as many as possible y based nodes and then ask the
+
+\newpage
 \subsection{Give psuedo code for the query algorithm}
-%TODO: sla lijnen op geen punten.
-
+\begin{algorithmic}[1]
+	\Function{select}{$v$, $x \times [y:y']$}
+		\If{$v$ is a leaf and $y < y_{val}(v) < y'$}
+			\State
+			\Return \Call{selectRange}{$subtree(v), x$}
+		\EndIf
+		\If{$y' < y_{val}(v)$}
+		\State
+		\Return \Call{select}{$c_{left}(v), x \times [y:y']$}
+		\EndIf
+		\If{$y > y_{val}(v)$}
+		\State
+		\Return \Call{select}{$c_{right}(v), x \times [y:y']$}
+		\EndIf
+		\State
+		\Comment{If we get here it means we have a split node}
+		\State
+		\Return $\Call{reportLeft}{v, x\times y, c_{right}}\cup
+		\Call{reportRight}{v,x\times y', c_{left}}$
+	\EndFunction
+	\Function{reportRight}{$v$, $x \times y$}
+		\If{$isLeaf(v)$}
+			\State
+			\Return $\emptyset$
+		\EndIf
+		\If{$y \leq y_{val}(v)$}
+		\State
+		\Return $\Call{reportRight}{c_{left}(v), x \times y} \cup
+		\Call{selectRange}{subtree(c_{right}(v)),x}$
+		\EndIf
+		\State
+		\Return $\Call{reportRight}{c_{right}(v), x \times y}$
+	\EndFunction
+	\Function{reportLeft}{$v$, $x \times y$}
+		\If{$isLeaf(v)$}
+			\State
+			\Return $\emptyset$
+		\EndIf
+		\If{$y > y_{val}(v)$}
+		\State
+		\Return $\Call{reportLeft}{c_{right}(v), x \times y} \cup
+		\Call{selectRange}{subtree(c_{left}(v)),x}$
+		\EndIf
+		\State
+		\Return $\Call{reportLeft}{c_{left}(v), x \times y}$
+	\EndFunction
+	\Function{selectRange}{$v$, $x$}
+	\If{$v$ is a leaf}
+		\State
+		\Return $\emptyset$
+	\EndIf
+	\If{$x < x_{mid}(v)$}
+		\State
+		$r \gets $ starting at the leftmost endpoint keep
+		returning all the intervals in $l_{left}(v)$ untill the
+		intervals no longer contain $x$. (basically a takewhile
+		operation in your favorite langauge).
+		\State
+		\Return $r \cup \Call{selectRange}{c_{left}(v),x}$
+	\EndIf
+	\State
+	$r \gets $ starting at the rightmost endpoint keep
+	returning all the intervals in $l_{right}(v)$ until the
+	intervals no longer contain $x$.
+	\State
+	\Return $r \cup \Call{selectRange}{c_{right}(v),x}$
+	\EndFunction
+\end{algorithmic}
 \subsection{Analyze the query time}
@@
 %TODO: we hebben ook bogen. Kun je representateren als een lijn en een punt.
+
+\begin{thebibliography}{9}
+\bibitem{book}
+de Berg et al,
+\textit{Computational geometry and applications }.
+3rd edition, 2008. Springer-Verlag Berlin Heidelberg.
+\end{thebibliography}
 \end{document}

### 2016-04-03 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
arguing

@@
 \subsection{Argue that all horizontal line segs are found}
-All line segments are found because the binary search tree will
-select as many as possible y based nodes and then ask the
+All lines are found because the outer tree just selects based on the y range
+as soon as a split node is encountered its starts reporting.
+Then the $x$ interval tree will filter out all lines that are not on
+the x at that level.
 

### 2016-04-03 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
generic approach
finished the algorithm

--- 2016-04-03 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
@@
 \subsection{Argue that all horizontal line segs are found}
-All line segments are found because the binary search tree will
-select as many as possible y based nodes and then ask the
+The outer tree will not start reporting untill it encounters a
+split node. Then it will only report the inner tree if the
+value of the nodes fall in the range of the query. So for
+the $[y:y']$ part of the query the algorithm is correct.
 
-\newpage
+The intervall tree for the $x$ part of the query will start
+reporting as soon as it encounters a decision. However it
+will only report the linesegements that intersect with $x$.
+
 \subsection{Give psuedo code for the query algorithm}
@@
 	\EndFunction
+	\algstore{tree}
+\end{algorithmic}
+\newpage
+\begin{algorithmic}[1]
+	\algrestore{tree}
 	\Function{reportRight}{$v$, $x \times y$}
+		\State
+		$\phi \gets y \leq y_{val}(v)$
 		\If{$isLeaf(v)$}
+			\If{$\phi$}
+				\State
+				\Return $\Call{selectRange}{subtree(v),x}$
+			\EndIf
 			\State
@@
 		\EndIf
-		\If{$y \leq y_{val}(v)$}
+		\If{$\phi$}
 		\State
@@
 	\Function{reportLeft}{$v$, $x \times y$}
+		\State
+		$\phi \gets y > y_{val}(v)$
 		\If{$isLeaf(v)$}
+			\If{$\phi$}
+				\State
+				\Return $\Call{selectRange}{subtree(v),x}$
+			\EndIf
 			\State
@@
 		\EndIf
-		\If{$y > y_{val}(v)$}
+		\If{$\phi$}
 		\State

### 2016-04-03 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
query time

@@
 Every level of the outer tree has an $n$ amount of items stored in its
-subtrees. Because at deeper level the subtrees get split further.
-This results in the following order of storage:
+subtrees, and there are $log(n)$ levels because no 2 line segments have the
+same y coordinate. This results in the following order of storage:
 
@@
 will only report the linesegements that intersect with $x$.
+
+The result also will not contain any duplicates, because
+when someting is repoted to the right in the main tree
+the algorithm will go to the left, so the next tree only
+contains things that weren't in the original report.
 
@@
 
+When we traverse the tree we spend at most $O(1+r_v)$ time at
+that node $v$ where $r_v$ is the time spend reporting the results.
+Summing over all $r_v$ results in $r$. The depth of the main tree
+is $log(n)$. we visit at most two nodes at any depth of the tree.
+This results in:
+
+\[ 2\cdot (log(n)+r)\]
+
+time to traverse it. Now we will fill in R by analysing the
+inner tree. At any node $v$ we visit we spend $O(1+k_v)$ time where
+$k_v$ is the number of linesegments we report at $v$. Summing
+over $k_v$ results in $k$. The depth of the tree is $log(n)$,
+we visit at most one node at any depth of the tree. This results in:
+
+\[ log(n)+k\]
+
+So now we can fill in $r$:
+
+\[ 2\cdot (log(n)+log(n)+k)\]
+
+Which results in:
+
+\[ O(log(n)+k) \]
+
+query time.
 \section{Voronoi nearest neigbour\emph{s}}
+\subsection{Storage}
+\subsection{Query time}
 %TODO: wtf? maybe with higher order voronoi?? But it won't be log n

### 2016-04-03 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
fixed temporal analysis.

@@
 
-When we traverse the tree we spend at most $O(1+r_v)$ time at
-that node $v$ where $r_v$ is the time spend reporting the results.
-Summing over all $r_v$ results in $r$. The depth of the main tree
-is $log(n)$. we visit at most two nodes at any depth of the tree.
-This results in:
-
-\[ 2\cdot (log(n)+r)\]
-
-time to traverse it. Now we will fill in R by analysing the
-inner tree. At any node $v$ we visit we spend $O(1+k_v)$ time where
+We'll start by analysing the inner tree.
+At any node $v$ we visit we spend $O(1+k_v)$ time where
 $k_v$ is the number of linesegments we report at $v$. Summing
@@
 
-So now we can fill in $r$:
+So in the main tree we spend $O(log(n)+r_v)$ per call, where $r_v$ is
+the result for node $v$. Its also worth noting that we visit each level
+of the tree a maximum two times. Hence the total time we spend is:
 
-\[ 2\cdot (log(n)+log(n)+k)\]
+\[ 2\cdot \sum_v(log(n)+r_v)\]
 
+Notice that $\sum_v r_v = r$ and that the search path of $y$ and $y'$
+in the main tree have depth $log(n)$. so $\sum_v O(log(n))=log^2(n)$.
 Which results in:
 
-\[ O(log(n)+k) \]
+\[ O(log^2(n)+r) \]
 
-query time.
 \section{Voronoi nearest neigbour\emph{s}}

### 2016-04-04 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
did the bonus and the storage of 3

@@
 
+\subsection{Bonus}
+Applying what we've learned above we know we can handle ranges by using
+binary trees. For single values we can use interval trees.
+So suppose we know that $X$ will be an interval. Then the main tree will
+be a binary tree sorting on $x$ where each node has a two sub-binary trees $y$
+and $z$.
+When we start reporting for x we'll choose the associated subtree which
+is the interval qeury.
+If $Y$ is an interval then we choose the $y$ subtree.
+In the $y$ tree nodes have ascsociated structures that are interval trees
+on $z$.
+If $Z$ is the interval then we choose the $z$ subtree. In $z$ subtree
+there is again an associated structure for each node which is an interval
+tree on $y$.
+We can create another structure for when $X$ is not an interval. Similiar
+to the one above so that we can handle any range query.
+
+The storage requirement is:
+
+\[O(n\cdot log^2(n))\]
+
+Because the main tree is amplified aditionally in size by an extra associated
+structure depth.
+The query time is:
+
+\[ O(log^3(n)+r) \]
+
+Because there are 3 trees to traverse and we still need to report the anwsers.
 \section{Voronoi nearest neigbour\emph{s}}
 \subsection{Storage}
+The original datastructure was already of order:
+
+\[O(n)\]
+
+storage, this was a given, now we will start analyising how the extension
+will impact this order.
+For every point $p$ we store a trapazoidal map, which is $O(n)$ size
+storage (Theorom 6.3 of the book \cite{book}) where $n$ equals the amount of
+voronoi neighbours $k_p$. So per point we now have a storage of the order:
+
+\[O(1+k_p)\]
+
+However we can put an expectation on $k_p$ by using theorom 7.3 of the book
+\cite{book}. It says the number of edges is at most $3n-6$, an edge is a
+division between neigbours so the expected amount of neigbours per point
+is $3$. So the expected storage complexity is:
+
+\[ O(3n) = O(n) \]
 \subsection{Query time}
-%TODO: wtf? maybe with higher order voronoi?? But it won't be log n
-%TODO: we hebben ook bogen. Kun je representateren als een lijn en een punt.
 

### 2016-04-04 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
finished 3??

@@
 division between neigbours so the expected amount of neigbours per point
-is $3$. So the expected storage complexity is:
+is $\frac{3n-6}{n} \approx 3$. So the storage complexity is:
 
 \[ O(3n) = O(n) \]
+
+Note that this is not an expected complexity because the theorom 7.3 puts
+a hard upperbound on the amount of neigbours as long as $n\geq3$.
 \subsection{Query time}
+The original datastructure had an expected query time of the follwoing order:
+
+\[O(log(n))\]
+
+This was given. Now we will analyse if the extension will impact this
+perormance. The substructure is a trapezoidal map which has an expected
+query time of $O(log(k))$ where $k$ is the amount of neigbours for a
+particular point. If we combine this with the existing structure
+we get an expected query time of:
+
+\[O(log(n) \cdot log(k)) \]
 

### 2016-04-04 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
spelling

--- 2016-04-03 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
@@
 Let $H_{i}$ be the set of the first $i$ constraints, and
-let $C_{i}$ be the feasible region defined by these constranits:
+let $C_{i}$ be the feasible region defined by these constraints:
 
@@
 \begin{enumerate}[(i)]
-	\item \emph{There is no disk that has its centor non the $x$ axis
+	\item \emph{There is no disk that has its center non the $x$ axis
 			that lies inside all half-planes.}
 
-		$C$ does not intersect the x axis. So its either an empty set or entirly above or below the x axis.
+		$C$ does not intersect the x axis. So its either an empty set or entirely above or below the x axis.
 
-	\item \emph{There lare arbitrarly large disk with the specifications.}
+	\item \emph{There arbitrary large disk with the specifications.}
 
-		$C$ has no parallel bounds to te x axis and
-		at most one bounding linesegment crossing the x axis.
+		$C$ has no parallel bounds to the x axis and
+		at most one bounding line segment crossing the x axis.
 
@@
 
-		$C$ bounding linesegments cross the x-axis 2 times and
+		$C$ bounding line segments cross the x-axis 2 times and
 			the disk border is intersected at least 2 times by non x axis
-			parralel bounding linesegments, one linesegment intersects the
+			parallel bounding line segments, one line segment intersects the
 			left part of the disk and another the right part.
-			An alternative defintion is not case 1,2 and 4.
+			An alternative definition is not case 1,2 and 4.
 
@@
 
-		$C$ produces an area with one or two bounding linesgments
-		parallel to the x-axis and these linesegment intersect the border
+		$C$ produces an area with one or two bounding line segments
+		parallel to the x-axis and these line segment intersect the border
 		of the possible disks.
@@
 \newpage
-\subsection{Deal with case ii in linear worstcase time}
+\subsection{Deal with case ii in linear worst case time}
 \begin{algorithmic}[1]
-	\Function{hasArbitrarlylargeDisc}{$H$}
+	\Function{hasArbitrarylargeDisc}{$H$}
 		\State
@@
 		$d_{i-1} \subset C_{i-1}$ this means that $d_{i-1} \subset C_i$.
-		Furtheremore,the optimal disk in $C_i$
+		Furthermore,the optimal disk in $C_i$
 		cannot be better than the optimal disk in $C_{i-1}$, since
@@
 		$C_i$ intersects the x-axis and that $b_i \cap l_i = \emptyset$.
-		Consider the linesegment $s$ between $center(d_i)$ and 
+		Consider the line segment $s$ between $center(d_i)$ and 
 		$center(d_{i-1})$.
@@
 				\EndIf
-				\If{$h_i$ is parralel to the x axis}
+				\If{$h_i$ is parallel to the x axis}
 				\State
@@
 
-% Don't forget to remove constants
 We will prove that a$calculateBiggestDisk(H)$ will run in $\Theta(n)$ expected
-time with help of backwards analises: Consider the situation after $h_i$ is
+time with help of backwards analyses: Consider the situation after $h_i$ is
 inserted, and $d_i$ is computed (either by $d_i = d_{i-1}$ or with help
@@
 There are $i$ half-planes that could have been defining $d_i$, and $i-2$
-of these are in random order. Since the order was reandom eacho of the $i-2$
+of these are in random order. Since the order was random each of the $i-2$
 half-planes has the same probability to be the last one added, and only
@@
 
-Excpected time for $i$-th addition:
+Expected time for $i$-th addition:
 
@@
 
-\subsection{Eplain why you can't use RIC for points}
+\subsection{Explain why you can't use RIC for points}
 To create a RIC algorithm with linear expected time we need to have the
@@
 \begin{itemize}
-	\item \emph{The test wether the next input object violates the current
+	\item \emph{The test of the next input object violates the current
 		optimum must be possible and fast.}
@@
 		is trivially easy.
-	\item \emph{If the next input boject violates the current optimum,
+	\item \emph{If the next input object violates the current optimum,
 		finding the new optimum must be an easier problem than the general
 		problem.}
-		This is the violating property. because you can still move
+		This is the violating property. Because you can still move
 		the disk having a point violate the current disk may not be a point
-		wich with the disks boundary has to intersect.
+		which with the disks boundary has to intersect.
 
@@
 	\item \emph{The analysis must work out.}
-		It doesn't work out because you can't gaurantee that you've just
+		It doesn't work out because you can't guarantee that you've just
 		found a local optimum instead of a global optimum. I mean you
-		could use this as a heuristic but not as an anwser.
+		could use this as a heuristic but not as an answer.
 \end{itemize}
-\section{A datastructure for horizontal line segments}
+\section{A data structure for horizontal line segments}
 \subsection{Analyze the storage requirements}
@@
 Now we have the outer tree that is a binary tree based on $y$ coordinates.
-For each interval node $v$ we have a canoconical subset $S_v \subseteq S$
-of all linesegments that appear in the subtree rooted at node $v$.
+For each interval node $v$ we have a canonical subset $S_v \subseteq S$
+of all line segments that appear in the subtree rooted at node $v$.
 Note that $subtree(c_{left}(v))\cup subtree(c_{right}(v)) = subtree(v)$
@@
 subtrees, and there are $log(n)$ levels because no 2 line segments have the
-same y coordinate. This results in the following order of storage:
+same $y$ coordinate. This results in the following order of storage:
 
@@
 
-\subsection{Argue that all horizontal line segs are found}
-The outer tree will not start reporting untill it encounters a
+\subsection{Argue that all horizontal line segments are found}
+The outer tree will not start reporting until it encounters a
 split node. Then it will only report the inner tree if the
@@
 
-The intervall tree for the $x$ part of the query will start
+The interval tree for the $x$ part of the query will start
 reporting as soon as it encounters a decision. However it
-will only report the linesegements that intersect with $x$.
+will only report the line segments that intersect with $x$.
 
 The result also will not contain any duplicates, because
-when someting is repoted to the right in the main tree
+when something is reported to the right in the main tree
 the algorithm will go to the left, so the next tree only
@@
 
-\subsection{Give psuedo code for the query algorithm}
+\subsection{Give pseudo code for the query algorithm}
 \begin{algorithmic}[1]
@@
 		$r \gets $ starting at the leftmost endpoint keep
-		returning all the intervals in $l_{left}(v)$ untill the
-		intervals no longer contain $x$. (basically a takewhile
-		operation in your favorite langauge).
+		returning all the intervals in $l_{left}(v)$ until the
+		intervals no longer contain $x$. (basically a takeWhile
+		operation in your favorite language).
 		\State
@@
 At any node $v$ we visit we spend $O(1+k_v)$ time where
-$k_v$ is the number of linesegments we report at $v$. Summing
+$k_v$ is the number of line segments we report at $v$. Summing
 over $k_v$ results in $k$. The depth of the tree is $log(n)$,
@@
 Notice that $\sum_v r_v = r$ and that the search path of $y$ and $y'$
-in the main tree have depth $log(n)$. so $\sum_v O(log(n))=log^2(n)$.
+in the main tree have depth $log(n)$. So $\sum_v O(log(n))=log^2(n)$.
 Which results in:
@@
 
-\section{Voronoi nearest neigbour\emph{s}}
+\subsection{Bonus}
+Applying what we've learned above we know we can handle ranges by using
+binary trees. For single values we can use interval trees.
+So suppose we know that $X$ will be an interval. Then the main tree will
+be a binary tree sorting on $x$ where each node has a two sub-binary trees $y$
+and $z$.
+When we start reporting for x we'll choose the associated subtree which
+is the interval query.
+If $Y$ is an interval then we choose the $y$ subtree.
+In the $y$ tree nodes have associated structures that are interval trees
+on $z$.
+If $Z$ is the interval then we choose the $z$ subtree. In $z$ subtree
+there is again an associated structure for each node which is an interval
+tree on $y$.
+We can create another structure for when $X$ is not an interval. Similar
+to the one above so that we can handle any range query.
+
+The storage requirement is:
+
+\[O(n\cdot log^2(n))\]
+
+Because the main tree is amplified additionally in size by an extra associated
+structure depth.
+The query time is:
+
+\[ O(log^3(n)+r) \]
+
+Because there are 3 trees to traverse and we still need to report the answers.
+\section{Voronoi nearest neighbour\emph{s}}
 \subsection{Storage}
+The original data structure was already of order:
+
+\[O(n)\]
+
+Storage, this was a given, now we will start analysing how the extension
+will impact this order.
+For every point $p$ we store a trapezoidal map, which is $O(n)$ size
+storage (Theorem 6.3 of the book \cite{book}) where $n$ equals the amount of
+Voronoi neighbours $k_p$. So per point we now have a storage of the order:
+
+\[O(1+k_p)\]
+
+However we can put an expectation on $k_p$ by using theorem 7.3 of the book
+\cite{book}. It says the number of edges is at most $3n-6$, an edge is a
+division between neighbours so the expected amount of neighbours per point
+is $\frac{3n-6}{n} \approx 3$. So the storage complexity is:
+
+\[ O(3n) = O(n) \]
+
+Note that this is not an expected complexity because the theorem 7.3 puts
+a hard upper bound on the amount of neighbours as long as $n\geq3$.
 \subsection{Query time}
-%TODO: wtf? maybe with higher order voronoi?? But it won't be log n
-%TODO: we hebben ook bogen. Kun je representateren als een lijn en een punt.
+The original data structure had an expected query time of the following order:
+
+\[O(log(n))\]
+
+This was given. Now we will analyse if the extension will impact this
+performance. The substructure is a trapezoidal map which has an expected
+query time of $O(log(k))$ where $k$ is the amount of neighbours for a
+particular point. If we combine this with the existing structure
+we get an expected query time of:
+
+\[O(log(n) \cdot log(k)) \]
 

### 2016-04-04 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
imporved the proofs of 1, also made psuedocode fit.

@@
 		in $C_i$ as well.
-%TODO: replace with using a line between center(d_{i-1}) and
-%		center(d_i) then reason about how d_i will keep growing
-%		 until it hits h_i.
 	\item Let $d_{i-1} \not\subset h_{i}$. Suppose for a contradiction that
@@
 		the line segment $s$ is contained within $C_{i-1}$.
-		Since $d_{i-1}$ is the optimal disk in $C_{i-1}$ and the optimum is to
-		maximize $r_i$ the radius of $d_i$. Consider $d_i$
-		where $s$ is maximized, now to increase $r_i$ we move $d_i$ along $s$,
+		Consider $d_i$ with the center at the end of $s$.
+		This is the smallest possible disk,
+		now to increase the radius we move $d_i$ along $s$,
 		we can keep doing this until $b_i \cap h_i$.
-		So $d_i$ needs to intersect $h_i$ otherwise $d_i$ is not the optimum
+		$d_i$ needs to intersect $h_i$ otherwise $d_i$ is not the optimum
 		disk in $C_i$. This shows it cannot be the case that $C_i$ intersects
@@
 
-We will prove that a$calculateBiggestDisk(H)$ will run in $\Theta(n)$ expected
+We will prove that $calculateBiggestDisk(H)$ will run in $\Theta(n)$ expected
 time with help of backwards analyses: Consider the situation after $h_i$ is
@@
 \[3\Theta(n) = \Theta(n)\]
+
+One $\Theta(n)$ for dealing with case ii, one $\Theta(n)$ for shuffling and
+the final $\Theta(n)$ for the algorithm.
 
@@
 \subsection{Analyze the storage requirements}
-As can be read from lemma 10.2 (and the suceeding proof) from
+As can be read from lemma 10.2 (and the succeeding proof) from
 the book \cite{book}, an interval tree uses $O(n)$ storage.
@@
 
+\newpage
 \subsection{Give pseudo code for the query algorithm}
@@
 	\EndFunction
-	\algstore{tree}
-\end{algorithmic}
-\newpage
-\begin{algorithmic}[1]
-	\algrestore{tree}
 	\Function{reportRight}{$v$, $x \times y$}
@@
 	\EndFunction
+	\algstore{tree}
+\end{algorithmic}
+\newpage
+\begin{algorithmic}[1]
+	\algrestore{tree}
 	\Function{selectRange}{$v$, $x$}

### 2016-04-04 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
extended the bonus to be more 1 point worthy :p

@@
 
+\newpage
 \subsection{Bonus}
-Applying what we've learned above we know we can handle ranges by using
-binary trees. For single values we can use interval trees.
-So suppose we know that $X$ will be an interval. Then the main tree will
-be a binary tree sorting on $x$ where each node has a two sub-binary trees $y$
-and $z$.
-When we start reporting for x we'll choose the associated subtree which
-is the interval query.
-If $Y$ is an interval then we choose the $y$ subtree.
-In the $y$ tree nodes have associated structures that are interval trees
-on $z$.
-If $Z$ is the interval then we choose the $z$ subtree. In $z$ subtree
-there is again an associated structure for each node which is an interval
-tree on $y$.
-We can create another structure for when $X$ is not an interval. Similar
-to the one above so that we can handle any range query.
+Let $S=\{s_1,\dots,s_n\}$ be a set of $n$ line segments in space each
+of which is parallel to the x-axis, y-axis or z-axis.
+\paragraph{Structure} We use a data-structure whose main tree is a balanced
+binary search tree on the $x$ coordinates of the line segments of $S$.
+For each interval node $v$, we have a canonical subset $S_v\subseteq S$
+of all line segments that appear in the subtree rooted at $v$. The
+set $S_v$ is stored in two associated structures with $v$ as well:
+these are again balanced binary trees, one is on the $y$ coordinates
+of the line segments of $S_v$ and the other one is the $z$ coordinates
+of the line segments of $S_v$. These associated structures are similar
+in structure to the main tree however the associated structures of their
+nodes are interval trees the $y$
+subtree has associated interval trees based on the $z$ interval and
+the $z$ tree has it based on the $y$ interval.
+This structure cannot treat $x$ intervals however. To also treat
+$x$ intervals we create another tree similar to the one above, however
+we substitute $x$ by $y$ and $y$ by $x$.
+\paragraph{Query algorithm} If $X$ is an interval we use the first tree.
+So we go down the tree until we encounter a split node. If we do we
+split up the algorithm, one side starts reporting left and the other one
+right. However we need to choose the right associated
+structure. We first want to do the range query and save the single value for
+last. So if $Y$ in this case is a range query we choose the $y$ associated
+structure when reporting.
+The report call will query the associated structures similarly
+to the select of `2.c'. Only in this case the dimension name of $x$ will be
+replaced by $z$.
+\paragraph{Query time}
+The query time is:
 
+\[ O(log^3(n)+r) \]
+
+Because there are 3 trees to traverse and we still need to report the answers.
+\paragraph{Storage}
 The storage requirement is:
@@
 structure depth.
-The query time is:
-
-\[ O(log^3(n)+r) \]
-
-Because there are 3 trees to traverse and we still need to report the answers.
 \section{Voronoi nearest neighbour\emph{s}}

### 2016-04-25 multiAgentLearning/summeries/InteractiveLearningProblem.tex
new subject!!

\documentclass{article}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{enumerate}
\renewcommand{\thesubsection}{\thesection.\alph{subsection}}
\algrenewcomment[1]{\(\triangleright\) #1}
\begin{document}
\author{Jappie Klooster}
\title{A summery of the interactive learning problem}

\maketitle

Equilibrium is very important to social systems. The nearest
thing we have as a fully developed framework for the study of
equilibria in social systems is Bayesian decision theory.

However there are difficulties in using this. Firstly it requires that
everyone has to anticipate correctly the behaviours of others.
Secondly it requires that individuals improve their strategies
at each point in time as events unfold.
Both these things are incredibility difficult if not impossible to do.

The reason why this only affects social systems and not physical systems
(which also desire equilibria) is that they're self referential.
The act of learning changes the thing to be learned.

\end{document}

### 2016-04-25 multiAgentLearning/summeries/InteractiveLearningProblem.tex
finished the first one

@@
 The act of learning changes the thing to be learned.
+% do double slit to confirm
 
+%now lets talk about not knowing about copy on write
+These difficulties posed by interactive learning effect many areas of economics.
+There is an enormous informational burden on the adjustment
+system of certain price convergence mechanism. This means
+that truly decentralized convergent mechanisms don't grow
+on trees.
+
+These issues also plague macroeconomics. The question
+is weather agents can learn the expected equilibrium from
+out not in equilibrium data.
+A proposed dumbed version system is a real valued variable,
+for example next periods price level, is determined by current
+and historical price level and also by people's expectation
+about next period.
+
+However convergence in such a system can be sensitive to the
+nature of the learning process.
+
+These issues also show up in game theory. Interaction
+between expectation and behaviors is even more starkly present.
+Some people think that games may be learned by simple updating
+procedures, while others think this is not possible, because
+this only works in special cases such as zero sum games.
+They also think the Bayesian methods that lead to learning
+are bad, because it requires to much computation and the
+coordination required to do this requires some pre equilibrium
+in the believe base. (The agents already assumed it was
+possible and were willing to do their bestest best).
+
+There will be Bayesian forced Nash equilibria, however there is
+a limitation to this, because it assumes that players already
+have partial knowledge of each others strategies.
+
+Uncertain games create many issues. Its hard for example
+to guess what the preference is based on just passed play
+if you try to coordinate simultaneously.
 \end{document}

### 2016-04-26 multiAgentLearning/summeries/learningAndTeaching.tex
summeries

\documentclass{article}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{enumerate}
\renewcommand{\thesubsection}{\thesection.\alph{subsection}}
\algrenewcomment[1]{\(\triangleright\) #1}
\begin{document}
\author{Jappie Klooster}
\title{A summery of the interactive learning problem}

\maketitle

A lot of AI involves learning performed by a single agent. In this case the
agent tries to figure out a function. Many techniques have been
developed for this.

However in a multi-agent system the environment contains other
agents that also try to learn a function. Thus the environment
changes without the agent interacting with it. So the learned
function may be wrong.

Simple rules can lead to complex behavior, this leads us to believe
that one cannot separate learning from teaching in multi agent systems.
When deciding what to do, an agent should consider both the things he
learned from other agents passed behavior and how he wants to
influence their future behavior.

A problem arises when both agents want to be play teacher and
both keep selecting uncoordinated actions. The question is if
we can think of a simple rule that will enable them
to coordinate without an external designation of a teacher.

We consider something learning if there is a temporal nature in the domain
and regularity across time. By doing this we can used
passed experience to improve future strategy.

Tit for That is an example of a very simple learning strategy.
You could think of more complex strategies in which you consider
more of the history to play in a more sophisticated manner.
Consider frequency analysis of the opponents play for example.
This is called fictitious play.

Stochastic games are another example of where learning can be
applied.

Other aspects involving both types are weather agents already
know the game. If the game is not known the agents can in
addition learn about the structure of the game.
Another question is if the game is observable, do the
players know each others actions/payoff while playing?

Another model that won't be treated is the investigation
of large populations, this is used in biology.

Learning strategies should be understood as choosing actions
and updating beliefs. A consequence however is that
using `accumulated knowledge' may not always be beneficial
because it pre commits to a given strategy (while others may
be better).

Descriptive theories try to tell us how learning takes place
in real life.
The ideal descriptive theory would have the following two properties:
Realism: `There should be a good match between the formal theory and the
natural phenomenon being studied.', Convergence: `The formal
theory should exhibit interesting behavioral properties, in particular
convergence of the strategy profile being played to some solution concept
(e.g., equilibrium) of the game being played.'.
To demonstrate realism we can use experiments.

The holy grail of convergence properties is demonstrating convergence to
stationary strategies which form a Nash equilibrium of the stage game.
Another approach recognizes Nash equilibria are rare as a result
of a learning process. It requires the empirical frequency of
play convergence to such an equilibrium.
The third approach is giving up on Nash equilibria. This uses
convergence to a correlated equilibrium of the stage game.
The last one gives up on convergence of stationary policies. Instead it
requires that the non-stationary policies converge to an interesting state.

Prescriptive theories describes how learning should happen.
For zero-sum games and even in repeated or stochastic
zero sum games it can be asked if learning is happening optimally.
But in general this won't work because the learning often depends on the
learning of other agents.

The equilibrium of learning strategies is an uncommon approach
instead we use the more common and more practical approach
of asking weather the payoff of a learning strategy is good enough.

There are multiple different payoff requirements:
Safety: `A learning rule is safe if it guarantees the agent at least
its maxmin payoff, or “security value.” (Recall that this is the payoff the agent
can guarantee to himself regardless of the strategies adopted by the opponents)'
, Rationality: `A learning rule is rational if whenever the opponent
settles on a stationary strategy of the stage game (i.e., the opponent adopts the
same mixed strategy each time, regardless of the past), the agent settles on a best
response to that strategy.', No-regret: `Against any set of opponents it yields a payoff that is no less than the
payoff the agent could have obtained by playing any one of his pure strategies
throughout. We give a more formal definition of this condition later in the chapter.'
\end{document}

### 2016-04-26 multiAgentLearning/notes.tex
notes...

\documentclass{article}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{enumerate}
\renewcommand{\thesubsection}{\thesection.\alph{subsection}}
\algrenewcomment[1]{\(\triangleright\) #1}
\begin{document}
\author{Jappie Klooster}
\title{A note on multi agent learning}

\maketitle

Almost surely or a.s for short \} $pr(\dots) = 1$, but this $\neq$ certainity,
because you cannot proof it.

11111111
1211111
41343434
5311111

DIagonal argument. You can use the diagonal to realize
that .... .... .... bla bla not equal subset is equal to 0??

if you can read everything and there is no possiblity of a tie than in a
voting game there is convergence in one round.

Peyton young does some qualititatively analyses.
\end{document}

### 2016-04-27 multiAgentLearning/summeries/reapeted.tex
another summery, someone is applying theory into practice

\documentclass{article}
\usepackage{amsmath}
\begin{document}
\author{Jappie Klooster}
\title{A brief summery of repeated}

\maketitle

Within the prisoners dilemma game the Pareto inefficient outcome
cannot be avoided in a Nash equilibrium. This summery considers
two person infinitely repeated games, subgame perfect
equilibrium and Nash equilibrium.

\[G_p= 
	\bordermatrix{~ & C & D \cr
                  C & 50,50 & 30,60 \cr
                  D & 60,30 & 40,40 \cr}
\]

$G_p$ is a prisoners dilemma game. The strict dominant strategy
for each player is $D$. So $(D,D)$ is the Nash equilibrium of this game.
If $G_p$ is played infinitely many times $t=0,1,2\dots$ and each player
learns what element of the set $\{(C,C),(C,D), (D,C),(D,D)\}$ has occurred.
This induces an infinite stream of associated payoffs and it assumed that
there is a common discount factor $0 < \delta < 1$ so that the final payoff
to each player is $\delta$-discounted value of the infinite stream of
payoffs. Player i(=1,2) obtains:

\[\sum^\infty_{t=0}\delta^t_i\cdot(\text{payoff from $t$-th play})\]

Where $\delta_i=\delta$ for $i=1,2$. A strategy should prescribe an
action for a player at each moment in the history of the game up
to time $t$ of the set $\{(C,C),(C,D),(D,C),(D,D)\}$. The infinite
extensive form game just defined is denoted by $G^\infty_p(\delta)$.
A natural solution to this game is the concept of subgame perfect
equilibrium, where each subgame is equal to the game $G^\infty_p(\delta)$
itself.

A trigger strategy remains the same until some action occurs which
will change the strategy permanently. So you keep playing a cooperating
strategy until your opponent deviates from cooperating, to which you respond
with eternal punishment after that. With common sense we can determine that
such a strategy can solve the prisoners dilemma. This result is an important
reason to study infinitely repeating games.

By using more complex Trigger strategies utility can be increased even further
Such as for example an alternating trigger strategy, where you keep alternating
between playing $C$ and $D$, but as soon as your moves align with the opponent
you play $D$ forever.

If $G$ is an arbitrary $n-$person game, and the strategy combination
$s=(s_1,\dots,s_n)$ is a Nash equilibrium in $G$, If $0<\delta<1$ then each
player $i$ playing $s_i$ at every moment $t$ is a subgame perfect
equilibrium in $G^\infty(\delta)$.

The Folk theorem describes what happens with trigger strategies, players
keep cooperating with each other under the threat of being punished forever.

Even if you relax the possible game by not creating a Nash equilibrium, it
still pays of for players to cooperate with each other (under the threat
of playing the lesser profitable mixed strategy), since when doing
this they can increase their utility.
\paragraph{Questions}
\begin{itemize}
	\item one
	\item two
\end{itemize}

\end{document}

### 2016-05-05 multiAgentLearning/summeries/reinforcementAndRegret.tex
a week of multiAgentLearning.

\documentclass{article}
\usepackage{amsmath}
\usepackage{tensor}
\begin{document}
\author{Jappie Klooster}
\title{Reinforcing regrets}

\maketitle

Reinforcment means you like actions that have had a higher payoff in the passed,
and thus will do them more often.

In every time period $t$, a subject must choose an action from a finite set $X$.
In each period $t$ the oponent generates an action $y \in Y$, which is
or isn't revealed. Let $(x^t, y^t)$ denote the realized action in period $t$.
The \emph{payoff} to teh subject in period $t$ is given by a function
$u(x^t, y^t)$.

Auxilaiarly variable $\theta^t_x \geq 0$ defines the \emph{propensitiy} of playing
action $x$ at time $t$. $q^t_x$ is the \emph{probability} of playing action $x$
at time $t$. A common assumption is that $q^t_x$ is \emph{proportional} to
$\theta^t_x$:
\[q^t_x=\theta^t_x/\sum_{x'\in X} \theta^t_{x'}\]

If $q^t$ and $\theta^t$ are vectors in $R^k$ then the $x^{th}$ components are
$q^t_x$ and $\theta^t_x$.
The payoff at time $t$ is $u^t=u(x^t,y^t)$ which is assumed to be positive.
If $e^t \in R^k$ then:

\[e^t_x = 
\begin{cases}
	1 \text{ if } x \text{ is played at time } t \\
	0 \text{ if } x \text{ is not played at time } t \\
\end{cases}\]

The \emph{linear updating model} takes form:

\[\forall t \geq, \theta^{t-1}=\lambda^t\theta^t+u^te^t+w^t\]

$\lambda^t$ is a discount factor $0 < \lambda^t < 1$. $w^t \in R^k_+$ is
a positive vector of random perturbations. The initial propensitiy vector
$\theta^1$ is strictly positive in all components. A simple model of this
called \emph{cumulative payoff matching}:
\[\theta^{t+1}=\theta^t+u^te^t\]
Here the subject playes action at $t+1$ with a probability proportional to
the action's cumulative payoff untill time $t$.

$\Delta q^t = q^{t+1} - q^t$ is the incremental change in the probability
vector between $t$ and $t+1$. The cumulative payoff in all periods up trough
time t is $\sum_s \leq \tensor*[_t]{\mathbf{u}}{^s}$. Add to this the sum
of the initial propensitiies $|\theta^1|=\sum_x\theta^1_x$ and the resulting
value: $v^t=|\theta^1|+\sum_{s\leq t}u^s$. It follows that:
\[\Delta q^t=[u^t/v^t][e^t-q^t]\]
Since the payoff in each period $u^t$ is bounded and cumulative payoff $v^t$
is unbounded, the incremental impact of payoff in any given period $u^t$ 
diminishes at a rate of $1/t$.

A variant:
\[\Delta q^t = [u^t/(Ct^p+u^t)][e^t-q^t]\]
Where C and p are positive constants. The current payoff is assumed to decrease
at the power of $t$ in this variant. This implies that that subjects that
use such a rule can lock into suboptimal play even against an opponent
who is using a fixed istribution in every period.

Another varient with $\lambda < 1$ and in each period $t$ there are random
perturbations $w^t$:
\[\theta^{t+1}=\lambda\theta^t+u^te^t+w^t\]
Ignoring the pertrubations we can rewrite this in terms of change of the probability
distribution:
\[\Delta q^t=[u^t/v^t(\lambda)][e^t-q^t]\]
\[v^t(\lambda)=\lambda^t|\theta^1|+\sum_{1\leq s \leq t}\lambda^{t-s}u^s\]
The denominator $v^t(\lambda)$ is bounded above, so the marginal impact
of period by period payoff levels eventually.

Another extension uses asspiration level $a^t$. Actions are positively
reinforced if the payoff exceed $a^t$ and negativly otherwise.
\paragraph{Questions}
\begin{itemize}
	\item What is propensitiy?
	\item Which of the proposed models is most popular/used and why?
\end{itemize}

\end{document}

### 2016-06-15 multiAgentLearning/paper/article.tex
a beginning

\abstract{
	%TODO abstract
}

%TODO Intro blurb, if any

\section{Introductions}

Satisficing play was originally introduced
to try and limit the computation power that earlier models required of human
behavior \citep{herbert}. Originating from the field of ecnomics, they
wanted to use a model that was not ``super intelligent''.

Such assumptions can include:

\begin{itemize}
  \item The structure of the game is known.
  \item Awareness of other player's decisions.
  \item Awareness of being in a game situation
  \item Being able to preform maximization calculations.
\end{itemize}

\section{Algorithm}
The core algorithm as presented by \citep{herbert} and formalized by
\citep{karandikar}
can be seen bellow:

\begin{enumerate}
  \item $ \begin{cases}
              \text{if } \pi_t \geq \alpha_t & A_{t+1} = A_t\\
              \text{else} & A_{t+1} \neq A_t
              \end{cases}$
  \item $ \alpha_{t+1} = \lambda \alpha_t + (1-\lambda) \pi_t $ \\
\end{enumerate}
\begin{tabular}{ll}
  Definition & symbol \\ \toprule
  Time & $t$ \\
  Action & $A_t$ \\
  Utility & $\pi_t$ \\
  Aspiration & $\alpha_t$ \\
  Learning rate & $\lambda$ \\ \bottomrule
\end{tabular}

In the first phase the next action is selected. If the action still holds
according to the satificing level or ambition it will remain the same, if not another
action will be selected. In the second phase the satisficing level is updated.
In here the constant learning rate is used. This is a variable that can be
modified to adjust the amount of new information being used or how much
information from memory is used.

\section{Iterated prisoners dilemma}

To demonstrate a potential use of this algorithm we will use at several example
cases. The first one being the iterated prisoners dilemma. Consider
the following normal form game:

\begin{tabular}{lll}
            & Cooperate   & Defect \\
  Cooperate & $(2,2)$     & $(0,3)$  \\
  Defect    & $(3,0)$     & $(1,1)$  \\
\end{tabular}

The Nash equilibrium of this games lies on Defect,Defect. However, all but
defect are pareto optimal actions.

\section{References to treat}
\begin{itemize}
  \item \citep{stimpson:2001}
  \item \citep{stimpson:2003}
  \item \citep{sandholm} % q-learning in iterated prisoners dilemma
  \item \citep{arthur} % maybe not, comparing algorithms against humans
  \item \citep{bowling}
  \item \citep{feltovich}
  \item \citep{crandall}
\end{itemize}

\bibliography{main}

\address{Jappie Klooster\\
  Dept. of Informatics\\
  Universiteit Utrecht\\
  The Netherlands\\}
\email{j.t.klooster@students.uu.nl}

### 2016-06-16 multiAgentLearning/paper/article.tex
Title and added a place for each paper

@@
   %TODO abstract
+	abstract..
 }
-
-%TODO Intro blurb, if any
 
@@
 wanted to use a model that was not ``super intelligent''.
-
-Such assumptions can include:
 
@@
 The Nash equilibrium of this games lies on Defect,Defect. However, all but
-defect are pareto optimal actions.
+that action are pareto optimal actions.
 
-\section{References to treat}
-\begin{itemize}
-  \item \citep{stimpson:2001}
-  \item \citep{stimpson:2003}
-  \item \citep{sandholm} % q-learning in iterated prisoners dilemma
-  \item \citep{arthur} % maybe not, comparing algorithms against humans
-  \item \citep{bowling}
-  \item \citep{feltovich}
-  \item \citep{crandall}
-\end{itemize}
+\section{Properties of satisifing play}
+Using both \citep{karandikar} and \citep{stimpson:2001} works to look deeper
+at the properties of satisifing play.
+
+karandikar used a theoretical proofing method, wehreas stimpson showed some
+experimental results.
+
+also showing the contrast graph which supports the work of \citep{arthur}
+
+\section{An extended satisficing play}
+
+In here we discuss the work of \citep{stimpson:2003}, where he extends the
+algorithm to be played with mutliple action choices.
+
+\section{Satisficing compared to other learning methods}
+With help of the work of \citep{crandall} saticificing play is compared
+to Q-learning and WOLF-PHC.
+Q-learning is explained briefly with help of \citep{sandholm} and WOLF-PHC
+with help of \citep{bowling}.
+
+Finally some graphs are shown how they compare
 

### 2016-06-23 multiAgentLearning/paper/article.tex
added a bunch of summeries and the paper concept

@@
 
-\section{Introductions}
+\section{Introduction}
 
@@
 to try and limit the computation power that earlier models required of human
-behavior \citep{herbert}. Originating from the field of ecnomics, they
+behavior \citep{herbert}. Originating from the field of economics, they
 wanted to use a model that was not ``super intelligent''.
 
+The classical models from the field of economics had at least one of the
+following assumptions:
+
 \begin{itemize}
+\itemsep0em 
   \item The structure of the game is known.
@@
 
-\section{Algorithm}
+These assumptions can be called common knowledge, and are a demanding to
+make \citep{binmore}.
+
+\section{Core algorithm}
 The core algorithm as presented by \citep{herbert} and formalized by
@@
 In the first phase the next action is selected. If the action still holds
-according to the satificing level or ambition it will remain the same, if not another
+according to the satisficing level or ambition it will remain the same, if not another
 action will be selected. In the second phase the satisficing level is updated.
@@
 \section{Iterated prisoners dilemma}
+[To be extended]
 
@@
 The Nash equilibrium of this games lies on Defect,Defect. However, all but
-that action are pareto optimal actions.
+that action are pareto optimal actions. So the Nash equilibrium would be the
+``right'' strategy to play if the game is only played once, but if you repeat
+the game against the same opponent, interestingly enough you can ``solve''
+the prisoners dilemma by ``teaching'' the opponent to cooperate.
 
+However strategies such as tit for tat or win stay lose shift already know
+which actions are winning and which are punishing to try and solve it.
+Satisficing play learns this however by updating the aspiration it got from
+playing a move. If we consider a game vs always cooperate for example:
+
+...
+
+But now if we start with a lower apsiration level:
+
+...
+
+And finally vs tit for tat, a game which can also be found in the
+\citep{stimpson:2001} paper:
+
+...
 \section{Properties of satisifing play}
+[To be extended]
+
 Using both \citep{karandikar} and \citep{stimpson:2001} works to look deeper
@@
 \section{An extended satisficing play}
+[To be extended]
 
@@
 algorithm to be played with mutliple action choices.
+Stimpson created a game where resources had to be allocated among agents.
+
+\begin{tabular}{ll}
+  Definition & symbol \\ \toprule
+  Time & $t$ \\
+  Action & $A_t$ \\
+  Utility & $\pi_t$ \\
+  Aspiration & $\alpha_t$ \\
+  Learning rate & $\lambda$ \\ \bottomrule
+\end{tabular}
 

### 2016-06-27 multiAgentLearning/paper/article.tex
todoing

@@
   %TODO abstract
-	abstract..
+  \todo[inline]{abstract missing}
 }
-
 \section{Introduction}
-
+\toReview
 Satisficing play was originally introduced
@@
 following assumptions:
-
 \begin{itemize}
@@
 The core algorithm as presented by \citep{herbert} and formalized by
-\citep{karandikar}
+\citep{karandikar} \todo[inline]{karandikars algorithm is different..}
 can be seen bellow:
@@
 \section{Iterated prisoners dilemma}
-[To be extended]
-
+\drafting
 To demonstrate a potential use of this algorithm we will use at several example
@@
 
-...
+\todo{copy table}
 
@@
 
-...
+\todo{copy table}
 \section{Properties of satisifing play}
-[To be extended]
+\drafting
 
@@
 \section{An extended satisficing play}
-[To be extended]
+\drafting
 
@@
 \section{Satisficing compared to other learning methods}
+\drafting
 With help of the work of \citep{crandall} saticificing play is compared

### 2016-06-27 multiAgentLearning/paper/article.tex
todos out of the comments

@@
 }
+\todo{be more critical? maybe it depends to much on its paremeters?
+  Or which game types are good for saticifing play?
+}
+\todo{misses callouts, intermezzi and quotes}
 \section{Introduction}
@@
 
-The Nash equilibrium of this games lies on Defect,Defect. However, all but
+The Nash equilibrium of this game lies on Defect,Defect. However, all but
 that action are pareto optimal actions. So the Nash equilibrium would be the
@@
 \drafting
+
 With help of the work of \citep{crandall} saticificing play is compared
@@
 
+\todo{check }
 \bibliography{main}

### 2016-06-27 multiAgentLearning/paper/article.tex
spelling and todos

@@
 The core algorithm as presented by \citep{herbert} and formalized by
-\citep{karandikar} \todo[inline]{karandikars algorithm is different..}
-can be seen bellow:
+\citep{karandikar},
+in here we present the simplified version of \citep{stimpson:2001} first:
 
@@
 \end{enumerate}
+\todo{add table caption}
 \begin{tabular}{ll}
@@
 
-\section{Iterated prisoners dilemma}
+\section{Iterated prisoner's dilemma}
 \drafting
 To demonstrate a potential use of this algorithm we will use at several example
-cases. The first one being the iterated prisoners dilemma. Consider
+cases. The first one being the iterated prisoner's dilemma. Consider
 the following normal form game:
 
+\todo{add table caption}
 \begin{tabular}{lll}
@@
 the game against the same opponent, interestingly enough you can ``solve''
-the prisoners dilemma by ``teaching'' the opponent to cooperate.
+the prisoner's dilemma by ``teaching'' the opponent to cooperate.
+\todo{can one really solve? Is it really teaching? should either add references
+or change wording}
 
@@
 Satisficing play learns this however by updating the aspiration it got from
-playing a move. If we consider a game vs always cooperate for example:
+playing a move. If we consider a game versus always cooperate for example:
 
@@
 
-And finally vs tit for tat, a game which can also be found in the
+And finally versus tit for tat, a game which can also be found in the
 \citep{stimpson:2001} paper:
@@
 \todo{copy table}
-\section{Properties of satisifing play}
+\section{Properties of satisficing play}
 \drafting
@@
 Using both \citep{karandikar} and \citep{stimpson:2001} works to look deeper
-at the properties of satisifing play.
+at the properties of satisficing play.
 
-karandikar used a theoretical proofing method, wehreas stimpson showed some
+karandikar used a theoretical proofing method, whereas Stimpson showed some
 experimental results.
@@
 
-\begin{tabular}{ll}
-  Definition & symbol \\ \toprule
-  Time & $t$ \\
-  Action & $A_t$ \\
-  Utility & $\pi_t$ \\
-  Aspiration & $\alpha_t$ \\
-  Learning rate & $\lambda$ \\ \bottomrule
-\end{tabular}
-
 \section{Satisficing compared to other learning methods}
@@
 
-\todo{check }
+\todo{add caption for tables?}
 \bibliography{main}

### 2016-06-28 multiAgentLearning/paper/article.tex
coppied tables

@@
 
-The classical models from the field of economics had at least one of the
+\toReview The classical models from the field of economics had at least one of the
 following assumptions:
@@
 
-These assumptions can be called common knowledge, and are a demanding to
+\toReview These assumptions can be called common knowledge, and are a demanding to
 make \citep{binmore}.
 
+\todo{give paper overview}
+
 \section{Core algorithm}
-The core algorithm as presented by \citep{herbert} and formalized by
+\toReview The core algorithm as presented by \citep{herbert} and formalized by
 \citep{karandikar},
@@
 
-In the first phase the next action is selected. If the action still holds
+\toReview In the first phase the next action is selected. If the action still holds
 according to the satisficing level or ambition it will remain the same, if not another
@@
 \section{Iterated prisoner's dilemma}
+\todo{make an intermezzo of this?}
 \drafting
@@
 
-The Nash equilibrium of this game lies on Defect,Defect. However, all but
+\drafting The Nash equilibrium of this game lies on Defect,Defect. However, all but
 that action are pareto optimal actions. So the Nash equilibrium would be the
@@
 
-However strategies such as tit for tat or win stay lose shift already know
+\drafting However strategies such as tit for tat or win stay lose shift already know
 which actions are winning and which are punishing to try and solve it.
@@
 
-\todo{copy table}
+\begin{tabular}{lllll}
+    t & Cooperate & $A_t$ & $\pi_t$ & $\alpha_t$ \\
+    0 & C     & C    & 3       & 4 \\
+    1 & C     & D    & 4       & 3.5 \\
+    2 & C     & D    & 4       & 3.75 \\
+\end{tabular}
 
+\drafting In $t=0$ satisficing play doesn't meet its ambition of 4, therefore, in the
+next game it modifies its ambition and changes its action. Here we can see
+that satisficing play does get more utility than its ambition, and is satisfied
+with its action.
 But now if we start with a lower apsiration level:
 
-\todo{copy table}
+\begin{tabular}{lllll}
+  t & Coop & $A_t$ & $\pi_t$ & $\alpha_t$ \\
+  0 & C     & C    & 3       & 2 \\
+  1 & C     & C    & 3       & 2.5 \\
+  2 & C     & C    & 3       & 2.75 \\
+\end{tabular}
 
-And finally versus tit for tat, a game which can also be found in the
+\drafting So here satisficing play finds a Pareto optimal action, but it doesn't find
+the best possible action it could get against this opponent. From this example
+we can also see how dependent satisficing play is on its starting parameters.
+\todo{here we can interject an intermezo, of how karandikar tried to solve
+this with trajection and random pertrubations}
+
+\drafting And finally versus tit for tat, a game which can also be found in the
 \citep{stimpson:2001} paper:
 
-\todo{copy table}
+\begin{tabular}{lllll}
+        t & TT & $A_t$ & $\pi_t$ & $\alpha_t$ \\
+        0 & C     & C    & 3       & 4  \\
+        1 & C     & D    & 4       & 3.5 \\
+        2 & D     & D    & 2       & 3.75 \\
+        3 & D     & C    & 1       & 2.87 \\
+        4 & C     & D    & 4       & 3.435 \\
+        5 & D     & D    & 2       & 2.715 \\
+        6 & D     & C    & 1       & 1.85875 \\
+\end{tabular}
 \section{Properties of satisficing play}

### 2016-06-28 multiAgentLearning/paper/article.tex
added the tables and descriptions

@@
 
-\section{Iterated prisoner's dilemma}
+\section{Satisficing play in action}
 \todo{make an intermezzo of this?}
@@
     t & Cooperate & $A_t$ & $\pi_t$ & $\alpha_t$ \\
-    0 & C     & C    & 3       & 4 \\
-    1 & C     & D    & 4       & 3.5 \\
-    2 & C     & D    & 4       & 3.75 \\
+    0 & C     & C    & 2       & 3 \\
+    1 & C     & D    & 3       & 2.5 \\
+    2 & C     & D    & 3       & 2.75 \\
 \end{tabular}
@@
   t & Coop & $A_t$ & $\pi_t$ & $\alpha_t$ \\
-  0 & C     & C    & 3       & 2 \\
-  1 & C     & C    & 3       & 2.5 \\
-  2 & C     & C    & 3       & 2.75 \\
+  0 & C     & C    & 2       & 1 \\
+  1 & C     & C    & 2       & 1.5 \\
+  2 & C     & C    & 2       & 1.75 \\
 \end{tabular}
@@
 we can also see how dependent satisficing play is on its starting parameters.
-\todo{here we can interject an intermezo, of how karandikar tried to solve
-this with trajection and random pertrubations}
+\todo{here we can interject an intermezzo, of how karandikar tried to solve
+this with trajection and random perturbations}
 
-\drafting And finally versus tit for tat, a game which can also be found in the
-\citep{stimpson:2001} paper:
+\drafting Now consider a game versus tit for tat (TT), Tit for Tat is a
+strategy that always copies the last action of its opponent.
+This game can also be found in the \citep{stimpson:2001} paper:
 
@@
         t & TT & $A_t$ & $\pi_t$ & $\alpha_t$ \\
-        0 & C     & C    & 3       & 4  \\
-        1 & C     & D    & 4       & 3.5 \\
-        2 & D     & D    & 2       & 3.75 \\
-        3 & D     & C    & 1       & 2.87 \\
-        4 & C     & D    & 4       & 3.435 \\
-        5 & D     & D    & 2       & 2.715 \\
-        6 & D     & C    & 1       & 1.85875 \\
+        0 & C     & C    & 2       & 3  \\
+        1 & C     & D    & 3       & 2.5 \\
+        2 & D     & D    & 1       & 2.75 \\
+        3 & D     & C    & 0       & 1.87 \\
+        4 & C     & D    & 3       & 2.435 \\
+        5 & D     & D    & 1       & 1.715 \\
+        6 & D     & C    & 0       & 0.85875 \\
 \end{tabular}
+
+In this table we can see satisficing play and tit for tat ending up in a cycle.
+at $t=0$ satisficing gained to little reward for its ambition level $\alpha_t$,
+and therefore switches action in the next round. In $t=1$ saticificing play
+is satisfied, but int $t=2$ it no longer is so it switches but this even ends
+up being worse at $t=3$. After that the cycle resets at $t=4$, the cycle becomes
+most obvious when looking at the $\pi_t$ column.
 \section{Properties of satisficing play}

### 2016-06-28 multiAgentLearning/paper/article.tex
added a intermezzo

@@
 To demonstrate a potential use of this algorithm we will use at several example
-cases. The first one being the iterated prisoner's dilemma. Consider
-the following normal form game:
-
-\todo{add table caption}
+cases. The first one being the iterated prisoner's dilemma.
+\fbox{
+  \begin{minipage}{2.5in}
+    \subsection{A prisoner's dillema introduction}
+    The prisoner's dillema is one of the classical game thoery examples.
+    Consider the following normal form game which is a prisoner's dillema:
     \begin{tabular}{lll}
@@
     \end{tabular}
-
-\drafting The Nash equilibrium of this game lies on Defect,Defect. However, all but
+    The Nash equilibrium of this game lies on (Defect,Defect). However, all but
     that action are pareto optimal actions. So the Nash equilibrium would be the
@@
     or change wording}
-
+	\end{minipage}
+	}
 \drafting However strategies such as tit for tat or win stay lose shift already know

### 2016-06-29 multiAgentLearning/paper/article.tex
added a bunch of text

@@
 \section{Introduction}
+\label{sec:introduction}
 \toReview
@@
 \section{Core algorithm}
+\label{sec:core-algorithm}
 \toReview The core algorithm as presented by \citep{herbert} and formalized by
@@
 \todo{make an intermezzo of this?}
-\drafting
+\toReview
 To demonstrate a potential use of this algorithm we will use at several example
@@
   \begin{minipage}{2.5in}
+    \toReview
     \subsection{A prisoner's dillema introduction}
-    The prisoner's dillema is one of the classical game thoery examples.
+    The prisoner's dillema is one of the classical game theory examples.
     Consider the following normal form game which is a prisoner's dillema:
@@
     that action are pareto optimal actions. So the Nash equilibrium would be the
-    ``right'' strategy to play if the game is only played once, but if you repeat
-    the game against the same opponent, interestingly enough you can ``solve''
-    the prisoner's dilemma by ``teaching'' the opponent to cooperate.
-    \todo{can one really solve? Is it really teaching? should either add references
-    or change wording}
+    ``right'' strategy to play if the game is only played once. However if you repeat
+    the game against the same opponent you can ``solve''
+    the prisoner's dilemma, meaning not playing the Nash equilibrium,
+    but trying to get pareto optimal rewards. This
+    is often done by starting defecting as soon as your opponent starts
+    defecting, both the tit for tat and win stay lose shift strategies do this.
   \end{minipage}
 }
-\drafting However strategies such as tit for tat or win stay lose shift already know
+\toReview Strategies such as tit for tat or win stay lose shift already know
 which actions are winning and which are punishing to try and solve it.
@@
 
-\drafting In $t=0$ satisficing play doesn't meet its ambition of 4, therefore, in the
+\toReview In $t=0$ satisficing play doesn't meet its ambition of 4, therefore, in the
 next game it modifies its ambition and changes its action. Here we can see
@@
 
-\drafting So here satisficing play finds a Pareto optimal action, but it doesn't find
+\toReview So here satisficing play finds a Pareto optimal action, but it doesn't find
 the best possible action it could get against this opponent. From this example
 we can also see how dependent satisficing play is on its starting parameters.
+
 \todo{here we can interject an intermezzo, of how karandikar tried to solve
@@
 up being worse at $t=3$. After that the cycle resets at $t=4$, the cycle becomes
-most obvious when looking at the $\pi_t$ column.
+most obvious when looking at the $\pi_t$ column. Also note about this opponent
+that if saticificing play would've started with a lower aspiration, it would
+not enter this cycle and just keep on cooperating.
+\fbox{
+  \begin{minipage}{2.5in}
+    \toReview
+    \subsection{Karandikar's algorithm}
+    In the \nameref{sec:core-algorithm} section we said we would use Karandikar's
+    notation, but use the simplified version of stimpson.
+    Karandikar version would not necessarily change action when being
+    disappointed (aspiration level not met), but would switch with a
+    probability $p$, which he called \emph{inertia}.
+
+    He also introduced another model, which used random perturbations on
+    $\alpha_t$. This ``solved'', the initial state problem, however to prevent
+    the random trembles to run out of hand relative to the game rewards the tremble
+    generating function needed a feasible payoff interval $\Lambda$. Which
+    requires more knowledge of the game, breaking the initial assumptions
+    presented in the \nameref{sec:introduction}.
+  \end{minipage}
+}
+
+\section{A saticificing comparison}
+\drafting So, until now we've seen thoroughly how saticificing play works, and even some
+strengths and weaknesses. However,
+
+\todo{
+With help of the work of \citep{crandall} saticificing play is compared
+to Q-learning and WOLF-PHC.
+Q-learning is explained briefly with help of \citep{sandholm} and WOLF-PHC
+with help of \citep{bowling}.
+
+Finally some graphs are shown how they compare
+}
+
 \section{Properties of satisficing play}
@@
 
+\todo{
 Using both \citep{karandikar} and \citep{stimpson:2001} works to look deeper
@@
 also showing the contrast graph which supports the work of \citep{arthur}
-
+}
 \section{An extended satisficing play}
@@
 
+\todo{
 In here we discuss the work of \citep{stimpson:2003}, where he extends the
@@
 Stimpson created a game where resources had to be allocated among agents.
-
-\section{Satisficing compared to other learning methods}
-\drafting
-
-With help of the work of \citep{crandall} saticificing play is compared
-to Q-learning and WOLF-PHC.
-Q-learning is explained briefly with help of \citep{sandholm} and WOLF-PHC
-with help of \citep{bowling}.
-
-Finally some graphs are shown how they compare
+}
 
@@
 \email{j.t.klooster@students.uu.nl}
+
+%  LocalWords:  Karandikar

### 2016-06-29 multiAgentLearning/paper/article.tex
added a fuckton of text

@@
 
-\todo{here we can interject an intermezzo, of how karandikar tried to solve
-this with trajection and random perturbations}
-
-\drafting Now consider a game versus tit for tat (TT), Tit for Tat is a
+\toReview Now consider a game versus tit for tat (TT), Tit for Tat is a
 strategy that always copies the last action of its opponent.
@@
 \drafting So, until now we've seen thoroughly how saticificing play works, and even some
-strengths and weaknesses. However,
+strengths and weaknesses. However, what we really want to see is how it does
+compare against other MAL algorithms. So to do this we'll first introduce the
+competitors (Q-learning and WOLF-PHC) and then we'll show per game how they
+do in comparison with help of the work of \citep{crandall}.
 
-\todo{
-With help of the work of \citep{crandall} saticificing play is compared
-to Q-learning and WOLF-PHC.
-Q-learning is explained briefly with help of \citep{sandholm} and WOLF-PHC
-with help of \citep{bowling}.
+Q-Learning is described by \citep{sandholm} in the following way: ``The
+Q-learning algorithm works by estimating the values of state-action pairs. The
+value $Q(s,q)$ is the expected discounted sum of future payoffs obtained by taking
+action $a$ from state $s$ and following an optimal policy thereafter.
+Once these values have been learned, the optimal action from any state
+is the one with the highest Q-value. After initializing the Q-values to arbitrary
+numbers they are estimated based on experience as follows: 1. Select an action
+$a$ from the current state $s$, causing an immediate payoff $r$ and arriving at
+state $s'$. 2. Update $Q(s,a)$ based on this experience as follows:
+\[\Delta Q(s,a) = \lambda[r+\gamma \text{ max }Q(s',b)-Q(s,a)]\]
+Where $\lambda$ is the learning rate and $0 \leq \gamma < 1$ is the discount
+factor.
+3. Go to 1.
+'' A look-up table is used to store the Q-values.
 
-Finally some graphs are shown how they compare
-}
+\todo{explain WOLF-PHC better?}
+WoLF-PHC stands or win or learn fast policy hill climbing. It is 
+introduced as an extension to q-learning. However with the addition that
+learning rate should be increased when ``losing''. A more precise definition
+can be found in the paper of \citep{bowling}.
+
+The actual comparison of the algorithms is done by \citep{crandall}, in here
+we will discuss several of his results. All the algorithms are compared in self
+play. The first game result we'll discuss
+is the already introduced prisoner's dillema:
+\includegraphics[width=2.5in]{crandall/prisoners}
+
+Where the S-Algorithm is saticificing play as defined by stimpson, WoLF-PHC is
+defined as above and QL is Q-learning. We can clearly see saticificing play
+performing a lot better than the other two algorithms in this instance.
+(Note that in this figure different values were used than our example, but
+the basic structure of the game is the same).
+
+A game of chicken is a game where 2 players drive in a straight line in their
+cars straight towards each other. If both players keep on driving straight,
+they'll crash and get a bad reward. If one of them moves out of the way, he'll
+be ashamed, but be better of than crashing, the other one will of course get
+some fame, and if both players move out of the way, they'll share the same.
+Formally it looks like this:
+
+\begin{tabular}{lll}
+              & Bend & Straight \\
+    Bend & $(3,3)$     & $(2,3.5)$  \\
+    Straight & $(3.5,2)$     & $(1,1)$  \\
+\end{tabular}
+
+And here is the result of Crandall:
+\includegraphics[width=2.5in]{crandall/chicken}
+
+In here we see again saticificing play performing better than Q-Learning and
+WoLF. However, the difference is a lot less refined.
+
+Shapley's game is a 3 action choice game. If saticificing play has to change
+action, instead of picking the other action, it will just pick uniformly random
+of the 2 other actions. The context of the game is left to the imagination
+of the reader and the normal form game looks like this:
+
+\begin{tabular}{llll}
+              & a & b & c \\
+    A & $(0,0)$ & $(0,1)$     & $(1.0)$  \\
+    B & $(1,0)$ & $(0,0)$     & $(0.1)$  \\
+    C & $(0,1)$ & $(1,0)$     & $(0.0)$  \\
+\end{tabular}
+
+This is the result of crandall:
+\includegraphics[width=2.5in]{crandall/shapleys}
+
+In here we can see saticifing play performing a lot worse than the other two
+algorithms. This is because, Crandall explains, the joint aspiration never
+falls below the Pareto frontier, thus the behavior is essentially random.
 

### 2016-06-30 multiAgentLearning/paper/article.tex
with parameters

@@
 \section{Satisficing play in action}
-\todo{make an intermezzo of this?}
 \toReview
@@
 cases. The first one being the iterated prisoner's dilemma.
+
+\noindent
 \fbox{
@@
 not enter this cycle and just keep on cooperating.
+
+\noindent
 \fbox{
@@
 we will discuss several of his results. All the algorithms are compared in self
-play. The first game result we'll discuss
+play. The algorithms use the following parameters:
+
+\begin{tabular}{ll}
+  Satisficing & $\alpha_0\in \mathbb{P}_0, \lambda=0.99$ \\
+  WoLF-PHC & $\lambda=1/(10+0.001 k^a_s), \gamma = 0.95, \epsilon-\text{greedy w}$ 
+             $/\epsilon = \tex{max}(0.2-0.0006t,0)$ \\
+\end{tabular}
+
+The first game result we'll discuss
 is the already introduced prisoner's dillema:
@@
 
-Shapley's game is a 3 action choice game. If saticificing play has to change
-action, instead of picking the other action, it will just pick uniformly random
-of the 2 other actions. The context of the game is left to the imagination
-of the reader and the normal form game looks like this:
+Shapley's game is a 3 action choice game. The context of the game is left to
+the imagination of the reader and the normal form game looks like this:
 
@@
 
+\noindent
+\fbox{
+  \begin{minipage}{2.5in}
+    \drafting
+    \subsection{Multi-action saticificing}
+    In their paper \citep{stimpson:2003} introduce a modification to
+    saticificing play.
+    Rather than just picking the other action, which it can't do because there
+    are more than that available, it will pick uniformly across the other
+    actions. This change allows saticificing play to play in $n$-action games
+    rather than just 2 action games.
+  \end{minipage}
+}
+
+The multi-agent social dillema (MASD) was invented by \citep{stimpson:2003}. This game
+is impossible to express in normal form, since it consists of $N$ players (well
+it is possible if $N=2$, but otherwise not really).
+Each player has a choice of allocating $M$ units to either his own goal, 
+or the group goal. The Nash equilibrium of this game lies when all agents
+contribute all of the resources towards their own goal. But the Nash bargaining
+solution lies on contributing everything towards to group goal. We will present
+the results of \citep{crandall}, since \citep{stimpson:2003} didn't do bar-charts,
+although they also compared against Q-Learning and came towards similar results:
+
+\includegraphics[width=2.5in]{crandall/masd}
+
+In here we can see saticificing play outperforming Q-Learning and WoLF quite
+starkly. But we can also see that this result looks quite a lot like the
+prisoners dillema result. This is because the MASD was basically a generalization
+of the prisoners dillema, so this result isn't surprising.
+
 \section{Properties of satisficing play}
+To analyse saticfing play further we use the work of \citep{karandikar} and
+\citep{stimpson:2001}. Karandikar gave some good theoretical foundations,
+whereas stimpson did some deep emepircal experiments on self play in the
+prisoners dillema with parameter perturbations.
 \drafting
@@
 also showing the contrast graph which supports the work of \citep{arthur}
-}
-\section{An extended satisficing play}
-\drafting
-
-\todo{
-In here we discuss the work of \citep{stimpson:2003}, where he extends the
-algorithm to be played with mutliple action choices.
-Stimpson created a game where resources had to be allocated among agents.
 }

### 2016-06-30 multiAgentLearning/paper/article.tex
added more text

@@
 \abstract{
-  %TODO abstract
-  \todo[inline]{abstract missing}
+  This paper takes a closer look at satisficing play. We explain the algorithm
+  thoroughly and then look at empirical comparisons with Q-Learning and WoLF-PHC. From
+  this we can conclude that satisficing play is relatively
+  good at games with a cooperation aspect.
+  Then we take a closer look at it in the prisoner's dillema. Where we
+  discover that even with variance in the paremeters satisficing play ends up
+  cooperating most of the time in self play. We conclude that satisficing play
+  is an algorithm of serious interest to MAL researches.
 }
 \todo{be more critical? maybe it depends to much on its paremeters?
-  Or which game types are good for saticifing play?
+  Or which game types are good for satisficing play?
 }
-\todo{misses callouts, intermezzi and quotes}
+\todo{misses callouts and quotes}
 \section{Introduction}
@@
 
-\toReview The classical models from the field of economics had at least one of the
+The classical models from the field of economics had at least one of the
 following assumptions:
@@
 
-\toReview These assumptions can be called common knowledge, and are a demanding to
+These assumptions can be called common knowledge, and are a demanding to
 make \citep{binmore}.
 
-\todo{give paper overview}
+In this paper we'll take a closer look at satisficing play. First we'll introduce
+the core algorithm. Then we see how it works with several examples, which will
+give us a better intuition about the algorithm. After that we see how it
+compares to more ``classical'' MAL algorithms. Finally we dive deep into the
+theoretical properties of satisficing play, and look at some empirical results
+of satisficing play in the prisoner's dillema.
 
@@
 
-\toReview In the first phase the next action is selected. If the action still holds
+In the first phase the next action is selected. If the action still holds
 according to the satisficing level or ambition it will remain the same, if not another
@@
 \section{Satisficing play in action}
+\label{sec:playaction}
 \toReview
@@
 \section{A saticificing comparison}
-\drafting So, until now we've seen thoroughly how saticificing play works, and even some
+\toReview So, until now we've seen thoroughly how saticificing play works, and even some
 strengths and weaknesses. However, what we really want to see is how it does
@@
 we will discuss several of his results. All the algorithms are compared in self
-play. The algorithms use the following parameters:
-
-\begin{tabular}{ll}
-  Satisficing & $\alpha_0\in \mathbb{P}_0, \lambda=0.99$ \\
-  WoLF-PHC & $\lambda=1/(10+0.001 k^a_s), \gamma = 0.95, \epsilon-\text{greedy w}$ 
-             $/\epsilon = \tex{max}(0.2-0.0006t,0)$ \\
-\end{tabular}
-
+play. 
 The first game result we'll discuss
@@
 
-In here we can see saticifing play performing a lot worse than the other two
+In here we can see satisficing play performing a lot worse than the other two
 algorithms. This is because, Crandall explains, the joint aspiration never
@@
   \begin{minipage}{2.5in}
-    \drafting
-    \subsection{Multi-action saticificing}
+    \toReview
+    \subsection{$N$-action saticificing}
     In their paper \citep{stimpson:2003} introduce a modification to
@@
 starkly. But we can also see that this result looks quite a lot like the
-prisoners dillema result. This is because the MASD was basically a generalization
-of the prisoners dillema, so this result isn't surprising.
+prisoner's dillema result. This is because the MASD was basically a generalization
+of the prisoner's dillema, so this result isn't surprising.
 
 \section{Properties of satisficing play}
+\toReview
 To analyse saticfing play further we use the work of \citep{karandikar} and
 \citep{stimpson:2001}. Karandikar gave some good theoretical foundations,
-whereas stimpson did some deep emepircal experiments on self play in the
-prisoners dillema with parameter perturbations.
-\drafting
+whereas stimpson did empirical experiments on self play in the
+prisoner's dillema with parameter perturbations. However as said in the
+\nameref{sec:core-algorithm} it should be noted
+that the algorithms of Karandikar and Stimpson are different. The main
+difference is the lack of \emph{inertia} in the algorithm of Stimpson, which
+makes the behavior deterministic.
 
-\todo{
-Using both \citep{karandikar} and \citep{stimpson:2001} works to look deeper
-at the properties of satisficing play.
+Karandikar sums up his work quite thoroughly in  the informal discussion section
+of his paper. First of all ``   First, the untrembled joint process of
+aspirations and actions always converges to a pure strategy state.''
+This means that even without trembles, saticfing play with inertia
+will end up in a state wherein aspiration levels of both players equals payoff.
+The second result I want to discuss is:
+``The limit invariant distribution places
+almost all weight on the cooperative outcome, provided that the persistence
+parameter is sufficiently close to unity.''. Where the persistence parameter
+is the learning rate, and the invariant distribution is the trembled process,
+in other words, the saticfing play algorithm with trembles. For more details
+the reader is referred towards that paper.
 
-karandikar used a theoretical proofing method, whereas Stimpson showed some
-experimental results.
+Before the results of Stimpson are discussed we first have to consider the
+possible outcomes. It is important to keep in mind that in his algorithm
+the result of karandikar doesn't hold, because inertia is missing. Therefore
+not all processes will converge to a pure strategy state.
+the possible outcomes are:
+\begin{enumerate}
+  \item Convergence to certain actions.
+  \item Action cycles.
+  \item Complete chaos.
+\end{enumerate}
+We've seen both one and two in the section \nameref{sec:playaction}. Three
+is mentioned because there lacks a theoretical proof of it being impossible.
+Stimpson studied the behavior of saticfing play empirically in the prisoner's
+dillema, so over 5000 trials he ended up with the following result:
 
-also showing the contrast graph which supports the work of \citep{arthur}
+\includegraphics[width=2.5in]{stimpson/strategies}
+
+Where the leters stand for the actions, for example DC is (Defect,Cooperate).
+If letters are have dashes in between they indicate action cycles, for example:
+$DD-CC-DC$ is (Defect, Defect), (Cooperate, Cooperate) and 
+(Defect,Cooperate).
+From here we can see that saticfing play ends up cooperating most of the time.
+The variations are because the parameters were set randomly, such as initial
+aspiration, learning rate or even initial action.
+
+They also made another graph depicting how initial aspiration levels influenced
+the perecentage of the time ending up with mutual cooperation, this happened
+over 1000 trials:
+
+\includegraphics[width=2.5in]{stimpson/aspiration}
+
+Alfa is the initial aspiration of player 1, Beta of player 2, and the colors
+indicate the percentage of the time that cooperation happened. From this we can
+see that if the goal is cooperation, having initially high aspiration is good
+most of the time. You can also see some kind of intrusion towards the center,
+from which you could conclude that having similiar aspirations is good for
+cooperation. But the authors do not do this. It is also somewhat crenelated,
+probably because of the low number of trials they used to make this graph.
+
+To further study satisficing play in the prisoner's dillema, a generalization
+has to be made. Therefore stimpson introduces a generalized dillema that looks
+like the following normal form game:
+\begin{tabular}{lll}
+            & Cooperate   & Defect \\
+  Cooperate & $(\sigma,\sigma)$     & $(0,1)$  \\
+  Defect    & $(1,0)$     & $(\delta,\delta)$  \\
+\end{tabular}
+Where $(0<\delta<\sigma<1)$ and $\sigma>0.5$. Sigma $\sigma$ for cooperate and
+Delta $\delta$ for defect. The constraint $(0<\delta<\sigma<1)$ exists to make
+sure the game is a prisoner's dillema and $\sigma>0.5$ is used to ensure the
+Pareto front of the prisoner's dillema.
+The following graph illustrates the effect of different reward contrasts,
+and was created over 1000 trials:
+
+\noindent
+\includegraphics[width=2.5in]{stimpson/contrast}
+
+The colors indicate the percentage of time a run resulted in cooperation.
+According to the authors, cooperation is most likely when $\delta$ is small
+and $\sigma$ is large. If they are close together distinction between 
+defection and cooperation becomes difficult. This is typical of non-optimizing
+algorithms, and Arthur writes that his algorithm and even human behavior
+
+\multicolinterrupt{
+\begin{center}
+\large\emph{``Beyond the perceptual threshold, non-optimal outcomes become likely''}
+\end{center}
+\vspace{-2em}\hbox{}\hfill --- \citep{arthur}
 }
+
+``appear to discover and exploit the optimal action with high probability,
+\emph{as long as it is not difficult to discriminate}. But beyond a perceptual
+threshold, where differences in alternatives become less pronounced,
+non-optimal outcomes become more likely'' \citep{arthur}
+
+The final result presented by stimpson is the learning rate. The learning
+rate determines how quickly aspiration should be updated. A higher learning
+rate means that values of the past are less relevant. The following graph
+was made over 1000 trials:
+
+\noindent
+\includegraphics[width=2.5in]{stimpson/learning-rate}
+
+Where lambda is the learning rate. As we can see the percentage of cooperation
+(the number on the $y$-axis), increases with higher learning rate. But starkly
+drops once it reaches one and aspiration isn't updated at all.
+
+\section{Conclusion}
+\toReview
+
+We've seen how saticfing play works and some potential issues with it.
+Then we compared it against other multi-agent
+algorithm's empircally. Where we've
+seen it be better in a lot of famous games, but in others such as shapely it
+performed relatively more poorly. We've also seen it been extended to play
+with more actions and against more agents in the MASD.
+After that we looked closer to the theoretical properties
+of saticfing play, where we've seen it with the introduction of random elements
+convergence to the coopertive outcome becomes likely.
+Finally we've also seen empirical tests
+on the deterministic variant in the prisoner's dillema. From this we've seen
+that parameters are important, but most of the time it results in the coopertive
+outcome anyway.
+
+From this we can conclude that saticfing play is a serious algorithm to be of
+interest of MAL researches.
 

### 2016-06-30 multiAgentLearning/paper/article.tex
satisficing play blah

@@
-\abstract{
+\cleared \abstract{
   This paper takes a closer look at satisficing play. We explain the algorithm
-  thoroughly and then look at empirical comparisons with Q-Learning and WoLF-PHC. From
-  this we can conclude that satisficing play is relatively
+  thoroughly and then look at empirical comparisons with Q-Learning and
+  WoLF-PHC. From this we can conclude that satisficing play is relatively
   good at games with a cooperation aspect.
-  Then we take a closer look at it in the prisoner's dillema. Where we
-  discover that even with variance in the paremeters satisficing play ends up
+  After that a more thorough inspection is done with satisficing play in the
+  prisoner's dilemma. Where we
+  discover that even with variance in the parameters satisficing play ends up
   cooperating most of the time in self play. We conclude that satisficing play
@@
 }
-\todo{be more critical? maybe it depends to much on its paremeters?
-  Or which game types are good for satisficing play?
-}
-\todo{misses callouts and quotes}
+
 \section{Introduction}
 \label{sec:introduction}
-\toReview
+\cleared
 Satisficing play was originally introduced
@@
 give us a better intuition about the algorithm. After that we see how it
-compares to more ``classical'' MAL algorithms. Finally we dive deep into the
+compares to several other MAL algorithms. Finally we dive deep into the
 theoretical properties of satisficing play, and look at some empirical results
-of satisficing play in the prisoner's dillema.
+of satisficing play in the prisoner's dilemma.
 
@@
 \label{sec:core-algorithm}
-\toReview The core algorithm as presented by \citep{herbert} and formalized by
+\cleared The core algorithm as presented by \citep{herbert} and formalized by
 \citep{karandikar},
-in here we present the simplified version of \citep{stimpson:2001} first:
+in here we present the simplified version of \citep{stimpson:2001}:
 
@@
 \end{enumerate}
-\todo{add table caption}
+
 \begin{tabular}{ll}
@@
 \label{sec:playaction}
-\toReview
-To demonstrate a potential use of this algorithm we will use at several example
-cases. The first one being the iterated prisoner's dilemma.
+\cleared
+To demonstrate a potential use of this algorithm we will use it in several
+example cases. The first one being the iterated prisoner's dilemma.
 
@@
   \begin{minipage}{2.5in}
-    \toReview
-    \subsection{A prisoner's dillema introduction}
-    The prisoner's dillema is one of the classical game theory examples.
-    Consider the following normal form game which is a prisoner's dillema:
+    \cleared
+    \subsection{A prisoner's dilemma introduction}
+    The prisoner's dilemma is one of the classical game theory examples.
+    Consider the following normal form game which is a prisoner's dilemma:
     \begin{tabular}{lll}
@@
     The Nash equilibrium of this game lies on (Defect,Defect). However, all but
-    that action are pareto optimal actions. So the Nash equilibrium would be the
-    ``right'' strategy to play if the game is only played once. However if you repeat
-    the game against the same opponent you can ``solve''
+    that action are Pareto optimal actions. So the Nash equilibrium would be the
+    ``right'' strategy to play if the game is only played once. However if you
+    repeat the game against the same opponent you can ``solve''
     the prisoner's dilemma, meaning not playing the Nash equilibrium,
-    but trying to get pareto optimal rewards. This
+    but trying to get Pareto optimal rewards. This
     is often done by starting defecting as soon as your opponent starts
-    defecting, both the tit for tat and win stay lose shift strategies do this.
+    defecting, or ``punishing'' non-cooperation,
+    both the tit for tat and win stay lose shift strategies do this.
   \end{minipage}
 }
-\toReview Strategies such as tit for tat or win stay lose shift already know
-which actions are winning and which are punishing to try and solve it.
+\cleared Strategies such as tit for tat or win stay lose shift already ``know''
+which actions are winning and which are punishing when solving the iterated
+prisoner's dilemma. Therefore they introduce the knowledge about the game
+structure axiom discussed in the \nameref{sec:introduction}.
 Satisficing play learns this however by updating the aspiration it got from
@@
 
-\toReview In $t=0$ satisficing play doesn't meet its ambition of 4, therefore, in the
-next game it modifies its ambition and changes its action. Here we can see
-that satisficing play does get more utility than its ambition, and is satisfied
-with its action.
-But now if we start with a lower apsiration level:
+\cleared In $t=0$ satisficing play doesn't meet its ambition of 4, therefore,
+in the next game it modifies its ambition and changes its action.
+Here we can see that satisficing play does get more utility than its ambition,
+and is satisfied with its action.
+But now if we start with a lower aspiration level:
 
@@
 
-\toReview So here satisficing play finds a Pareto optimal action, but it doesn't find
+\cleared Here satisficing play finds a Pareto optimal action, but it doesn't find
 the best possible action it could get against this opponent. From this example
@@
 
-\toReview Now consider a game versus tit for tat (TT), Tit for Tat is a
+\cleared Now consider a game versus tit for tat (TT), Tit for Tat is a
 strategy that always copies the last action of its opponent.
-This game can also be found in the \citep{stimpson:2001} paper:
+This game can also be found in the \citep{stimpson:2001} paper, although
+this version is extended more:
 
@@
 
-In this table we can see satisficing play and tit for tat ending up in a cycle.
+\cleared In this table we can see satisficing play and tit for tat ending up in a cycle.
 at $t=0$ satisficing gained to little reward for its ambition level $\alpha_t$,
-and therefore switches action in the next round. In $t=1$ saticificing play
-is satisfied, but int $t=2$ it no longer is so it switches but this even ends
-up being worse at $t=3$. After that the cycle resets at $t=4$, the cycle becomes
+and therefore switches action in the next round. In $t=1$ satisficing play
+is satisfied, but in $t=2$ it no longer is satisfied, so it switches but this ends
+up being worse at $t=3$, after that the cycle resets at $t=4$. The cycle becomes
 most obvious when looking at the $\pi_t$ column. Also note about this opponent
-that if saticificing play would've started with a lower aspiration, it would
+that if satisficing play would've started with a lower aspiration, it would
 not enter this cycle and just keep on cooperating.
@@
   \begin{minipage}{2.5in}
-    \toReview
+    \cleared
     \subsection{Karandikar's algorithm}
     In the \nameref{sec:core-algorithm} section we said we would use Karandikar's
-    notation, but use the simplified version of stimpson.
+    notation, but use the simplified version of Stimpson.
     Karandikar version would not necessarily change action when being
@@
 
-    He also introduced another model, which used random perturbations on
+    \cleared He also introduced another model, which used random perturbations on
     $\alpha_t$. This ``solved'', the initial state problem, however to prevent
@@
     requires more knowledge of the game, breaking the initial assumptions
-    presented in the \nameref{sec:introduction}.
+    (somewhat) presented in the \nameref{sec:introduction}.
   \end{minipage}
@@
 
-\section{A saticificing comparison}
-\toReview So, until now we've seen thoroughly how saticificing play works, and even some
+\section{A satisficing comparison}
+\cleared Until now we've seen thoroughly how satisficing play works, and even some
 strengths and weaknesses. However, what we really want to see is how it does
@@
 
-Q-Learning is described by \citep{sandholm} in the following way: ``The
+\cleared Q-Learning is described by \citep{sandholm} in the following way: ``The
 Q-learning algorithm works by estimating the values of state-action pairs. The
@@
 3. Go to 1.
-'' A look-up table is used to store the Q-values.
+''. A look-up table is used to store the Q-values. The experiments use the
+following parameters for Q-Learning:
 
-\todo{explain WOLF-PHC better?}
-WoLF-PHC stands or win or learn fast policy hill climbing. It is 
+\begin{tabular}{ll}
+  Variable & Value \\
+  $\gamma$ & 0 \\
+  $\lambda$ & $1/(10+0.01 k_s^a)$ \\
+  $\epsilon$-greedy w/$\epsilon$ & 0.1
+\end{tabular}
+
+\cleared Where $k^a_s$ is the number of times that action $a$ has been played
+in state $s$.
+
+\cleared WoLF-PHC stands or win or learn fast policy hill climbing. It is 
 introduced as an extension to q-learning. However with the addition that
 learning rate should be increased when ``losing''. A more precise definition
-can be found in the paper of \citep{bowling}.
+can be found in the paper of \citep{bowling}. The experiments used the following
+parameters for WoLF-PHC:
+\begin{tabular}{ll}
+  Variable & Value \\
+  $\gamma$ & 0.95 \\
+  $\lambda$ & $1/(10+0.01 k_s^a)$ \\
+  $\epsilon$-greedy w/$\epsilon$ & max($0.2-0.0006 t, 0.1$)
+\end{tabular}
 
-The actual comparison of the algorithms is done by \citep{crandall}, in here
-we will discuss several of his results. All the algorithms are compared in self
-play. 
+\cleared The parameters that were used for satisficing play by Crandall were the
+following:
+
+\begin{tabular}{ll}
+  Variable & Value \\
+  $\alpha$ & Random payoff \\
+  $\lambda$ & 0.99 \\
+\end{tabular}
+
+\cleared The actual comparison of the algorithms is done by \citep{crandall},
+in here we will discuss several of his results.
+All the algorithms are compared in self play. 
 The first game result we'll discuss
-is the already introduced prisoner's dillema:
+is the already introduced prisoner's dilemma:
 \includegraphics[width=2.5in]{crandall/prisoners}
 
-Where the S-Algorithm is saticificing play as defined by stimpson, WoLF-PHC is
-defined as above and QL is Q-learning. We can clearly see saticificing play
+\cleared Where the S-Algorithm is satisficing play as defined by stimpson,
+WoLF-PHC is defined as above and QL is Q-learning.
+We can clearly see satisficing play
 performing a lot better than the other two algorithms in this instance.
@@
 
-A game of chicken is a game where 2 players drive in a straight line in their
+\cleared A game of chicken is a game where 2 players drive in a straight line in their
 cars straight towards each other. If both players keep on driving straight,
@@
 
-And here is the result of Crandall:
+\cleared This is the result presented by Crandall:
+
+\noindent
 \includegraphics[width=2.5in]{crandall/chicken}
 
-In here we see again saticificing play performing better than Q-Learning and
+\cleared In here we see again satisficing play performing better than Q-Learning and
 WoLF. However, the difference is a lot less refined.
 
-Shapley's game is a 3 action choice game. The context of the game is left to
+\cleared Shapley's game is a 3 action choice game. The context of the game is left to
 the imagination of the reader and the normal form game looks like this:
@@
 
-This is the result of crandall:
+\cleared This is the result of Crandall:
+
+\noindent
 \includegraphics[width=2.5in]{crandall/shapleys}
 
-In here we can see satisficing play performing a lot worse than the other two
+\cleared In here we can see satisficing play performing a lot worse than the other two
 algorithms. This is because, Crandall explains, the joint aspiration never
 falls below the Pareto frontier, thus the behavior is essentially random.
+Basically Q-Learning and WoLF learn to coordinate, however satisficing play
+can't do this, because it lacks the extensive memory structure the other two
+algorithms have.
 
@@
   \begin{minipage}{2.5in}
-    \toReview
-    \subsection{$N$-action saticificing}
+    \cleared  
+    \subsection{$N$-action satisficing}
     In their paper \citep{stimpson:2003} introduce a modification to
-    saticificing play.
-    Rather than just picking the other action, which it can't do because there
-    are more than that available, it will pick uniformly across the other
-    actions. This change allows saticificing play to play in $n$-action games
-    rather than just 2 action games.
+    satisficing play.
+    Rather than just picking the other action, it will pick uniformly across
+    the other actions. This change allows satisficing play to play in
+    $n$-action games rather than just 2 action games.
   \end{minipage}
@@
 
-The multi-agent social dillema (MASD) was invented by \citep{stimpson:2003}. This game
-is impossible to express in normal form, since it consists of $N$ players (well
-it is possible if $N=2$, but otherwise not really).
+\cleared The multi-agent social dilemma (MASD) was invented by
+\citep{stimpson:2003}. This game
+is difficult to express in normal form, since it consists of $N$ players.
 Each player has a choice of allocating $M$ units to either his own goal, 
@@
 solution lies on contributing everything towards to group goal. We will present
-the results of \citep{crandall}, since \citep{stimpson:2003} didn't do bar-charts,
-although they also compared against Q-Learning and came towards similar results:
+the results of \citep{crandall}, since \citep{stimpson:2003} didn't do
+bar-charts, although they also compared against Q-Learning and came towards
+similar results:
 
+\noindent
 \includegraphics[width=2.5in]{crandall/masd}
 
-In here we can see saticificing play outperforming Q-Learning and WoLF quite
+\cleared In here we can see satisficing play outperforming Q-Learning and WoLF quite
 starkly. But we can also see that this result looks quite a lot like the
-prisoner's dillema result. This is because the MASD was basically a generalization
-of the prisoner's dillema, so this result isn't surprising.
+prisoner's dilemma result. This is because the MASD was developed
+to try and capture the key features of the prisoner's dilemma,
+so this result isn't surprising.
 
 \section{Properties of satisficing play}
-\toReview
-To analyse saticfing play further we use the work of \citep{karandikar} and
+\cleared 
+To analyse satisficing play further we use the work of \citep{karandikar} and
 \citep{stimpson:2001}. Karandikar gave some good theoretical foundations,
 whereas stimpson did empirical experiments on self play in the
-prisoner's dillema with parameter perturbations. However as said in the
+prisoner's dilemma with parameter perturbations. However as said in the
 \nameref{sec:core-algorithm} it should be noted
 that the algorithms of Karandikar and Stimpson are different. The main
-difference is the lack of \emph{inertia} in the algorithm of Stimpson, which
-makes the behavior deterministic.
+difference is the lack of \emph{inertia} in the algorithm of Stimpson, without 
+it, the behavior becomes deterministic.
 
-Karandikar sums up his work quite thoroughly in  the informal discussion section
-of his paper. First of all ``   First, the untrembled joint process of
+\cleared Karandikar sums up his work quite thoroughly in  the informal
+discussion section of his paper. First of all
+``   the untrembled joint process of
 aspirations and actions always converges to a pure strategy state.''
-This means that even without trembles, saticfing play with inertia
-will end up in a state wherein aspiration levels of both players equals payoff.
+This means that even without trembles, satisficing play with inertia
+will end up in a state wherein aspiration levels of both players equals payoff
+per action.
 The second result I want to discuss is:
@@
 is the learning rate, and the invariant distribution is the trembled process,
-in other words, the saticfing play algorithm with trembles. For more details
-the reader is referred towards that paper.
+by trembled process he means the satisficing play algorithm with trembles.
+For more details the reader is referred towards that paper.
 
-Before the results of Stimpson are discussed we first have to consider the
-possible outcomes. It is important to keep in mind that in his algorithm
-the result of karandikar doesn't hold, because inertia is missing. Therefore
+\cleared Before the results of Stimpson are discussed we first have to consider
+the possible outcomes. It is important to keep in mind that in his algorithm
+the result of Karandikar doesn't hold, because inertia is missing. Therefore
 not all processes will converge to a pure strategy state.
@@
 \end{enumerate}
-We've seen both one and two in the section \nameref{sec:playaction}. Three
+\cleared We've seen both one and two in the section \nameref{sec:playaction}. Three
 is mentioned because there lacks a theoretical proof of it being impossible.
-Stimpson studied the behavior of saticfing play empirically in the prisoner's
-dillema, so over 5000 trials he ended up with the following result:
+Stimpson studied the behavior of satisficing play empirically in the prisoner's
+dilemma, the parameters used in all of his experiments were 
+uniformly selected per trial in the following ranges:
 
+\begin{tabular}{lll}
+  Variable & Min & Max \\
+  $\alpha, \beta$ & 0.5 & 2  \\
+  $\lambda$ & 0.1 & 0.9  \\
+  $\sigma$ & 0.51 & 1.0  \\
+  $\delta$ & 0.1 & $\sigma$
+\end{tabular}
+
+Where $\alpha, \beta$ are the initial aspiration levels of both players.
+$\lambda$ is the learning rate, $\sigma$ is the payoff for cooperation and
+$\delta$ for defection, more on that later.
+The initial actions were randomly selected.
+Over 5000 trials he ended up with the following result:
+
+\noindent
 \includegraphics[width=2.5in]{stimpson/strategies}
 
-Where the leters stand for the actions, for example DC is (Defect,Cooperate).
+\cleared Where the letters stand for the actions, for example DC is (Defect,Cooperate).
 If letters are have dashes in between they indicate action cycles, for example:
@@
 (Defect,Cooperate).
-From here we can see that saticfing play ends up cooperating most of the time.
+From here we can see that satisficing play ends up cooperating most of the time.
 The variations are because the parameters were set randomly, such as initial
@@
 
-They also made another graph depicting how initial aspiration levels influenced
-the perecentage of the time ending up with mutual cooperation, this happened
-over 1000 trials:
+\cleared They made another graph depicting how initial aspiration levels
+influenced the percentage of the time ending up with mutual cooperation,
+this happened over 1000 trials:
 
+\noindent
 \includegraphics[width=2.5in]{stimpson/aspiration}
 
-Alfa is the initial aspiration of player 1, Beta of player 2, and the colors
+\cleared Alpha is the initial aspiration of player 1, Beta of player 2, and the colors
 indicate the percentage of the time that cooperation happened. From this we can
@@
 most of the time. You can also see some kind of intrusion towards the center,
-from which you could conclude that having similiar aspirations is good for
-cooperation. But the authors do not do this. It is also somewhat crenelated,
-probably because of the low number of trials they used to make this graph.
+from which you could conclude that having similar aspirations is good for
+cooperation. But the authors do not do this and therefore we won't either.
+It is also somewhat crenelated, probably because of the low number of trials
+they used to make this graph.
 
-To further study satisficing play in the prisoner's dillema, a generalization
-has to be made. Therefore stimpson introduces a generalized dillema that looks
+\cleared To further study satisficing play in the prisoner's dilemma, a generalization
+has to be made. Therefore Stimpson introduces a generalized dilemma that looks
 like the following normal form game:
@@
 \end{tabular}
-Where $(0<\delta<\sigma<1)$ and $\sigma>0.5$. Sigma $\sigma$ for cooperate and
+\cleared Where $(0<\delta<\sigma<1)$ and $\sigma>0.5$. Sigma $\sigma$ for cooperate and
 Delta $\delta$ for defect. The constraint $(0<\delta<\sigma<1)$ exists to make
-sure the game is a prisoner's dillema and $\sigma>0.5$ is used to ensure the
-Pareto front of the prisoner's dillema.
+sure the game is a prisoner's dilemma and $\sigma>0.5$ is used to ensure the
+Pareto front of the prisoner's dilemma.
 The following graph illustrates the effect of different reward contrasts,
@@
 
-The colors indicate the percentage of time a run resulted in cooperation.
+\cleared The colors indicate the percentage of time a run resulted in cooperation.
 According to the authors, cooperation is most likely when $\delta$ is small
@@
 
-\multicolinterrupt{
-\begin{center}
-\large\emph{``Beyond the perceptual threshold, non-optimal outcomes become likely''}
-\end{center}
-\vspace{-2em}\hbox{}\hfill --- \citep{arthur}
-}
-
-``appear to discover and exploit the optimal action with high probability,
+\cleared ``appear to discover and exploit the optimal action with high probability,
 \emph{as long as it is not difficult to discriminate}. But beyond a perceptual
@@
 
-The final result presented by stimpson is the learning rate. The learning
+\cleared The final result presented by stimpson is the learning rate. The learning
 rate determines how quickly aspiration should be updated. A higher learning
@@
 
-Where lambda is the learning rate. As we can see the percentage of cooperation
+\cleared Where lambda is the learning rate. As we can see the percentage of cooperation
 (the number on the $y$-axis), increases with higher learning rate. But starkly
@@
 \section{Conclusion}
-\toReview
-
-We've seen how saticfing play works and some potential issues with it.
+\cleared 
+We've seen how satisficing play works and some potential issues with it.
 Then we compared it against other multi-agent
@@
 seen it be better in a lot of famous games, but in others such as shapely it
-performed relatively more poorly. We've also seen it been extended to play
+performed relatively poorly. We've also seen it been extended to play
 with more actions and against more agents in the MASD.
 After that we looked closer to the theoretical properties
-of saticfing play, where we've seen it with the introduction of random elements
+of satisficing play, where we've seen it with the introduction of random elements
 convergence to the coopertive outcome becomes likely.
 Finally we've also seen empirical tests
-on the deterministic variant in the prisoner's dillema. From this we've seen
+on the deterministic variant in the prisoner's dilemma. From this we've seen
 that parameters are important, but most of the time it results in the coopertive
@@
 
-From this we can conclude that saticfing play is a serious algorithm to be of
+\cleared From this we can conclude that satisficing play is a serious algorithm to be of
 interest of MAL researches.
 
-\todo{add caption for tables?}
 \bibliography{main}

### 2016-11-07 datamining/summery.tex
summery dataming, finished the trees

\documentclass[12pt,landscape]{extarticle}
\usepackage{multicol}
\usepackage{calc}
\usepackage{ifthen}
\usepackage[landscape]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage{color,graphicx,overpic}
\usepackage{hyperref}
\usepackage{graphicx}
\newcommand{\indep}{\rotatebox[origin=c]{90}{$\vDash$}}

\pdfinfo{
  /Title (EC cheatsheet)
  /Creator (TeX)
  /Producer (pdfTeX 1.40.0)
  /Author (Jappie Klooster)
  /Subject (Evulotionary computing)
  /Keywords (Evolutionary, Computing, Gentic, Algorithm, Heuristics, Strategy)}

% This sets page margins to .5 inch if using letter paper, and to 1cm
% if using A4 paper. (This probably isn't strictly necessary.)
% If using another size paper, use default 1cm margins.
\ifthenelse{\lengthtest { \paperwidth = 11in}}
    { \geometry{top=.5in,left=.5in,right=.5in,bottom=.5in} }
    {\ifthenelse{ \lengthtest{ \paperwidth = 297mm}}
        {\geometry{top=0.5cm,left=0.5cm,right=0.5cm,bottom=0.5cm} }
        {\geometry{top=0.5cm,left=0.5cm,right=0.5cm,bottom=0.5cm} }
    }

% Turn off header and footer
\pagestyle{empty}

% Redefine section commands to use less space
\makeatletter
\renewcommand{\section}{\@startsection{section}{1}{0mm}%
                                {-1ex plus -.5ex minus -.2ex}%
                                {0.5ex plus .2ex}%x
                                {\normalfont\large\bfseries}}
\renewcommand{\subsection}{\@startsection{subsection}{2}{0mm}%
                                {-0.8ex plus -.5ex minus -.2ex}%
                                {0.2ex plus .2ex}%
                                {\normalfont\normalsize\bfseries}}
\renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{0mm}%
                                {-0ex plus -.5ex minus -.2ex}%
                                {0.1ex plus .2ex}%
                                {\normalfont\small\bfseries}}
\makeatother

% Define BibTeX command
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

% Don't print section numbers
\setcounter{secnumdepth}{0}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0pt plus 0.5ex}

%My Environments
\newtheorem{example}[section]{Example}
% -----------------------------------------------------------------------

\begin{document}
\raggedright
\footnotesize
\begin{multicols}{3}

% multicol parameters
% These lengths are set only within the two main columns
%\setlength{\columnseprule}{0.25pt}
\setlength{\premulticols}{1pt}
\setlength{\postmulticols}{1pt}
\setlength{\multicolsep}{1pt}
\setlength{\columnsep}{2pt}

\begin{center}
     \Large{\underline{Datamining}} \\
\end{center}
\section{Classification trees}
\subsection{Building classification trees}
The objective of tree construction is to finally obtain nodes that are pure,
in the sense that they contain cases of a single class only.
\subsubsection{Impurity and quality of a split}
Impurity is a function:
\[i(t)=\phi(p_1,p_2,...,p_J)\]
where $p_j(j=1,...,J)$ is the relative frequency of the $J$ different classes.
$\phi$ is the impurity function with the following constraints:
\begin{enumerate}
  \item An impurity measure of a node should be at maximum when the observations
    are distributed evenly over all classes in that node.
  \item It should be at a minimum when all observations belong to a single class.
  \item $\phi$ is a symmetric function of $p_1,...,p_J$.
    a symmetric function of n variables is one whose value at any n-tuple of
    arguments is the same as its value at any permutation of that n-tuple.
\end{enumerate}
The quality of a split is defined as follows:
\[\Delta i(s,t)=i(t)-\phi(l)i(l)-\phi(r)i(r)\]
Where $\phi$ is a function that determines the proportion of cases for a node
and $l$ and $r$ are the left and right nodes of $t$.
\paragraph{Resubstitution error}
It measures what fraction of the class in a node is classified incorrectly if
we assign every case to the majority class in that node.
\paragraph{Gini} $i(t)=\sum_j p(j|t)(1-p(j|t))$
\paragraph{entropy}$i(t)=\sum_j p(j|t)log p(j|t)$

\subsubsection{Overfiting and pruning}
There are two strategies to avoid overfiting in tree construction:
Stopping rules, don't expand a node if impurity reduction below some threshold
and pruning, grow a large tree than later merge back nodes.
Stopping rules have a disadvantage that they sometimes require you to make
a weak split to get a good split.
\paragraph{Cost complexity pruning}
A problem with pruning is that the number of pruned subtree is $O(e^n)$.
So cost complexity pruning only considers trees that are
``the best of their kind''.
$T$ denotes the tree with leaf nodes $\overset{\sim}{T}$  and leaf node count
$|\overset{\sim}{T}|$
$R(T)$ denotes the fraction of cases in the training sample that are
misclassified by $T$, the re-substitution error of $T$.
We define the total cost: $C_\alpha(T)=R(T)+\alpha |\overset{\sim}{T}|$
The total cost of tree $T$ consists of the $R(T)$ and a penalty for complexity
$\alpha|\overset{\sim}{T}|$ where $\alpha$ is the parameter that determines the
complexity penalty.
a complex tree that has no errors may have a higher cost than a mall tree with
a number of errors depending on the value of $\alpha (\geq 0)$.
$T_{max}$ denotes the large tree that should be pruned to the right size.
If we fix the value of $\alpha$, therer is a smallest minimizing subtree $T(\alpha)$
of $T_{max}$ that fulfills:
$C_\alpha(T(\alpha))=min T \leq T_{max} C_\alpha(T)$ which says that there is no
subtree of $T_[max]$ with lower cost than $T(\alpha)$, and
If $C_\alpha (T) = C_\alpha(T(\alpha))$ then $T(\alpha) \leq T$ which says that
fi there is a tie then we pick the smallest tree (the one that is a subtree of
all other that achieve the minimum).
It can be shown that for every value of $\alpha$ there is  such a smallest
minimizing subtree. But this is non trivial.
Even though $\alpha$ goes trough a continum of values there is only a finite
numer of sutrees of $T_{max}$. We can construct a descreasing sequence of
subtrees of $T_{max}$: $T_1 > T_2 > ... > {t_1} $ where $t_1$ is the root node,
such tat $T_k$ is the smallest minimizing subtree for
$\alpha \in (\alpha_k, \alpha_{k+1})$. $T_1$ is the smallest subtree of $T_{max}$
with the same resubstitution error as $T_{max}$ ($T_1=T(\alpha=0)$).
the algorithm to find $T_1$ is: find any pair of leaf nodes with a common parent
that can be merged back withoutincreasing te resubtition error. Continue until
no more such pair can be found.
\[g(t)=\frac{R(t)-R(T_{1,t})}{|\overset{\sim}{T}_{1,t}|-1}\]
is a function to calculate the next value for $\alpha$ at which $T_k - T_t$
becomes better than $T_k$.
The algorithm for cost complexity is as follows: We start by computing $T_1$ from $T_{max}$
by using $\alpha = 0$. Then we repeat until we reach the root node:
For each node $t$ in he current tree $T_k$ we compute $g_k(t)$. Then we prune
$T_k$ in all nodes for which $g_k$ achieves the minimum to obtain $T_{k+1}$.
\subsubsection{Selection of the final tree}
$R^{ts}(T)$ denotes the estimated error rate of tree $T$. You can use
the entire set and just pick the one with the lowest estimated error.
There is also cross validation, which can be used if the data set is smaller.

\subsection{Missing data}
The easiest way is  to throw away records with missing data.
This has the following disadvantages: Potential bias and loss of power (which
means that our precision is reduced because we loose a lot of examples).
If we want to use a tree with incomplete data we have the following problems:
How to determine the quality of a spiet? Which way gets an observation send
with a missing value?
If the class labe misses its useless anyway so it should be thrown away.
A solution is ignoring the quality aspect and add surogate splits.
\subsection{Computational efficiency}
A categorical attribute with $L$ is distinct values has $2^{L-1}-1$ possible
distinct splits.
For binary class problems and impurity measures that belong to class $\mathcal{F}$
there is a more efficient algorithm for finding the optimal split.
$p(0|x=b_l)$  denotes the realitve frequency of class 0 for observations in the
current node with $x=b_l$. Order $p(0|x=b_l)$. Then one of the $L$ subsets
is the optimal split. Thus the search is reduced to $O(L-1)$
\subsubsection{Tree construction}
   Asuming all $p$ attributes $x_1, ..., x_p$ are numeric we have to sort
   them per attribute, which takes $O(n log n)$ time.
   Best case (where each split is a perfect subdivision) we traverse the tree
   and have to consider $O(n)$ splits per level resulting in $O(n log n)$ time.
   Worst case it becomes $O(n^2)$
\section{Graphical models}
\subsection{Probability rules}
sum rule $P(X)=\sum_y P(X,Y)$, product rule $P(X,Y)=P(X|Y)P(Y)$.
\subsection{Independence}
Random vectors $X$ and $Y$ are independent iff $P(x,y)=P(x)P(y) \text{ for all } (x,y)$
Therefore $P(x|y)=P(x)$. Independence can be written as $X \indep Y$
\section{Baysian networks}
\section{Freq item set mining}
\end{multicols}
\end{document}

### 2016-11-08 datamining/summery.tex
summery stuff

@@
 \pdfinfo{
-  /Title (EC cheatsheet)
+  /Title (Datamining cheatsheet)
   /Creator (TeX)
@@
 sum rule $P(X)=\sum_y P(X,Y)$, product rule $P(X,Y)=P(X|Y)P(Y)$.
+To estimate the joint distrubtion of $X$ and $Y$ we use $\overset{\wedge}{P}(x,y)=\frac{n(x,y)}{n}$
 \subsection{Independence}
 Random vectors $X$ and $Y$ are independent iff $P(x,y)=P(x)P(y) \text{ for all } (x,y)$
-Therefore $P(x|y)=P(x)$. Independence can be written as $X \indep Y$
+Therefore $P(x|y)=P(x)$. Independence can be written as $X \indep Y$.
+This gives the factorisation criterion for independent randomvectors:
+random vectors $X$ and $Y$ are independent iff there exists two functions $g$
+and $h$ such that: $P(x,y) = g(x)h(y)$ for all $(x,y)$.
+Random vectors $X$ and $Y$ re independent given $Z$ iff
+$P(x,y,z)=P(x,z)P(y,z)/P(z)$.
 \section{Baysian networks}

### 2016-11-08 datamining/summery.tex
added more graph stuff

@@
 \paragraph{Gini} $i(t)=\sum_j p(j|t)(1-p(j|t))$
-\paragraph{entropy}$i(t)=\sum_j p(j|t)log p(j|t)$
+\paragraph{entropy}$i(t)=\sum_j p(j|t)\log p(j|t)$
 
@@
    Asuming all $p$ attributes $x_1, ..., x_p$ are numeric we have to sort
-   them per attribute, which takes $O(n log n)$ time.
+   them per attribute, which takes $O(n \log n)$ time.
    Best case (where each split is a perfect subdivision) we traverse the tree
-   and have to consider $O(n)$ splits per level resulting in $O(n log n)$ time.
+   and have to consider $O(n)$ splits per level resulting in $O(n \log n)$ time.
    Worst case it becomes $O(n^2)$
@@
 Therefore $P(x|y)=P(x)$. Independence can be written as $X \indep Y$.
-This gives the factorisation criterion for independent randomvectors:
+This gives the factorization criterion for independent random vectors:
 random vectors $X$ and $Y$ are independent iff there exists two functions $g$
 and $h$ such that: $P(x,y) = g(x)h(y)$ for all $(x,y)$.
+A ``log version'' is: $\log P(x,y)=g'(x)+h'(y)$ for all $(x,y)$.
 Random vectors $X$ and $Y$ re independent given $Z$ iff
+$P(x,y|z)=P(x|z)P(y|z)$ for all $(x,y)$ and for all $z$ for which $P(z)>0$.
+This is the same as:
 $P(x,y,z)=P(x,z)P(y,z)/P(z)$.
+With marginal independence we can state a simple factorization criterion
+to establish conditional independence: random vectors $X$ and $Y$ are
+conditionally independent given $Z,X \indep Y | Z$ iff there exists
+functions $g$ and $h$ such that $P(x,y,z)=g(x,z)h(y,z)$.
+\subsection{Independence graphs}
+Conditional independence relations can be represented as a conditional
+independence graph. Let $X=(X_1,X_2,\dots, X_k)$ be a $k$-dimensional random
+vector. The conditional independence graph of $X$ is the undirected graph
+$G=(K,E)$ with $K={1,2,...,k}$ and where ${i,j}$ is not in the edge set $E$ 
+iff $X_i \indep X_j|rest$.
+Global markov property: $a$ seperates $b$ from $c$ ($a,b,c$ disjoint) iff
+$X_b \indep X_c | X_a$ where $X_a=(X_i;i\in a)$ and $a$ seperates $b$ from
+$c$ if for all $i \in b, j\in c:$ $a$ seperates $i$ from $j$.
+Is equal to local markov property: $X_i \indep$ rest | boundary$(i)$ where
+the boundary of a vertex $i$ is simply the set of adjecent vertices.
+\subsection{Log linear models}
+\subsubsection{for binary data}
+A random experiment that only disguises between two possible outcomes is
+called a \emph{Bernoulli} experiment. Random variable $X$ denotes the number of
+successes in a Bernoulli experiment; $X$ therefore has possible values 0 and 1.
+The probability distribution of $X$ is completely determined by the probablity of
+success which we denote by $p$  and is $P(X=0) = 1 - p $ and $ P(X=1)=p$.
+$X$ has the probability function $P(x)=p^x(1-p)^{1-x}$ for $x=0,1$ and
+$0 \leq p \leq 1$.
+log properities: $\log a^b=b \log a$, $\log ab = \log a + \log b$,
+$\log \frac{a}{b}=\log a - \log b$.
+The cartesion product 
+${0,1} \times {0,1}$ creates a table of the bivariate Bernoulli random
+vector $(X_1,X_2)$.
+\begin{tabular}{llll}
+  $P(x_1,x_2)$ & $x_2=0$ & $x_1=1$ & Total \\
+  $x_1=0$ & $p(0,0)$ & $p(0,1)$ & $p_1(0)$ \\
+  $x_1=1$ & $p(1,0)$ & $p(1,1)$ & $p_1(1)$ \\
+  Total & $p_2(0)$ & $p_2(1)$ & $1$ \\
+\end{tabular}
+which can be written as
+$P(x_1,x_2)=p(0,0)^{(1-x_1)(1-x_2)}p(0,1)^{(1-x_1)x_2}p(1,0)^{x_1(1-x_2)}p(1,1)^{x_1x_2}$
+for $x_1=0,1$ and $x_2=0,1$.
+Taking logarithms of this identity for $P$ and collecting terms in$x_1$ and
+$x_2$ gives:
+$\log P(x_1,x_2)= \begin{matrix}
+    \log p(0,0) + x_1 \log \frac{p(1,0)}{p(0,0)} +  \\
+    x_2 \log \frac{p(0,1)}{p(0,0)} + x_1 x_2 \log \frac{p(1,1)p(0,0)}{p(0,1)p(1,0)}
+    \end{matrix}
+$
+reparameterizing the right hand side leads tothe so-caled \emph{log-linear expension}
+$\log P(x_1,x_2)=u_\emptyset+ x_1u_1 + x_2 u_2 + x_1 x_2 u_{12}$
+the $u$ coefficients are called $u$-terms, they substitue the log part.
+The coefficient of the product $x_1x_2$ is the logarithm of the cross product
+ratio:
+$u_{12}=\log cpr(X_1,X_2)$
+The random variables $X_1$ and $X_2$ are independent iff $u_{12}=0$
+If $u_{12}=0$ we can take $g(x_1)=u_\emptyset + x_1 u_1$ and $h(x_2)= u_2 x_2$.
+This is not posible with $u_{12}\neq 0$.
+We can enforce (conditional) independence constraints by etting the right
+$u$-terms to zero.
+\subsection{non binary data}
+To work with non binary data we make the u terms functions rather than constants:
+$log P(x_1,x_2)=u_\emptyset + u_1(x_1) + u_2(x_2) + u_{12}(x_1,x_2)$
+This produces to many parameters, and to identify them we have to impose
+an extra constraint: $u_a(x_a)=0$ whenever $x_i=0$ and $i\in a$. Here we
+assume that if $x_i$ has $d_i$ possible values, these are numbered
+$0,1,...,d_i-1$, but not ordered.
+\subsection{Hierarchical and graphical}
+% TODO: perhaps delete this section?
+The log-linear expension of the crossclassified multinominal density function
+$P_K$ is $\log P_K(x)=\underset{a \subseteq K}{\sum} u_a(x_a)$ where the sum is
+taken over all possible subsets of $K={1,2,...,k}$ and where the $u$-terms are
+coorinate projection functions, so that $u_a(x)=u_a(x_a)$ and also satisfy the
+constraint that $u_a(x)=0$ whenever $x_i = 0$ and $i\in a$.
+\subsubsection{Independence and the u-terms}
+$If(X_a,X_b,X_c)$ is a partitioned multinominal random vector then
+$X_b \indep X_c|X_a$ if and only if all u-terms in the log-linear expension with
+one or more coordinate in $b$ and one or more coordinate in $c$ are zero.
+The proof is direct application of the factorisation theorom for conditional
+independence.
+\subsubsection{Graphical model}
+Given an independence graph $G=(K,E)$ the cross-classified multinomal
+distrubtion for the random vector $X$ is a graphical model for $X$ if the
+distribution of $X$ is an arbitrary apart from constraints of the form that for
+all pairs of coordinates not in the edge set $E$ of $G$ the $u$ terms containing
+the selected coordinates are identically zero.
+\subsection{Maximum likelhood estimation}
+The maximum likelihood estimator of gaphical log linear model $M$ satisfies the
+likelihood equations: $\overset{\wedge}{n}_a=N\overset{\wedge}{P}_a=n_a$
 \section{Baysian networks}

### 2016-11-09 datamining/summery.tex
finished graphical models

@@
-\documentclass[12pt,landscape]{extarticle}
+\documentclass[10pt,landscape]{extarticle}
 \usepackage{multicol}
@@
     {\ifthenelse{ \lengthtest{ \paperwidth = 297mm}}
-        {\geometry{top=0.5cm,left=0.5cm,right=0.5cm,bottom=0.5cm} }
-        {\geometry{top=0.5cm,left=0.5cm,right=0.5cm,bottom=0.5cm} }
+        {\geometry{top=0.5cm,left=0.2cm,right=0.2cm,bottom=0.5cm} }
+        {\geometry{top=0.5cm,left=0.2cm,right=0.2cm,bottom=0.5cm} }
     }
@@
 likelihood equations: $\overset{\wedge}{n}_a=N\overset{\wedge}{P}_a=n_a$
+\subsubsection{Iterative proportional fitting}
+Is a converging algorithm
+Start with a uniform table. Then set the row margins to the fitted counts
+and distribute them proportionally. Then do the same for the columns, keep
+doing this until the values converged.
+
+\subsection{Decomosable graphical models}
+Decomposable models are graphical models that have explicit formulas for the
+maximum likelihood estimates.
+This is a convenient property from a computational viewpoint.
+If we only have to fit one model this is perhaps not so important,
+but when we have little prior knowledge we typically have to search a
+potentially large space of possible models.
+Decomposable models are very easy to characterize by their independence graphs.
+They have triangulated independence graphs: their independence graphs have no
+chordless cycles of length greater than three. A cycle is called chordless
+if no other than successive pairs of vertices in the cycle are adjacent. 
+\subsection{Deviance and likelhood ratio test}
+The deviance of a fitted model compares the log-likelihood of the fitted model
+to the log-likelihood of the saturated model.
+The larger the model deviance, the poorer the fit.
+The likelihood of a model $M$ is
+$L(\overset{\wedge}{P}(x);n(x))=\underset{x}{\prod}\overset{\wedge}{P}(x)^{n(x)}$
+and the log likelihood of a model $M$ is
+$\underset{x}{\sum} n(x)\log \overset{\wedge}{P}(x)$
+$L^i=L(\overset{\wedge}{P}^{M_i})$ is the value of the log likelihood function
+valutated at $\overset{\wedge}{P}^{M_i}$; the ML estimates of $P$ under $M_i$.
+Let $M_0 \subseteq M_1$; ie $M_0$ can be obtained from $M_1$ by setting
+additional $u$-terms to zero (more restrictions). The deviance between $M_0$ and
+$M_1$ is $dev(M_0) - dev(M_1) = 2(L^1 - L^0)$.
+For large $N$ $2(L^1-L^0) \approx M_0 \chi^2_v$ where the degrees of freedom $v$
+is equal to the number of additional restrictions of $M_0$. 
+We reject the null hypothesis that $M_0$ is the true model when:
+$2(L^1-L^0)>\chi^2_{v;\alpha}$. 
 \section{Baysian networks}
+
 \section{Freq item set mining}

### 2016-11-09 datamining/summery.tex
added definitions of baysian

@@
 \section{Baysian networks}
-
-\section{Freq item set mining}
+\subsection{Definition and notation}
+A directed graph is a pair $G = (K, E)$, where $K$ is a set of vertices and $E$
+is a set of edges with ordered pairs of vertices.
+If there is an arrow from $i$ to $j$,
+then we write this as $i \to j$, or equivalently as $(ij) \in E$.
+We restrict attention to directed graphs with no directed cycles,
+i.e. acyclic directed graphs (DAGs). If $i \to j$,
+then $X_i$ is called a parent of $X_j$ ,
+and $X_j$ is called a child of $X_i$ .
+The set of coordinates of the parents of $X_j$ is denoted $pa(j)$, so
+$X_{pa(j)}$ denotes the set of parents of $X_j$ .
+If there is a directed path from $i$ to $j$,
+then $X_i$ is called an ancestor of $X_j$ .
+The set of coordinates of the ancestors of $X_j$ is denoted an $(j)$.
+The absence of any directed cycles is equivalent to the existence of an
+ordering of the nodes ${1, 2, \dots , k}$ such that $i \to j$ only when $i < j$.
+In other words, there exists an ordering of the nodes such that arrows point
+only from lower-numbered nodes to higher-numbered nodes.
+Suppose that a priori knowledge tells us the variables can be labeled
+$X_1 , X_2 , \dots , X_k$ such that $X_i$ is prior to $X_{i+1}$.
+Corresponding to this ordering we can factorize the joint density of
+$X_1 , X_2 , \dots , X_k$ as $P(X)=P(X_1)P(X_2|X_1)\dots P(X_k|X_{k-1},X_{k-2},\dots,X_1)$
+In contructing a DAG, an arrow is drawn from $i$ to $j$, where $i<j$, unless
+$P(X_j|X_{j-1}\dots,X_1)$ does not depend on $X_i$.
+This is the key difference between DAGs and undirected graphs.
+In both types of graph a missing edge between $X_i$ and $X_j$ is equivalent to
+a conditional independence relation between $X_i$ and $X_j$.
+In undirected graphs they are conditionally independent given all the remaining
+variables, whereas in DAGs they are conditionally independent given all prior
+variables. 
+We can write the joint density:
+$P(X)=\underset{i=1}{\overset{k}{\prod}}P(X_i|X_{pa(i)})$.
+and the conditional indepencence as: $i \indep j|an({i,j})$
 \end{multicols}

### 2016-11-09 datamining/summery.tex
fixed geometry and added most baysian stuff I think

--- 2016-11-08 datamining/summery.tex
@@
 % If using another size paper, use default 1cm margins.
-\ifthenelse{\lengthtest { \paperwidth = 11in}}
-    { \geometry{top=.5in,left=.5in,right=.5in,bottom=.5in} }
-    {\ifthenelse{ \lengthtest{ \paperwidth = 297mm}}
-        {\geometry{top=0.5cm,left=0.5cm,right=0.5cm,bottom=0.5cm} }
-        {\geometry{top=0.5cm,left=0.5cm,right=0.5cm,bottom=0.5cm} }
-    }
+\geometry{top=0.2cm,left=0.2cm,right=0.2cm,bottom=0.2cm}
 
@@
 likelihood equations: $\overset{\wedge}{n}_a=N\overset{\wedge}{P}_a=n_a$
+\subsubsection{Iterative proportional fitting}
+Is a converging algorithm
+Start with a uniform table. Then set the row margins to the fitted counts
+and distribute them proportionally. Then do the same for the columns, keep
+doing this until the values converged.
+
+\subsection{Decomosable graphical models}
+Decomposable models are graphical models that have explicit formulas for the
+maximum likelihood estimates.
+This is a convenient property from a computational viewpoint.
+If we only have to fit one model this is perhaps not so important,
+but when we have little prior knowledge we typically have to search a
+potentially large space of possible models.
+Decomposable models are very easy to characterize by their independence graphs.
+They have triangulated independence graphs: their independence graphs have no
+chordless cycles of length greater than three. A cycle is called chordless
+if no other than successive pairs of vertices in the cycle are adjacent. 
+\subsection{Deviance and likelhood ratio test}
+The deviance of a fitted model compares the log-likelihood of the fitted model
+to the log-likelihood of the saturated model.
+The larger the model deviance, the poorer the fit.
+The likelihood of a model $M$ is
+$L(\overset{\wedge}{P}(x);n(x))=\underset{x}{\prod}\overset{\wedge}{P}(x)^{n(x)}$
+and the log likelihood of a model $M$ is
+$\underset{x}{\sum} n(x)\log \overset{\wedge}{P}(x)$
+$L^i=L(\overset{\wedge}{P}^{M_i})$ is the value of the log likelihood function
+valutated at $\overset{\wedge}{P}^{M_i}$; the ML estimates of $P$ under $M_i$.
+Let $M_0 \subseteq M_1$; ie $M_0$ can be obtained from $M_1$ by setting
+additional $u$-terms to zero (more restrictions). The deviance between $M_0$ and
+$M_1$ is $dev(M_0) - dev(M_1) = 2(L^1 - L^0)$.
+For large $N$ $2(L^1-L^0) \approx M_0 \chi^2_v$ where the degrees of freedom $v$
+is equal to the number of additional restrictions of $M_0$. 
+We reject the null hypothesis that $M_0$ is the true model when:
+$2(L^1-L^0)>\chi^2_{v;\alpha}$. 
 \section{Baysian networks}
-\section{Freq item set mining}
+\subsection{Definition and notation}
+A directed graph is a pair $G = (K, E)$, where $K$ is a set of vertices and $E$
+is a set of edges with ordered pairs of vertices.
+If there is an arrow from $i$ to $j$,
+then we write this as $i \to j$, or equivalently as $(ij) \in E$.
+We restrict attention to directed graphs with no directed cycles,
+i.e. acyclic directed graphs (DAGs). If $i \to j$,
+then $X_i$ is called a parent of $X_j$ ,
+and $X_j$ is called a child of $X_i$ .
+The set of coordinates of the parents of $X_j$ is denoted $pa(j)$, so
+$X_{pa(j)}$ denotes the set of parents of $X_j$ .
+If there is a directed path from $i$ to $j$,
+then $X_i$ is called an ancestor of $X_j$ .
+The set of coordinates of the ancestors of $X_j$ is denoted an $(j)$.
+The absence of any directed cycles is equivalent to the existence of an
+ordering of the nodes ${1, 2, \dots , k}$ such that $i \to j$ only when $i < j$.
+In other words, there exists an ordering of the nodes such that arrows point
+only from lower-numbered nodes to higher-numbered nodes.
+Suppose that a priori knowledge tells us the variables can be labeled
+$X_1 , X_2 , \dots , X_k$ such that $X_i$ is prior to $X_{i+1}$.
+Corresponding to this ordering we can factorize the joint density of
+$X_1 , X_2 , \dots , X_k$ as $P(X)=P(X_1)P(X_2|X_1)\dots P(X_k|X_{k-1},X_{k-2},\dots,X_1)$
+In contructing a DAG, an arrow is drawn from $i$ to $j$, where $i<j$, unless
+$P(X_j|X_{j-1}\dots,X_1)$ does not depend on $X_i$.
+This is the key difference between DAGs and undirected graphs.
+In both types of graph a missing edge between $X_i$ and $X_j$ is equivalent to
+a conditional independence relation between $X_i$ and $X_j$.
+In undirected graphs they are conditionally independent given all the remaining
+variables, whereas in DAGs they are conditionally independent given all prior
+variables. 
+We can write the joint density:
+$P(X)=\underset{i=1}{\overset{k}{\prod}}P(X_i|X_{pa(i)})$.
+and the conditional indepencence as: $i \indep j|an({i,j})$
+\subsection{Interpretation}
+Given a DAG we construct the moral graph $G^m$ by marrying parents and deleting
+directions, that is:
+For each $i \in K$ we connect all vertices in $pa(i)$ with lines, we replace all
+arrows in $E$ with lines.
+\subsection{Maximum likelyhood estimation}
+\subsubsection{ML estimation for Multinomal distribution}
+We want to estimate the probabilities $p1 , p2 , \dots , p_J$
+of getting outcomes $1, 2, \dots , J$ If in $n$ trials,
+we observe n1 outcomes $1, n_2$ of $2, \dots, n_J$ of $J$,
+then the obvious guess is to estimate $p_j , j = 1, \dots , J$, by $n_j /n$.
+This is also the maximum likelihood estimate.
+\subsection{ML estimation for bayesian nets}
+The joint probability for $n$ independent observation in log likelyhood becomes
+$L=\underset{i=1}{\overset{k}{\sum}}\underset{x_i,x_{pa(i)}}{\sum}n(x_i,x_{pa(i)})
+\log p(x_i|x_{pa(i)})$
+where $p$ stands for network parameters, $n(x_i,x_{pa(i)})$ for the number of
+obsevations with $X_i=x_i$ and $X_{pa(i)}=x_{pa(i)}$.
+If the parameters are not related this is just a bunch of independent
+multinomalestimation problems. So we can say
+$\overset{\wedge}{p}(x_i|x_{pa(i)})=\frac{n(x_i,x_{pa(i)})}{n(x_{pa(i))}}$,
+which can be substituted in the $L$ formula.
+\subsection{Estimation from incomplete data}
+We paritition the complete data in an observed part $X_{obs}$ and a missing part
+$X_{miss}$. For observation $j$ we write $X^{(j)}=(X_{obs}^{(j)},X_{mis}^{(j)})$.
+We want to find those parameter values that maximize the probability of the
+observed data.
+This means that if some values are missing,
+we have to obtain the marginal probability of the observed data by summing out
+the missing data. For observation $j$,
+the probability thus is:
+$P(X^{j}_{obs})=\underset{X^{(j)}_{mis}}{\sum}P(X^{(j)})$
+For all observations it becomes a product.
+Direct maximization of the observed data likelihood is complicated because we get
+a sum of parameters inside a log.
+There is an algorithm iterative scheme to compute the ML estimate called
+expectation maximization, it alternates between expectation step and maximization
+steps. 
+\subsection{Model selection with complete data}
+To prevent overfitting we introduce a penalty. $L^M$ denotes the value of the
+log-likelyhood function evaluated at $\overset{\wedge}{p}^M$; the ML estimates
+of $p$ under model $M$. Akaikes information criterion would ive the follwoing
+quality form model $M$: $AIC(M)=-2L^M+2dim(M)$ where $dim(M)$ is thenumber of
+parameters of the model. which we divide by $-2$ to get higher value for higher
+quality: $AIC(M)=L^M-dim(M)$. AIC gives a relativly low penalty for complex
+models therefore Baysian infromation criterion is more popular:
+$BIC(M)=L^M-\frac{\log n}{2} dim(M)$
+So now we have a well-defined optimzation problem: Given training data, scoring
+function (BIC), space of possible models (all DAG's) find a network that
+maximaizes the score.
+This is however NP hard therefore heuristics such as hillclimbing are used.
+The score is decomposable, its a sum of terms where each term contains the
+variables $i \cup pa(i)$. This means we only have to recalculate for variables
+of which the parent set has changed.
+\section{Freq paternset minig}
 \end{multicols}

### 2016-11-09 datamining/summery.tex
Finished with freq pattern mining

@@
-\documentclass[10pt,landscape]{extarticle}
+\documentclass[9pt,landscape]{extarticle}
 \usepackage{multicol}
@@
 \newcommand{\indep}{\rotatebox[origin=c]{90}{$\vDash$}}
+\usepackage{amssymb}
+\usepackage{algorithmicx}
+\usepackage{algpseudocode}
 
@@
 Impurity is a function:
-\[i(t)=\phi(p_1,p_2,...,p_J)\]
+$i(t)=\phi(p_1,p_2,...,p_J)$
 where $p_j(j=1,...,J)$ is the relative frequency of the $J$ different classes.
 $\phi$ is the impurity function with the following constraints:
-\begin{enumerate}
-  \item An impurity measure of a node should be at maximum when the observations
-    are distributed evenly over all classes in that node.
-  \item It should be at a minimum when all observations belong to a single class.
-  \item $\phi$ is a symmetric function of $p_1,...,p_J$.
-    a symmetric function of n variables is one whose value at any n-tuple of
+An impurity measure of a node should be at maximum when the observations
+are distributed evenly over all classes in that node,
+it should be at a minimum when all observations belong to a single class,
+$\phi$ is a symmetric function of $p_1,...,p_J$.
+A symmetric function of n variables is one whose value at any n-tuple of
 arguments is the same as its value at any permutation of that n-tuple.
-\end{enumerate}
 The quality of a split is defined as follows:
@@
 \section{Freq paternset minig}
+\subsection{Association rules: The binary case}
+The traditional setting for association rules consists of a set of transactions
+in which each transaction is a set of items.
+Translated to a relational setting,
+we have a table db with schema $R = {I_1 , \dots, I_n }$,
+in which each $I_i$ is a binary attribute.
+The attributes correspond with the items,
+the rows in the table with the transactions
+; a row has value 1 for an attribute if and only if the transaction contains
+that item.
+For $X,Y \subseteq R$, with $X\cap Y=\emptyset$ let $s(X)$ denote the support of
+$X$, the number of tuples that have value 1 for all items in $X$. for an
+association rule $X \to Y$ define the support is $s(X\cup Y)$, the confidence is
+$s(X\cup Y)/s(X)$
+The problem is to find all association rules that match or exceed the user
+defined lower thresholds for confidence, \emph{minconf}, and support
+\emph{minsup}.
+In a major abuse of notation,
+you could say that the support of X (when expressed as a fraction) corresonds
+to $P (X)$ in probability terms,
+and the confidence of $X \to Y$ with the conditional probability $P (Y | X)$.
+There are two tesholds we have to satisfy:
+find all sets $Z$ whoese support exceeds the minimal threshold. These sets are
+called frequent. Then test for all non-empty subsets $X$ of frequent sets $Z$
+wheter the rul $X \to Z \setminus X(=Y)$ holds with sufficient confidence.
+\subsubsection{A priori alogirhtm}
+Finding all frequents sets is problematic since all sets of $R$ are
+$|\mathcal{P}(R)|=2^{|R|}$. But since $Z$ can only be frequent if all its
+(non-empty) subsets are frequent we can do it.
+This is known as the A-Priori property. In other words, we can search level wise
+for the frequent sets.
+The level is the number of items in the set. Denote by $C(i)$ the sets of $i$
+items that are potentially frequent (the candidate sets) and by $F (i)$
+the frequent sets of $i$ items.
+Start with level one all entries of $R$ (in random order),
+filter on the level $C(i)$ each entry
+that doesn't have enough support, these are the frequent items $F(i)$.
+Then $i$ is increased, and we start creating the
+next level of candidates $C(i)$ based on $F(i-1)$ (which we just made).
+Then they do this:
+For each $X \in F (i − 1)$ do
+For each $Y \in F (i − 1)$ that shares the first $i − 2$ items with X do
+$C(i) := C(i) \cup {X \cup Y }$
+If we assume the db is sparse the complexity becomes $O(|R|^{k+1})$ where $k$
+is the maximal size of frequent items sets that is $k << |R|$.
+To generate association rules:
+\emph{flatmap} each freq set $X$ \emph{with} each non empty $Y \subset X$ filtering
+$s(X)/s(X\setminus Y)\geq minconf$ as $X \setminus Y \to Y $
+\subsection{Drowining in assoc rules}
+\subsubsection{Postprocessing the result set}
+We can order the resultss we can use conidence and support: they define a
+partial order on the set of rules. Given rules $r_1$ and $r_2, r_1A$ iff 
+$s(r_1) \leq s(r_2) \wedge conf(r_1) < con (r_2)$ or
+$s(r_1) < s(r_2) \wedge conf(r_1) \leq conf(r_2)$
+A rule is interesting if it gives usefull infromation. Lots of measures have
+been defined as an attempt to do that: lift, interest, conviction, colective
+strength, gain, gini, entropy, $\chi^2$ This is lift:
+$lift(A\to C)=\frac{P(C|A)}{P(C)}=\frac{P(A,C)}{P(A)P(C)}$
+The lift should be bigger than one for a rle to be itneresting.
+\subsubsection{Generating less rules}
+Closed frequent itemsets are itemsets that completely characterise their
+associated set of transactions.
+That is, a frequent itemset $I$ is closed if $I$ contains all items that occur in
+all transactions in the support of $I$.
+For an itemset $I$, denote by $\sigma(I)$ the set of tuples in which all items
+in $I$ are bought, that is $\sigma(I)=\{t \in db | \forall i \in I, i \in t\}$
+For a set of transaction $T$ let $f(T)$ denote the set of items that are bought
+in all transaction in $T$, that is $f(T)=\{i \in R | \forall t \in T, i \in t\}$
+The closure can be obtained by $c(I)=f(\sigma(I))$.
+To give a procedural description of applying the closure operator to an itemset
+$I$:
+first get the transactions in which all items in $I$ are bought,
+and then see whether there are any more items that are common to all these
+transactions.
+If so, add them to $I$ and return the result.
+Clearly, $I \subseteq c(I)$ and $I$ has the same support as $c(I)$.
+An itemset $I$ is closed if and only if $c(I) = I$.
+An itemset $I$ is a closed frequent itemset iff it is both frequent and closed.
+The A-Close algorithm for finding al frequent closed itemsets consists of 2
+phases: One, discover all freq closed itemsets in db. Derive all frequent
+itemsets from the freqeunt closed itemsets found in phase 1.
+In phase 1 the algorithm determines a set of so-called generators that will
+produce all frequent closed itemsets by application of the closure operator $c$.
+An itemset $I$ is a generator of a closed itemset $J$ if it is one of the
+smallest itemsets with $c(I) = J$.
+A-Close performs a levelwise search like Apriori:
+$G_{i+1}$ (the set of generators at level $i + 1$) is constructed using $G_i$.
+Using their support, and the support of their $i$-subsets in $G_i$,
+infrequent itemsets and itemsets that have the same support as one of their
+subsets are deleted from $G_{i+1}$. The rational of the secon pruning rule is
+that if an itemset has a subset with the same suppport, then this subset also
+has the same closure.
+Now that the generators have been determined, we compute their closures,
+where $c(I)=\cap t \in db: I \subseteq t$. Different generators may produce the
+same closure, duplicates are removed.
+In phase 2 we determine all frequent itemsets and their support.
+We use the following properties:
+1. All maximal frequent itemsets are closed.
+2. Every frequent itemset is a subset of a maximal frequent itemset.
+3. The support of an itemset equals the support of the smallest closed itemset
+in which it is contained (its closure).
+Our strategy is therefore to select the maximal itemsets,
+generate all their sub- sets,
+and determine their support using the third property.
 \end{multicols}

### 2016-11-09 datamining/summery.tex
Added a large portion of frequent pattern mining

@@
 of which the parent set has changed.
+You have the reverse opperation because hillclimbing only looks one step ahead,
+therefore adding and deleting can't emulate it.
 \section{Freq paternset minig}
@@
 and determine their support using the third property.
+\section{Frequent pattern mining}
+$\Sigma$ is the set of labels, $s=s_1,_2,\dots,s_n$ where $s_i \in \Sigma$,
+prefix $s[1:i]=s_1,s_2,\dots,s_1, 0 \leq i \leq n$. suffix $s[i:n]$.
+Let $s=s_1,s_2\dots s_n$ and $r=r_1,r_2,\dots r_m$ be two sequences over
+$\Sigma$. We say $r$ is a subsequence of $s$ denoted $r \subseteq s$, if there
+exists a one-to-one mapping $\phi:[1,m]\to[1,n]$ such that $r[i]=s[\phi(i)]$ and
+$i< j \Rightarrow \phi(i)<\phi(j)$.
+Each position in $r$ is mapped to a position in $s$ with the same label and the
+orer of labels is preserved. There may however be interverning gapsbetween
+consecutive elements of $r$ in the mapping.
+Given a database $D={s_1,\dots,s_N}$ of $N$ sequences and given some sequenc $r$
+the suport is defined as the total number of seuences in $D$ that contain $r$:
+$sup(r)=|\{s_i\in D:r\subseteq s_i\}|$. Given a minimum support threshold minsup,
+compute $\mathcal{F}(minsup,D)=\{r|sup(r)\leq minsup\}$.
+For two sequences $r_1$ and $r_2$ we have $r_1 \subseteq r_2 \Rightarrow
+sup(r_1) \leq sup(r_2)$ becuase $\forall s \in D: r_2 \subseteq s \Rightarrow
+r_1 \subseteq s$
+Hence in a level wise search for frequent sequences there is no point in
+expending infrequent ones.
+\subsection{GSP algorthm}
+Perform level wise search. Don't extend infrequent sequences.
+Candiadate generation for level $k+1$: take two frequent sequences $r_a$ and
+$r_b$ of length $k$ with $r_a[q:k-1]=r_b[1:k-1]$ and generate pre-candadate
+$r_{ab}=r_a+r_b[k]$. Pre-candidate $r_{ab}$ becomes a candidate (has to be
+counted) if all its subsequences of length $k$ are frequent. Note that we allow
+$r_a = r_b$.
+\subsection{Node labeled Graph}
 \end{multicols}

### 2016-11-10 datamining/summery.tex
Freq pattern mining

@@
-\documentclass[9pt,landscape]{extarticle}
+\documentclass[8pt,landscape]{extarticle}
 \usepackage{multicol}
@@
 \subsection{Node labeled Graph}
+A node labled graph is a quadruple $G=(V,E,\Sigma,L)$ where: $V$ is the set of
+nodes, $E$ is the set of edges, $\Sigma$ is the set of labels, and $\L: V->
+\Sigma$ is a labeling function.
+A labeled rooted unordered tree $U = (V , E , \Sigma, L, v^r )$ is an $G$ with a
+special node $v^r$ called the root of the tree such that there exists exactly
+one path betweeen the root node and any other node in $V$.
+A labeled rooted rodered tree $T = (V, E, L, v^r, \leq)$ is an unordered tree
+$U$ where between all the iblings an orderer $\leq$ is dfined. to every node in
+an ordered tree a preorder $(pre(v))$ number is assigned according to the depth
+first traversel of the tree.
+\subsubsection{Tree inclusion relations}
+An induced subtree preserves parent child relationships.
+$\pi(v)$ denotes the parent of node $v$.
+Given two oreered trees $D$ and $T$ we call $T$ an induced subtree of $D$ if
+there exists an injective matching function $\phi$ of $V_T$ into $V_d$ with the
+following conditions: $\phi$ preservers the labels: $L_T(v)=L_D(\phi(v))$,
+$\phi$ preservers the parent child relation:
+$v_i = \pi_T(v_j)\Leftrightarrow \phi(v_i)=\pi_D(\phi(v_j))$
+$\phi$ preservers the left to right orderbetween the nodes: $pre(v_i) <
+pre(v_j ) \Leftrightarrow pre(\phi(v_i)) < pre(phi(v_j))$  
+An induced subtreee $T$ can be obtained from a tree $D$ by rpeatedly removing
+leaf nodes, or possibly the root node if it has only one child.
+This definition is the same for embedded subtree, except the equal sign become
+$\in$ in the relation constiant and $\pi$ talks about the ancestor set.
+An embedded subtree preserves ancestor decendent realtionships. 
+\subsubsection{Freq tree mining task}
+Given a database of trees $D=\{d_1, d_2, \dots, d_n\}$ and a tree inclusion
+relation $\preceq$ we define the suppot of a tree $T$ as
+$sup(T,D)=|\{d\in D | T \preceq d\}|$. Given a minimum support treshehold
+minsup, compute $\mathcal{F}(minsup,D)=\{T|sup(T,D)\leq minsup\}$
+\emph{Anti monoticity property}: For a database of trees $D$ and two trees $T_1$
+and $T_2$ we have $T_1 \preceq T_2 \Rightarrow sup(T_1, D) \leq sup(T_2, D)$,
+because $\forall d \in D: T_2 \preceq d \Rightarrow T_1 \preceq d$
+Hence in a level wise search fro frequenttreees there is no point in expanding
+infrequent trees.
+\subsubsection{Right most extension}
+Let $T_k$ denote a tree of size $k$.
+Consider the node numbering of $T_k$ according to pre-order (depth-first)
+traversal of the tree.
+The right-most branch of the tree is the path from the root node to the
+right-most leaf (i.e. the node with number $k$).
+To expand the tree $T_k$,
+it is only allowed to add a node as the right-most child of a node on
+the right-most branch of $T_k$.
+This node gets number $k + 1$,
+as it is the last node in the pre-order traversal of $T_{k+1}$.
+The right-most extension technique generates each tree at most once.
+Also, the right-most expansion technique generates every possible tree, so each
+tree is generated exactly once.
+\emph{occurene list}:
+To determine whether a pattern tree occurs in a data tree, an occurrence list
+is maintained that contains the list of nodes in the data tree to which the
+nodes in the pattern tree can be mapped.
+FREQT only stores the nodes of the data tree to which the right-most node in
+the pattern tree can be mapped.
+This is sufficient since only the nodes on the right-most branch are needed
+for future extension.
+\subsection{Frequent pattern and clasification}
+Frequent pattern mining can also be used to extract features for classification
+tasks: 1 Find frequent patterns per class.
+2 Define discriminating patterns, for example,
+as patterns that are frequent in one class but not in the other.
+3 Use the presence/absence of such a discriminating pattern as a (binary)
+feature for constructing a classifier (e.g. classification tree!).
+4 In this way we can include non-tabular data (sequences, trees, graphs)
+into an algorithm that requires a tabular data structure.
 \end{multicols}

### 2016-11-10 datamining/summery.tex
Fitted count stuff

@@
 sum rule $P(X)=\sum_y P(X,Y)$, product rule $P(X,Y)=P(X|Y)P(Y)$.
-To estimate the joint distrubtion of $X$ and $Y$ we use $\overset{\wedge}{P}(x,y)=\frac{n(x,y)}{n}$
+To estimate the joint distrubtion of $X$ and $Y$ we use
+$\overset{\wedge}{P}(x,y)=\frac{n(x,y)}{n}$
+\subsection{Fitted counts}
+These are simply obtained by multiplying
+the estimated probabilities with the total number of observations. 
+To determine whether the independence assumption is justified, we compare the
+observed counts with the fitted counts of the independence model.
 \subsection{Independence}

### 2016-11-10 datamining/summery.tex
Final notes for exam

@@
 
+It is not correct to only consider splits on the border between
+segments in the presence of a minleaf constraint.
 \subsubsection{Overfiting and pruning}
@@
 doing this until the values converged.
-
+The equation for the cells than is:
+$\overset{\wedge}{n}(x)^{\{s+1\}}=\overset{\wedge}{n}(x)^{\{s\}}\frac{n_1(x_1)}{\overset{\wedge}{n}(x)^{\{s\}}}$
+where $s$ is the current step in ipf. The margins are replaced for every
+iteration by the desired value, whereas the others are calulated, this gives
+convergene because you alternate.
 \subsection{Decomosable graphical models}
@@
 therefore adding and deleting can't emulate it.
+\subsection{Chow and lius algorithm}
+that maximizes log likelyhood score:
+Compute the mutual information between each pair of variables.
+Make a complete (undirected) graph where each node corresponds to a variable.
+Put the mutual information between a pair of variables as a weight on the edge
+between that pair. Compute a maximum weight spanning tree.
+Pick a root node, and let all the edges point away from the root.
 \section{Freq paternset minig}
@@
 and determine their support using the third property.
+A-close is more efficient then Apriori on dense, highly correlated data sets.
 \section{Frequent pattern mining}
@@
 into an algorithm that requires a tabular data structure.
+\section{Social networks}
+What progblem do we run into when we want to classify the objects in the test
+set of social networks?
+To predict the class label of one object,
+we need the class labels of the linked objects,
+because they determine the value of the link-attributes.
+But the class labels of these linked objects are also unknown.
+To bootstrap the whole process, we can start with
+an initial classification that only uses the node-attributes and disregards the
+link- attributes.
+Starting from this initial classification we can iteratively improve
+the solution.
 \end{multicols}
